### 2019-03-05 README.Rmd
updates setting

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "Title of your paper goes here"
```

# master.paper

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/LiYingWang/master.paper/archive/master.zip>

Or you can install this compendium as an R package, master.paper, from GitHub with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("LiYingWang/master.paper")
```
### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2019-03-05 analysis/paper/paper.Rmd
updates setting

---
title: "Title Goes Here"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(master.paper) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

# Introduction

Here is a citation [@Marwick2017]

# Background

```{r demo, eval = FALSE}
1 + 1
```

# Methods

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
git2r::repository(here::here())
```

### 2019-03-24 analysis/paper/paper.Rmd
read in data
updates for plot

@@
 
-```{r demo, eval = FALSE}
-1 + 1
+```{r read-in-the-data, eval = FALSE}
+library(here)
+kwl_upper <- readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
 ```
@@
 ```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- readr::read_csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))
+# summarise the ornaments by layers
+library(tidyverse)
+library(ggplot2)
+
+# L5 and L6 represent pre-European contact period, 
+# L4 was the time of European contact, and 
+# L2 and L1 was the period of Chinese contact.
+
+ornaments_period <- 
+kwl_upper %>% 
+  filter(!is.na(`6-layers`)) %>% 
+  mutate(period = case_when(
+    `6-layers` %in% 1:2 ~ "Chinese Contact",
+    `6-layers` == 4 ~ "European Contact",
+    `6-layers` %in% 5:6 ~ "Before European Contact",
+     TRUE ~ "other"
+  )) %>% 
+  filter(period != "other")
+
+# need to reorder, try https://forcats.tidyverse.org/reference/index.html
+# reorder factor for categories (by frequency) and period (manually)
+
+ornaments_count <-
+  ornaments_period %>% 
+  filter(!is.na(`6-layers`)) %>% 
+  group_by(period, Categories) %>%
+  count() %>% 
+  ungroup() %>% 
+  mutate(Categories = fct_lump(as.factor(Categories), n = 5)) %>% 
+  mutate(Categories = fct_reorder(Categories, n))
+
+ggplot(transform(ornaments_count,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Categories, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency")
 ```

### 2019-03-24 analysis/paper/paper.Rmd
stash

--- 2019-03-05 analysis/paper/paper.Rmd
@@
 
-library(master.paper) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
 ```
@@
 # Results
-
-```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- readr::read_csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))
-```
 

### 2019-03-24 testingfile-1234.Rmf.Rmd
stash

Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

Liying Wang
Ben Marwick

```{r}
library(ggplot2)
ggplot(iris, 
       aes(Sepal.Width, 
           Sepal.Length)) +
  geom_point()
```

Introduction and Research Question

Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display (Joyce 2005). Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appear in Taiwan around 1,800 years ago through regional trade network in Southeast Asia (Wang and Liu 2007). Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the Metal Age and reached a high peak of quantity and variety in the early 17th century when the Europeans arrived and involved the island in an international trade network on a large scale. Foreign trade goods introduced into the local indigenous societies in an European colonial or imperial contexts usually had an impact on local culture and societies in many parts of the world that usually involved in the transformation of indigenous economic, cultural, and socio-political systems (Dietler 1997; Junker 1993; Mitchell 2000; Silliman 2005). Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that stress active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices (Given 2004; Mullins 2011; Silliman 2001; Rubertone 2000; Torrence 2000; Torrence and Clarke 2000; Voss 2005).

Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced. Direct colonial rule was rare, but the question of long-lasting indirect impacts on local communities remains unanswered. Indirect effects of colonialism is addressed recently to discuss the impact on the indigenous societies in the periphery of colonial control (Trabert 2017). Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where Northeastern Taiwan is an especially unique example for discussing indirect effects due to its physically isolated location that made it not easily accessible to the Europeans that the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012). In addition, prior to the European contact, there were small-scale regional exchange networks between China and Taiwan. Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century. This made northeastern Taiwan a good case to explore culture contacts at different periods and their indirect impacts by taking a long-term perspective.

We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan.  Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact. We hypothesise that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan. We hypothesise that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.

Cultural context of Ornaments in Northeastern Taiwan

Historical records from Spanish and Dutch provide rich information about the historical and cultural background both for Europeans and indigenous people in the early 17th century. Spanish occupied and built forts in Northern Taiwan since 1624, from which Spanish missionaries often went to indigenous settlement and left behind numerous records from those places. The report of the Dominican Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) as a materials to exchange with necessities they need with other indigenous people. This trend then gradually spread to other areas, and even the Spanish soldiers used agate as bargaining chips for gambling, since agate equaled money (Li 2006: 132-149).  European contact was focused on trading with indigenous communities in northeastern Taiwan. Trading activity of many types of goode among these indigenous communities increased with the arrival of European colonisers. Thus, we predict that this trading activity resulted in an increase in the amount and diversity of ornaments coming into indigenous communities in northeastern Taiwan.
Father Jacinto Esquivel recorded how people in Tamsui entered marriages. A man who wanted to take a wife had to give her parents cuentas. If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled (Li 2006: 151). Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals (Pao 2008: 122, 143, 151). Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity. They put a mat into the grave to protect the body from moisture and then put cooked rice on both sides of the head as provision. The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need. The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave (Li 2006: 153). 
Documents from the Qing dynasty usually describe the decorative purposes of the ornaments. In the Kavalan Subprefecture Gazetteer (Chen 2006[1852]: 308), the Kavalan Zhi Lue (Ko 1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao 1996[1829]: 77), it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments. Their number was small and their value high, hence only the rich possessed them. Ordinary people wore agate beads or glass beads on their head or neck during various festivals. In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori (Kanori 1996: 227-232) made field surveys among various ethnic groups in Taiwan. He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.
Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans (Liu 2008: 133-134). Interview records describe agate beads as valuable objects passed down from mother to daughter. When they were not used, they were often put on roof beams so that they were not seen by other people. But the origin of these beads is still not known. Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer (Hsu 1992: 22) and carp-shaped ornaments had completely disappeared. Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination. It was usually the female shamans that were responsible for divination rites, a technique passed down by ancestors. Most aborigines today don’t know the origin of agate beads any more, but their worth is still highlighted by oral history and through their scarcity.

Those historical documents from the early 17th century to the modern ethnography described how local indigenous people use those ornaments in local cultural contexts that represents some social roles or status. However, compared to European contact period, there is less mention of beads in Chinese contact period and the description about ornaments is only limited to their dressing culture.

Kiwulan
Kiwulan (Figure 1) is located at Yilan, north of Lanyang Plain in northeastern Taiwan, surrounded by the Central Mountain Range in the west and the Pacific Ocean in the east. Kiwulan was excavated from 2001 to 2004, and the total excavation area was 3,814 m2. Most beads were found during the excavation and some of them through screens with 20 mm and 1.5 mm mesh. The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures. The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates[a], which coordinate to the late Iron Age and Proto-historical period in Taiwan. This paper focuses on the Upper Layer Culture which overlaps with the contact periods with the Europeans and Chinese. It provides a good examples to discuss the foreign impacts on local indigenous society through body ornaments that usually convey social and cultural meaning by displaying.    
The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local settlements were attacked by the Spanish according to the official documents (add citation; Borao Mateo 2009). Later in 1647, the Dutch attacked the indigenous villages in Yilan and forced them to accept colonial rules and economic demands by paying annual tribute (add citation; Borao Mateo 2009). According to Dutch census reports in 1650, Kiwulan was the biggest indigenous community in the Yilan Plain, with a population of 840 adults (Chen 2007; Kang 2012; Li 2014). The diagnostic artifacts during European contact found at Kiwulan including An-ping jars and glass beads that were largely introduced to Taiwan during the early 17th century. 
The European colonization ended in 1662 when they were defeated by the kingdom of Tungning found by Koxinga from China. Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century. The evidence of Chinese migration can be identified from Chinese official records [b]and large amount of Chinese blue-and-white porcelain found at Kiwulan.[c]
The ornaments sampled in this study come from 40 adjacent 4m x 4m units out of 262 units in total since those sampled units were less disturbed by modern constructions and remains intact. The Upper Layer Culture can be divided into six sub layers spanning from 14 century to 19 century according to current radiocarbon dates, excavation depth, types of ceramics, and sediment texture and color (Hsieh 2008; Wang 2011). Layers are assigned numbers from 1 to 6 for which L1 corresponds to the upper context and L6, also the bottom layer, refers to lower context. Based on Hsieh’s (2008) chronological study of Chinese porcelains, L5 and L6 represent pre-European contact period, L4 was the time of European contact, and L2 and L1 was the period of Chinese contact. 
In general, intact ornaments with clear context were mostly found in burials that helps to understand how people wear them and any relationships between shapes of ornaments and gender or age. In addition to burials, ornaments were also found in middens and living space that can be identified by post holes. The frequency of ornaments from burials is more than those from the living space due to well preserved condition. There is a wide variety of raw materials for ornaments  (Table 1) of which ornaments glass, agate and metal ornaments that account for 99% of all ornaments were thought to be introduced by trading with the Europeans and Chinese traders, while ornaments made of shells, woods, and bones were believed locally made that is about only about 1% of all ornaments. The reason for the large quantity of glass beads is that they usually have been found in clusters because of the original form such as necklaces. The shape of ornaments include beads, bracelets, rings, bells, pendants, and knitted objects that shows shape are closely related to raw material due to its characteristic. For example, beads are only made of glass, agate, and shell, and bone is the main material for knitted objects that used to connect beads. Also, metal ornaments covers more shapes such bells, pendants, and rings  (Chen 2007).

Results

chronological distribution
* Compare the three analytical units

spatial distribution
* Compare spatial distribution of ornaments three analytical units

Discussion and Conclusion 
European contact

Chinese contact

________________

Notes on drafting

What is the hypothesis?
1. Ornaments in the archaeological record at Kiwulan indicate cultural changes after European contact 
2. Ornaments in the archaeological record at Kiwulan also indicate cultural change after Chinese contact

Background:

some debate about origins of social inequality in Taiwan… KWL as a important case study… centre of the debates…

H1
   1. We compare ornaments UCL-Pre-E and LCL
   2. Raw material, shape/type, spatial distribution, chronological distribution
H2
   1. We compare ornaments UCL-Pre-E and UCL-Post-E
   2. Raw material, shape/type, spatial distribution, chronological distribution
Results

Discussion and conclusion:

* Social values, gender, burials

Relevant papers to cite

https://www.tandfonline.com.sci-hub.tw/doi/full/10.1080/0734578X.2018.1471655?scroll=top&needAccess=true

________________

A Research of the Imported Ornaments Excavated from the Kiwulan site, I-lan
Introduction
Culture contact is a big issue which explores the interaction and relationship between two groups from different cultural contexts. In archaeology, the culture contact between European and indigenous people in the postcolonial period is the common studies which explore the roles of European and indigenous people, and how indigenous people understood and reacted to the dynamic processes of colonization (Decorse 1992; Given 2004; Lyons and Papadopoulos 2002; Torrence and Clarke 2000). Despite Taiwan was colonized by Europeans, the Dutch and Spanish, from 1624 to 1662 (Israel 1989: 173-174), little has been discovered in archaeology about the interaction between indigenous groups and Europeans, and how indigenous people reacted in the postcolonial period. 
In addition to historical documents, European materials or the imported materials which are associated with exchange network stimulated by Europeans provide important evidence for understanding the effects of culture contact. One of the common imported artifacts excavated from the late Iron Age (1000-400BP) sites to historical period (400-100BP) in Taiwan are glass beads and agate beads, in which most are imported materials which might enter Taiwan through trade network. Ornaments are usually viewed as materials which convey social meaning in a culture because their function as decoration. Researches about ornaments in this period can be divided into two categories: The first is concern about the technology of beads and another one is interested in the regional exchange system (Chen 2008, Hong 2005, Wang and Liu 2007, Liu 2005). However, these studies tend to view imported materials as an evidence of diffusion of cultural traits and focus on their sourcing and regional circulation. The issues such as how do indigenous people view these imported materials and what is the meaning of imported materials in local settlements are seldom explored. 
In this paper, I will explore the relationship between Europeans and indigenous people by examining the imported materials from late Iron Age Site in Northeastern Taiwan. The Upper Culture Layer (600-100BP) of the Kiwulan site in I-lan was excavated abundant imported materials, including glass beads, agate beads, and metal ornaments, which indicates the trade network stimulated by the Europeans. This article examines historical records and the archaeological contexts in order to know functions of imported ornaments, and their differences and universalities in various subgroups through quantitative analysis and ethnographic records. The aim is to identify the principles of ornaments usage in the Kiwulan society, explore their core social culture through the construction of a value system of various ornaments, and suggest that the ubiquitous ornaments might reflect the results of negotiation between European and indigenous people.
Historical and Ethnographic Background 
Historical records from Spanish and Dutch provide rich information about the historical and cultural background both for Europeans and indigenous people in the early 17th century. Here I focus on how were imported ornaments exchanged between different groups of people, and what is the meaning for those ornaments in local settlements by examining the written records from the observations of Europeans. Spanish occupied and built forts in Northern Taiwan since 1624, from which Spanish missionaries often went to indigenous settlement and left behind numerous records from those places. The report of the Dominican Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) as a materials from exchange the necessities they need with other indigenous people. This shows that agate beads were kind of medium of exchange in the trade network. This trend then gradually spread to other areas, and even the Spanish soldiers used agate as bargaining chips for gambling, since agate equaled money (Li 2006: 132-149).
Father Jacinto Esquivel recorded how people in Tamsui entered marriages. A man who wanted to take a wife had to give her parents cuentas. If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled (Li 2006: 151). Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals (Pao 2008: 122, 143, 151). Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity. They put a mat into the grave to protect the body from moisture and then put cooked rice on both sides of the head as provision. The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need. The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave (Li 2006: 153). 
We can see from the literature that agate beads had a certain value; they were used as medium of exchange, as dowry and as payment for settling problems. At the same time they played an important role in healing rites and as grave goods, being an indicator for reputation or a high position. 
Documents from the Qing dynasty usually describe the decorative purposes of the ornaments. In the Kavalan Subprefecture Gazetteer (Chen 2006[1852]: 308), the Kavalan Zhi Lue (Ko 1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao 1996[1829]: 77), it is mentioned that aborigines in I-lan used metal threads to weave so called golden carp-shaped ornaments. Their number was small and their value high, hence only the rich possessed them. Ordinary people wore agate beads or glass beads on their head or neck during various festivals. Women unearthed from the burials of the Ki-Wu-Lan site did possess more agate beads and glass beads than men, which are verified by documents and records. At the same time, fish-shape ornaments were also unearthed at the site. But, due to the poor condition of the burials, no inferences can be made regarding gender and age of bodies that were given fish-shape ornaments as burial objects. However, from the fact that bodies found with fish-shape ornaments also had comparatively more other grave goods, one can conclude that they were of high value. 
In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori (Kanori 1996: 227-232) made field surveys among various ethnic groups in Taiwan. He describes the Plains Aborigines from I-lan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.
Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans (Liu 2008: 133-134). Interview records describe agate beads as valuable objects passed down from mother to daughter. When they were not used, they were often put on roof beams so that they were not seen by other people. But the origin of these beads is still not known. Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer (Hsu 1992: 22) and carp-shaped ornaments had completely disappeared. Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination. It was usually the female shamans that were responsible for divination rites, a technique passed down by ancestors. Most aborigines today don’t know the origin of agate beads any more, but their worth is still highlighted by oral history and through their scarcity. 
Agate beads, glass beads and copper bracelets were not only an ornament for decoration, but also possessed social significance. In the 17th century agate beads were used as currency, as a medium of exchange and were used in rites. In the 19th century they were used as decoration and tribes now usually perceive them as objects for rites and divination. They have become ever scarcer and their origin is already forgotten by aborigines today. We can see that their nature as object of divination has not changed since the 17th century, and even future generations will remember that they were used by female shamans. Previous literature mentions golden beads and agate beads as divination beads of the Kavalan tribe, and both were often worn as ornaments. What value and significance did they actually have, and did they have any relationship? Did the indigenous people accept those different kind of imported ornaments in the same way or treat them separately? Those questions could be answered by examining the archaeological evidence, which shows that different kind of ornaments, such as golden beads and agate beads might have totally different meaning and they are incorporated into indigenous societies based on some existing principles or social rules in their culture.
Archaeological Background 
Kiwulan site in northeastern Taiwan provides a good framework for understanding the interaction between Europeans and indigenous people. The Kiwulan site (Figure 1) is located at I-lan, north of Lanyang Plain in northeastern Taiwan, surrounded by the Central Mountain Range in the west and the Pacific Ocean in the east. Kiwulan site was excavated from 2001 to 2004, and the total excavation area was 3,814 m2. The time period of Kiwulan site can be divided into Lower Layer Culture (1300-800BP) and Upper Layer Culture (600-100BP) based on a series of 32 radiocarbon dates, which coordinate to the Iron Age and Proto-historical period in Taiwan. The archaeological evidence includes a rich amount of artifacts, burials, ash pits, wooden pillars and stone piles. The AD district is the area which has been unearthed most of the artifacts and remains (Figure 2). Preliminary investigation suggests that the wooden pillars were part of a building structure of a main living place. The highest concentration of remains in the Upper Culture Layer indicates a large living space that was used over a long time (Chen et al., 2007, Chiu 2004: 30).

Figure 1: Location of the Kiwulan site

 AD區

Figure 2: The distribution of archaeological remains in AD district

Ornaments found in Kiwulan site
The total number of ornaments found at the Ki-Wu-Lan site amounts to 12,597[1]. They are made of agate, glass, metal, shells, wood and bones (Table 1) and the categories includes beads, rings, bells, pieces and knitted objects. Most of the ornaments come from the Upper Culture Layer and just few found from Lower Culture Layer.

Figure 1: Frequency of different ornaments

Most ornaments are glass beads, followed by metal ornaments and agate beads, together accounting for 99% of all ornaments. Ornaments made of shell, wood or bone only account for 1% of all ornaments. The reason for the high quantity of glass beads is that they usually have been found in clusters because of their original form, for example, necklaces. 
    The frequency of ornaments from burials is more than those from the living space. There is a total of 125 burials at the Ki-Wu-Lan site, 90 burials belong to the Upper Culture Layer and 35 from the Lower Culture Layer. Burial goods include potteries, stonewares, imported ceramics, metal bracelets, glass beads, wooden combs, coins and bone artifacts. Among the ornaments, metal bracelets and beads have often been found in burial as grave goods (Chen, 2007).

Ornaments from Iron Age
Research about ornaments from the Iron Age in Taiwan can be roughly divided into two research areas: The first is concerned about the production system of artifacts and the regional interaction system. A component analysis of part of the glass beads from the Ki-Wu-Lan site has been carried out. With Lead Isotope Analysis, Chen (2008) found out that the composition of golden beads among them is similar to lead ore produced in mainland China. The production of agate and metal ornaments from the Ki-Wu-Lan site has not yet been discovered, but initial research suggests that most of the ornaments were foreign. Wang and Liu’s (2007) study of imported materials point out that the 16th and 17th century trade with Spaniards and Dutch was very different from the comparatively inter-insular trade in the prehistoric period. These imported materials after circulating for a long period of time in Taiwan have become important part of the indigenous material culture. 
Regarding the regional interaction system, Hong (2005: 81-940) examines the development of external beads from the Philippine Archipelago and makes inferences about the changing patterns of glass beads excavated in Taiwan. Hong believes that both areas developed from nephrite beads to Indo-Pacific beads and eventually to Chinese coil beads, and suggests that the trade on the South China Sea between Taiwan, Philippines and Southeast Asia can be traced back to the Neolithic Age and becoming more developed and complex during the Iron Age. As for the transfer and changes of domestic and external materials on the island of Taiwan, Liu (2005:211) points out that the appearance of small numbers of imported materials like metal, glass or agate has began from the late period of Neolithic Age (3000-1800BP). The diffusion of the use of metal tools was from north to south on the east coast, with agate and glass beads gradually replacing jade burial objects from 1800 years ago. This process of replacement of decoration system is related with the foreign exchange system. At the end of the prehistoric period, the long-distance trade with Asia area such as India, the Indochina Peninsula and China was incorporated into a world trading system by European expansion in the 16th to 17th century. 
The second research area approaches the excavation of ornaments from a social cultural point of view. People in traditional societies rarely decorate their bodies solely for aesthetic purposes; most ornaments possess religious, economic or political significance and indicate a certain social behavior. From an archeological perspective these artifacts are scarce but of great importance (Lien 2003: 161). Ornaments are an important part of in an ancient society; hence archaeological excavations of ornaments can help to understand their social meaning. This kind of research is concerned about the significance of imported materials in the context of a certain site and the role these materials played in the society. Hong (2005: 94) notes that every single glass bead has to be understood in the context of its excavation site and analyzed with other non-perishable materials such as imported ceramics, iron, gold or coins to clarify the socio economic changes that the beads reflect in a particular community. 
When imported materials enter the local community they can experience a change in meaning. There has been a long period of development in anthropology about the meaning transformation of imported materials. Chen (2004: 63) discusses the production of traditional ceremonial dresses of the Nanwang Puyuma tribe, and describes the development of the traditional “weave” and the later “embroidery”. He believes that although the skill of “embroidery” was imported, it still became indigenous craft that ancestors passed down through re-contextualization. Dresses and personal adornments that originally did not possess any ceremonial value or taboo meaning were given traditional significance in order to conform to the original cultural identity that existed in the society among different ages or genders. This concept of transformation of significance can also be used for archaeological data. Hu (2007) examines glass beads of the Austronesian society of different areas of Taiwan and considers cultural value and change of significance through their material characteristics and fluctuations. Hu points out that different areas exhibit a different logic of usage towards imported material, and the value and significance of materials is continually constructed and reconstructed. 
Ornaments are a part of material culture and through observation and interpretation of the material it becomes possible for archaeologists to interpret the past. From this perspective Hsu (1992) studies glass beads of the Paiwan tribe. On the basis of structural principles, material combinations and legends of the beads, she discusses the underlying social organization. Hsu believes that if we see ornaments only as objects for decorative purposes, we only see their functional significance. But if we consider the ideas and values that people attached to them, we can understand their real significance. Deetz (1977) thinks, especially in regard to historical archaeology, that material culture is not culture but a product of culture. This culture refers to social norms and patterns of thinking and behavior. Miller’s (1987) consumer theory further stresses the independence of the object itself and the close connection of social culture and individuals. He examines the relationship between society and material and believes that the material is a strong manifestation of culture. These objects not only reflect personal tastes and practical characteristics, but at the same time moral standards and social concepts. However, Douglas and Asherwood (1996) point out that the significance of materials is closely related to social practices, symbolizing potential social distinctions and shaping of social classes. For those ornaments that possess less practical value, the examination of materials highlights the close relationship between the object itself and the social culture.
Research Methods
This research conducted a quantitative analysis of ornaments and compared the ornaments from different archaeological contexts such as settlement area and the burials. In burials, the relationships between the ornaments and gender or age will be examined. The temporal changes of the ornaments in different archaeological contexts will also be studies in this research.
The Upper Cultural Layer of the Ki-Wu-Lan site is belongs to the Iron Age in Taiwan (2000-400BP), but because of the Age of Discovery from the 16th century, it actually entered the historical ages, leaving behind a number of foreign descriptions and writings. Therefore this research takes into account literatures and records from Spain, the Netherlands, China and Japan.
Past research about ornaments solely concentrated on specific objects like glass beads, because the ornaments are well preserved and the materials are still precious and important to aborigines. However, the ornaments excavated at the Ki-Wu-Lan site are numerous (including glass beads, agate and metal objects), well preserved and have research potential. Therefore, I believe that view of all the objects in their entirety will enable a better understanding of the usage and characteristics of ornaments.
Results and Discussion
1. Ornament Shapes and Types
This research applies the already existing ornament classification principles of the archaeological reports of Kiwulan site because their shape classification is more detailed. For quantitative reasons, I combine ornaments made of the same material and analyze them accordingly. The following shapes and types of ornaments were excavated at the site:
1. Agate Beads
A total of 783 pieces were found. Natural carnelian was first grinded and then perforated from both sides, however, the holes are often askew. According to their appearance the ornaments can be divided into eight categories. From these eight groups, hexagonal cylinder beads were most common in the burials, followed by cylindrical beads and big barrel beads.

 Figure 2: Number of various agate bead shapes

1. Glass Beads
A total of 10,821 pieces were found. The beads from the Upper Culture Layer can be divided into 39 categories (UL1-UL39). Category 1(UL1), golden bead category, was most common (898 pieces), followed by category 7 (UL7), the orange cylindrical bead category (573 pieces) and category 20 (UL20), the red opaque small bead category (88 pieces). All other categories have less than 50 pieces. Golden beads and orange cylindrical beads were the most common categories in the burials.
1. Metal Ornaments
A total of 930 pieces were found. Metal ornaments show a wide range of shapes, are mostly made of copper, but there were iron, gold and silver materials as well. They can be divided into various shapes such as bells, rings or pieces. Metal rings are most common, accounting for 66.8% of all metal ornaments. Metal fish-shape knits were only found in burials (M003 and M011) but not in the settlement area. Small-sized bells were most common in burials, large-sized bells most common in the settlement area. In general, metal rings were the ornaments most commonly found, and they can be subdivided into bracelets and ear rings.

Figure 3: Number of various metal ornament shapes
1. Shell, wooden, and bone ornaments
Shell ornaments are scarce, only one piece of earring was excavated from the Upper Culture Layer and none were found in the burials. Four of the 16 perforated bone and horn ornaments were excavated from M020 and one from M094 (both Upper Culture Layer). The shapes of the objects found in the burials are short and small and don’t show any carvings. We speculate that they were used together with beads.
2. Settlement Area
As part of the area was destroyed by the engineering project, quantitative analysis was only carried out in areas that were well preserved. I used Hsieh (2008: 16) as reference, and divided the Upper Culture Layer into six layers according to her area sampling standard and layer classification principle[2]. Statistical analysis was done on 25 test pits in the AD district (Figure 2); each pit size was 4m x 4m meters, with an accumulated size of 400 square meters.

 未命名.jpg

Figure 3: Sampling area of AD district
From the overall number of ornaments one can see that their number peaked at L4 and then started to decrease, showing a unimodal distribution (Table 4). At a closer look at the different types of ornaments, I combined ornaments made of similar materials and categorized them into metal ornaments, glass beads, golden beads and agate beads. 
The results show that golden beads start to appear in large numbers in L5, more than other ornaments, reaching a peak at L4 and then descending up to L2. The tendency of glass beads and metal ornaments is similar. These three types are most numerous in L4 and then begin to descend. Only agate beads show a slightly different tendency with a peak in L3.

Figure 4: Number of ornaments and distribution in sampling area 
(with L6 the earliest layer and L1 the latest)

Figure 5: Distribution of four ornaments in sampling area
(with L6 as the earliest layer and L1 the latest)

3. Quantitative Analysis[3]
1. Method
Objects of the quantitative analysis are glass beads, agate beads and metal ornaments from burials of the Upper Culture Layer. The purpose is to examine whether there are differences between age groups and gender, hence only samples where gender and age could be determined were taken. This research applied the chi-squared test and the Mann-Whitney test from non-parametric statistics. The chi-square test was carried out for the qualitative analysis, mainly to test the relation between the two variables in order to examine the relation of ornaments with gender and age. The Mann-Whitney test was carried out for the quantitative analysis and used for the examination of inter-gender differences in regard to the number of ornaments.
15 male bodies and 19 female bodies were excavated from the burials. First, the chi-square test examines if gender shows a significant relation with specific ornaments. In the case that both men and women have ornaments, from the number of ornaments we can infer if there are any differences. From 61 bodies the age could be determined, and there are six age distribution: two infants (0-1 years), 12 children (3-12 years), two adolescents (12-20 years), 10 adults (20-35), seven middle-aged (35-50) and 2 elderly (over 50 years). In addition, 18 of the bodies were over 18.
There are problems of samples from age groups being too small, therefore percentage observation could be an additional explanation, and I also combined the infants, children and adolescents into a group (pre-adult group) and the adults, middle-aged and elderly into a group (adult group). Now the pre-adult group consists of 16 samples and the adult group of 37 samples.

1. SPSS results
The results suggest that the external ornaments excavated from the Ki-Wu-Lan site possess the following distribution characteristics:
(1) Gender shows a significant difference of ornaments, however age group does not have a significant difference, and only the percentage shows a difference in proportion.
(2) Regarding gender, most of the agate beads were found with female bodies, only one male body was found with agate beads[4], indicating that they were burial objects only given to women. Glass beads appear to be also given primarily to women. Metal rings were found with both men and women, however the number found with women was greater than men. For earrings and golden beads we could not find any significant gender-related difference.
(3) Wide metal rings were found with adults and all adolescents, suggesting a relationship between thick metal bracelet and adolescents. Bells were found with each body between 6 and eleven years, showing that they were a burial object for children.
(4) Regarding their possessions of ornaments and their number of possessions, adults and pre-adults don’t show any differences, suggesting that there was no distinction between adults and pre-adults.
Conclusion
The following features can be summarized to explain the characteristics of the ornaments found at Kiwulan site:

1. Most ornaments were excavated from the settlement area but also burials; however, the ornaments from the burials were more elaborate. Metal weavings in fish-shape were only excavated from two burials out of 90 burials of the Upper Culture Layer. According to historical literatures, these metal weavings were called “golden carp-shape” ornament, which is valuable in their society. The most common ornaments found in the burials are metal rings, excavated from 66 of 90 burials.
2. The distribution of golden beads, glass beads and metal ornaments in the different layers is similar, all of them peaking at L4. Golden beads, however, are most numerous. Agate beads differ from the distribution of the three aforementioned ornaments. They are less numerous and their number increases when the number of other ornaments decreases, indicating a possible different origin.
3. Initial statistical analysis suggests that agate beads were burial objects for women. The usage of metal rings and glass beads is similar among the whole population. But women possessed more metal rings and glass beads than men. Adolescents wore wide metal rings, and children, bells.
   Archaeological data show that agate beads given as burial objects show a gender difference. Agate beads seem to have a special meaning for women and modern ethnographic records suggest that they were used by the Kavalan people for divination, which was usually practiced by women. Spanish literature describes agate beads as medium of exchange and dowry. Some ethnographic records show that golden beads were used as an object for divination. But their meaning was very different which can be explained with the archaeological evidence. The universality of golden beads found in burials may be related to an influx in large numbers during that period or to the preference of the residents for these beads. They were easy to obtain, very popular in this area and more numerous than agate beads. When other ornaments prevailed, the number of agate beads declined, but they were still used as grave goods. The agate beads found were small in number, but still they were used as burial objects. Compared to golden beads, they might indicate a higher rank, and the fact that they were only found with women indicates their uniqueness in the society.
  The ornaments found at Ki-Wu-Lan society indicate different values and patterns of usage based on the burials and living area data. This reveals that the ornaments were not only used as decoration, they were items that carried cultural significance, which also means Kiwulan people transformed the meaning the imported material, and then recontextualize them into local materials by daily usage and ritual practice. In addition, the abundance of ornaments present the preference of local people and the interaction and negotiation between Kiwulan people and European or Chinese traders. The exchange of beads seems based on the pre-existing principles of exchange, which shows the agency of local indigenous people.

________________
[1] This number contains ornaments from the Upper Culture Layer, Lower Culture Layer, soil layer, old riverbed stratified deposit, surface collection and disrupted layer. 
[2] The elevation of the Upper Culture Layer is between -80 and 110 centimeters, Hsieh divided it into six major layers according to the contents of the cultural layer with cross-pit burial data. 
[3] The α value is set at 0.1 in SPSS.
[4] We assume that the M108 is a male burial, but because of the poor condition of the bones we cannot confirm this assumption. 
[a]put these in a table here
[b]citation
[c]citation

### 2019-03-24 analysis/paper/paper.Rmd
resolved
plot of major ornament types

--- 2019-03-24 analysis/paper/paper.Rmd
@@
   ornaments_period %>% 
-  filter(!is.na(`6-layers`)) %>% 
+  filter(!is.na(`6-layers`),
+         Categories != "Unknown Metal") %>% 
+  mutate(Categories = fct_lump(Categories, 
+                               n = 4))  %>% 
   group_by(period, Categories) %>%
@@
   ungroup() %>% 
-  mutate(Categories = fct_lump(as.factor(Categories), n = 5)) %>% 
   mutate(Categories = fct_reorder(Categories, n))
@@
   coord_flip() +
-  labs(y = "Frequency")
+  labs(y = "Frequency",
+       x = "") +
+  theme_minimal(base_size = 14)
 ```

### 2019-03-25 analysis/paper/paper.Rmd
more plots for beads

--- 2019-03-24 analysis/paper/paper.Rmd
@@
 
+```{r get-data, eval = FALSE}
+# summarise the ornaments by layers
+library(tidyverse)
+library(ggplot2)
+
+# L5 and L6 represent pre-European contact period, 
+# L4 was the time of European contact, and 
+# L2 and L1 was the period of Chinese contact.
+
+ornaments_period <- 
+kwl_upper %>% 
+  filter(!is.na(`6-layers`)) %>% 
+  mutate(period = case_when(
+    `6-layers` %in% 1:2 ~ "Chinese Contact",
+    `6-layers` == 4 ~ "European Contact",
+    `6-layers` %in% 5:6 ~ "Before European Contact",
+     TRUE ~ "other"
+  )) %>% 
+  filter(period != "other")
+
+# need to reorder, try https://forcats.tidyverse.org/reference/index.html
+# reorder factor for categories (by frequency) and period (manually)
+ornaments_count <-
+  ornaments_period %>% 
+  filter(!is.na(`6-layers`),
+         Categories != "Unknown Metal") %>% 
+  mutate(Categories = fct_lump(Categories, 
+                               n = 5)) %>% 
+  group_by(period, Categories) %>%
+  count() %>% 
+  ungroup() %>% 
+  mutate(Categories = fct_reorder(Categories, n))
+
+ggplot(transform(ornaments_count,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Categories, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "") +
+  theme_minimal(base_size = 14)
+
+# plot for intact golden bead by period
+ornaments_intact_gold <-
+  ornaments_period %>% 
+  filter(Condition == 'complete',
+         Categories == 'Golden bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
+
+ggplot(transform(ornaments_intact_gold,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_point(aes(Length, Width, col = Perforation_average)) +
+  facet_wrap(~ period) +
+  labs(x = "Length(mm)",
+       y = "Width(mm)",
+       title = "Golden beads") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+
+# plot for intact agate bead by period
+ornaments_intact_agate <-
+  ornaments_period %>% 
+  filter(Condition == c('complete','near-complete'),
+         Categories == 'Agate bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
+
+ggplot(ornaments_intact_agate, 
+       aes(Length, Width, col = period))+
+  geom_point() +
+  labs(x = "Length(mm)",
+       y = "Width(mm)",
+       title = "Agate beads") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+```
+
 # Discussion

### 2019-03-25 analysis/paper/paper.Rmd
some updates for plots

--- 2019-03-24 analysis/paper/paper.Rmd
@@
 
+```{r get-data, eval = FALSE}
+# summarise the ornaments by layers
+library(tidyverse)
+library(ggplot2)
+
+# L5 and L6 represent pre-European contact period, 
+# L4 was the time of European contact, and 
+# L2 and L1 was the period of Chinese contact.
+
+ornaments_period <- 
+kwl_upper %>% 
+  filter(!is.na(`6-layers`)) %>% 
+  mutate(period = case_when(
+    `6-layers` %in% 1:2 ~ "Chinese Contact",
+    `6-layers` == 4 ~ "European Contact",
+    `6-layers` %in% 5:6 ~ "Before European Contact",
+     TRUE ~ "other"
+  )) %>% 
+  filter(period != "other")
+
+# need to reorder, try https://forcats.tidyverse.org/reference/index.html
+# reorder factor for categories (by frequency) and period (manually)
+# total number of ornaments for each categories 
+ornaments_count_test <-
+  ornaments_period %>% 
+  filter(!is.na(`6-layers`),
+         Categories != "Unknown Metal") %>% 
+  mutate(Categories = fct_lump(Categories, 
+                               n = 5)) %>% 
+  group_by(Categories) %>%
+  count() %>% 
+  ungroup() %>% 
+  mutate(Categories = fct_reorder(Categories, n))
+
+ggplot(ornaments_count_test,
+       aes(Categories, n)) +
+  geom_col() +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Categories") +
+  theme_minimal(base_size = 14)
+
+# categories of ornaments by period
+ornaments_count <-
+  ornaments_period %>% 
+  filter(!is.na(`6-layers`),
+         Categories != "Unknown Metal") %>% 
+  mutate(Categories = fct_lump(Categories, 
+                               n = 5)) %>% 
+  group_by(period, Categories) %>%
+  count() %>% 
+  ungroup() %>% 
+  mutate(Categories = fct_reorder(Categories, n))
+
+ggplot(transform(ornaments_count,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Categories, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Categories") +
+  theme_minimal(base_size = 14)
+```
+
+```{r get-data, eval = FALSE}
+# plot for intact golden bead by period
+ornaments_intact_gold <-
+  ornaments_period %>% 
+  filter(Condition == 'complete',
+         Categories == 'Golden bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
+
+ggplot(transform(ornaments_intact_gold,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_point(aes(Length, Width, col = Perforation_average)) +
+  facet_wrap(~ period) +
+  labs(x = "Length (mm)",
+       y = "Width (mm)",
+       title = "Golden beads") +
+  coord_fixed(ratio = 1) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+
+# plot for intact agate bead by period
+ornaments_intact_agate <-
+  ornaments_period %>% 
+  filter(Condition == c('complete','near-complete'),
+         Categories == 'Agate bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
+
+ggplot(ornaments_intact_agate, 
+       aes(Length, Width, col = period))+
+  geom_point() +
+  labs(x = "Length (mm)",
+       y = "Width (mm)",
+       title = "Agate beads") +
+  coord_fixed(ratio = 1) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+```
+
 # Discussion

### 2019-03-25 README.Rmd
Update README.Rmd

@@
 
-# master.paper
+# kwlornaments
 
@@
 # install.packages("devtools")
-devtools::install_github("LiYingWang/master.paper")
+devtools::install_github("LiYingWang/kwlornaments")
 ```

### 2019-03-25 analysis/paper/paper.Rmd
get ornament types in the right order

@@
 
-```{r get-data, eval = FALSE}
+```{r tidy-data, eval = FALSE}
 # summarise the ornaments by layers
@@
   filter(period != "other")
+```
 
+```{r plot-counts}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 # total number of ornaments for each categories
+
 ornaments_count_test <-
@@
   theme_minimal(base_size = 14)
+```
 
+```{r plot-counts-by-period}
 # categories of ornaments by period
@@
                                n = 5)) %>% 
+  mutate(Categories = fct_infreq(Categories)) %>% 
+  mutate(Categories = fct_rev(Categories)) %>% 
   group_by(period, Categories) %>%
   count() %>% 
-  ungroup() %>% 
-  mutate(Categories = fct_reorder(Categories, n))
+  ungroup()
 

### 2019-03-28 analysis/paper/paper.Rmd
updates for plots

@@
 
+```{r get-data, eval = FALSE}
+# plot for different shape of agate bead by period
+agate_bead_shape <-
+  ornaments_period %>% 
+  filter(Categories == 'Agate bead') %>% 
+  mutate(Type = ifelse(Type == 'prism/octagonal', 
+                       str_remove(Type, '/octagonal'), Type)) %>% 
+  mutate(Type = ifelse(Type == 'prism/pentagonal', 
+                       str_remove(Type, '/pentagonal'), Type)) %>% 
+  mutate(Type = ifelse(Type == 'prism/hexagonal', 
+                       str_remove(Type, 'prism/'), Type)) %>% 
+  mutate(Type = fct_infreq(Type)) %>% 
+  mutate(Type = fct_rev(Type)) %>% 
+  group_by(period, Type) %>% 
+  count()
+
+ggplot(transform(agate_bead_shape,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Type, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Shape of Agate bead") +
+  theme_minimal(base_size = 14)
+```
+
+ggplot(data = diamonds) + 
+  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
+
 # Discussion

### 2019-03-29 analysis/paper/paper.Rmd
add spatial plot

--- 2019-03-25 analysis/paper/paper.Rmd
@@
 
-```{r get-data, eval = FALSE}
+```{r scatter-plot-bead-by-period}
 # plot for intact golden bead by period
@@
   ornaments_period %>% 
-  filter(Condition == 'complete',
+  filter(Condition %in% c('complete', 'near-complete'),
          Categories == 'Golden bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
+
+ggplot(transform(ornaments_intact_gold,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_point(aes(Length, Width, col = Perforation_average)) +
+  facet_wrap(~ period) +
+  labs(x = "Length (mm)",
+       y = "Width (mm)",
+       title = "Golden beads") +
+  coord_fixed(ratio = 1) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+
+# plot for intact glass bead by period
+ornaments_intact_glass <-
+  ornaments_period %>% 
+  filter(Condition %in% c('complete', 'near-complete'),
+         Categories == 'Glass bead') %>% 
   mutate(Length = as.numeric(`Length(mm)`),
@@
   ornaments_period %>% 
-  filter(Condition == c('complete','near-complete'),
+  filter(Condition %in% c('complete','near-complete'),
          Categories == 'Agate bead') %>% 
@@
   theme(plot.title = element_text(hjust=0.5))
+```
+
+```{r plot-count-type-by-period}
+# plot for different shape of agate bead by period
+agate_bead_shape <-
+  ornaments_period %>% 
+  filter(Categories == 'Agate bead') %>% 
+  mutate(Type = ifelse(Type == 'prism/octagonal', 
+                       str_remove(Type, '/octagonal'), Type)) %>% 
+  mutate(Type = ifelse(Type == 'prism/pentagonal', 
+                       str_remove(Type, '/pentagonal'), Type)) %>% 
+  mutate(Type = ifelse(Type == 'prism/hexagonal', 
+                       str_remove(Type, 'prism/'), Type)) %>% 
+  mutate(Type = fct_infreq(Type)) %>% 
+  mutate(Type = fct_rev(Type)) %>% 
+  group_by(period, Type) %>% 
+  count()
+
+ggplot(transform(agate_bead_shape,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Type, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Shape of Agate bead") +
+  theme_minimal(base_size = 14)
+
+# plot for different shape of metal bell by period
+bell_shape <-
+  ornaments_period %>% 
+  filter(Categories == 'Bell',
+         Type != 'unclassified') %>% 
+  mutate(Type = fct_infreq(Type)) %>% 
+  mutate(Type = fct_rev(Type)) %>% 
+  group_by(period, Type) %>% 
+  count()
+
+ggplot(transform(bell_shape,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Type, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Shape of Bell") +
+  theme_minimal(base_size = 14)
+
+# plot for different shape of metal ring by period
+ring_shape <-
+  ornaments_period %>% 
+  filter(Categories == 'Metal ring') %>% 
+  mutate(Type = fct_infreq(Type)) %>% 
+  mutate(Type = fct_rev(Type)) %>% 
+  group_by(period, Type) %>% 
+  count()
+
+ggplot(transform(ring_shape,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_col(aes(Type, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Shape of Metal ring") +
+  theme_minimal(base_size = 14)
+```
+
+```{r scatter-plot-metal-by-period}
+# plot for intact bell by period
+ornaments_intact_bell <-
+  ornaments_period %>% 
+  filter(Condition %in% c('complete', 'near-complete'),
+         Categories == 'Bell',
+         Type != 'unclassified') %>% 
+  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`)), #remove +
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`))
+
+ggplot(transform(ornaments_intact_bell,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_point(aes(Length, Width, col = Type)) +
+  facet_wrap(~ period) +
+  labs(x = "Length (mm)",
+       y = "Width (mm)",
+       title = "Bell") +
+  coord_fixed(ratio = 1) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+
+# plot for intact ring by period
+ornaments_intact_ring <-
+  ornaments_period %>% 
+  filter(Condition %in% c('complete', 'near-complete'),
+         Categories == 'Metal ring') %>% 
+  mutate(Length = as.numeric(gsub("\\+|\\-", "", `Length(mm)`)), #remove + -
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`))
+
+ggplot(transform(ornaments_intact_ring,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_point(aes(Length, Width, col = Type)) +
+  facet_wrap(~ period) +
+  labs(x = "Length (mm)",
+       y = "Width (mm)",
+       title = "Metal ring") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5))
+```
+
+```{r read-in-the-spatial-data, eval = FALSE}
+library(sf)
+library(cartography)
+burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
+AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
+AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
+location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
+units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
+```
+
+```{r tidy-data, eval = FALSE}
+# 40 sampling units
+sample_units <-
+  AD_data %>% 
+  semi_join(units_40, by = c (`坑號` = "the_sq"))
+
+# join two dataset to get coordinate
+ornaments_period_join <-
+  ornaments_period %>%
+  mutate(period = as.factor(period)) %>% # seems need this line first 
+  group_by(Pit_No, period) %>% 
+  count(Categories) %>% 
+  inner_join(AD_data, by = c("Pit_No" = "坑號"))
+
+# filter golden bead
+ornaments_period_join_gold <-
+  ornaments_period_join %>% 
+  filter(Categories %in% 'Golden bead')
+
+# plot by period
+ggplot(transform(ornaments_period_join_gold,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
+  geom_sf(data = sample_units) +
+  geom_sf(data = ornaments_period_join_gold, aes(fill = n)) +
+  facet_wrap(~ period) +
+  labs(title = "Spatial Distribution of Golden Beads",
+       fill = "frequency") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust=0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank())
 ```

### 2019-03-30 analysis/paper/paper.Rmd
updates for code, add mmore spatial plots and try animation

@@
   )) %>% 
-  filter(period != "other")
+  filter(period != "other") %>% 
+  mutate(period = factor(period,
+                       level = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))
 ```
@@
 
-ggplot(transform(ornaments_count,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(ornaments_count) +
   geom_col(aes(Categories, n)) +
@@
          Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`)) %>% 
+         Width = as.numeric(`Width(mm)`),
+         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
   rowwise() %>% 
-  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+  mutate(Perforation_average = mean(c(`Perforation1`, `Perforation2(mm)`), 
                                na.rm = TRUE))
 
-ggplot(transform(ornaments_intact_gold,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(ornaments_intact_gold) +
   geom_point(aes(Length, Width, col = Perforation_average)) +
@@
 # plot for intact glass bead by period
-ornaments_intact_glass <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Glass bead') %>% 
-  mutate(Length = as.numeric(`Length(mm)`),
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`)) %>% 
-  rowwise() %>% 
-  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
-                               na.rm = TRUE))
-
-ggplot(transform(ornaments_intact_gold,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
-  geom_point(aes(Length, Width, col = Perforation_average)) +
-  facet_wrap(~ period) +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Golden beads") +
-  coord_fixed(ratio = 1) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
+# need to deal with the original data
 
@@
 
-ggplot(transform(agate_bead_shape,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(agate_bead_shape) +
   geom_col(aes(Type, n)) +
@@
 
-ggplot(transform(bell_shape,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(bell_shape) +
   geom_col(aes(Type, n)) +
@@
 
-ggplot(transform(ring_shape,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(ring_shape) +
   geom_col(aes(Type, n)) +
@@
 
-ggplot(transform(ornaments_intact_bell,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(ornaments_intact_bell) +
   geom_point(aes(Length, Width, col = Type)) +
@@
 
-ggplot(transform(ornaments_intact_ring,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
+ggplot(ornaments_intact_ring) +
   geom_point(aes(Length, Width, col = Type)) +
@@
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
+```
 
+```{r plot-spatial-distribution}
 # filter golden bead
@@
 # plot by period
-ggplot(transform(ornaments_period_join_gold,
-                 period = factor(period, 
-                                     levels = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact")))) +
-  geom_sf(data = sample_units) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n)) +
+ggplot(ornaments_period_join_gold) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
   facet_wrap(~ period) +
@@
        fill = "frequency") +
+  scale_fill_gradient(high = "black", low = "grey") +
   theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5),
+  theme(plot.title = element_text(hjust = 0.5),
         axis.text.x = element_blank(),
-        axis.text.y = element_blank())
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+
+# filter glass bead
+ornaments_period_join_gold <-
+  ornaments_period_join %>% 
+  filter(Categories %in% 'Glass bead')
+
+# plot by period
+ggplot(ornaments_period_join_gold) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
+  facet_wrap(~ period) +
+  labs(title = "Spatial Distribution of Glass Beads",
+       fill = "frequency") +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+
+# filter agate bead
+ornaments_period_join_gold <-
+  ornaments_period_join %>% 
+  filter(Categories %in% 'Agate bead')
+
+# plot by period
+ggplot(ornaments_period_join_gold) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
+  facet_wrap(~ period) +
+  labs(title = "Spatial Distribution of Agate Beads",
+       fill = "frequency") +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+```
+
+```{r-gganimate}
+devtools::install_github('thomasp85/gganimate')
+library(gganimate)
+
+anim_gold_bead <-
+  ggplot(ornaments_intact_gold,
+       aes(Length, Width)) +
+  geom_point() +
+  transition_states(period, 3, 1) +
+  labs(title = 'Period: {closest_state}',
+       x ='Length (mm) of golden bead',
+       y ='Width (mm) of golden bead')
+
+anim_gold_bead_spatial <-
+  ggplot(ornaments_period_join_gold) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
+  transition_states(period, 3, 1) +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+
 ```

### 2019-03-30 analysis/paper/paper.Rmd
move some code to the tidy-data chunk

@@
                                                 "European Contact", 
-                                                "Chinese Contact")))
+                                                "Chinese Contact"))) %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`),
+         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
 ```
@@
          Categories != "Unknown Metal") %>% 
+  mutate(Categories = as.factor(Categories)) %>% 
   mutate(Categories = fct_lump(Categories, 
@@
   filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Golden bead') %>% 
-  mutate(Length = as.numeric(`Length(mm)`),
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`),
-         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
-  rowwise() %>% 
-  mutate(Perforation_average = mean(c(`Perforation1`, `Perforation2(mm)`), 
-                               na.rm = TRUE))
+         Categories == 'Golden bead')
 
@@
   filter(Condition %in% c('complete','near-complete'),
-         Categories == 'Agate bead') %>% 
-  mutate(Length = as.numeric(`Length(mm)`),
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`)) %>% 
-  rowwise() %>% 
-  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
-                               na.rm = TRUE))
+         Categories == 'Agate bead')
 
@@
          Type != 'unclassified') %>% 
-  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`)), #remove +
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`))
+  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`))) #remove +
 
@@
          Categories == 'Metal ring') %>% 
-  mutate(Length = as.numeric(gsub("\\+|\\-", "", `Length(mm)`)), #remove + -
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`))
+  mutate(Length = as.numeric(gsub("\\+|\\-", "", `Length(mm)`))) #remove + -
 
@@
 ```{r-gganimate}
+# test gganimate for measurement of beads by period
 devtools::install_github('thomasp85/gganimate')
@@
 
+# does not work
 anim_gold_bead_spatial <-

### 2019-03-30 analysis/paper/paper.Rmd
Update paper.Rmd
fix the problem of reorder for factor by adding ungroup()

@@
          Categories != "Unknown Metal") %>%
-  mutate(Categories = as.factor(Categories)) %>% 
+  ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
   mutate(Categories = fct_lump(Categories, 
@@
          Categories != "Unknown Metal") %>%
+  ungroup() %>% 
   mutate(Categories = fct_lump(Categories, 
@@
   group_by(period, Categories) %>%
-  count() %>% 
-  ungroup()
+  count()
 
@@
   filter(Categories == 'Agate bead') %>%
+  ungroup() %>% 
   mutate(Type = ifelse(Type == 'prism/octagonal', 
@@
          Type != 'unclassified') %>% 
-  mutate(Type = fct_infreq(Type)) %>% 
+  ungroup() %>% 
   mutate(Type = fct_rev(Type)) %>% 
@@
   filter(Categories == 'Metal ring') %>%
+  ungroup() %>% 
   mutate(Type = fct_infreq(Type)) %>%

### 2019-03-31 analysis/paper/paper.Rmd
Update paper.Rmd

--- 2019-03-30 analysis/paper/paper.Rmd
@@
 
-```{r tidy-data, eval = FALSE}
+```{r tidy-data2, eval = FALSE}
 # 40 sampling units
@@
 # test gganimate for measurement of beads by period
-devtools::install_github('thomasp85/gganimate')
+# devtools::install_github('thomasp85/gganimate')
 library(gganimate)

### 2019-03-31 analysis/paper/paper.Rmd
Update paper.Rmd

@@
 
-```{r read-in-the-data, eval = FALSE}
+```{r read-in-the-data}
 library(here)
@@
 
-```{r tidy-data, eval = FALSE}
+```{r tidy-data}
 # summarise the ornaments by layers
@@
 
-```{r read-in-the-spatial-data, eval = FALSE}
+```{r read-in-the-spatial-data}
 library(sf)
@@
 
-```{r tidy-data2, eval = FALSE}
+```{r tidy-data2}
 # 40 sampling units

### 2019-04-22 analysis/paper/paper.Rmd
stash

@@
   )) %>% 
-  filter(period != "other") %>% 
-  mutate(period = factor(period,
-                       level = c("Before European Contact", 
-                                                "European Contact", 
-                                                "Chinese Contact"))) %>% 
-  mutate(Length = as.numeric(`Length(mm)`),
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`),
-         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
-  rowwise() %>% 
-  mutate(Perforation_average = mean(c(`Perforation1`, `Perforation2(mm)`), 
-                               na.rm = TRUE))
+  filter(period != "other")
 ```
@@
          Categories != "Unknown Metal") %>% 
-  mutate(Categories = as.factor(Categories)) %>% 
   mutate(Categories = fct_lump(Categories, 
@@
 
-ggplot(ornaments_count) +
+ggplot(transform(ornaments_count,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
   geom_col(aes(Categories, n)) +
@@
 
-```{r scatter-plot-bead-by-period}
+```{r get-data}
 # plot for intact golden bead by period
@@
   ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Golden bead')
+  filter(Condition == 'complete',
+         Categories == 'Golden bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
 
-ggplot(ornaments_intact_gold) +
+ggplot(transform(ornaments_intact_gold,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
   geom_point(aes(Length, Width, col = Perforation_average)) +
@@
 
-# plot for intact glass bead by period
-# need to deal with the original data
-
 # plot for intact agate bead by period
@@
   ornaments_period %>% 
-  filter(Condition %in% c('complete','near-complete'),
-         Categories == 'Agate bead')
+  filter(Condition == c('complete','near-complete'),
+         Categories == 'Agate bead') %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average = mean(c(`Perforation1(mm)`, `Perforation2(mm)`), 
+                               na.rm = TRUE))
 
@@
 
-```{r plot-count-type-by-period}
+```{r get-data}
+
+  ornaments_period %>% 
+  filter(Categories == 'Agate bead') %>% 
+  mutate(Type = ifelse(Type == c('prism/octagonal', 'prism/pentagonal'),
+                       'prism', Type)
+
 # plot for different shape of agate bead by period
@@
 
-ggplot(agate_bead_shape) +
+ggplot(transform(agate_bead_shape,
+                 period = factor(period, 
+                                     levels = c("Before European Contact", 
+                                                "European Contact", 
+                                                "Chinese Contact")))) +
   geom_col(aes(Type, n)) +
@@
   theme_minimal(base_size = 14)
-
-# plot for different shape of metal bell by period
-bell_shape <-
-  ornaments_period %>% 
-  filter(Categories == 'Bell',
-         Type != 'unclassified') %>% 
-  mutate(Type = fct_infreq(Type)) %>% 
-  mutate(Type = fct_rev(Type)) %>% 
-  group_by(period, Type) %>% 
-  count()
-
-ggplot(bell_shape) +
-  geom_col(aes(Type, n)) +
-  facet_wrap(~ period) +
-  coord_flip() +
-  labs(y = "Frequency",
-       x = "Shape of Bell") +
-  theme_minimal(base_size = 14)
-
-# plot for different shape of metal ring by period
-ring_shape <-
-  ornaments_period %>% 
-  filter(Categories == 'Metal ring') %>% 
-  mutate(Type = fct_infreq(Type)) %>% 
-  mutate(Type = fct_rev(Type)) %>% 
-  group_by(period, Type) %>% 
-  count()
-
-ggplot(ring_shape) +
-  geom_col(aes(Type, n)) +
-  facet_wrap(~ period) +
-  coord_flip() +
-  labs(y = "Frequency",
-       x = "Shape of Metal ring") +
-  theme_minimal(base_size = 14)
 ```
 
-```{r scatter-plot-metal-by-period}
-# plot for intact bell by period
-ornaments_intact_bell <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Bell',
-         Type != 'unclassified') %>% 
-  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`))) #remove +
-
-ggplot(ornaments_intact_bell) +
-  geom_point(aes(Length, Width, col = Type)) +
-  facet_wrap(~ period) +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Bell") +
-  coord_fixed(ratio = 1) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
-
-# plot for intact ring by period
-ornaments_intact_ring <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Metal ring') %>% 
-  mutate(Length = as.numeric(gsub("\\+|\\-", "", `Length(mm)`))) #remove + -
-
-ggplot(ornaments_intact_ring) +
-  geom_point(aes(Length, Width, col = Type)) +
-  facet_wrap(~ period) +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Metal ring") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
-```
-
-```{r read-in-the-spatial-data}
-library(sf)
-library(cartography)
-burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
-AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
-AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
-location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
-units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
-```
-
-```{r tidy-data2}
-# 40 sampling units
-sample_units <-
-  AD_data %>% 
-  semi_join(units_40, by = c (`坑號` = "the_sq"))
-
-# join two dataset to get coordinate
-ornaments_period_join <-
-  ornaments_period %>%
-  mutate(period = as.factor(period)) %>% # seems need this line first 
-  group_by(Pit_No, period) %>% 
-  count(Categories) %>% 
-  inner_join(AD_data, by = c("Pit_No" = "坑號"))
-```
-
-```{r plot-spatial-distribution}
-# filter golden bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Golden bead')
-
-# plot by period
-ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Golden Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-# filter glass bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Glass bead')
-
-# plot by period
-ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Glass Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-# filter agate bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Agate bead')
-
-# plot by period
-ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Agate Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-```
-
-```{r-gganimate}
-# test gganimate for measurement of beads by period
-# devtools::install_github('thomasp85/gganimate')
-library(gganimate)
-
-anim_gold_bead <-
-  ggplot(ornaments_intact_gold,
-       aes(Length, Width)) +
-  geom_point() +
-  transition_states(period, 3, 1) +
-  labs(title = 'Period: {closest_state}',
-       x ='Length (mm) of golden bead',
-       y ='Width (mm) of golden bead')
-
-# does not work
-anim_gold_bead_spatial <-
-  ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
-  transition_states(period, 3, 1) +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-```
+ggplot(data = diamonds) + 
+  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
 

### 2019-04-22 analysis/paper/paper.Rmd

--- 2019-03-31 analysis/paper/paper.Rmd
@@
 ```{r scatter-plot-bead-by-period}
+
 # plot for intact golden bead by period

### 2019-04-22 analysis/paper/paper.Rmd
move text in from google doc

@@
 ---
-title: "Title Goes Here"
+title: "Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan"
 author:
-  - Author One
-  - Author Two
+  - Li-Ying Wang
+  - Ben Marwick
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-# Introduction
+# Introduction and Research Question
 
-Here is a citation [@Marwick2017]
+Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display (Joyce 2005).
+Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appear in Taiwan around 1,800 years ago through regional trade network in Southeast Asia (Wang and Liu 2007).
+Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the Metal Age and reached a high peak of quantity and variety in the early 17th century when the Europeans arrived and involved the island in an international trade network on a large scale.
+Foreign trade goods introduced into the local indigenous societies in an European colonial or imperial contexts usually had an impact on local culture and societies in many parts of the world that usually involved in the transformation of indigenous economic, cultural, and socio-political systems (Dietler 1997; Junker 1993; Mitchell 2000; Silliman 2005).
+Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that stress active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices (Given 2004; Mullins 2011; Silliman 2001; Rubertone 2000; Torrence 2000; Torrence and Clarke 2000; Voss 2005).
 
-# Background
+Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
+Direct colonial rule was rare, but the question of long-lasting indirect impacts on local communities remains unanswered.
+Indirect effects of colonialism is addressed recently to discuss the impact on the indigenous societies in the periphery of colonial control (Trabert 2017).
+Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where Northeastern Taiwan is an especially unique example for discussing indirect effects due to its physically isolated location that made it not easily accessible to the Europeans that the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012).
+In addition, prior to the European contact, there were small-scale regional exchange networks between China and Taiwan.
+Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century.
+This made northeastern Taiwan a good case to explore culture contacts at different periods and their indirect impacts by taking a long-term perspective.
+
+We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan.
+Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact.
+We hypothesise that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
+We hypothesise that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.
+
+# Cultural context of Ornaments in Northeastern Taiwan
+
+Historical records from Spanish and Dutch provide rich information about the historical and cultural background both for Europeans and indigenous people in the early 17th century.
+Spanish occupied and built forts in Northern Taiwan since 1624, from which Spanish missionaries often went to indigenous settlement and left behind numerous records from those places.
+The report of the Dominican Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) as a materials to exchange with necessities they need with other indigenous people.
+This trend then gradually spread to other areas, and even the Spanish soldiers used agate as bargaining chips for gambling, since agate equaled money (Li 2006: 132-149).
+European contact was focused on trading with indigenous communities in northeastern Taiwan.
+Trading activity of many types of goode among these indigenous communities increased with the arrival of European colonisers.
+Thus, we predict that this trading activity resulted in an increase in the amount and diversity of ornaments coming into indigenous communities in northeastern Taiwan.
+
+Father Jacinto Esquivel recorded how people in Tamsui entered marriages.
+A man who wanted to take a wife had to give her parents cuentas.
+If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled (Li 2006: 151).
+Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals (Pao 2008: 122, 143, 151).
+Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity.
+They put a mat into the grave to protect the body from moisture and then put cooked rice on both sides of the head as provision.
+The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need.
+The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave (Li 2006: 153).
+
+Documents from the Qing dynasty usually describe the decorative purposes of the ornaments.
+In the Kavalan Subprefecture Gazetteer (Chen 2006[1852]: 308), the Kavalan Zhi Lue (Ko 1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao 1996[1829]: 77), it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments.
+Their number was small and their value high, hence only the rich possessed them.
+Ordinary people wore agate beads or glass beads on their head or neck during various festivals.
+In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori (Kanori 1996: 227-232) made field surveys among various ethnic groups in Taiwan.
+He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.
+
+Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans (Liu 2008: 133-134).
+Interview records describe agate beads as valuable objects passed down from mother to daughter.
+When they were not used, they were often put on roof beams so that they were not seen by other people.
+But the origin of these beads is still not known.
+Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer (Hsu 1992: 22) and carp-shaped ornaments had completely disappeared.
+Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination.
+It was usually the female shamans that were responsible for divination rites, a technique passed down by ancestors.
+Most aborigines today don’t know the origin of agate beads any more, but their worth is still highlighted by oral history and through their scarcity.
+
+Those historical documents from the early 17th century to the modern ethnography described how local indigenous people use those ornaments in local cultural contexts that represents some social roles or status.
+However, compared to European contact period, there is less mention of beads in Chinese contact period and the description about ornaments is only limited to their dressing culture.
+
+# Kiwulan in northeastern Taiwan
+
+Kiwulan (Figure 1) is located at Yilan, north of Lanyang Plain in northeastern Taiwan, surrounded by the Central Mountain Range in the west and the Pacific Ocean in the east.
+Kiwulan was excavated from 2001 to 2004, and the total excavation area was 3,814 m2.
+Most beads were found during the excavation and some of them through screens with 20 mm and 1.5 mm mesh.
+The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures.
+The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates[a], which coordinate to the late Iron Age and Proto-historical period in Taiwan.
+This paper focuses on the Upper Layer Culture which overlaps with the contact periods with the Europeans and Chinese.
+It provides a good examples to discuss the foreign impacts on local indigenous society through body ornaments that usually convey social and cultural meaning by displaying.
+
+The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local settlements were attacked by the Spanish according to the official documents (add citation; Borao Mateo 2009).
+Later in 1647, the Dutch attacked the indigenous villages in Yilan and forced them to accept colonial rules and economic demands by paying annual tribute (add citation; Borao Mateo 2009).
+According to Dutch census reports in 1650, Kiwulan was the biggest indigenous community in the Yilan Plain, with a population of 840 adults (Chen 2007; Kang 2012; Li 2014).
+The diagnostic artifacts during European contact found at Kiwulan including An-ping jars and glass beads that were largely introduced to Taiwan during the early 17th century.
+
+The European colonization ended in 1662 when they were defeated by the kingdom of Tungning found by Koxinga from China.
+Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
+The evidence of Chinese migration can be identified from Chinese official records [b]and large amount of Chinese blue-and-white porcelain found at Kiwulan.[c]
+
+The ornaments sampled in this study come from 40 adjacent 4m x 4m units out of 262 units in total since those sampled units were less disturbed by modern constructions and remains intact.
+The Upper Layer Culture can be divided into six sub layers spanning from 14 century to 19 century according to current radiocarbon dates, excavation depth, types of ceramics, and sediment texture and color (Hsieh 2008; Wang 2011).
+Layers are assigned numbers from 1 to 6 for which L1 corresponds to the upper context and L6, also the bottom layer, refers to lower context.
+Based on Hsieh’s (2008) chronological study of Chinese porcelains, L5 and L6 represent pre-European contact period, L4 was the time of European contact, and L2 and L1 was the period of Chinese contact.
+
+Graph of 40 units and chronology
+
+In general, intact ornaments with clear context were mostly found in burials that helps to understand how people wear them and any relationships between shapes of ornaments and gender or age.
+In addition to burials, ornaments were also found in middens and living space that can be identified by post holes.
+The frequency of ornaments from burials is more than those from the living space due to well preserved condition.
+There is a wide variety of raw materials for ornaments of which ornaments glass, agate and metal ornaments that account for 99% of all ornaments were thought to be introduced by trading with the Europeans and Chinese traders, while ornaments made of shells, woods, and bones were believed locally made that is about only about 1% of all ornaments (Table 1) .
+The reason for the large quantity of glass beads is that they usually have been found in clusters because of the original form such as necklaces.
+The shape of ornaments include beads, bracelets, rings, bells, pendants, and knitted objects that shows shape are closely related to raw material due to its characteristic.
+For example, beads are only made of glass, agate, and shell, and bone is the main material for knitted objects that used to connect beads.
+Also, metal ornaments covers more shapes such bells, pendants, and rings  (Chen 2007).
+
+# Results
+
+This research compares the chronological and spatial distribution of trade ornaments during different time periods with foreign contacts, including European and Chinese Contacts.
+The principle of classification for ornaments is based on the material and shape that was used in the excavation report.
+For quantitative comparison between different time periods, minor types of ornaments with similar shapes from the same material are combined as one category that will be explained in each material section.
+The results are presented by the general distribution across the 40 sample squares, followed by each major types of materials from three analytical units that are before European contact, during European contact, and Chinese contact.
+
+chronological distribution
+
+Sketch of results section: 
+H1
+   1. We compare Raw material, shape/type and counts of ornaments in three time periods, and by spatial distribution 
+   2.
 
@@
 ```
-
-# Methods
-
-# Results
 

### 2019-04-22 analysis/paper/paper.Rmd
sketch of results section

@@
 Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact.
-We hypothesise that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
+We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
 We hypothesise that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.
@@
 
-Sketch of results section: 
-H1
-   1. We compare Raw material, shape/type and counts of ornaments in three time periods, and by spatial distribution 
-   2.
+Sketch of results section, should have some graphs and tables for these:
+
+1. Overall summary of  Raw material, shape/type and counts of ornaments
+2. Ornaments by: 
+ - time period
+ -- for each time period, spatial patterning
 

### 2019-04-22 analysis/paper/paper.Rmd
undates

--- 2019-04-22 analysis/paper/paper.Rmd
@@
 
-```{r read-in-the-data}
+```{r read-in-the-data, eval = TRUE}
 library(here)
@@
 
-```{r tidy-data}
+```{r tidy-data, eval = TRUE}
 # summarise the ornaments by layers
@@
 ```{r scatter-plot-bead-by-period}
-
 # plot for intact golden bead by period
@@
 
-```{r read-in-the-spatial-data}
+```{r read-in-the-spatial-data, eval = TRUE}
 library(sf)
@@
 
-```{r tidy-data2}
+```{r tidy-data2, eval = TRUE}
 # 40 sampling units

### 2019-04-22 analysis/paper/paper.Rmd
fix conflict

--- 2019-04-22 analysis/paper/paper.Rmd
@@
 
-```{r tidy-data}
+```{r tidy-data, eval = TRUE}
+
 # summarise the ornaments by layers
@@
 
-```{r read-in-the-spatial-data}
+```{r read-in-the-spatial-data, eval = TRUE}
+
 library(sf)

### 2019-05-18 analysis/paper/paper.Rmd
updates

@@
 
-Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display (Joyce 2005).
-Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appear in Taiwan around 1,800 years ago through regional trade network in Southeast Asia (Wang and Liu 2007).
+Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005].
+Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appear in Taiwan around 1,800 years ago through regional trade network in Southeast Asia [@WangandLiu2007].
 Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the Metal Age and reached a high peak of quantity and variety in the early 17th century when the Europeans arrived and involved the island in an international trade network on a large scale.
@@
 Direct colonial rule was rare, but the question of long-lasting indirect impacts on local communities remains unanswered.
-Indirect effects of colonialism is addressed recently to discuss the impact on the indigenous societies in the periphery of colonial control (Trabert 2017).
+Indirect effects of colonialism is addressed recently to discuss the impact on the indigenous societies in the periphery of colonial control [@Trabert2017].
 Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where Northeastern Taiwan is an especially unique example for discussing indirect effects due to its physically isolated location that made it not easily accessible to the Europeans that the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012).
@@
 
-In general, intact ornaments with clear context were mostly found in burials that helps to understand how people wear them and any relationships between shapes of ornaments and gender or age.
+In general, complete ornaments with clear context were mostly found in burials that helps to understand how people wear them and any relationships between shapes of ornaments and gender or age.
 In addition to burials, ornaments were also found in middens and living space that can be identified by post holes.
@@
 
-# Results
+# Methods
 
-This research compares the chronological and spatial distribution of trade ornaments during different time periods with foreign contacts, including European and Chinese Contacts.
-The principle of classification for ornaments is based on the material and shape that was used in the excavation report.
-For quantitative comparison between different time periods, minor types of ornaments with similar shapes from the same material are combined as one category that will be explained in each material section.
-The results are presented by the general distribution across the 40 sample squares, followed by each major types of materials from three analytical units that are before European contact, during European contact, and Chinese contact.
-
-chronological distribution
-
-Sketch of results section, should have some graphs and tables for these:
-
-1. Overall summary of  Raw material, shape/type and counts of ornaments
-2. Ornaments by: 
- - time period
- -- for each time period, spatial patterning
+To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research exmines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
+Trade ornaments were analyzed by three analytical units, including pre-European contact, European contact, and Chinese contact.
+Because the appearance of ornament is highly correlated to the raw material itself, which is selected as the first criteria for classification.
+Under each category of raw material, ornaments are further classified into subtypes according to shapes when variation are found in each category.
+To compare the general pattern of changes in major ornaments, less frequent ornaments are combined as one category named other, which will be discussed separately.
 
@@
 ```{r tidy-data, eval = TRUE}
-
 # summarise the ornaments by layers
@@
   rowwise() %>% 
-  mutate(Perforation_average = mean(c(`Perforation1`, `Perforation2(mm)`), 
+  mutate(Perforation_average =
+           mean(c(`Perforation1`,`Perforation2(mm)`), 
                 na.rm = TRUE))
+
+sample_size <- nrow(ornaments_period)
+total_AD <- nrow(kwl_upper)
+sample_pro <- sample_size/total_AD
 ```
 
-```{r plot-counts}
+# Results
+
+1. Overall summary of  Raw material, shape/type and counts of ornaments
+There are `r sample_size` ornaments excavated from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the excavation section where they were from.
+Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
+Ornaments in some categories are divided into sub-types due to the variations of shape, such as agate bead, metal ring, and metal bell.
+Table ?
+shows the subtypes for the major ornaments.
+
+To investigate whether the types of ornaments changes resulted from culture contact and how they changed, we compare ornaments from pre-European, European contact period, and Chinese contact period.
+The results a general pattern that most ornaments appeared in the pre-European contact period, reached a peak in quantity during the European contact and dropped down during the Chinese contact.
+This trend can be seen on the ornaments including golden bead, agate bead, metal ring, and bell.
+However, glass bead shows different pattern that higher frequency in the pre-European contact, and then the quantity goes down over time in the European and Chinese contact periods.
+
+2. Ornaments by: 
+ - time period
+
+ - for each time period, spatial patterning
+
+```{r plot-all-ornaments-count, fig.cap= "Frequency of ornaments"}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 
-ornaments_count_test <-
+ornaments_count <-
   ornaments_period %>% 
-  filter(!is.na(`6-layers`),
-         Categories != "Unknown Metal") %>%
+  filter(Categories != "Unknown Metal") %>%
   ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
@@
 
-ggplot(ornaments_count_test,
+ggplot(ornaments_count,
        aes(Categories, n)) +
@@
 ```
+
+```{r tab-all}
+ornaments_shape <-
+  ornaments_period %>% 
+  ungroup() %>% 
+  select(Categories, Type) %>% 
+  distinct(Categories, Type, .keep_all = TRUE) %>% 
+  filter(Categories != c("Unknown Metal", "Bone Ornament"))
+
+knitr::kable(ornaments_shape)
+```
+
+%>% 
+  spread(Categories, value)
+
+ filter(Categories != "Unknown Metal") %>% 
+%>% 
+  mutate(Categories = factor(Categories,
+                         level = c("Agate bead", 
+                                 "European Contact", 
+                                 "Chinese Contact"))) %>%
 

### 2019-05-18 analysis/paper/paper.Rmd
work on table of counts

@@
 output: 
-    bookdown::html_document2:
+    bookdown::word_document2:
       fig_caption: yes
@@
 Ornaments in some categories are divided into sub-types due to the variations of shape, such as agate bead, metal ring, and metal bell.
-Table ?
-shows the subtypes for the major ornaments.
+Table \@ref(tab:tab-all) shows the subtypes for the major ornaments.
 
@@
 
-```{r tab-all}
+```{r taball}
+# clean some 
+ornaments_period_clean <- 
+ornaments_period %>% 
+  mutate(Type = str_remove(Type, "prism/"),
+         Type = ifelse(Type == "glass bead", "medium bead", Type))
+
 ornaments_shape <-
-  ornaments_period %>% 
-  ungroup() %>% 
-  select(Categories, Type) %>% 
-  distinct(Categories, Type, .keep_all = TRUE) %>% 
-  filter(Categories != c("Unknown Metal", "Bone Ornament"))
+  ornaments_period_clean %>% 
+  group_by(period, Categories, Type ) %>% 
+  tally() %>% 
+  spread(period, n, fill = 0) %>% 
+  filter(!Categories %in% c("Unknown Metal",
+
+                            "Perforated plate", 
+                            "Knited item",
+                            "Hair pin",
+                            "Bone Ornament",
+                            "Metal Piece",
+                            "Glass ring"))
 
@@
 ```
+
+Table \@ref(tab:taball) shows the subtypes for the major ornaments.
 

### 2019-05-25 analysis/paper/paper.Rmd
updates

@@
 Table \@ref(tab:tab-all) shows the subtypes for the major ornaments.
+## Comparisons between time periods
 
-To investigate whether the types of ornaments changes resulted from culture contact and how they changed, we compare ornaments from pre-European, European contact period, and Chinese contact period.
-The results a general pattern that most ornaments appeared in the pre-European contact period, reached a peak in quantity during the European contact and dropped down during the Chinese contact.
-This trend can be seen on the ornaments including golden bead, agate bead, metal ring, and bell.
-However, glass bead shows different pattern that higher frequency in the pre-European contact, and then the quantity goes down over time in the European and Chinese contact periods.
+To investigate whether the types of ornaments changes resulted from culture contact and how they changed, we compared ornaments from pre-European, European contact period, and Chinese contact period.
+The results show a general pattern that most ornaments appeared in the pre-European contact period, their frequencies reach a peak during the European contact and dropped significantly in the Chinese contact period, especially for golden bead.
+This trend can be also seen on ornaments including agate bead, metal ring, and bell.
+However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
+
+To further understand whether there are some changes in the frequency of subtypes mainly defined by the shape, we compared the distribution of frequency for each major categories. For the
+
+For golden bead, we focus on their length that
 
@@
   mutate(Type = str_remove(Type, "prism/"),
-         Type = ifelse(Type == "glass bead", "medium bead", Type))
+         Type = ifelse(Type == "glass bead", "medium bead", Type),
+         Type = ifelse(Type == "small I", "plain small", Type),
+         Type = ifelse(Type == "small II", "thin small", Type))
 
@@
   filter(!Categories %in% c("Unknown Metal", 
-
                             "Perforated plate", 
@@
 ```
-
-Table \@ref(tab:taball) shows the subtypes for the major ornaments.
-
-%>% 
-  spread(Categories, value)
-
- filter(Categories != "Unknown Metal") %>% 
-%>% 
-  mutate(Categories = factor(Categories,
-                         level = c("Agate bead", 
-                                 "European Contact", 
-                                 "Chinese Contact"))) %>%
 

### 2019-06-10 analysis/paper/paper.Rmd
rearrange the section order

@@
 We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
-We hypothesise that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.
+We hypothesize that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.
 
@@
 The frequency of ornaments from burials is more than those from the living space due to well preserved condition.
-There is a wide variety of raw materials for ornaments of which ornaments glass, agate and metal ornaments that account for 99% of all ornaments were thought to be introduced by trading with the Europeans and Chinese traders, while ornaments made of shells, woods, and bones were believed locally made that is about only about 1% of all ornaments (Table 1) .
+There is a wide variety of raw materials for ornaments of which ornaments glass, agate and metal ornaments that account for 99% of all ornaments were thought to be introduced by trading with the Europeans and Chinese traders, while ornaments made of shells, woods, and bones were believed locally made that is about only about 1% of all ornaments (Table 1).
 The reason for the large quantity of glass beads is that they usually have been found in clusters because of the original form such as necklaces.
@@
 Also, metal ornaments covers more shapes such bells, pendants, and rings (Chen 2007).
-
-# Methods
-
-To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research exmines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
-Trade ornaments were analyzed by three analytical units, including pre-European contact, European contact, and Chinese contact.
-Because the appearance of ornament is highly correlated to the raw material itself, which is selected as the first criteria for classification.
-Under each category of raw material, ornaments are further classified into subtypes according to shapes when variation are found in each category.
-To compare the general pattern of changes in major ornaments, less frequent ornaments are combined as one category named other, which will be discussed separately.
 
@@
 
+# Samples and Methods
+
+To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research examines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
+Trade ornaments were analyzed by three analytical units, including pre-European contact, European contact, and Chinese contact.
+Because the appearance of ornament is highly correlated to the raw material itself, which is selected as the first criteria for classification.
+Under each category of raw material, ornaments are further classified into subtypes according to shapes when variation are found in each category.
+To compare the general pattern of changes in major ornaments, less frequent ornaments are combined as one category named other, which will be discussed separately.
+
+There are `r sample_size` ornaments excavated from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the excavation section where they were from.
+Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
+Ornaments in some categories are divided into sub-types due to the variations of shape, including agate bead, metal ring, and metal bell.
+The frequency of the subtypes are summarized in \@ref(tab:tidy-data-for-table).
+
 # Results
 
-1. Overall summary of  Raw material, shape/type and counts of ornaments
-There are `r sample_size` ornaments excavated from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the excavation section where they were from.
-Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
-Ornaments in some categories are divided into sub-types due to the variations of shape, such as agate bead, metal ring, and metal bell.
-Table \@ref(tab:tab-all) shows the subtypes for the major ornaments.
-## Comparisons between time periods
+## Frequency between time periods
 
-To investigate whether the types of ornaments changes resulted from culture contact and how they changed, we compared ornaments from pre-European, European contact period, and Chinese contact period.
-The results show a general pattern that most ornaments appeared in the pre-European contact period, their frequencies reach a peak during the European contact and dropped significantly in the Chinese contact period, especially for golden bead.
-This trend can be also seen on ornaments including agate bead, metal ring, and bell.
+The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden bead.
+This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
 However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
-
-To further understand whether there are some changes in the frequency of subtypes mainly defined by the shape, we compared the distribution of frequency for each major categories. For the
+The distribution of frequency for subtypes in each major category are presented in (make a figure).
+Agate bead can be divided into 7 subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and the number increased in the Eruopean contact and then declined in the Chinese contact.
+Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
+There are more subtypes during European contact during the European contact compared to other time periods.
 
@@
 
-```{r taball}
+```{r tidy-data-for-table}
 # clean some 
@@
   ornaments_period_clean %>% 
-  group_by(period, Categories, Type ) %>% 
+  group_by(period, Categories, Type) %>% 
   tally() %>% 
@@
 
-knitr::kable(ornaments_shape)
+knitr::kable(ornaments_shape,
+             caption = "A table of ornament subtypes")
 ```

### 2019-06-14 analysis/paper/paper.Rmd
some updates

@@
 Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005].
-Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appear in Taiwan around 1,800 years ago through regional trade network in Southeast Asia [@WangandLiu2007].
-Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the Metal Age and reached a high peak of quantity and variety in the early 17th century when the Europeans arrived and involved the island in an international trade network on a large scale.
-Foreign trade goods introduced into the local indigenous societies in an European colonial or imperial contexts usually had an impact on local culture and societies in many parts of the world that usually involved in the transformation of indigenous economic, cultural, and socio-political systems (Dietler 1997; Junker 1993; Mitchell 2000; Silliman 2005).
-Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that stress active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices (Given 2004; Mullins 2011; Silliman 2001; Rubertone 2000; Torrence 2000; Torrence and Clarke 2000; Voss 2005).
+Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@WangandLiu2007].
+Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age.
+We observed a greater quantity and variety of foreign goods in the early 17th century during the European canonization period that involved the whole island in an international trade network on a large scale.
+Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems (Dietler 1997; Junker 1993; Mitchell 2000; Silliman 2005).
+Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices (Given 2004; Mullins 2011; Silliman 2001; Rubertone 2000; Torrence 2000; Torrence and Clarke 2000; Voss 2005).
 
 Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
-Direct colonial rule was rare, but the question of long-lasting indirect impacts on local communities remains unanswered.
-Indirect effects of colonialism is addressed recently to discuss the impact on the indigenous societies in the periphery of colonial control [@Trabert2017].
-Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where Northeastern Taiwan is an especially unique example for discussing indirect effects due to its physically isolated location that made it not easily accessible to the Europeans that the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012).
-In addition, prior to the European contact, there were small-scale regional exchange networks between China and Taiwan.
+Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local communities remains unanswered.
+Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017].
+Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012).
+Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period.
 Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century.
-This made northeastern Taiwan a good case to explore culture contacts at different periods and their indirect impacts by taking a long-term perspective.
+Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.
 
 We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan.
-Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact.
+Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network.
 We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
@@
 
+Q: Is the difference in the frequency of ornaments related to the scale of exchange network, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
+
+Q: what is the scale of those exchange network: 
+- pre-European contact: middle scale, regional trade network 
+- European contact: global scale, international trade network 
+- Chinsese contact: small scale, mutual trade network
+
+Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
+- pre-European contact: indirect 
+- European contact: indirect, colonial context
+- Chinsese contact: direct, Han people migration
+
+ and intense of contact
+
 # Cultural context of Ornaments in Northeastern Taiwan
+
+<--Add a paragraph about exchange network before European contact-->
 
@@
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The evidence of Chinese migration can be identified from Chinese official records [b]and large amount of Chinese blue-and-white porcelain found at Kiwulan.[c]
+The evidence of Chinese migration can be identified from Chinese official records [b]and large amount of Chinese blue-and-white porcelain found at Kiwulan.
 
-The ornaments sampled in this study come from 40 adjacent 4m x 4m units out of 262 units in total since those sampled units were less disturbed by modern constructions and remains intact.
+The ornaments studied in this paper came from 40 adjacent 4m x 4m units out of 262 units in the middle part of excavation.
+The sampling principle is based on the extent of intact contexts that the least disturbed by modern constructions are the preference.
 The Upper Layer Culture can be divided into six sub layers spanning from 14 century to 19 century according to current radiocarbon dates, excavation depth, types of ceramics, and sediment texture and color (Hsieh 2008; Wang 2011).
@@
 
-Graph of 40 units and chronology
+<--add a graph of 40 units and chronology-->
 
@@
 
-# Samples and Methods
-
-To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research examines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
-Trade ornaments were analyzed by three analytical units, including pre-European contact, European contact, and Chinese contact.
-Because the appearance of ornament is highly correlated to the raw material itself, which is selected as the first criteria for classification.
-Under each category of raw material, ornaments are further classified into subtypes according to shapes when variation are found in each category.
-To compare the general pattern of changes in major ornaments, less frequent ornaments are combined as one category named other, which will be discussed separately.
-
-There are `r sample_size` ornaments excavated from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the excavation section where they were from.
-Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
-Ornaments in some categories are divided into sub-types due to the variations of shape, including agate bead, metal ring, and metal bell.
-The frequency of the subtypes are summarized in \@ref(tab:tidy-data-for-table).
-
-# Results
-
-## Frequency between time periods
-
-The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden bead.
-This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
-However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
-The distribution of frequency for subtypes in each major category are presented in (make a figure).
-Agate bead can be divided into 7 subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and the number increased in the Eruopean contact and then declined in the Chinese contact.
-Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
-There are more subtypes during European contact during the European contact compared to other time periods.
-
-For golden bead, we focus on their length that
-
-2. Ornaments by: 
- - time period
-
- - for each time period, spatial patterning
-
 ```{r plot-all-ornaments-count, fig.cap= "Frequency of ornaments"}
@@
 
+# Samples and Methods
+
+To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research examines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
+Trade ornaments were analyzed by three analytical units including pre-European contact, European contact, and Chinese contact that indicates three scale of exchange network from less to intense contact subsequently.
+Because the appearance of ornaments is highly related to the original raw material couple with the function such as bead and bracelet, they are combined as the first criterion of classification which determines five major categories.
+Under each category of raw material, ornaments are further classified into several subtype according to the shape if any variations were identified in each category.
+In order to compare the general pattern of changes for major ornaments, less frequent ornaments are combined as one category named other, which will be discussed separately.
+
+Ornaments with clear contexts were mainly found from the post-hole area, burials, and middens, where ornaments from the post-hole area accounts for the majority of the amount in total.
+Post-hole area is also the contexts that we mainly focus on because it represents house units accross the site that can further provide a basis for comparison to see if there is any spatial pattern of distribution.
+Ornaments from middens and burials will also be discussed in the later section.
+For the post-hole area, there were `r sample_size` ornaments recovered from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the same excavation section including all contexts.
+Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
+Ornaments in most categories can be divided into sub-types according to the variation of shape.
+The variation was defined by the difference identified in at least two variables, for example dimensional variables such as length, or stylish variables such as patterns.
+The frequency and feature of subtypes for major categories are summarized in \@ref(tab:tidy-data-for-table).
+
+# Results
+
+## Frequency between time periods
+
+The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
+This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
+However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
+
+The distribution of frequency for subtypes in each major category are presented in (make a figure to combine all).
+Agate beads can be divided into 7 subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and the number increased in the Eruopean contact and then declined in the Chinese contact.
+Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
+There are more subtypes of agate beads during European contact compared to other time periods.
+For golden bead, we focus on their length
+
+## Frequency for spatial pattern
+2. Ornaments by: 
+ - time period
+
+ - for each time period, spatial patterning
+
 ```{r plot-counts-by-period}
@@
 
-```{r scatter-plot-bead-by-period}
+```{r scatter-plot-bead-by-perio, eval= FALSE}
 # plot for intact golden bead by period
@@
 ```{r plot-count-type-by-period}
+# tidy the data with categories and shapes
+ornaments_shape_count <-
+  ornaments_period_clean %>% 
+  group_by(period, Categories, Type) %>% 
+  tally() %>% 
+  filter(!Categories %in% c("Unknown Metal", 
+                            "Perforated plate", 
+                            "Knited item",
+                            "Hair pin",
+                            "Bone Ornament",
+                            "Metal Piece",
+                            "Glass ring",
+                            "Golden bead"))
+
 # plot for different shape of agate bead by period

### 2019-06-14 analysis/paper/paper.Rmd
updates based on Ben's suggestions

@@
 We hypothesize that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.
-
-Q: Is the difference in the frequency of ornaments related to the scale of exchange network, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
-
-Q: what is the scale of those exchange network: 
-- pre-European contact: middle scale, regional trade network 
-- European contact: global scale, international trade network 
-- Chinsese contact: small scale, mutual trade network
-
-Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
-- pre-European contact: indirect 
-- European contact: indirect, colonial context
-- Chinsese contact: direct, Han people migration
-
- and intense of contact
 
@@
 
-```{r plot-count-type-by-period}
+```{r tidy-shape-by-period}
 # tidy the data with categories and shapes
@@
                             "Glass ring",
-                            "Golden bead"))
+                            "Golden bead"))  %>% 
+  ungroup() %>% 
+  mutate(Type = as.factor(Type)) 
+```
 
+```{r agate-bead-fre-plot}
 # plot for different shape of agate bead by period
 agate_bead_shape <-
-  ornaments_period %>% 
-  filter(Categories == 'Agate bead') %>%
-  ungroup() %>% 
-  mutate(Type = ifelse(Type == 'prism/octagonal', 
-                       str_remove(Type, '/octagonal'), Type)) %>% 
-  mutate(Type = ifelse(Type == 'prism/pentagonal', 
-                       str_remove(Type, '/pentagonal'), Type)) %>% 
-  mutate(Type = ifelse(Type == 'prism/hexagonal', 
-                       str_remove(Type, 'prism/'), Type)) %>%
-  mutate(Type = fct_infreq(Type)) %>%
-  mutate(Type = fct_rev(Type)) %>% 
-  group_by(period, Type) %>% 
-  count()
+  ornaments_shape_count %>% 
+  filter(Categories == 'Agate bead')
 
+agate_bead_shape_plot <- 
 ggplot(agate_bead_shape) +
-  geom_col(aes(Type, n)) +
+  geom_col(aes(reorder(Type, n), n)) +
   facet_wrap(~ period) +
@@
   theme_minimal(base_size = 14)
+```
 
+```{r metal-fre-plot}
+# plot for different shape of metal ring by period
+ring_shape <-
+  ornaments_shape_count %>% 
+  filter(Categories == 'Metal ring')
+
+ring_shape_plot <-
+ggplot(ring_shape) +
+  geom_col(aes(reorder(Type, n), n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency",
+       x = "Shape of Metal ring") +
+  theme_minimal(base_size = 14)
+```
+
+```{r bell-fre-plot}
 # plot for different shape of metal bell by period
 bell_shape <-
-  ornaments_period %>% 
+  ornaments_shape_count %>% 
   filter(Categories == 'Bell',
-         Type != 'unclassified') %>% 
-  ungroup() %>% 
-  mutate(Type = fct_rev(Type)) %>% 
-  group_by(period, Type) %>% 
-  count()
+         Type != 'unclassified')
 
+bell_shape_plot <-
 ggplot(bell_shape) +
-  geom_col(aes(Type, n)) +
+  geom_col(aes(reorder(Type,n), n)) +
   facet_wrap(~ period) +
@@
   theme_minimal(base_size = 14)
+```
 
-# plot for different shape of metal ring by period
-ring_shape <-
-  ornaments_period %>% 
-  filter(Categories == 'Metal ring') %>%
-  ungroup() %>% 
-  mutate(Type = fct_infreq(Type)) %>%
-  mutate(Type = fct_rev(Type)) %>% 
-  group_by(period, Type) %>% 
-  count()
+```{r glass-bead-fre-plot}
+# plot for different shape of glass bead by period
+glass_bead_shape <-
+  ornaments_shape_count %>% 
+  filter(Categories == 'Glass bead')
 
-ggplot(ring_shape) +
+glass_bead_plot <-
+ggplot(glass_bead_shape) +
   geom_col(aes(Type, n)) +
@@
   labs(y = "Frequency",
-       x = "Shape of Metal ring") +
+       x = "Shape of Glass Bead") +
   theme_minimal(base_size = 14)
+```
+
+```{r}
+library(cowplot)
+plot_grid(agate_bead_shape_plot, ring_shape_plot,
+          labels = c("A", "B"))
+
+plot_grid(bell_shape_plot, glass_bead_plot,
+          labels = c("C", "D"))
 ```
@@
 
+Q: Is the difference in the frequency of ornaments related to the scale of exchange network, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
+
+Q: what is the scale of those exchange network: 
+- pre-European contact: middle scale, regional trade network 
+- European contact: global scale, international trade network 
+- Chinese contact: small scale, mutual trade network
+
+Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
+- pre-European contact: indirect 
+- European contact: indirect, colonial context
+- Chinese contact: direct, Han people migration
+
+ and intense of contact
+
 # Conclusion

### 2019-06-17 analysis/paper/paper.Rmd
added some text and one spatial plot

@@
          Type = ifelse(Type == "small I", "plain small", Type),
-         Type = ifelse(Type == "small II", "thin small", Type))
+         Type = ifelse(Type == "small II", "thin small", Type),
+         Type = ifelse(Type == "large thin string", "thin large", Type))
 
@@
 The frequency and feature of subtypes for major categories are summarized in \@ref(tab:tidy-data-for-table).
-
-# Results
-
-## Frequency between time periods
-
-The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
-This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
-However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
-
-The distribution of frequency for subtypes in each major category are presented in (make a figure to combine all).
-Agate beads can be divided into 7 subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and the number increased in the Eruopean contact and then declined in the Chinese contact.
-Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
-There are more subtypes of agate beads during European contact compared to other time periods.
-For golden bead, we focus on their length
-
-## Frequency for spatial pattern
-2. Ornaments by: 
- - time period
-
- - for each time period, spatial patterning
 
@@
   ornaments_shape_count %>% 
-  filter(Categories == 'Agate bead')
+  filter(Categories == 'Agate bead',
+         Type != 'other')
 
@@
   ornaments_shape_count %>% 
-  filter(Categories == 'Metal ring')
+  filter(Categories == 'Metal ring',
+         !Type %in% c('large thick string',
+                       'small thin string'))
 
@@
 
-```{r}
+```{r plot-altogether, fig.cap= "Frequency of "}
 library(cowplot)
 plot_grid(agate_bead_shape_plot, ring_shape_plot,
-          labels = c("A", "B"))
+          bell_shape_plot, glass_bead_plot,
+          labels = c("A", "B", "C", "D"), ncol=2)
+```
+# Results
 
-plot_grid(bell_shape_plot, glass_bead_plot,
-          labels = c("C", "D"))
-```
+## Frequency between time periods
+The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
+This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
+However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
 
-```{r scatter-plot-metal-by-period}
+The distribution of frequency for subtypes in each major category are presented in \@ref(fig:plot-all-ornaments-count).
+The distribution shows that agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period.
+Agate beads can be divided into eight subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased during the European contact and then declined in the Chinese contact.
+Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
+There are more subtypes of agate beads during European contact compared to other time periods.
+Similarly metal rings can be classified into seven subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape.
+The greater varieties for those two categories might indicate the multiple sources due to global trade network brought by the Europeans.
+In contrast, copper bell and glass bead have smaller variety, but glass bead has larger number due to they were usually found in a cluster that reflect the original use as strings.
+Most common shape for copper bell is the large shape with human face as motif, while most common shape for glass bead is small that were identified as Indo-Pacific beads, the most common type that are widespread in Southeast Asian sites since 300 BC [@Francis2002].
+
+## Frequency for spatial pattern
+Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period.
+It shows that ornaments concentrate on one specific square in the pre-European contact, and more ubiquitous during the European contact and
+
+ - for each time period, spatial patterning
+
+```{r scatter-plot-metal-by-period, eval= FALSE}
 # plot for intact bell by period
@@
   theme(plot.title = element_text(hjust=0.5))
-
-# plot for intact ring by period
-ornaments_intact_ring <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Metal ring') %>% 
-  mutate(Length = as.numeric(gsub("\\+|\\-", "", `Length(mm)`))) #remove + -
-
-ggplot(ornaments_intact_ring) +
-  geom_point(aes(Length, Width, col = Type)) +
-  facet_wrap(~ period) +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Metal ring") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
 ```
@@
 ```{r read-in-the-spatial-data, eval = TRUE}
-
 library(sf)
@@
 
-```{r tidy-data2, eval = TRUE}
+```{r tidy-spatial-data, eval = TRUE}
 # 40 sampling units
@@
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
+```
+
+```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of ornaments"}
+ggplot(ornaments_period_join) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join, aes(fill =n), alpha = 0.9) +
+  facet_wrap(~ period) +
+  labs(fill = "frequency") +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
 ```
@@
 
-```{r-gganimate}
+```{r-gganimate, eval= FALSE}
 # test gganimate for measurement of beads by period
@@
 ```
+
+Word count: `r wordcountaddin::word_count()`

### 2019-06-19 analysis/paper/paper.Rmd
more texts and one plot

@@
 We observed a greater quantity and variety of foreign goods in the early 17th century during the European canonization period that involved the whole island in an international trade network on a large scale.
-Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems (Dietler 1997; Junker 1993; Mitchell 2000; Silliman 2005).
-Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices (Given 2004; Mullins 2011; Silliman 2001; Rubertone 2000; Torrence 2000; Torrence and Clarke 2000; Voss 2005).
+Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997;@Junker1993;@Mitchell2000;@Silliman2005).
+Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004;@Mullins2011;@Silliman2001;@Rubertone2000;@Torrence2000;@TorrenceandClarke2000;@Voss2005).
 
@@
 
+```{r read-in-the-spatial-data, eval = TRUE}
+library(sf)
+library(cartography)
+burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
+AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
+AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
+location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
+units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
+```
+
+```{r tidy-spatial-data, eval = TRUE}
+# 40 sampling units
+sample_units <-
+  AD_data %>% 
+  semi_join(units_40, by = c (`坑號` = "the_sq"))
+
+# join two dataset to get coordinate
+ornaments_period_join <-
+  ornaments_period %>%
+  mutate(period = as.factor(period)) %>% # seems need this line first 
+  group_by(Pit_No, period) %>% 
+  count(Categories) %>% 
+  inner_join(AD_data, by = c("Pit_No" = "坑號"))
+```
+
 The ornaments studied in this paper came from 40 adjacent 4m x 4m units out of 262 units in the middle part of excavation.
@@
 Also, metal ornaments covers more shapes such bells, pendants, and rings (Chen 2007).
+
+# Samples and Methods
 
@@
 
-# Samples and Methods
-
 To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research examines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
@@
   coord_flip() +
-  labs(y = "Frequency",
-       x = "Categories") +
+  labs(y = "Frequency", x = "Categories") +
   theme_minimal(base_size = 14)
+
+# stack histogram
+ggplot(ornaments_count) +
+  geom_col(aes(Categories, n, fill = period)) +
+  labs(y = "Frequency", x = "Categories") +
+  coord_flip() +
+  theme_minimal(base_size = 12) +
+  scale_fill_viridis_d()
 ```
@@
 Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period.
-It shows that ornaments concentrate on one specific square in the pre-European contact, and more ubiquitous during the European contact and
+It shows that one square unearthed more ornaments than others in the pre-European contact.
+During the European contact period, the distribution is more ubiquitous with some concentration on northern part, and more even during Chinese contact period.
+Further look at the distribution for each major category individually, it shows that some squares have higher number of ornaments during the European contact period, including golden bead, agate bead, metal ring, and bell.
+However, there is no consistent pattern of concentration for different ornaments across the area.
+The squares with high number of ornaments distributed separately and independently without clear correlation.
+For example, more golden beads were found at the northern part, while more agate beads concentrated in the middle part.
 
- - for each time period, spatial patterning
+Golden beads and
 
@@
 
-```{r read-in-the-spatial-data, eval = TRUE}
-library(sf)
-library(cartography)
-burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
-AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
-AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
-location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
-units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
-```
-
-```{r tidy-spatial-data, eval = TRUE}
-# 40 sampling units
-sample_units <-
-  AD_data %>% 
-  semi_join(units_40, by = c (`坑號` = "the_sq"))
-
-# join two dataset to get coordinate
-ornaments_period_join <-
-  ornaments_period %>%
-  mutate(period = as.factor(period)) %>% # seems need this line first 
-  group_by(Pit_No, period) %>% 
-  count(Categories) %>% 
-  inner_join(AD_data, by = c("Pit_No" = "坑號"))
-```
-
 ```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of ornaments"}
@@
 
-```{r plot-spatial-distribution}
+```{r plot-spatial-pattern-each}
 # filter golden bead
@@
   labs(title = "Spatial Distribution of Agate Beads",
+       fill = "frequency") +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+
+# filter metal ring
+ornaments_period_join_metal <-
+  ornaments_period_join %>% 
+  filter(Categories %in% 'Metal ring')
+
+# plot by period
+ggplot(ornaments_period_join_metal) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_metal, aes(fill = n), alpha = 0.9) +
+  facet_wrap(~ period) +
+  labs(title = "Spatial Distribution of Metal Ring",
+       fill = "frequency") +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+
+# filter metal ring
+ornaments_period_join_bell <-
+  ornaments_period_join %>% 
+  filter(Categories %in% 'Bell')
+
+# plot by period
+ggplot(ornaments_period_join_bell) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_bell, aes(fill = n), alpha = 0.9) +
+  facet_wrap(~ period) +
+  labs(title = "Spatial Distribution of Copper Bell",
        fill = "frequency") +

### 2019-06-20 analysis/paper/paper.Rmd
update plot and some spatial analysis

@@
 
+```{r read-in-the-data}
+library(here)
+library(tidyverse)
+library(ggplot2)
+kwl_upper <- readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
+```
+
+```{r tidy-data, eval = TRUE}
+# summarise the ornaments by layers
+# L5 and L6 represent pre-European contact period, 
+# L4 was the time of European contact, and 
+# L2 and L1 was the period of Chinese contact.
+
+ornaments_period <- 
+kwl_upper %>% 
+  filter(!is.na(`6-layers`)) %>% 
+  mutate(period = case_when(
+    `6-layers` %in% 1:2 ~ "Chinese Contact",
+    `6-layers` == 4 ~ "European Contact",
+    `6-layers` %in% 5:6 ~ "Before European Contact",
+     TRUE ~ "other"
+  )) %>% 
+  filter(period != "other") %>% 
+  mutate(period = factor(period,
+                       level = c("Before European Contact", 
+                                 "European Contact", 
+                                 "Chinese Contact"))) %>% 
+  mutate(Length = as.numeric(`Length(mm)`),
+         Thick = as.numeric(`Thick(mm)`),
+         Width = as.numeric(`Width(mm)`),
+         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
+  rowwise() %>% 
+  mutate(Perforation_average =
+           mean(c(`Perforation1`,`Perforation2(mm)`), 
+                na.rm = TRUE))
+
+sample_size <- nrow(ornaments_period)
+total_AD <- nrow(kwl_upper)
+sample_pro <- sample_size/total_AD
+```
+
 ```{r read-in-the-spatial-data, eval = TRUE}
@@
 # Samples and Methods
-
-```{r read-in-the-data}
-library(here)
-kwl_upper <- readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
-```
-
-```{r tidy-data, eval = TRUE}
-# summarise the ornaments by layers
-library(tidyverse)
-library(ggplot2)
-
-# L5 and L6 represent pre-European contact period, 
-# L4 was the time of European contact, and 
-# L2 and L1 was the period of Chinese contact.
-
-ornaments_period <- 
-kwl_upper %>% 
-  filter(!is.na(`6-layers`)) %>% 
-  mutate(period = case_when(
-    `6-layers` %in% 1:2 ~ "Chinese Contact",
-    `6-layers` == 4 ~ "European Contact",
-    `6-layers` %in% 5:6 ~ "Before European Contact",
-     TRUE ~ "other"
-  )) %>% 
-  filter(period != "other") %>% 
-  mutate(period = factor(period,
-                       level = c("Before European Contact", 
-                                 "European Contact", 
-                                 "Chinese Contact"))) %>% 
-  mutate(Length = as.numeric(`Length(mm)`),
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`),
-         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
-  rowwise() %>% 
-  mutate(Perforation_average =
-           mean(c(`Perforation1`,`Perforation2(mm)`), 
-                na.rm = TRUE))
-
-sample_size <- nrow(ornaments_period)
-total_AD <- nrow(kwl_upper)
-sample_pro <- sample_size/total_AD
-```
 
@@
   coord_flip() +
-  labs(y = "Frequency",
-       x = "Shape of Agate bead") +
-  theme_minimal(base_size = 14)
+  labs(y = "",
+       x = "Agate bead") +
+  theme_minimal(base_size = 14) +
+  theme(strip.background = element_blank(), 
+        strip.text = element_blank())
 ```
@@
          !Type %in% c('large thick string',
-                       'small thin string'))
+                       'small thin string')) %>% 
+  ungroup() %>% 
+  mutate(Type = fct_drop(Type)) %>% 
+  mutate(Type = fct_infreq(Type)) %>% 
+  mutate(Type = fct_rev(Type))
 
@@
 ggplot(ring_shape) +
-  geom_col(aes(reorder(Type, n), n)) +
+  geom_col(aes(Type, n)) +
+  coord_flip() +
+  labs(y = "",
+       x = "Metal ring") +
   facet_wrap(~ period) +
-  coord_flip() +
-  labs(y = "Frequency",
-       x = "Shape of Metal ring") +
   theme_minimal(base_size = 14) 
@@
   coord_flip() +
-  labs(y = "Frequency",
-       x = "Shape of Bell") +
-  theme_minimal(base_size = 14)
+  labs(y = "",
+       x = "Bell") +
+  theme_minimal(base_size = 14) +
+  theme(strip.background = element_blank(), 
+        strip.text = element_blank())
 ```
@@
   labs(y = "Frequency",
-       x = "Shape of Glass Bead") +
-  theme_minimal(base_size = 14)
+       x = "Glass bead") +
+  theme_minimal(base_size = 14) +
+  theme(strip.background = element_blank(), 
+        strip.text = element_blank())
 ```
@@
 library(cowplot)
-plot_grid(agate_bead_shape_plot, ring_shape_plot,
-          bell_shape_plot, glass_bead_plot,
-          labels = c("A", "B", "C", "D"), ncol=2)
+plot_grid(ring_shape_plot,
+          agate_bead_shape_plot, 
+          bell_shape_plot, 
+          glass_bead_plot,
+          ncol=1,
+          align = "h",
+          axis = "lr",
+          rel_widths = 1,
+          rel_heights = c(1,1,0.7,0.62))
 ```
@@
 Most common shape for copper bell is the large shape with human face as motif, while most common shape for glass bead is small that were identified as Indo-Pacific beads, the most common type that are widespread in Southeast Asian sites since 300 BC [@Francis2002].
-
-## Frequency for spatial pattern
-Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period.
-It shows that one square unearthed more ornaments than others in the pre-European contact.
-During the European contact period, the distribution is more ubiquitous with some concentration on northern part, and more even during Chinese contact period.
-Further look at the distribution for each major category individually, it shows that some squares have higher number of ornaments during the European contact period, including golden bead, agate bead, metal ring, and bell.
-However, there is no consistent pattern of concentration for different ornaments across the area.
-The squares with high number of ornaments distributed separately and independently without clear correlation.
-For example, more golden beads were found at the northern part, while more agate beads concentrated in the middle part.
-
-Golden beads and
 
@@
 
+## Frequency for spatial pattern
+Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all classes of ornaments for each time period.
+It shows that one square unearthed more ornaments than others in the pre-European contact.
+During the European contact period, the presence of ornaments is more ubiquitous with some clusters on northern part, and more randomly distributed during the Chinese contact period.
+When looking at the distribution for each major class individually, it shows that some clusters of different class of ornaments across the area during the European contact period, including golden bead, agate bead, metal ring, and bell.
+However, it seems no clear consistent pattern of concentration for those ornaments.
+Each class shows its own pattern that squares with higher number of ornaments distributed separately and independently.
+For example, the cluster of golden bead were found at the northern part, while the cluster of agate bead were found in the middle part.
+In contrast, there are more than one cluster of metal rings that distribute separately across the research area.
+Copper bells were usually found solely and seem randomly across the area.
+Before European contact, the map shows greater amount of glass beads were found at the northern and middle part of the research area, and more widespread during the European contact period.
+In the Chinese Contact period, both the amount and density of different classes of ornaments decreases that they were only found in a few squares, for example, golden bead and glass bead were only found in a couple squares at the middle part.
+
+```{r point-pattern-data-setup}
+# get centroid 
+ornaments_period_join_cen <-
+  ornaments_period_join %>% 
+  mutate(lon = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
+         lat = map_dbl(geometry, ~st_point_on_surface(.x)[[2]]))
+
+# plot the centroid
+ggplot() +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_point(data = ornaments_period_join_cen, 
+             aes(x = lon, y = lat))
+
+# transform data to PPP format and sp SpatialPoints
+library(spatstat)
+
+#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points
+lon <- ornaments_period_join_cen$lon
+lat <- ornaments_period_join_cen$lat
+xrange <- range(lon, na.rm=T)
+yrange <- range(lat, na.rm=T)
+
+# create ppp
+ornaments_point <- ppp(lon, lat, xrange, yrange)
+
+```
+
+To access whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods.
+
 ```{r-gganimate, eval= FALSE}

### 2019-06-21 analysis/paper/paper.Rmd
added density and kernel density plot

@@
 library(spatstat)
+library(maptools)
+library(raster)
 
@@
 
-# create ppp
-ornaments_point <- ppp(lon, lat, xrange, yrange)
+# create ppp and sp(spacial points) format
+ornaments_point_ppp <- ppp(lon, lat, xrange, yrange)
+ornaments_point_sp <- as(ornaments_point_ppp, "SpatialPoints")
+r_ornaments <- raster(sample_units)
 
+# plot density
+den_ornaments <- rasterize(coordinates(ornaments_point_sp), r_ornaments, fun = 'count', background = 0)
+plot(den_ornaments)
+
+# kernel density estimation
+ds <- density(ornaments_point_ppp)
+plot(ds, main = "Ornament density")
+
+kernel_or <- density(ornaments_point_ppp)
+kernel_or %>%  # Using the default bandwidth
+  plot(main= "Ornament density", las=1)
+contour(kernel_or, add=TRUE)
 ```

### 2019-06-22 analysis/paper/paper.Rmd
code for nndist hypothesis testing and similation, make repetitive code shorter by map()function

--- 2019-06-20 analysis/paper/paper.Rmd
@@
 library(cartography)
-burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
-AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
-AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
-location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
-units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
+burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
+AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
+AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
+location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
+units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
 ```
@@
 ```{r point-pattern-data-setup}
-# get centroid 
+# get centroid and make a list of three periods
+ornaments_period_join_cen_all <-
+  ornaments_period_join %>% 
+  mutate(lon = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[1]])),
+         lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]])))
+
 ornaments_period_join_cen <- 
-  ornaments_period_join %>% 
-  mutate(lon = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
-         lat = map_dbl(geometry, ~st_point_on_surface(.x)[[2]]))
+ ornaments_period_join_cen_all %>% 
+  group_by(period) %>% 
+  group_split() %>% 
+  set_names(c("Pre-European",
+              "European",
+              "Chinese"))
+```
 
-# plot the centroid
-ggplot() +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_point(data = ornaments_period_join_cen, 
-             aes(x = lon, y = lat))
-
+```{r}
 # transform data to PPP format and sp SpatialPoints
 library(spatstat)
+library(maptools)
+library(raster)
 
-#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points
-lon <- ornaments_period_join_cen$lon
-lat <- ornaments_period_join_cen$lat
-xrange <- range(lon, na.rm=T)
-yrange <- range(lat, na.rm=T)
+#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points, by period
 
-# create ppp
-ornaments_point <- ppp(lon, lat, xrange, yrange)
+# use common range for each period
+lon_all <- ornaments_period_join_cen_all$lon
+lat_all <- ornaments_period_join_cen_all$lat
+xrange <- range(lon_all, na.rm=T)
+yrange <- range(lat_all, na.rm=T)
 
+# create a function to apply to each period
+density_analysis <- function(x, ...) {
+  # create ppp and sp(spacial points) format
+lon_tmp <- x$lon
+lat_tmp <- x$lat
+ornaments_point_ppp_tmp <- ppp(lon_tmp, lat_tmp, xrange, yrange)
+ornaments_point_sp_tmp <- as(ornaments_point_ppp_tmp, "SpatialPoints")
+
+# kernel density estimation with contour
+kernel_tmp <- density(ornaments_point_ppp_tmp)
+kernel_to_plot <- 
+  plot(kernel_tmp, ...)
+  contour(kernel_tmp, add = TRUE)
+
+return(list(ornaments_point_ppp = ornaments_point_ppp_tmp,
+            kernel_to_plot = kernel_to_plot))
+}
 ```
 
-To access whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods.
+```{r plot-all-densities}
+par(mfrow = c(1, 3))
+# apply function to each data frame in the list
+titles <- names(ornaments_period_join_cen)
+
+kernels_for_plotting <- 
+  map2(ornaments_period_join_cen, 
+       titles,
+      ~density_analysis(.x, main = .y))
+```
+
+```{r}
+# function for hypothesis testing
+nndist_hypo_testing <- function(x, n = 1e3L,  ... ) {
+
+ann_tmp <- mean(nndist(x$ornaments_point_ppp, k = 1))
+
+# object for storing simulated ANN values
+ann_r_tmp <- vector(length = n)
+
+for (i in 1:n) {
+  # Generate random point locations
+  rand_tmp <-  rpoint(n = x$ornaments_point_ppp$n,
+                      win = as.owin(as_Spatial(sample_units)))
+  ann_r_tmp[i] <-
+    mean(nndist(rand_tmp, k = 1))  # Tally the ANN values
+}
+
+the_plot <- 
+hist(
+  ann_r_tmp,
+  xlab = NULL,
+  las = 1,
+  breaks = 40,
+  col = "bisque",
+  xlim = range(ann_tmp, ann_r_tmp),
+  ...)
+
+abline(v = ann_tmp, col = "blue")
+
+return(the_plot)
+}
+```
+
+```{r}
+
+par(mfrow = c(3, 1))
+invisible(
+map2(kernels_for_plotting, 
+    titles,
+    ~invisible(nndist_hypo_testing(.x, 
+                         main = .y)))
+)
+```
+
+To assess whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods.
+We assume a uniform point density across the research area before European contact period, that is, randomly distributed, and test the null hypothesis that is there is no cluster of ornaments across the sapce during the European contact periods.
 

### 2019-06-22 analysis/paper/notes.R

@@ -1,0 +1,9 @@
+# plot the centroid
+map(ornaments_period_join_cen,
+    ~ggplot() +
+      geom_sf(data = sample_units, fill = NA) +
+      geom_point(data = .x,
+                 aes(x = lon,
+                     y = lat))) %>%
+  plot_grid(plotlist  = .)
+

### 2019-06-26 analysis/paper/paper.Rmd
paragraphs for kernel plots and hypo-testing

@@
 
-```{r point-pattern-data-setup}
+```{r point-pattern-data-tidy}
 # get centroid and make a list of three periods
@@
 
-```{r}
+```{r transformation-sptial-data}
 # transform data to PPP format and sp SpatialPoints
@@
 
-```{r plot-all-densities}
+```{r plot-all-densities, fig.cap= "Kernal density map"}
 par(mfrow = c(1, 3))
@@
 ```
+
+Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period.
+The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution.
+We randomly assigned points, here the location of ornaments, for each square according to the data, subset three groups for three time periods, and compute the kernal density for each time period for comparison.
+Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
+Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier.etal].
+Density value is assigned to each cell, of which the unit is meter.
+The results \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period.
+Those maps present three consistent sub-region with shifting core area over time.
+The distribution might indicate different social groups who own more ornaments, and the multiple groups during European contact period might reflect a more corporate mode of social organization.
+However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious core area.
+Next step is to test whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.
 
@@
 
-```{r}
+```{r hypo-test-distrubution, fig.cap= "Histogram of simulated ANN values"}
 
@@
 
-To assess whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods.
-We assume a uniform point density across the research area before European contact period, that is, randomly distributed, and test the null hypothesis that is there is no cluster of ornaments across the sapce during the European contact periods.
-
-```{r-gganimate, eval= FALSE}
-# test gganimate for measurement of beads by period
-# devtools::install_github('thomasp85/gganimate')
-library(gganimate)
-
-anim_gold_bead <-
-  ggplot(ornaments_intact_gold,
-       aes(Length, Width)) +
-  geom_point() +
-  transition_states(period, 3, 1) +
-  labs(title = 'Period: {closest_state}',
-       x ='Length (mm) of golden bead',
-       y ='Width (mm) of golden bead')
-
-# does not work
-anim_gold_bead_spatial <-
-  ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
-  transition_states(period, 3, 1) +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-```
+Here we use Monte Carlo method to test for the complete spatial randomness of spatial events based on the average nearest-neighbor distance method.
+We hypothesize that the distribution of ornaments is consistent with a completely random process, and then simulate the presence of ornaments across the space for one thousand times for each time periods.
+We compare our data to the random pattern generated by simulated processes to see if the our data is randomly distributed or not.
+The results \@ref(fig:hypo-test-distrubution) are represented by histogram showing one thousand simulation with the blue line indicating our observed value.
+The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments are clustered distributed.
+Similar result are also observed during the pre-European period but less extreme.
+The distribution of ornaments was random during the Chinese contact period and about one third of the simulated values are greater than our observed ANN value.
+This testing explains that the distribution of ornaments during the pre-European and European contact period might reflect the presence of different social groups.
+Moreover, the cluster of ornaments way more significant during the European contact period that might indicate some extent of control of ornaments.
 

### 2019-06-26 analysis/paper/notes.R

@@ -7,3 +7,31 @@ map(ornaments_period_join_cen,
                      y = lat))) %>%
   plot_grid(plotlist  = .)
 
+```{r-gganimate, eval= FALSE}
+# test gganimate for measurement of beads by period
+# devtools::install_github('thomasp85/gganimate')
+library(gganimate)
+
+anim_gold_bead <-
+  ggplot(ornaments_intact_gold,
+         aes(Length, Width)) +
+  geom_point() +
+  transition_states(period, 3, 1) +
+  labs(title = 'Period: {closest_state}',
+       x ='Length (mm) of golden bead',
+       y ='Width (mm) of golden bead')
+
+# does not work
+anim_gold_bead_spatial <-
+  ggplot(ornaments_period_join_gold) +
+  geom_sf(data = sample_units, fill = NA) +
+  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
+  transition_states(period, 3, 1) +
+  scale_fill_gradient(high = "black", low = "grey") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_text(size = 11))
+
+```

### 2019-06-26 analysis/paper/paper.Rmd

@@
 Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017].
-Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012).
+Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012).
 Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period.
@@
 We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
-We hypothesize that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.
+We hypothesize that there are clustered distribution of ornaments at Kiwulan relating to these international contacts.
 
@@
 A man who wanted to take a wife had to give her parents cuentas.
-If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled (Li 2006: 151).
-Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals (Pao 2008: 122, 143, 151).
+If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled [@Li2006:151].
+Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals [@Pao 2008: 122, 143, 151].
 Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity.
@@
 The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need.
-The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave (Li 2006: 153).
+The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave [@Li 2006: 153].
 
 Documents from the Qing dynasty usually describe the decorative purposes of the ornaments.
-In the Kavalan Subprefecture Gazetteer (Chen 2006[1852]: 308), the Kavalan Zhi Lue (Ko 1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao 1996[1829]: 77), it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments.
+In the Kavalan Subprefecture Gazetteer [@Chen2006[1852]: 308], the Kavalan Zhi Lue [@Ko1993[1837]: 11, 126) and the Dong Cha Ji Lue [@Yao1996[1829]: 77], it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments.
 Their number was small and their value high, hence only the rich possessed them.
 Ordinary people wore agate beads or glass beads on their head or neck during various festivals.
-In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori (Kanori 1996: 227-232) made field surveys among various ethnic groups in Taiwan.
+In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori [@Kanori1996: 227-232] made field surveys among various ethnic groups in Taiwan.
 He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.
 
-Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans (Liu 2008: 133-134).
+Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans [@Liu2008: 133-134].
 Interview records describe agate beads as valuable objects passed down from mother to daughter.
@@
 But the origin of these beads is still not known.
-Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer (Hsu 1992: 22) and carp-shaped ornaments had completely disappeared.
+Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer [@Hsu1992: 22] and carp-shaped ornaments had completely disappeared.
 Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination.
@@
 
+## Spatial distribution
 Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period.
@@
 
+The introduction of ornaments to Northeastern Taiwan during different periods could inform how local indigenous people perceive and distribute the trading goods in different contexts of culture contacts by examining their frequency and spatial distribution across space.
+Taiwan island had been involved in regional exchange networks of Southeast Asia since Neolithic period, a global trade network during the European contact period in the early 17th century, and later intense but smaller scale network with Han Chinese communities since 19th century [@Chen200?].
+Kiwulan demonstrates continuous occupation from the late Iron Age to one hundred years ago over 600 years that covers these three different culture contacts that enables a comparative research to explore the indirect influence of culture contact especially in the global colonial context.
+These three analytic chronology at Kiwulan represent different scale of exchange network and intensity of culture contact.
+For the scale of trade network, Pre-European contact can be defined as a middle scale represented by the prehistoric regional trade network in Southeast Asia.
+International trade network during the European contact presents a global scale, while local trade network during the Chinese contact presents a small scale.
+For the intensity of culture contact, both pre-European contact and European contact period could be defined as indirect contact, but the colonial power underlying the contact with the Europeans should be also considered.
+The culture contact during the Chinese period is more direct due to the migration of Han Chinese to Kiwulan.
+
+The results show there was a greater diversity of ornaments types and materials at Kiwulan during European contact.
+This might result from the large scale exchange network that stimulate the circulation of different ornaments classes.
+Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads, which might indicate the diverse origins.
+However, the frequency of overall ornaments and each subtype declines significantly during the Chinese contact period.
+The reason could be related to smaller scale of network and the overall decline of Kiwulan population at that time.
+The spatial pattern of ornaments shows that there were some clusters appeared since pre-European and European contact period.
+The hypothesis testing for spatial pattern further indicate more concentration of ornaments during European contact.
+Because trade ornaments had been viewed as prestige goods in local cultural context in the prehistoric Northeastern Taiwan, this differentiation could reflect some level of social inequality appeared before European contact and then it was reinforced during European period.
+
+Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
+
 Q: Is the difference in the frequency of ornaments related to the scale of exchange network, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
 
-Q: what is the scale of those exchange network: 
-- pre-European contact: middle scale, regional trade network 
-- European contact: global scale, international trade network 
-- Chinese contact: small scale, mutual trade network
+# Conclusion
 
-Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
-- pre-European contact: indirect 
-- European contact: indirect, colonial context
-- Chinese contact: direct, Han people migration
-
- and intense of contact
-
-# Conclusion
+Indirect influence of colonialism has been proposed recently and some students in many parts of the world [@Trabert2017].
+Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts.
 

### 2019-06-27 analysis/paper/paper.Rmd

@@
 ```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of ornaments"}
-ggplot(ornaments_period_join) +
+ggplot() +
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join, aes(fill =n), alpha = 0.9) +
+  geom_sf(data = ornaments_period_join, 
+          aes(geometry = geometry, 
+              fill =n), 
+          alpha = 0.9) +
   facet_wrap(~ period) +
@@
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
+  geom_sf(data = ornaments_period_join_gold, 
+          aes(geometry = geometry,
+              fill = n), alpha = 0.9) +
   facet_wrap(~ period) +
@@
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
+  geom_sf(data = ornaments_period_join_gold, 
+          aes(geometry = geometry, 
+              fill = n), alpha = 0.9) +
   facet_wrap(~ period) +
@@
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_metal, aes(fill = n), alpha = 0.9) +
+  geom_sf(data = ornaments_period_join_metal, 
+          aes(geometry = geometry,
+              fill = n), alpha = 0.9) +
   facet_wrap(~ period) +
@@
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_bell, aes(fill = n), alpha = 0.9) +
+  geom_sf(data = ornaments_period_join_bell, 
+          aes(geometry = geometry,
+              fill = n), alpha = 0.9) +
   facet_wrap(~ period) +

### 2019-06-27 analysis/paper/paper.Rmd
conclusion and a paragraph in background section

@@
 
-# Introduction and Research Question
+# Introduction
 
@@
 Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age.
-We observed a greater quantity and variety of foreign goods in the early 17th century during the European canonization period that involved the whole island in an international trade network on a large scale.
+We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.
 Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997;@Junker1993;@Mitchell2000;@Silliman2005).
@@
 
-<--Add a paragraph about exchange network before European contact-->
+Although the northeastern Taiwan forms a circumscribed environment due to the surrounding mountains and sea, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups in northern and southern Taiwan, foreign traders such as the Han people of China, and Europeans.
+This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan (Chen 2005; Wang and Liu 2007).
+The groups involved in the exchange network consisted by the Chinese traders, indigenous groups called Basay in northern Taiwan, and indigenous groups, Kavalan, in northeastern Taiwan.
+Historical records indicate that the Kavalan people were one of the Basay’s trading partners, and they exchanged rice, deer skin, and gold for metal tools or ornaments, beads, and Chinese ceramics (Chen 2005; Hsieh 2009a; Li and Wu 2006).
 
@@
 
-Indirect influence of colonialism has been proposed recently and some students in many parts of the world [@Trabert2017].
+Indirect influence of colonialism has been proposed recently and some studies reveal how peripheral areas had been influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017].
 Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts.
+Foreign trade ornaments were viewed as prestige goods and played an important role in the cultural context of indigenous societies in northeastern Taiwan according to the use context as grave goods and historical documents written by foreign record keeper.
+As a status marker, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
+Greater amount and diversity ornament types during the European contact period reflects the international stimulation of exchange in a colonial context.
+Those ornaments may convey different meaning with exotic and colonizer image that increases more competitions between aggrandized individuals for prestige and wealth accumulation, which might result in a increasing social inequality.
+The spatial pattern of ornaments at Kiwulan during European contact presents multiple clusters that is significantly different from a random distribution pattern.
+This study demonstrate that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization.
+It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system.
+During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwalan population.
+This is a case study that focuses on trade ornaments and more studies can be done to have a comparison with other archaeological evidence.
+Further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.
 

### 2019-06-29 analysis/paper/paper.Rmd
edited references and citation

@@
 Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005].
-Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@WangandLiu2007].
+Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@Wang2007].
 Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age.
 We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.
-Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997;@Junker1993;@Mitchell2000;@Silliman2005).
-Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004;@Mullins2011;@Silliman2001;@Rubertone2000;@Torrence2000;@TorrenceandClarke2000;@Voss2005).
+Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
+Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004; @Mullins2011; @Silliman2001; @Rubertone2000; @Torrence2000; @Torrence2000a; @Voss2005].
 
@@
 Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017].
-Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012).
+Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
 Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period.
@@
 
-We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan.
+We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan[@Chen2007].
 Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network.
@@
 Although the northeastern Taiwan forms a circumscribed environment due to the surrounding mountains and sea, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups in northern and southern Taiwan, foreign traders such as the Han people of China, and Europeans.
-This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan (Chen 2005; Wang and Liu 2007).
+This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan [@Chen2005; @Wang2007].
 The groups involved in the exchange network consisted by the Chinese traders, indigenous groups called Basay in northern Taiwan, and indigenous groups, Kavalan, in northeastern Taiwan.
-Historical records indicate that the Kavalan people were one of the Basay’s trading partners, and they exchanged rice, deer skin, and gold for metal tools or ornaments, beads, and Chinese ceramics (Chen 2005; Hsieh 2009a; Li and Wu 2006).
+Historical records indicate that the Kavalan people were one of the Basay’s trading partners, and they exchanged rice, deer skin, and gold for metal tools or ornaments, beads, and Chinese ceramics [@Chen2005; @Hsieh2009; @LiandWu2006].
 
@@
 The report of the Dominican Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) as a materials to exchange with necessities they need with other indigenous people.
-This trend then gradually spread to other areas, and even the Spanish soldiers used agate as bargaining chips for gambling, since agate equaled money (Li 2006: 132-149).
+This trend then gradually spread to other areas, and even the Spanish soldiers used agate as bargaining chips for gambling, since agate equaled money [@LiandWu2006, pp.
+132-149].
 European contact was focused on trading with indigenous communities in northeastern Taiwan.
@@
 A man who wanted to take a wife had to give her parents cuentas.
-If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled [@Li2006:151].
-Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals [@Pao 2008: 122, 143, 151].
+If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled [@LiandWu2006, pp.
+151].
+Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals [@Pao2008, pp.
+122-151].
 Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity.
@@
 The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need.
-The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave [@Li 2006: 153].
+The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave [@LiandWu2006, pp.
+153].
 
@@
 Ordinary people wore agate beads or glass beads on their head or neck during various festivals.
-In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori [@Kanori1996: 227-232] made field surveys among various ethnic groups in Taiwan.
+In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori [@Kanori1996, pp.
+227-232] made field surveys among various ethnic groups in Taiwan.
 He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.
 
-Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans [@Liu2008: 133-134].
+Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans [@Liu2008, pp.
+133-134].
 Interview records describe agate beads as valuable objects passed down from mother to daughter.
@@
 But the origin of these beads is still not known.
-Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer [@Hsu1992: 22] and carp-shaped ornaments had completely disappeared.
+Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer [@Hsu1992, pp.
+22] and carp-shaped ornaments had completely disappeared.
 Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination.
@@
 
-The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local settlements were attacked by the Spanish according to the official documents (add citation; Borao Mateo 2009).
-Later in 1647, the Dutch attacked the indigenous villages in Yilan and forced them to accept colonial rules and economic demands by paying annual tribute (add citation; Borao Mateo 2009).
-According to Dutch census reports in 1650, Kiwulan was the biggest indigenous community in the Yilan Plain, with a population of 840 adults (Chen 2007; Kang 2012; Li 2014).
+The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local settlements were attacked by the Spanish according to the official documents [add citation; Borao Mateo 2009].
+Later in 1647, the Dutch attacked the indigenous villages in Yilan and forced them to accept colonial rules and economic demands by paying annual tribute [add citation; Borao Mateo 2009].
+According to Dutch census reports in 1650, Kiwulan was the biggest indigenous community in the Yilan Plain, with a population of 840 adults [@Chen2007; @Kang2012; @Li2014].
 The diagnostic artifacts during European contact found at Kiwulan including An-ping jars and glass beads that were largely introduced to Taiwan during the early 17th century.
@@
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The evidence of Chinese migration can be identified from Chinese official records [b]and large amount of Chinese blue-and-white porcelain found at Kiwulan.
+The evidence of Chinese migration can be identified from Chinese official records and large amount of Chinese blue-and-white porcelain found at Kiwulan.
 
@@
 The sampling principle is based on the extent of intact contexts that the least disturbed by modern constructions are the preference.
-The Upper Layer Culture can be divided into six sub layers spanning from 14 century to 19 century according to current radiocarbon dates, excavation depth, types of ceramics, and sediment texture and color (Hsieh 2008; Wang 2011).
+The Upper Layer Culture can be divided into six sub layers spanning from 14 century to 19 century according to current radiocarbon dates, excavation depth, types of ceramics, and sediment texture and color [@Hsieh2009; @Wang2011].
 Layers are assigned numbers from 1 to 6 for which L1 corresponds to the upper context and L6, also the bottom layer, refers to lower context.
-Based on Hsieh’s (2008) chronological study of Chinese porcelains, L5 and L6 represent pre-European contact period, L4 was the time of European contact, and L2 and L1 was the period of Chinese contact.
+Based on @Hsieh2009 chronological study of Chinese porcelains, L5 and L6 represent pre-European contact period, L4 was the time of European contact, and L2 and L1 was the period of Chinese contact.
 
@@
 For example, beads are only made of glass, agate, and shell, and bone is the main material for knitted objects that used to connect beads.
-Also, metal ornaments covers more shapes such bells, pendants, and rings (Chen 2007).
+Also, metal ornaments covers more shapes such bells, pendants, and rings [@Chen2007].
 
@@
 ## Spatial distribution
-Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period.
+Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period [@Ducke2015].
 The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution.
@@
 Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
-Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier.etal].
+Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier2019, @Cortegoso2016].
 Density value is assigned to each cell, of which the unit is meter.
@@
 The introduction of ornaments to Northeastern Taiwan during different periods could inform how local indigenous people perceive and distribute the trading goods in different contexts of culture contacts by examining their frequency and spatial distribution across space.
-Taiwan island had been involved in regional exchange networks of Southeast Asia since Neolithic period, a global trade network during the European contact period in the early 17th century, and later intense but smaller scale network with Han Chinese communities since 19th century [@Chen200?].
+Taiwan island had been involved in regional exchange networks of Southeast Asia since Neolithic period, a global trade network during the European contact period in the early 17th century, and later intense but smaller scale network with Han Chinese communities since 19th century [@Chen2005].
 Kiwulan demonstrates continuous occupation from the late Iron Age to one hundred years ago over 600 years that covers these three different culture contacts that enables a comparative research to explore the indirect influence of culture contact especially in the global colonial context.

### 2019-06-29 analysis/paper/notes.R

@@ -7,7 +7,7 @@ map(ornaments_period_join_cen,
                      y = lat))) %>%
   plot_grid(plotlist  = .)
 
-```{r-gganimate, eval= FALSE}
+
 # test gganimate for measurement of beads by period
 # devtools::install_github('thomasp85/gganimate')
 library(gganimate)

### 2019-07-01 analysis/paper/paper.Rmd
added map in the script file

@@
 ## Spatial distribution
-Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period [@Ducke2015].
+Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period [@Bevan2016, @Ducke2015].
 The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution.

### 2019-07-01 analysis/paper/map.R
added map in the script file

@@ -1,0 +1,53 @@
+library("ggplot2")
+theme_set(theme_bw())
+library("sf")
+library("rnaturalearth")
+library("rnaturalearthdata")
+library("here")
+
+world <- ne_countries(scale = "medium", returnclass = "sf")
+class(world)
+world_points <- st_centroid(world)
+world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))
+
+# read Taiwan administrative areas
+TW_adm <- st_read(here("analysis", "data", "raw_data","TWN_adm2.shp"))
+TW_map <- fortify(TW_adm)
+
+# add site location
+Kiwulan <- data.frame(lon = c(121.7809), lat = c(24.8066))
+
+TW_SE_Asia <-
+ggplot(data = world) +
+  geom_sf() +
+  geom_rect(xmin = 119.8, xmax = 122.2, ymin = 21.7, ymax = 25.5,
+            fill = NA, colour = "red", size = 0.5) +
+  coord_sf(xlim = c(107, 135), ylim = c(13, 35), expand = FALSE, datum = NA)
+
+TW_SE_Asia_void <-
+  TW_SE_Asia +
+  theme_void() +
+  theme(panel.border = element_rect(colour = "black",
+                                    fill = NA))
+
+library("ggspatial")
+TW_map_with_site <-
+ggplot(data = world) +
+  geom_sf(data = TW_map) +
+  geom_point(data = Kiwulan, aes(x = lon, y = lat),
+             size = 3, color = "red") +
+  geom_text(aes(x = lon, y = lat, label = "Kiwulan"),
+            data = Kiwulan, position = position_nudge(y = -0.1)) +
+  coord_sf(xlim = c(118, 122.2), ylim = c(21.9, 25.4), expand = FALSE) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true",
+                         pad_x = unit(0.25, "in"), pad_y = unit(0.45, "in"),
+                         style = north_arrow_fancy_orienteering)
+
+TW_map_with_site +
+  annotation_custom(
+    grob = ggplotGrob(TW_SE_Asia),
+    xmin = 118,
+    xmax = 120,
+    ymin = 23.5,
+    ymax = 25.5)

### 2019-07-01 analysis/paper/map.R
update for map

@@ -19,30 +19,30 @@ Kiwulan <- data.frame(lon = c(121.7809), lat = c(24.8066))
 
 TW_SE_Asia <-
 ggplot(data = world) +
-  geom_sf() +
+  geom_sf(fill= "antiquewhite") +
   geom_rect(xmin = 119.8, xmax = 122.2, ymin = 21.7, ymax = 25.5,
             fill = NA, colour = "red", size = 0.5) +
-  coord_sf(xlim = c(107, 135), ylim = c(13, 35), expand = FALSE, datum = NA)
-
-TW_SE_Asia_void <-
-  TW_SE_Asia +
-  theme_void() +
-  theme(panel.border = element_rect(colour = "black",
-                                    fill = NA))
+  coord_sf(xlim = c(107, 135), ylim = c(13, 35), expand = FALSE, datum = NA) +
+  theme(panel.background = element_rect(fill = "azure"))
 
 library("ggspatial")
 TW_map_with_site <-
 ggplot(data = world) +
-  geom_sf(data = TW_map) +
+  geom_sf(data = TW_map, fill= "antiquewhite") +
   geom_point(data = Kiwulan, aes(x = lon, y = lat),
              size = 3, color = "red") +
-  geom_text(aes(x = lon, y = lat, label = "Kiwulan"),
-            data = Kiwulan, position = position_nudge(y = -0.1)) +
+  geom_text(data = Kiwulan,
+            aes(x = lon, y = lat, label = "Kiwulan"),
+            fontface = "bold",
+            position = position_nudge(y = -0.15)) +
   coord_sf(xlim = c(118, 122.2), ylim = c(21.9, 25.4), expand = FALSE) +
   annotation_scale(location = "bl", width_hint = 0.5) +
   annotation_north_arrow(location = "bl", which_north = "true",
                          pad_x = unit(0.25, "in"), pad_y = unit(0.45, "in"),
-                         style = north_arrow_fancy_orienteering)
+                         style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "azure"),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
 
 TW_map_with_site +
   annotation_custom(
@@ -50,4 +50,4 @@ TW_map_with_site +
     xmin = 118,
     xmax = 120,
     ymin = 23.5,
-    ymax = 25.5)
+    ymax = 25.7)

### 2019-07-01 analysis/paper/map.R
map with topo

@@ -4,9 +4,13 @@ library("sf")
 library("rnaturalearth")
 library("rnaturalearthdata")
 library("here")
+library("ggspatial")
+# devtools::install_github('3wen/legendMap')
+library(legendMap)
+
 
 world <- ne_countries(scale = "medium", returnclass = "sf")
-class(world)
+# class(world)
 world_points <- st_centroid(world)
 world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))
 
@@ -20,7 +24,7 @@ Kiwulan <- data.frame(lon = c(121.7809), lat = c(24.8066))
 TW_SE_Asia <-
 ggplot(data = world) +
   geom_sf() +
-  geom_rect(xmin = 119.8, xmax = 122.2, ymin = 21.7, ymax = 25.5,
+  geom_rect(xmin = 121, xmax = 122.0, ymin = 24.3, ymax = 25.2,
             fill = NA, colour = "red", size = 0.5) +
   coord_sf(xlim = c(107, 135), ylim = c(13, 35), expand = FALSE, datum = NA)
 
@@ -30,24 +34,63 @@ TW_SE_Asia_void <-
   theme(panel.border = element_rect(colour = "black",
                                     fill = NA))
 
-library("ggspatial")
+library(ggmap)
+library(tmaptools)
+library(shadowtext)
+# we don't want to download every time, so let's save the map locally
+# from https://stackoverflow.com/a/52710855/1036500
+# tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(121.0, 24.3,
+#                                                  122.0, 25.2))), zoom = 10))
+# saveRDS(tw_map, here("analysis", "data", "raw_data", "tw_map.rds"))
+tw_map <- readRDS(here("analysis", "data", "raw_data", "tw_map.rds"))
+pg <- ggplot_build(tw_map)
+
 TW_map_with_site <-
-ggplot(data = world) +
-  geom_sf(data = TW_map) +
-  geom_point(data = Kiwulan, aes(x = lon, y = lat),
-             size = 3, color = "red") +
-  geom_text(aes(x = lon, y = lat, label = "Kiwulan"),
-            data = Kiwulan, position = position_nudge(y = -0.1)) +
-  coord_sf(xlim = c(118, 122.2), ylim = c(21.9, 25.4), expand = FALSE) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true",
[…]

### 2019-07-01 analysis/paper/map.R
map tiles are too big for git

@@ -34,10 +34,10 @@ library(tmaptools)
 library(shadowtext)
 # we don't want to download every time, so let's save the map locally
 # from https://stackoverflow.com/a/52710855/1036500
-# tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(121.0, 24.3,
-#                                                  122.0, 25.2))), zoom = 10))
+ tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(121.0, 24.3,
+                                                  122.0, 25.2))), zoom = 10))
 # saveRDS(tw_map, here("analysis", "data", "raw_data", "tw_map.rds"))
-tw_map <- readRDS(here("analysis", "data", "raw_data", "tw_map.rds"))
+# tw_map <- readRDS(here("analysis", "data", "raw_data", "tw_map.rds"))
 pg <- ggplot_build(tw_map)
 
 TW_map_with_site <-

### 2019-07-05 analysis/paper/paper.Rmd
updates for introduction and references

@@
 
-Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005].
-Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@Wang2007].
-Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age.
-We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.
-Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
-Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004; @Mullins2011; @Silliman2001; @Rubertone2000; @Torrence2000; @Torrence2000a; @Voss2005].
+1. background of the topic
 
+Taiwan island became part of a global trading network during the early 17th century because of western European expansion that connected European and Asia.
 Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
-Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local communities remains unanswered.
-Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017].
-Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
-Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period.
-Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century.
-Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.
+Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local indigenous communities remains unanswered.
+Indirect effects of colonialism is emphasized increasingly to discuss the colonial impact on the indigenous people in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
+In many parts of the world, foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts might cause significant transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
+Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed to emphasize the ability of active adoptions of foreign effects, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
+Studying the consumption of foreign goods is one way to understand indigenous experience of the colonial encounter and explore the agency of indigenous people by examining how those goods were distributed and used when obtained [@Dietler2005; @Mullins2011; @Scaramelli2005].
+
+Taiwan was colonized by the Spanish and the Dutch subsequently from 1624 to 1662, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
+However, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans.
+Large-scale and frequent trade activities brought a wide variety of trade objects into indigenous communities in Yilan through local regional exchange network.
+Among those trade goods, ornament such as glass and stone beads was one of the common foreign goods found at the settlement sites during this period [@Chen2007; @Li2014; @NMTH2005; @Wang2007].
+Body adornment is usually viewed as a signal in indigenous social life to represent and reinforce their status through the body as the scene to display [@Joyce2005; @Scaramelli2005].
+The consumption of trade beads in Southeast Asia during the Iron Age were usually associated with increasing social stratification or socio-political complexity that influenced [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].
+
+identify the value of the beads
+
+@Theunissen2000 <---This has never been explored in any depth despite the fact it has been recognized for some time that the introduction of new prestige goods from external sources may cause restructuring of existing prestige goods exchange and subsequent increases in social complexity (Higham 1984: 250-4).
+--->
+
+@Mullins2011 <---suggest that beads were one class of goods that continued and embellished indigenous value systems that stressed the powers embedded in shiny objects such as quartz crystals, and the European use of beads in items such as rosaries may have reinforced their powers in the eyes of indigenous people.--->
+
+@Carter2016 <---has also proposed that these lower quality beads may instead be related to intensified trade with South Asia and the importation of mass-produced beads from this location. --->
+
+My research will explore the evidence at Kiwulan by addressing the question of whether European contact and influence on local trade networks stimulated an increase in social inequality among the inhabitants of
+Kiwulan. In this project, I intend to test the hypothesis that European contact in northeastern Taiwan stimulated a change in indigenous social organization at Kiwulan.
+
+2. debate or problem have not solved yet
+3. how this case can answer it
+4. what are you going to do
 
@@
 # Cultural context of Ornaments in Northeastern Taiwan
+
+Before European arrival, the native people had already established their own trading networks but the arrival of Europeans and their trade goods in the 17th century built on, but ...
+
+Foreign ornaments including glass beads, stone beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@Wang2007].
+Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the transition from the Neolithic to the Iron Age.
+We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.
+
+Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period.
+Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century.
+Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.
+
+@Mullins2011---<They examine the ways gift-giving worked in the Middle Orinoco region of Venezuela, where missionaries hoped to secure indigenous alliances by providing goods for which they perceived a native demand.
+One class of goods missionaries provided was alcohol, which was used commonly in feasts and ritual contexts, but interjecting externally produced alcohol unseated existing production and consumption systems.
+Prior to colonization, drink had been produced by indigenous domestic units to enhance a family's prestige, but the introduction of imported spirits compelled hosts to obtain them through exchange, which forced increasing reliance on the production of cash crops and increased debt among indigenous consumers.
+This shift rendered drink “a commodity—a product that had to be bought, rather than produced,” so even though European alcohol was consumed within existing consumer contexts it led to dependency and exploitation--->
 
@@
 Foreign trade ornaments were viewed as prestige goods and played an important role in the cultural context of indigenous societies in northeastern Taiwan according to the use context as grave goods and historical documents written by foreign record keeper.
-As a status marker, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
+As a marker for status, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
 Greater amount and diversity ornament types during the European contact period reflects the international stimulation of exchange in a colonial context.

### 2019-07-05 analysis/paper/paper.Rmd
added map and references, and revised introduction

@@
 
-1. background of the topic
-
-Taiwan island became part of a global trading network during the early 17th century because of western European expansion that connected European and Asia.
+Taiwan had encountered multiple episodes of culture contact with China and Southeast Asia since Neolithic and became part of a global trading network during the early 17th century because of western European expansion that connected European and Asia.
 Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
@@
 
-Taiwan was colonized by the Spanish and the Dutch subsequently from 1624 to 1662, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
-However, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans.
+Despite colonization of Taiwan by the Spanish and the Dutch from 1624 to 1662, little has been discovered in archaeology about the The reaction of indigenous people to the European colonization.
+Northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
+Although the location is relatively peripheral, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans.
 Large-scale and frequent trade activities brought a wide variety of trade objects into indigenous communities in Yilan through local regional exchange network.
-Among those trade goods, ornament such as glass and stone beads was one of the common foreign goods found at the settlement sites during this period [@Chen2007; @Li2014; @NMTH2005; @Wang2007].
+Among those trade goods, ornament such as glass and stone beads was one of the common foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
 Body adornment is usually viewed as a signal in indigenous social life to represent and reinforce their status through the body as the scene to display [@Joyce2005; @Scaramelli2005].
-The consumption of trade beads in Southeast Asia during the Iron Age were usually associated with increasing social stratification or socio-political complexity that influenced [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].
+For example, the consumption of stone beads in Southeast Asia during Iron Age were usually associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].
 
-identify the value of the beads
-
-@Theunissen2000 <---This has never been explored in any depth despite the fact it has been recognized for some time that the introduction of new prestige goods from external sources may cause restructuring of existing prestige goods exchange and subsequent increases in social complexity (Higham 1984: 250-4).
---->
-
-@Mullins2011 <---suggest that beads were one class of goods that continued and embellished indigenous value systems that stressed the powers embedded in shiny objects such as quartz crystals, and the European use of beads in items such as rosaries may have reinforced their powers in the eyes of indigenous people.--->
-
-@Carter2016 <---has also proposed that these lower quality beads may instead be related to intensified trade with South Asia and the importation of mass-produced beads from this location. --->
-
-My research will explore the evidence at Kiwulan by addressing the question of whether European contact and influence on local trade networks stimulated an increase in social inequality among the inhabitants of
-Kiwulan. In this project, I intend to test the hypothesis that European contact in northeastern Taiwan stimulated a change in indigenous social organization at Kiwulan.
-
-2. debate or problem have not solved yet
-3. how this case can answer it
-4. what are you going to do
-
-We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan[@Chen2007].
-Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network.
-We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
-We hypothesize that there are clustered distribution of ornaments at Kiwulan relating to these international contacts.
+In this paper, I explore the evidence from a site spans from late Iron Age to historical period by addressing the question of whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
+Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
+Northeastern Taiwan provides a good example to explore the indirect impacts on indigenous societies caused by comparing different episodes of culture contact based on a long-term perspective.
+My hypotheses are that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network, and the consumption of ornaments present some spatial patterns that indicate increasing social inequality.
+I also hypothesize that the use of ornaments declined during Chinese contact period.
 
@@
 
-Foreign ornaments including glass beads, stone beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@Wang2007].
+Foreign ornaments including glass beads, stone beads, and metal objects first appeared in Taiwan around 1,800 BCE through the regional exchange network in Southeast Asia [@Wang2007].
 Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the transition from the Neolithic to the Iron Age.
 We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.
-
-Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period.
-Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century.
-Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.
 
@@
 
-Kiwulan (Figure 1) is located at Yilan, north of Lanyang Plain in northeastern Taiwan, surrounded by the Central Mountain Range in the west and the Pacific Ocean in the east.
+```{r, KWL-map, fig.cap = "Kiwulan location in Southeast Asia", out.width= "400px", echo=FALSE}
+knitr::include_graphics("kiwulan-location-map.jpeg")
+```
+
+Kiwulan \@ref(fig:KWL-map) is located at Yilan, north of Lanyang Plain in northeastern Taiwan, surrounded by the Central Mountain Range in the west and the Pacific Ocean in the east.
 Kiwulan was excavated from 2001 to 2004, and the total excavation area was 3,814 m2.

### 2019-07-05 analysis/paper/map.R

@@ -51,3 +51,6 @@ TW_map_with_site +
     xmax = 120,
     ymin = 23.5,
     ymax = 25.7)
+
+ggsave("kiwulan-location-map.jpeg")
+

### 2019-07-06 analysis/paper/paper.Rmd
making map for sampling area and revise

@@
 
-Taiwan had encountered multiple episodes of culture contact with China and Southeast Asia since Neolithic and became part of a global trading network during the early 17th century because of western European expansion that connected European and Asia.
+Taiwan had been involved in regional exchange network of East Asia and Southeast Asia since Neolithic and became part of a global trading network during the early 17th century due to western European expansion that connected European and Asia.
 Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
@@
 
-In this paper, I explore the evidence from a site spans from late Iron Age to historical period by addressing the question of whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
+In this paper, I explore the evidence from a site that spans the time from late Iron Age to historical period by addressing the question of whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
 Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
-Northeastern Taiwan provides a good example to explore the indirect impacts on indigenous societies caused by comparing different episodes of culture contact based on a long-term perspective.
-My hypotheses are that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network, and the consumption of ornaments present some spatial patterns that indicate increasing social inequality.
-I also hypothesize that the use of ornaments declined during Chinese contact period.
+As @Lape2003 points out based on case studies in island Southeast Asia, the study of multiple episodes of culture contact has potential to expand our understanding of the causes and effects of culture contact both locally and worldwide.
+Northeastern Taiwan provides a good example to explore the indirect impacts on indigenous societies by comparing archaeological evidence from different episodes of culture contact based on a long-term perspective.
+The variety and cultural context of trade ornaments will be reviewed, and their capacity to inform on how the distribution of ornaments reflect social changes will be examined.
+My hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network, and the consumption of ornaments present some spatial pattern that indicates increasing social inequality.
+I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
 
-# Cultural context of Ornaments in Northeastern Taiwan
+# Cultural context of ornaments in northeastern Taiwan, Yilan
 
-Before European arrival, the native people had already established their own trading networks but the arrival of Europeans and their trade goods in the 17th century built on, but ...
-
-Foreign ornaments including glass beads, stone beads, and metal objects first appeared in Taiwan around 1,800 BCE through the regional exchange network in Southeast Asia [@Wang2007].
-Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the transition from the Neolithic to the Iron Age.
-We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.
-
-@Mullins2011---<They examine the ways gift-giving worked in the Middle Orinoco region of Venezuela, where missionaries hoped to secure indigenous alliances by providing goods for which they perceived a native demand.
-One class of goods missionaries provided was alcohol, which was used commonly in feasts and ritual contexts, but interjecting externally produced alcohol unseated existing production and consumption systems.
-Prior to colonization, drink had been produced by indigenous domestic units to enhance a family's prestige, but the introduction of imported spirits compelled hosts to obtain them through exchange, which forced increasing reliance on the production of cash crops and increased debt among indigenous consumers.
-This shift rendered drink “a commodity—a product that had to be bought, rather than produced,” so even though European alcohol was consumed within existing consumer contexts it led to dependency and exploitation--->
-
-Although the northeastern Taiwan forms a circumscribed environment due to the surrounding mountains and sea, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups in northern and southern Taiwan, foreign traders such as the Han people of China, and Europeans.
-This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan [@Chen2005; @Wang2007].
-The groups involved in the exchange network consisted by the Chinese traders, indigenous groups called Basay in northern Taiwan, and indigenous groups, Kavalan, in northeastern Taiwan.
-Historical records indicate that the Kavalan people were one of the Basay’s trading partners, and they exchanged rice, deer skin, and gold for metal tools or ornaments, beads, and Chinese ceramics [@Chen2005; @Hsieh2009; @LiandWu2006].
+Before European arrival, the indigenous people in northern Taiwan had already established their own exchange network that defined as "inter-insular trade" by @Chen2005 [p.
+12] to refer to small-scale regional trade on irregular basis .
+Despite being on the periphery of major trade ports at northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
+Yilan, also called "Kavalan" by local indigenous people since the prehistoric time, is a alluvial plain circumscribed by the Pacific and mountains on the sides.
+Rivers and seas provide the way for the interaction between local indigenous people and another native group called Basay in northern Taiwan and Han Chinese.
+Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
+This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan, and was intensified in the 17th century when the arrival of Europeans and their trade goods [@Chen2005; @Wang2007].
 
@@
 European contact was focused on trading with indigenous communities in northeastern Taiwan.
-Trading activity of many types of goode among these indigenous communities increased with the arrival of European colonisers.
+Trading activity of many types of goode among these indigenous communities increased with the arrival of European colonizers.
 Thus, we predict that this trading activity resulted in an increase in the amount and diversity of ornaments coming into indigenous communities in northeastern Taiwan.
+
+@Mullins2011---<They examine the ways gift-giving worked in the Middle Orinoco region of Venezuela, where missionaries hoped to secure indigenous alliances by providing goods for which they perceived a native demand.
+One class of goods missionaries provided was alcohol, which was used commonly in feasts and ritual contexts, but interjecting externally produced alcohol unseated existing production and consumption systems.
+Prior to colonization, drink had been produced by indigenous domestic units to enhance a family's prestige, but the introduction of imported spirits compelled hosts to obtain them through exchange, which forced increasing reliance on the production of cash crops and increased debt among indigenous consumers.
+This shift rendered drink “a commodity—a product that had to be bought, rather than produced,” so even though European alcohol was consumed within existing consumer contexts it led to dependency and exploitation>
 
@@
 Documents from the Qing dynasty usually describe the decorative purposes of the ornaments.
-In the Kavalan Subprefecture Gazetteer [@Chen2006[1852]: 308], the Kavalan Zhi Lue [@Ko1993[1837]: 11, 126) and the Dong Cha Ji Lue [@Yao1996[1829]: 77], it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments.
+In the Kavalan Subprefecture Gazetteer (Chen2006[1852]: 308), the Kavalan Zhi Lue (Ko1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao1996[1829]: 77), it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments.
 Their number was small and their value high, hence only the rich possessed them.
@@
 But the origin of these beads is still not known.
-Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer [@Hsu1992, pp.
+Later, because they were bought by Japanese and people in Taipei, agate beads became rarer [@Hsu1992, pp.
 22] and carp-shaped ornaments had completely disappeared.
@@
 ```{r, KWL-map, fig.cap = "Kiwulan location in Southeast Asia", out.width= "400px", echo=FALSE}
-knitr::include_graphics("kiwulan-location-map.jpeg")
+knitr::include_graphics("kiwulan-location-map.png")
 ```
 
-Kiwulan \@ref(fig:KWL-map) is located at Yilan, north of Lanyang Plain in northeastern Taiwan, surrounded by the Central Mountain Range in the west and the Pacific Ocean in the east.
-Kiwulan was excavated from 2001 to 2004, and the total excavation area was 3,814 m2.
-Most beads were found during the excavation and some of them through screens with 20 mm and 1.5 mm mesh.
+Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was excavated from 2001 to 2004 that creates 262 squares with the total area of 3,814 meter squares.
+Most beads were found in situ while digging and some were collected through screens with 20 mm and 1.5 mm mesh.
 The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures.
-The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates[a], which coordinate to the late Iron Age and Proto-historical period in Taiwan.
+The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates, which coordinate to the late Iron Age and Proto-historical period in Taiwan.
 This paper focuses on the Upper Layer Culture which overlaps with the contact periods with the Europeans and Chinese.
-It provides a good examples to discuss the foreign impacts on local indigenous society through body ornaments that usually convey social and cultural meaning by displaying.
 
@@
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
+
+# mapping sampling area
+ggplot(data = AD_data) +
+  geom_sf(fill = NA) +
+  geom_sf(data = sample_units, color = 'red', fill = NA) +
+  geom_sf_text(aes(label = `坑號`), alpha = 0.7) +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank())
 ```

### 2019-07-06 analysis/paper/map.R

@@ -89,6 +89,5 @@ TW_map_with_site +
     ymin = 24.7,
     ymax = 25.2)
 
-ggsave("kiwulan-location-map.jpeg")
-
+ggsave("analysis/figures/kiwulan-location-map.png")
 

### 2019-07-06 analysis/paper/paper.Rmd
update for the map of sampling area

@@
 # mapping sampling area
-ggplot(data = AD_data) +
-  geom_sf(fill = NA) +
-  geom_sf(data = sample_units, color = 'red', fill = NA) +
+ggplot(sample_units) +
+  geom_sf(data = AD_data, fill = NA) +
+  geom_sf(color = 'red') +
   geom_sf_text(aes(label = `坑號`), alpha = 0.7) +
   theme_minimal() +
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank())
+  theme_void() 
 ```

### 2019-07-06 analysis/paper/paper.Rmd
update for Taiwan map

@@
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
+```
 
+```{r, KWL-sam-area, fig.cap = "Sampling area with 40 squares", out.width= "400px", echo=FALSE}
 # mapping sampling area
@@
 
-The ornaments studied in this paper came from 40 adjacent 4m x 4m units out of 262 units in the middle part of excavation.
+The ornaments studied in this paper came from 40 adjacent 4m x 4m units out of 262 units in the middle part of excavation (Figure \@ref(fig:KWL-sam-area).
 The sampling principle is based on the extent of intact contexts that the least disturbed by modern constructions are the preference.
@@
 
-<--add a graph of 40 units and chronology-->
+<--add a chart of chronology-->
 

### 2019-07-06 analysis/paper/map.R

@@ -84,10 +84,10 @@ tw_map +
 TW_map_with_site +
   annotation_custom(
     grob = ggplotGrob(TW_SE_Asia),
-    xmin = 120.4,
-    xmax = 121.2,
-    ymin = 24.7,
-    ymax = 25.2)
+    xmin = 118,
+    xmax = 120,
+    ymin = 23.5,
+    ymax = 25.7)
 
 ggsave("analysis/figures/kiwulan-location-map.png")
 

### 2019-07-08 analysis/paper/paper.Rmd
add image to figure folder and some more editing

@@
 
-Taiwan had been involved in regional exchange network of East Asia and Southeast Asia since Neolithic and became part of a global trading network during the early 17th century due to western European expansion that connected European and Asia.
+Taiwan had been involved in regional exchange network of East Asia and Southeast Asia since the tenth century and became part of a global trade systmes during the early 17th century due to western European expansion that connected European and Asia [@Liu2017].
 Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
@@
 
-Despite colonization of Taiwan by the Spanish and the Dutch from 1624 to 1662, little has been discovered in archaeology about the The reaction of indigenous people to the European colonization.
+Despite colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, little had been discussed in archaeology about the reaction of indigenous people to the European colonization.
+@Berrocal2018 study the long-term record of Heping Dao in northern Taiwan and explore the effects of European colonization on the local populations by focusing on the environmental change and introduction of diseases, with a little mention of indigenous social complexity.
 Northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
@@
 
-In this paper, I explore the evidence from a site that spans the time from late Iron Age to historical period by addressing the question of whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
-Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
+In this paper, I explore the evidence from a site spanning from late Iron Age to historical period by addressing the question of whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
+Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems, and later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
 As @Lape2003 points out based on case studies in island Southeast Asia, the study of multiple episodes of culture contact has potential to expand our understanding of the causes and effects of culture contact both locally and worldwide.
@@
 The variety and cultural context of trade ornaments will be reviewed, and their capacity to inform on how the distribution of ornaments reflect social changes will be examined.
-My hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network, and the consumption of ornaments present some spatial pattern that indicates increasing social inequality.
+I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network, and the consumption of ornaments present some spatial pattern that indicates increasing social inequality.
 I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
@@
 
-Before European arrival, the indigenous people in northern Taiwan had already established their own exchange network that defined as "inter-insular trade" by @Chen2005 [p.
-12] to refer to small-scale regional trade on irregular basis .
-Despite being on the periphery of major trade ports at northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
+Prior to European colonization, the indigenous people in northern Taiwan had already established their own exchange network that defined as "inter-insular trade" by @Chen2005 [p.
+12] to refer to small-scale regional trade between China and Taiwan on irregular basis.
+This local exchange network of northeastern Taiwan had increasingly influenced by long distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since 15th century due to its location on the shipping routes, and was intensified in the 17th century when the arrival of the Europeans and their trade goods [@Chen2005; @Liu2017; @Wang2007].
+Despite being on the periphery of major trade ports, Keelung and Tanshui, in northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
 Yilan, also called "Kavalan" by local indigenous people since the prehistoric time, is a alluvial plain circumscribed by the Pacific and mountains on the sides.
-Rivers and seas provide the way for the interaction between local indigenous people and another native group called Basay in northern Taiwan and Han Chinese.
+Rivers and seas provide the way for the interaction between local indigenous people and another native group called Basai in northern Taiwan and Han Chinese.
 Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
-This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan, and was intensified in the 17th century when the arrival of Europeans and their trade goods [@Chen2005; @Wang2007].
 
-Historical records from Spanish and Dutch provide rich information about the historical and cultural background both for Europeans and indigenous people in the early 17th century.
-Spanish occupied and built forts in Northern Taiwan since 1624, from which Spanish missionaries often went to indigenous settlement and left behind numerous records from those places.
-The report of the Dominican Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) as a materials to exchange with necessities they need with other indigenous people.
-This trend then gradually spread to other areas, and even the Spanish soldiers used agate as bargaining chips for gambling, since agate equaled money [@LiandWu2006, pp.
+Historical documents from the Spanish and Dutch in the early 17th century provide some information about the social life and culture of indigenous people in Northern Taiwan when the Europeans arrived.
+The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purpose [@Blusse2000, pp.
+343].
+The report of the Dominican priest Fr.
+Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) in exchange for necessities with other indigenous group.
+This form of exchange spread widely and even the Spanish soldiers learned to use agate as bargaining chips for gambling, since agate equaled money [@LiandWu2006, pp.
 132-149].
-European contact was focused on trading with indigenous communities in northeastern Taiwan.
-Trading activity of many types of goode among these indigenous communities increased with the arrival of European colonizers.
-Thus, we predict that this trading activity resulted in an increase in the amount and diversity of ornaments coming into indigenous communities in northeastern Taiwan.
-
-@Mullins2011---<They examine the ways gift-giving worked in the Middle Orinoco region of Venezuela, where missionaries hoped to secure indigenous alliances by providing goods for which they perceived a native demand.
-One class of goods missionaries provided was alcohol, which was used commonly in feasts and ritual contexts, but interjecting externally produced alcohol unseated existing production and consumption systems.
-Prior to colonization, drink had been produced by indigenous domestic units to enhance a family's prestige, but the introduction of imported spirits compelled hosts to obtain them through exchange, which forced increasing reliance on the production of cash crops and increased debt among indigenous consumers.
-This shift rendered drink “a commodity—a product that had to be bought, rather than produced,” so even though European alcohol was consumed within existing consumer contexts it led to dependency and exploitation>
-
-Father Jacinto Esquivel recorded how people in Tamsui entered marriages.
-A man who wanted to take a wife had to give her parents cuentas.
-If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled [@LiandWu2006, pp.
+Father Jacinto Esquivel also recorded how indigenous viewed it as prestige goods in their culture.
+A indigenous male who wanted to get married had to pay agate beads to the parents of his future wife.
+Also, agate or golden beads could be used as to solve conflicts in their daily lives [@LiandWu2006, pp.
 151].
-Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals [@Pao2008, pp.
+Other records mention that the female shamans (majuorbol) in the tribe would use agate beads as magical items in ritual practice for body healing[@Borao2009, pp.
 122-151].
-Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity.
-They put a mat into the grave to protect the body from moisture and then put cooked rice on both sides of the head as provision.
-The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need.
-The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave [@LiandWu2006, pp.
+A funeral process in indigenous culture was recorded about how agate beads used commonly in ritual contexts, where more agate beads, pottery and cloths were be placed into the grave of more influential people to indicate a family's prestige [@LiandWu2006, pp.
 153].
 
-Documents from the Qing dynasty usually describe the decorative purposes of the ornaments.
-In the Kavalan Subprefecture Gazetteer (Chen2006[1852]: 308), the Kavalan Zhi Lue (Ko1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao1996[1829]: 77), it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments.
-Their number was small and their value high, hence only the rich possessed them.
-Ordinary people wore agate beads or glass beads on their head or neck during various festivals.
-In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori [@Kanori1996, pp.
-227-232] made field surveys among various ethnic groups in Taiwan.
-He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.
+Those documents indicate that agate beads or golden beads were already treated as prestige goods before the arrival of the Europeans, that is believed to be introduced earlier from Chinese traders.
+In 1642, the Dutch VOC defeated the Spanish and take over the northern Taiwan.
+They introduced feudal symbolism in an attempt to control the indigenous communities by asking indigenous leaders to attend annual ceremony for feudal practices [@Andrade2007, ch.
+9; @Kang2016, ch.
+4].
+More beads entered indigenous settlements through the gift-giving that the Dutch provided goods based on the demand of native people to secure indigenous alliances in the annual ceremony or on the way of traveling [@Kang2016, ch.
+6].
+Thus, I assume that the way of gift giving resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities within existing consumer contexts during this period.
 
-Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans [@Liu2008, pp.
+The documents written by Chinese historians in Qing dynasty usually described the decorative purposes of the ornaments and some documents had similar texts that might be copied from previous ones.
+According to those records, [@Chen2006[1852]: 308], [@Ko1993[1837]: 11, 126] and [@Yao1996[1829]: 77], it is mentioned that indigenous people in Yilan used ornaments in ceremonial contexts to display their status.
+Amoung those ornaments, golden carp-shaped ornaments made from metal threads had higher value due to its delicacy and rare material that only possessed by rich people.
+Other people wore agate beads or glass beads on their head or neck to participating various ceremonies.
+In 1895, at the beginning of Japanese colonization, Ino Kanori [@Kanori1996, pp.
+227-232] conducted a field survey for several ethnic groups and found that carp-shaped ornaments had not been used in Yilan that time, but elder people still used beads as ornaments.
+
+Recent ethnographic research of the Kavalan tribe describes that agate beads were used in divination practice, called "subli", by female shamans [@Liu2008, pp.
 133-134].
-Interview records describe agate beads as valuable objects passed down from mother to daughter.
-When they were not used, they were often put on roof beams so that they were not seen by other people.
-But the origin of these beads is still not known.
-Later, because they were bought by Japanese and people in Taipei, agate beads became rarer [@Hsu1992, pp.
+Interview records mention agate bead is valuable object that are usually passed down from mother to daughter as a heirloom in shaman families without knowing the exact origin.
+Were not been using, beads would be often hided at invisible place such as on the roof beams to make sure no one sees that.
+During the Japanese colonization period, the number of agate beads declines due to the trade with Japanese who live in Taipei for collecting purpose [@Hsu1992, pp.
 22] and carp-shaped ornaments had completely disappeared.
-Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination.
-It was usually the female shamans that were responsible for divination rites, a technique passed down by ancestors.
-Most aborigines today don’t know the origin of agate beads any more, but their worth is still highlighted by oral history and through their scarcity.
-
-Those historical documents from the early 17th century to the modern ethnography described how local indigenous people use those ornaments in local cultural contexts that represents some social roles or status.
+Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination that is consistent with the observation in European colonial period.
+It was usually the female shamans conducting divination ritual that passed down from their ancestors.
+Despite most aborigines today don’t know the origin of agate beads, the value of beads is still be highlighted through oral history and their scarcity.
+Those historical documents from the early 17th century to the modern ethnography described how local indigenous people use ornaments in local cultural contexts that represents some social roles or status.
 However, compared to European contact period, there is less mention of beads in Chinese contact period and the description about ornaments is only limited to their dressing culture.
@@
 ```{r, KWL-map, fig.cap = "Kiwulan location in Southeast Asia", out.width= "400px", echo=FALSE}
-knitr::include_graphics("kiwulan-location-map.png")
+library(here)
+knitr::include_graphics(here("analysis", "figures", "kiwulan-location-map.png"))
 ```
 
-Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was excavated from 2001 to 2004 that creates 262 squares with the total area of 3,814 meter squares.
+excavate an open area for the first time reaching a surface a bit larger than 109 m2.
+
+Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was excavated from 2001 to 2004 that creates an open area with 262 squares (4m by 4m) reaching 3,814 meter squares in total.
 Most beads were found in situ while digging and some were collected through screens with 20 mm and 1.5 mm mesh.
@@
 ```{r read-in-the-data}
-library(here)
 library(tidyverse)
@@
                             "Metal Piece",
-                            "Glass ring"))
+                            "Glass ring")) %>%
+  filter(!Type %in% c("other",
+                      "unclassified")) %>% 
+  arrange(Categories, desc(`European Contact`))
 
@@
 We hypothesize that the distribution of ornaments is consistent with a completely random process, and then simulate the presence of ornaments across the space for one thousand times for each time periods.
-We compare our data to the random pattern generated by simulated processes to see if the our data is randomly distributed or not.
+The data was compared to the random pattern generated by simulated processes to see if the our data is randomly distributed or not.
 The results \@ref(fig:hypo-test-distrubution) are represented by histogram showing one thousand simulation with the blue line indicating our observed value.

### 2019-07-08 analysis/paper/map.R

@@ -5,9 +5,8 @@ library("rnaturalearth")
 library("rnaturalearthdata")
 library("here")
 library("ggspatial")
-# devtools::install_github('3wen/legendMap')
+#devtools::install_github('3wen/legendMap')
 library(legendMap)
-
 
 world <- ne_countries(scale = "medium", returnclass = "sf")
 # class(world)
@@ -29,16 +28,16 @@ ggplot(data = world) +
   coord_sf(xlim = c(107, 135), ylim = c(13, 35), expand = FALSE, datum = NA) +
   theme(panel.background = element_rect(fill = "azure"))
 
-
+# Topographic map
 library(ggmap)
 library(tmaptools)
 library(shadowtext)
 # we don't want to download every time, so let's save the map locally
 # from https://stackoverflow.com/a/52710855/1036500
- tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(121.0, 24.3,
-                                                  122.0, 25.2))), zoom = 10))
-# saveRDS(tw_map, here("analysis", "data", "raw_data", "tw_map.rds"))
-# tw_map <- readRDS(here("analysis", "data", "raw_data", "tw_map.rds"))
+tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(121.0, 24.3,
+                                                 122.0, 25.2))), zoom = 10))
+#saveRDS(tw_map, here("analysis", "data", "raw_data", "tw_map.rds"))
+#tw_map <- readRDS(here("analysis", "data", "raw_data", "tw_map.rds"))
 pg <- ggplot_build(tw_map)
 
 TW_map_with_site <-
@@ -80,14 +79,12 @@ tw_map +
     # size of letter 'N' on N arrow, in km
     arrow_north_size = 5)
 
-
 TW_map_with_site +
-  annotation_custom(
-    grob = ggplotGrob(TW_SE_Asia),
-    xmin = 118,
-    xmax = 120,
-    ymin = 23.5,
-    ymax = 25.7)
+annotation_custom(
+grob = ggplotGrob(TW_SE_Asia),
+xmin = 120.4,
+xmax = 121.2,
+ymin = 24.7,
+ymax = 25.2)
 
 ggsave("analysis/figures/kiwulan-location-map.png")
-

### 2019-07-09 analysis/paper/paper.Rmd
revise and edit reference

@@
 According to those records, [@Chen2006[1852]: 308], [@Ko1993[1837]: 11, 126] and [@Yao1996[1829]: 77], it is mentioned that indigenous people in Yilan used ornaments in ceremonial contexts to display their status.
-Amoung those ornaments, golden carp-shaped ornaments made from metal threads had higher value due to its delicacy and rare material that only possessed by rich people.
+Amoung those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production that usually possessed by rich people.
 Other people wore agate beads or glass beads on their head or neck to participating various ceremonies.
 In 1895, at the beginning of Japanese colonization, Ino Kanori [@Kanori1996, pp.
-227-232] conducted a field survey for several ethnic groups and found that carp-shaped ornaments had not been used in Yilan that time, but elder people still used beads as ornaments.
+227-232] conducted a field survey for several aboriginal groups and found that golden necklaces had not been used in Yilan that time, but elder people still used beads.
 
@@
 During the Japanese colonization period, the number of agate beads declines due to the trade with Japanese who live in Taipei for collecting purpose [@Hsu1992, pp.
-22] and carp-shaped ornaments had completely disappeared.
+22] and golden fish-shaped necklaces had completely disappeared.
 Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination that is consistent with the observation in European colonial period.
@@
 
-excavate an open area for the first time reaching a surface a bit larger than 109 m2.
+Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was a rescue archaeology projects that carried out from 2001 to 2004 in advance of construction.
+The total area includes eight open area sections that created 262 squares (4m by 4m) reaching 3,814 meter squares in total [@Chen2007].
+Most ornaments were found in situ in the context while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
+The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures that indicates it was a settlement.
+The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates [@Chen2007 ].
+This paper focuses on the Upper Layer Culture spanning from the late Iron Age that also called proto-historical period [@Liu2017], and historical period starting from the 17th century with European contact in Taiwan.
 
-Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was excavated from 2001 to 2004 that creates an open area with 262 squares (4m by 4m) reaching 3,814 meter squares in total.
-Most beads were found in situ while digging and some were collected through screens with 20 mm and 1.5 mm mesh.
-The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures.
-The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates, which coordinate to the late Iron Age and Proto-historical period in Taiwan.
-This paper focuses on the Upper Layer Culture which overlaps with the contact periods with the Europeans and Chinese.
-
-The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local settlements were attacked by the Spanish according to the official documents [add citation; Borao Mateo 2009].
-Later in 1647, the Dutch attacked the indigenous villages in Yilan and forced them to accept colonial rules and economic demands by paying annual tribute [add citation; Borao Mateo 2009].
-According to Dutch census reports in 1650, Kiwulan was the biggest indigenous community in the Yilan Plain, with a population of 840 adults [@Chen2007; @Kang2012; @Li2014].
-The diagnostic artifacts during European contact found at Kiwulan including An-ping jars and glass beads that were largely introduced to Taiwan during the early 17th century.
-
+The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
+163].
+Later in 1647, the Dutch attacked the indigenous villages and forced them to accept colonial rules and economic demands by paying annual tribute [@Andrade2007].
+According to Dutch census reports in 1650, Kiwulan was the largest indigenous settlement in the Yilan Plain, with a population of 840 adults [@Nakamura1938, pp.
+12].
+The diagnostic artifacts of European contact found at Kiwulan including An-ping jars, stonewares that were largely introduced to Taiwan during the early 17th century.
+Those artifacts coupled with the original radiocarbon dates were used to identify the pre-European and post-European periods.
 The European colonization ended in 1662 when they were defeated by the kingdom of Tungning found by Koxinga from China.
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The evidence of Chinese migration can be identified from Chinese official records and large amount of Chinese blue-and-white porcelain found at Kiwulan.
+The evidence of Chinese migration can be identified from Chinese official records and large amount of Chinese daily life porcelains found at Kiwulan.
 
@@
 
-The ornaments studied in this paper came from 40 adjacent 4m x 4m units out of 262 units in the middle part of excavation (Figure \@ref(fig:KWL-sam-area).
-The sampling principle is based on the extent of intact contexts that the least disturbed by modern constructions are the preference.
-The Upper Layer Culture can be divided into six sub layers spanning from 14 century to 19 century according to current radiocarbon dates, excavation depth, types of ceramics, and sediment texture and color [@Hsieh2009; @Wang2011].
-Layers are assigned numbers from 1 to 6 for which L1 corresponds to the upper context and L6, also the bottom layer, refers to lower context.
-Based on @Hsieh2009 chronological study of Chinese porcelains, L5 and L6 represent pre-European contact period, L4 was the time of European contact, and L2 and L1 was the period of Chinese contact.
+The ornaments studied in this paper found from 40 adjacent square of the largest open area that was located in the middle part of excavation (Figure \@ref(fig:KWL-sam-area).
+They were sampled because those units were more stratigraphically intact with the least disturbance by modern constructions.
+The Upper Layer Culture component could be divided into six sub contexts spanning from 14 century to 19 century with a hundred year interval according to original radiocarbon dates, excavation depth, and types of diagnostic porcelain [@Hsieh2009; @Wang2011].
+The contexts were assigned numbers from 1 to 6 where L1 corresponds to the youngest context and L6, the bottom layer, refers to the oldest context.
+To obtain a more reliable chronology for comparison between episodes of culture contact,I reevaluate those contexts for each sampled square by focusing on Anping jars and stonewares that are believed manufactured and distributed in Southeast Asia through European trade route during the early 17th century [@Ketel2011].
+Those two categories of ceramic coupled with radiocarbon dates were used to identify the pre-European and post-European periods.
+Following the previous numbering principle, I separated three episodes of time including pre-European contact period(L5, L6), European contact period(L4), and Chinese contact period (L2 and L1).
 
-<--add a chart of chronology-->
-
-In general, complete ornaments with clear context were mostly found in burials that helps to understand how people wear them and any relationships between shapes of ornaments and gender or age.
-In addition to burials, ornaments were also found in middens and living space that can be identified by post holes.
-The frequency of ornaments from burials is more than those from the living space due to well preserved condition.
-There is a wide variety of raw materials for ornaments of which ornaments glass, agate and metal ornaments that account for 99% of all ornaments were thought to be introduced by trading with the Europeans and Chinese traders, while ornaments made of shells, woods, and bones were believed locally made that is about only about 1% of all ornaments (Table 1).
-The reason for the large quantity of glass beads is that they usually have been found in clusters because of the original form such as necklaces.
-The shape of ornaments include beads, bracelets, rings, bells, pendants, and knitted objects that shows shape are closely related to raw material due to its characteristic.
-For example, beads are only made of glass, agate, and shell, and bone is the main material for knitted objects that used to connect beads.
-Also, metal ornaments covers more shapes such bells, pendants, and rings [@Chen2007].
+In general, complete ornaments with clear context were mostly found in burials that helps to understand how people wear them.
+In addition to burials, large amount of ornaments were found at household area that were identified by a series of post holes across the space.
+Here I focus on the ornaments found at post-hole area considering the overall quantity and the comparison across the space.
+Ornaments were made from a wide variety of raw materials, in which glass, agate and copper are the common materials that were thought to be introduced through trading with the Europeans and Chinese traders since no clear evidence of waste by-products in the archaeological record [@Chen2007].
+Although a possibility of small workshops that individuals were able to produce golden beads from prepared glass materials by heating [@Cheng2008], no evidence of accompanying tools or materials in proportion to the large amount of golden beads were found.
+The shape of ornaments include beads, bracelets, rings, bells, pendants, and knitted objects.
+It shows that the shape is closely related to what kind of raw material it was made from.
+For example, beads were only made of glass and agate, while copper ornaments were in the form such as bells, bracelets, and rings [@Chen2007].
 
@@
 
-To explore whether body ornaments can reflect the colonial influences on indigenous populations, this research examines the frequency and spatial distribution of trade ornaments at Kiwulan during different time periods with foreign contacts, including European and Chinese contacts.
-Trade ornaments were analyzed by three analytical units including pre-European contact, European contact, and Chinese contact that indicates three scale of exchange network from less to intense contact subsequently.
-Because the appearance of ornaments is highly related to the original raw material couple with the function such as bead and bracelet, they are combined as the first criterion of classification which determines five major categories.
-Under each category of raw material, ornaments are further classified into several subtype according to the shape if any variations were identified in each category.
-In order to compare the general pattern of changes for major ornaments, less frequent ornaments are combined as one category named other, which will be discussed separately.
+To explore whether body ornaments can reflect the colonial influences on indigenous populations, this paper examines the variety, frequency, and spatial distribution of trade ornaments across different time periods, including pre-European contact, European contact, and Chinese contact.
+Those three analytic units could represent three scales of exchange network from small to large, and three degree of contact from weak to intense.
+Because the appearance of ornaments is highly related to the original raw material, the type of raw material is the first criterion of classification that determined five major categories.
+In each category, ornaments were classified into several subtype according to their shapes when variations were present in each category.
+The variation was defined by the difference identified in at least two variables, for example dimensional variables such as length, or stylish variables such as patterns.
+In order to compare the general pattern of changes for subtypes, less frequent subtypes were left out from the analysis, but will be discussed separately in later section.
 
-Ornaments with clear contexts were mainly found from the post-hole area, burials, and middens, where ornaments from the post-hole area accounts for the majority of the amount in total.
-Post-hole area is also the contexts that we mainly focus on because it represents house units accross the site that can further provide a basis for comparison to see if there is any spatial pattern of distribution.
-Ornaments from middens and burials will also be discussed in the later section.
-For the post-hole area, there were `r sample_size` ornaments recovered from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the same excavation section including all contexts.
+Ornaments from the post-hole area accounts for the majority of the amount in total.
+Post-hole area provides a clear contexts and basis for comparison to see if there is any spatial pattern of distribution.
+For the post-hole area, there were `r sample_size` ornaments recovered from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the same open area section.
 Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
-Ornaments in most categories can be divided into sub-types according to the variation of shape.
-The variation was defined by the difference identified in at least two variables, for example dimensional variables such as length, or stylish variables such as patterns.
-The frequency and feature of subtypes for major categories are summarized in \@ref(tab:tidy-data-for-table).
+The variety of subtype and their individual frequency are summarized in \@ref(tab:tidy-data-for-table) in terms of three time episodes.
+[insert image here]
 

### 2019-07-09 analysis/paper/paper.Rmd
fixed references, added figure captions, checked every chunks

@@
 
-Prior to European colonization, the indigenous people in northern Taiwan had already established their own exchange network that defined as "inter-insular trade" by @Chen2005 [p.
+Prior to European colonization, the indigenous people in northern Taiwan had already established their own exchange network that defined as "inter-insular trade" by @Chen2005 [pp.
 12] to refer to small-scale regional trade between China and Taiwan on irregular basis.
@@
 
-The documents written by Chinese historians in Qing dynasty usually described the decorative purposes of the ornaments and some documents had similar texts that might be copied from previous ones.
-According to those records, [@Chen2006[1852]: 308], [@Ko1993[1837]: 11, 126] and [@Yao1996[1829]: 77], it is mentioned that indigenous people in Yilan used ornaments in ceremonial contexts to display their status.
-Amoung those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production that usually possessed by rich people.
+The documents written by Chinese historians in Qing dynasty usually described the decorative purposes of the ornaments and some documents had similar texts that might be copied from previous ones without actual investigation.
+According to those records, indigenous people in Yilan had a custom to wear ornaments in ceremonial contexts to display their wealth and status [@Chen1963, pp.
+228, 308; @Ke1993, pp.
+11, 126; @Yao1996, pp.
+77].
+Among those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production that usually possessed by rich people.
 Other people wore agate beads or glass beads on their head or neck to participating various ceremonies.
-In 1895, at the beginning of Japanese colonization, Ino Kanori [@Kanori1996, pp.
-227-232] conducted a field survey for several aboriginal groups and found that golden necklaces had not been used in Yilan that time, but elder people still used beads.
+In 1895, at the beginning of Japanese colonization, @Ino2002 conducted a field survey for several aboriginal groups and found that golden necklaces had not been used in Yilan that time, but elder people still used beads.
 
@@
 Were not been using, beads would be often hided at invisible place such as on the roof beams to make sure no one sees that.
-During the Japanese colonization period, the number of agate beads declines due to the trade with Japanese who live in Taipei for collecting purpose [@Hsu1992, pp.
-22] and golden fish-shaped necklaces had completely disappeared.
 Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination that is consistent with the observation in European colonial period.
@@
 The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures that indicates it was a settlement.
-The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates [@Chen2007 ].
+The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates [@Chen2007].
 This paper focuses on the Upper Layer Culture spanning from the late Iron Age that also called proto-historical period [@Liu2017], and historical period starting from the 17th century with European contact in Taiwan.
@@
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The evidence of Chinese migration can be identified from Chinese official records and large amount of Chinese daily life porcelains found at Kiwulan.
+The evidence of Chinese migration can be identified both from Chinese official records and large amount of Chinese daily life porcelains found at Kiwulan [@Chen1963].
 
@@
 total_AD <- nrow(kwl_upper)
-sample_pro <- sample_size/total_AD
+sample_pro <- round(sample_size/total_AD, 1)
 ```
@@
 
-```{r, KWL-sam-area, fig.cap = "Sampling area with 40 squares", out.width= "400px", echo=FALSE}
+```{r, KWL-sam-area, fig.cap = "Forty squares were sampled from the largest open area (in red with square ID number)", out.width= "70%", echo=FALSE}
 # mapping sampling area
@@
   geom_sf(color = 'red') +
-  geom_sf_text(aes(label = `坑號`), alpha = 0.7) +
+  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
   theme_minimal() +
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of ornaments"}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of ornaments", out.width='60%'}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 
+```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major category.
+A: agate beads, B: bells, C: glass beads and golden beads, D: metal rings.
+Photographs are presented in the same order as those subtypes in the table but from left to right instead.
+The photographs of B, C, D categories were from original excavation report.", out.width= "400px", echo=FALSE}
+knitr::include_graphics(here("analysis", "figures", "ornament.png"))
+```
+
 To explore whether body ornaments can reflect the colonial influences on indigenous populations, this paper examines the variety, frequency, and spatial distribution of trade ornaments across different time periods, including pre-European contact, European contact, and Chinese contact.
@@
 The variety of subtype and their individual frequency are summarized in \@ref(tab:tidy-data-for-table) in terms of three time episodes.
-[insert image here]
+The image for each subtype in major category shows in figure \@ref(fig:ornaments-image).
 
-```{r plot-counts-by-period}
+```{r, plot-counts-by-period, fig.cap = "Frequency of each category of ornament with a separate plot showing different time periods individually"}
 # categories of ornaments by period
@@
 
+# histogram facetting for three periods
+orna_periods_facet <-
 ggplot(ornaments_count) +
@@
   coord_flip() +
-  labs(y = "Frequency", x = "Categories") +
-  theme_minimal(base_size = 14)
+  labs(y = "Frequency") +
+  theme_minimal(base_size = 12) +
+  theme(axis.text.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.title.x = element_blank())
 
 # stack histogram
+orna_all_viridis <-
 ggplot(ornaments_count) +
@@
   scale_fill_viridis_d()
+
+library(cowplot)
+library(viridis)
+# add two plots together
+ggdraw() +
+  draw_plot(orna_all_viridis + 
+            theme(legend.justification = "top"), 0, 0, 1, 1) +
+  draw_plot(orna_periods_facet + scale_color_viridis(discrete = TRUE) + 
+              theme(legend.justification = "bottom"), 0.5, 0.15, 0.5, 0.5) +
+  draw_plot_label(c("A", "B"), c(0, 0.5), c(1, 0.7), size = 15)
 ```
+
+# Results
+
+## Frequency between time periods
+Figure \@ref(fig:plot-counts-by-period) shows the comparison between different time periods that indicates a general pattern that most ornaments existed already before European contact and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
+This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
+However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even much less in the Chinese contact period.
 
@@
 ```{r plot-altogether, fig.cap= "Frequency of "}
-library(cowplot)
 plot_grid(ring_shape_plot,
@@
 ```
-# Results
-
-## Frequency between time periods
-The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
-This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
-However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.
 
@@
 The distribution shows that agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period.
-Agate beads can be divided into eight subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased during the European contact and then declined in the Chinese contact.
+Agate beads can be divided into seven subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased significantly during the European contact and then declined in the Chinese contact.
 Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
-There are more subtypes of agate beads during European contact compared to other time periods.
-Similarly metal rings can be classified into seven subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape.
-The greater varieties for those two categories might indicate the multiple sources due to global trade network brought by the Europeans.
-In contrast, copper bell and glass bead have smaller variety, but glass bead has larger number due to they were usually found in a cluster that reflect the original use as strings.
-Most common shape for copper bell is the large shape with human face as motif, while most common shape for glass bead is small that were identified as Indo-Pacific beads, the most common type that are widespread in Southeast Asian sites since 300 BC [@Francis2002].
+Higher variety of agate beads were found in European contact compared to other time periods.
+Similarly, metal rings can be classified into eight subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape.
+The greater varieties for those two categories might indicate the multiple origins due to global trade network stimulated by the Europeans.
+In contrast, copper bell and glass bead have less variety, but glass bead has larger number due to they were usually found in a cluster that indicates the original use as strings or necklaces.
+The common shape for copper bell is the large shape with a wide variety of human face as a motif, while the common shape for glass bead is bead with length less than 1cm that are identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].
 
@@
 
-```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of ornaments"}
+```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
 ggplot() +
@@
 
-```{r plot-spatial-pattern-each}
+```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for each class of ornament by time periods"}
 # filter golden bead
@@
 # plot by period
+dis_golden <-
 ggplot(ornaments_period_join_gold) +
@@
   facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Golden Beads",
+  labs(title = "Golden Beads",
        fill = "frequency") +
@@
 # plot by period
+dis_glass <-
   ggplot(ornaments_period_join_gold) +
@@
   facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Glass Beads",
+  labs(title = "Glass Beads",
        fill = "frequency") +
@@
 # plot by period
+dis_agate <-
 ggplot(ornaments_period_join_gold) +
@@
   facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Agate Beads",
+  labs(title = "Agate Beads",
        fill = "frequency") +
@@
 # plot by period
+dis_metal <-
 ggplot(ornaments_period_join_metal) +
@@
   facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Metal Ring",
+  labs(title = "Metal Ring",
        fill = "frequency") +
@@
 # plot by period
+dis_bell <-
 ggplot(ornaments_period_join_bell) +
@@
   facet_wrap(~ period) +
-  labs(title = "Spatial Distribution of Copper Bell",
+  labs(title = "Copper Bell",
        fill = "frequency") +
@@
         strip.text.x = element_text(size = 11))
+
+plot_grid(dis_golden,
+          dis_glass,
+          dis_agate, 
+          dis_metal, 
+          dis_bell,
+          align = "v",
+          nrow = 5)
 ```
@@
 ## Frequency for spatial pattern
-Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all classes of ornaments for each time period.
+Figure \@ref(fig:plot-spatial-pattern-each) presents the spatial distribution of all categories for each time period.
 It shows that one square unearthed more ornaments than others in the pre-European contact.
-During the European contact period, the presence of ornaments is more ubiquitous with some clusters on northern part, and more randomly distributed during the Chinese contact period.
-When looking at the distribution for each major class individually, it shows that some clusters of different class of ornaments across the area during the European contact period, including golden bead, agate bead, metal ring, and bell.
-However, it seems no clear consistent pattern of concentration for those ornaments.
-Each class shows its own pattern that squares with higher number of ornaments distributed separately and independently.
+During the European contact period, ornaments was more ubiquitous with some clusters on northern part, and less presence during the Chinese contact period.
+When looking at the distribution for each major class individually, some clusters across the area were observed during the European contact including golden bead, agate bead, metal ring, and bell.
+However, it seems no clear consistent clustered pattern across those different ornaments.
+Each class shows its own pattern where the squares with higher number of ornaments distributed separately and independently.
 For example, the cluster of golden bead were found at the northern part, while the cluster of agate bead were found in the middle part.
-In contrast, there are more than one cluster of metal rings that distribute separately across the research area.
-Copper bells were usually found solely and seem randomly across the area.
+In contrast, there are more than one cluster of metal rings that distributed separately across the research area.
+Copper bells were usually found solely and seem randomly distributed across the area.
 Before European contact, the map shows greater amount of glass beads were found at the northern and middle part of the research area, and more widespread during the European contact period.
-In the Chinese Contact period, both the amount and density of different classes of ornaments decreases that they were only found in a few squares, for example, golden bead and glass bead were only found in a couple squares at the middle part.
+In the Chinese Contact period, both the amount and density of different classes of ornaments decreased that they were only found in a few squares, for example, golden bead and glass bead were only found in a couple squares at the middle part.
 
@@
 
-```{r plot-all-densities, fig.cap= "Kernal density map"}
+```{r plot-all-densities, fig.cap= "Kernal density map by periods"}
 par(mfrow = c(1, 3))
@@
 ## Spatial distribution
-Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period [@Bevan2016, @Ducke2015].
-The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution.
-We randomly assigned points, here the location of ornaments, for each square according to the data, subset three groups for three time periods, and compute the kernal density for each time period for comparison.
+To examine whether the distribution of ornament is uneven that could reflect the presence of social inequality, point pattern analysis is used to examine the distribution.
+Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some "hotspots" that imply social processes [@Bevan2016; @Ducke2015], and in this case the pattern of ornaments could be able to indicate the changes due to culture contact.
+The location where ornaments from were square-based recorded, thus the intensity approach was used that focuses on the average density of points across space to explore their distribution.
+Points were be randomly assigned to each square according to the frequency data.
+Next is to subset three groups for three time periods, and then compute the kernal density for each time period for comparison.
 Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
-Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier2019, @Cortegoso2016].
+Here we use KDE to visualize the event of the presence of ornaments that were represented by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
 Density value is assigned to each cell, of which the unit is meter.
@@
 Those maps present three consistent sub-region with shifting core area over time.
-The distribution might indicate different social groups who own more ornaments, and the multiple groups during European contact period might reflect a more corporate mode of social organization.
-However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious core area.
-Next step is to test whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.
+The distribution might indicate different social groups who possessed more ornaments, and the multiple groups during European contact period might reflect some presence of unequal consumption.
+However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious hot spot.
+To further evaluate the pattern, we test the hypothesis that whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.
 
@@
 
-```{r hypo-test-distrubution, fig.cap= "Histogram of simulated ANN values"}
+```{r hypo-test-distrubution, fig.cap= "Histogram of simulated ANN values with the blue line indicating the observed value"}
 
@@
 
-Here we use Monte Carlo method to test for the complete spatial randomness of spatial events based on the average nearest-neighbor distance method.
+Here we use Monte Carlo method to test the complete spatial randomness of spatial events based on the average nearest-neighbor distance method (ANN).
 We hypothesize that the distribution of ornaments is consistent with a completely random process, and then simulate the presence of ornaments across the space for one thousand times for each time periods.
-The data was compared to the random pattern generated by simulated processes to see if the our data is randomly distributed or not.
-The results \@ref(fig:hypo-test-distrubution) are represented by histogram showing one thousand simulation with the blue line indicating our observed value.
+The data was compared to the random pattern generated by simulated processes to see if our data is randomly distributed or not.
+The results \@ref(fig:hypo-test-distrubution) are represented by histogram showing one thousand simulation.
 The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments are clustered distributed.
-Similar result are also observed during the pre-European period but less extreme.
-The distribution of ornaments was random during the Chinese contact period and about one third of the simulated values are greater than our observed ANN value.
+Similar result is also observed during the pre-European period but less extreme.
+The distribution of ornaments is more random during the Chinese contact period and about one third of the simulated values are greater than our observed ANN value.
 This testing explains that the distribution of ornaments during the pre-European and European contact period might reflect the presence of different social groups.
@@
 Taiwan island had been involved in regional exchange networks of Southeast Asia since Neolithic period, a global trade network during the European contact period in the early 17th century, and later intense but smaller scale network with Han Chinese communities since 19th century [@Chen2005].
-Kiwulan demonstrates continuous occupation from the late Iron Age to one hundred years ago over 600 years that covers these three different culture contacts that enables a comparative research to explore the indirect influence of culture contact especially in the global colonial context.
+Kiwulan demonstrates continuous occupation covers those episode of culture contacts that enables a comparative research to explore the indirect influence of culture contact especially in the global colonial context by examining prestige goods.
 These three analytic chronology at Kiwulan represent different scale of exchange network and intensity of culture contact.
@@
 
-Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
-
-Q: Is the difference in the frequency of ornaments related to the scale of exchange network, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
-
 # Conclusion
@@
 The spatial pattern of ornaments at Kiwulan during European contact presents multiple clusters that is significantly different from a random distribution pattern.
-This study demonstrate that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization.
+This study demonstrates that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization.
 It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system.
 During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwalan population.
-This is a case study that focuses on trade ornaments and more studies can be done to have a comparison with other archaeological evidence.
-Further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.
-
+This is a case study that focuses on trade ornaments and more studies can be done to include other trade goods like porcelains.
+Also, further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.
 # Acknowledgements

### 2019-07-31 analysis/paper/paper.Rmd
adjust the paragraph
first editing, add comments to the file

@@
 
-Taiwan had been involved in regional exchange network of East Asia and Southeast Asia since the tenth century and became part of a global trade systmes during the early 17th century due to western European expansion that connected European and Asia [@Liu2017].
-Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced.
-Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local indigenous communities remains unanswered.
-Indirect effects of colonialism is emphasized increasingly to discuss the colonial impact on the indigenous people in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
-In many parts of the world, foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts might cause significant transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
-Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed to emphasize the ability of active adoptions of foreign effects, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
-Studying the consumption of foreign goods is one way to understand indigenous experience of the colonial encounter and explore the agency of indigenous people by examining how those goods were distributed and used when obtained [@Dietler2005; @Mullins2011; @Scaramelli2005].
+The direct impacts of European colonialism on indigenous communities in East Asia were much less pronounced than island Southeast Asia and Oceania.
+Throughout East Asia direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
+Understanding the indirect effects of colonialism are important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
+In many parts of the world, the introduction of foreign trade goods by colonial traders into local Indigenous societies caused significant transformations of Indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
+The agency of Indigenous peoples in the colonial contexts has been broadly discussed to understand the negotiations between colonized and colonizer, and the resistance and accommodations of Indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
+Studying the consumption of foreign goods is one way to understand Indigenous experiences of colonial encounters and explore the agency of Indigenous peoples by examining how goods were distributed and used [@Dietler2005; @Mullins2011; @Scaramelli2005].
 
-Despite colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, little had been discussed in archaeology about the reaction of indigenous people to the European colonization.
-@Berrocal2018 study the long-term record of Heping Dao in northern Taiwan and explore the effects of European colonization on the local populations by focusing on the environmental change and introduction of diseases, with a little mention of indigenous social complexity.
-Northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012].
-Although the location is relatively peripheral, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans.
-Large-scale and frequent trade activities brought a wide variety of trade objects into indigenous communities in Yilan through local regional exchange network.
-Among those trade goods, ornament such as glass and stone beads was one of the common foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
-Body adornment is usually viewed as a signal in indigenous social life to represent and reinforce their status through the body as the scene to display [@Joyce2005; @Scaramelli2005].
-For example, the consumption of stone beads in Southeast Asia during Iron Age were usually associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].
+Taiwan provides an ideal case study of indirect effects of colonialism because it has archaeological evidence spanning its initial involvement in regional exchange networks of East Asia and Southeast Asia since the tenth century.
+It became part of global trade systems during the early 17th century due to western European expansions that connected Europe and Asia [@Liu2017].
+Despite colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of Indigenous responses to the European colonization.
+@Berrocal2018 study \the long-term record of Heping Dao\ in northern Taiwan and explore the effects of European colonization on the local populations by focusing on the environmental change and introduction of diseases, with a little mention of indigenous social complexity.
+Yilan in Northeastern Taiwan is an especially unique location for exploring indirect effects due to its physically isolated location.
+The rugged \XXX mountains to the west of the Yilan Plain that made it not easily accessible to Europeans and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
+The sea provides a way through which Indigenous people in Yilan could interact with other groups from different regions, including other native Indigenous groups, foreign traders such as the Han people of China, and Europeans.
+Large-scale and frequent trade activities brought a wide variety of trade objects into Indigenous communities in Yilan through local regional exchange network.
+One of the most common types of trade goods in this region were personal ornaments such as glass and stone beads was one of the common foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
+Personal adornments are a social signal of an individual’s status [@Joyce2005; @Scaramelli2005].
+For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].
 
-In this paper, I explore the evidence from a site spanning from late Iron Age to historical period by addressing the question of whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
-Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems, and later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
-As @Lape2003 points out based on case studies in island Southeast Asia, the study of multiple episodes of culture contact has potential to expand our understanding of the causes and effects of culture contact both locally and worldwide.
-Northeastern Taiwan provides a good example to explore the indirect impacts on indigenous societies by comparing archaeological evidence from different episodes of culture contact based on a long-term perspective.
-The variety and cultural context of trade ornaments will be reviewed, and their capacity to inform on how the distribution of ornaments reflect social changes will be examined.
-I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network, and the consumption of ornaments present some spatial pattern that indicates increasing social inequality.
-I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan.
+In this paper, I explore the evidence from Yilan spanning from late Iron Age to historical period to address the question of whether indirect colonial influences on Indigenous populations can be detected.
+I focus on  the ornaments recovered from the archaeological excavations at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
+Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems, and \later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
+Northeastern Taiwan provides a good example to explore the indirect colonial impacts on Indigenous societies by comparing archaeological evidence from different episodes of culture contact [\cf.@Lape2003].
+I will describe the variety and cultural context of trade ornaments in this region, and examine their potential of the ornaments to inform on social changes relating to trade with colonial groups.
+I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to participation in a larger scale exchange network.
+I also predict that after European contact there will be new spatial patterns in the distributions of ornaments at Kiwulan that indicates increasing social inequality.
+I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan due to \???.
 
-# Cultural context of ornaments in northeastern Taiwan, Yilan
+# Cultural context of ornaments in Yilan, northeastern Taiwan
 
-Prior to European colonization, the indigenous people in northern Taiwan had already established their own exchange network that defined as "inter-insular trade" by @Chen2005 [pp.
+Prior to European colonization, the indigenous people in northern Taiwan had already established their own exchange network that has been described as "inter-insular trade" by @Chen2005 [pp.
 12] to refer to small-scale regional trade between China and Taiwan on irregular basis.
-This local exchange network of northeastern Taiwan had increasingly influenced by long distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since 15th century due to its location on the shipping routes, and was intensified in the 17th century when the arrival of the Europeans and their trade goods [@Chen2005; @Liu2017; @Wang2007].
-Despite being on the periphery of major trade ports, Keelung and Tanshui, in northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
-Yilan, also called "Kavalan" by local indigenous people since the prehistoric time, is a alluvial plain circumscribed by the Pacific and mountains on the sides.
-Rivers and seas provide the way for the interaction between local indigenous people and another native group called Basai in northern Taiwan and Han Chinese.
+This local exchange network of northeastern Taiwan was increasingly influenced by long distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since 15th century due to its location on the shipping routes.
+This was intensified in the 17th century with the arrival of the Europeans and their trade goods [@Chen2005; @Liu2017; @Wang2007].
+Due to its location on the periphery of major trade ports, \Keelung and \Tamsui, in northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
+Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
+Rivers and seas provide the way for the interaction between local indigenous settlements and further \afield communities, such as the Basai in northern Taiwan and Han Chinese.
 Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
@@
 The report of the Dominican priest Fr.
-Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) in exchange for necessities with other indigenous group.
-This form of exchange spread widely and even the Spanish soldiers learned to use agate as bargaining chips for gambling, since agate equaled money [@LiandWu2006, pp.
+Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) in exchange for necessities with other indigenous groups.
+This form of exchange was widespread and even the Spanish soldiers learned to use agate as bargaining chips for gambling, since agate equaled money [@LiandWu2006, pp.
 132-149].
-Father Jacinto Esquivel also recorded how indigenous viewed it as prestige goods in their culture.
-A indigenous male who wanted to get married had to pay agate beads to the parents of his future wife.
-Also, agate or golden beads could be used as to solve conflicts in their daily lives [@LiandWu2006, pp.
+Father Jacinto Esquivel also recorded how Indigenous people viewed agate pieces as prestige goods in their culture.
+A indigenous man who wanted to get married had to pay agate beads to the parents of his future wife.
+Also, agate or golden beads could be used as to resolve conflicts in their daily lives [@LiandWu2006, pp.
 151].
@@
 122-151].
-A funeral process in indigenous culture was recorded about how agate beads used commonly in ritual contexts, where more agate beads, pottery and cloths were be placed into the grave of more influential people to indicate a family's prestige [@LiandWu2006, pp.
+An Indigenous funeral process was recorded, documenting the use of agate beads in ritual contexts, with more agate beads, pottery and cloths placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
 153].
+Those historical accounts indicate that agate beads or golden beads were already treated as prestige goods before the arrival of the Europeans.
+These were likely introduced earlier from Chinese traders.
 
-Those documents indicate that agate beads or golden beads were already treated as prestige goods before the arrival of the Europeans, that is believed to be introduced earlier from Chinese traders.
-In 1642, the Dutch VOC defeated the Spanish and take over the northern Taiwan.
-They introduced feudal symbolism in an attempt to control the indigenous communities by asking indigenous leaders to attend annual ceremony for feudal practices [@Andrade2007, ch.
+In 1642, the Dutch VOC defeated the Spanish and took over their forts in northern Taiwan.
+They introduced a feudal system in an attempt to control the Indigenous communities by asking Indigenous leaders to attend annual ceremony for demonstrate their loyalty [@Andrade2007, ch.
 9; @Kang2016, ch.
 4].
-More beads entered indigenous settlements through the gift-giving that the Dutch provided goods based on the demand of native people to secure indigenous alliances in the annual ceremony or on the way of traveling [@Kang2016, ch.
+The Dutch provided beads and other goods based on the demands of indigenous people to secure alliances in the annual ceremony or during traveling [@Kang2016, ch.
 6].
-Thus, I assume that the way of gift giving resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities within existing consumer contexts during this period.
+Thus, I assume that the Dutch feudal system resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.
 
-The documents written by Chinese historians in Qing dynasty usually described the decorative purposes of the ornaments and some documents had similar texts that might be copied from previous ones without actual investigation.
-According to those records, indigenous people in Yilan had a custom to wear ornaments in ceremonial contexts to display their wealth and status [@Chen1963, pp.
+Chinese historical records from the \Qing dynasty contain some notes on the purposes of ornaments from Yilan.
+According to those records, Indigenous people in Yilan had a custom to wear ornaments in ceremonial contexts to display their wealth and status [@Chen1963, pp.
 228, 308; @Ke1993, pp.
@@
 77].
-Among those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production that usually possessed by rich people.
-Other people wore agate beads or glass beads on their head or neck to participating various ceremonies.
-In 1895, at the beginning of Japanese colonization, @Ino2002 conducted a field survey for several aboriginal groups and found that golden necklaces had not been used in Yilan that time, but elder people still used beads.
+Among those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production.
+These were usually possessed by wealthy people.
+Other people wore agate beads or glass beads on their head or neck to participate in ceremonies.
+In 1895, at the beginning of Japanese colonization, @Ino2002 conducted a field survey for several indigenous groups and found that golden necklaces were not used in Yilan at that time, but elderly people still used beads.
 
-Recent ethnographic research of the Kavalan tribe describes that agate beads were used in divination practice, called "subli", by female shamans [@Liu2008, pp.
+Recent ethnographic research with the living members of the Kavalan tribe describes how agate beads were used in divination practices, called “subli”, by female shamans [@Liu2008, pp.
 133-134].
-Interview records mention agate bead is valuable object that are usually passed down from mother to daughter as a heirloom in shaman families without knowing the exact origin.
-Were not been using, beads would be often hided at invisible place such as on the roof beams to make sure no one sees that.
-Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination that is consistent with the observation in European colonial period.
-It was usually the female shamans conducting divination ritual that passed down from their ancestors.
-Despite most aborigines today don’t know the origin of agate beads, the value of beads is still be highlighted through oral history and their scarcity.
-Those historical documents from the early 17th century to the modern ethnography described how local indigenous people use ornaments in local cultural contexts that represents some social roles or status.
-However, compared to European contact period, there is less mention of beads in Chinese contact period and the description about ornaments is only limited to their dressing culture.
+Interview records mention agate beads as valuable objects that are usually passed down from mother to daughter as a heirlooms in shaman families, without knowing the exact origin.
+When not in use, beads were often hidden in safe places, such as on the roof beams.
+Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination, consistent with observations from the European colonial period.
+It was usually the female shamans conducting divination rituals that passed the beads down from their ancestors.
+Despite most Kavalan people today not knowing the specific origin of their agate beads, the value of beads is still highlighted through oral history and by their scarcity.
+In sum, we find multiple sources describing how local Indigenous people used ornaments in cultural contexts that represents some social roles or high status.
+However, compared to European contact period, there are fewer mentions of beads in Chinese contact period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
 
-# Kiwulan in northeastern Taiwan
+# Excavations at Kiwulan in northeastern Taiwan
 
-```{r, KWL-map, fig.cap = "Kiwulan location in Southeast Asia", out.width= "400px", echo=FALSE}
+```{r, KWL-map, fig.cap = "Map showing the relative location of Kiwulan, Keelung, and Tamsui in northeastern Taiwan", out.width= "400px", echo=FALSE}
 library(here)
@@
 
-Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was a rescue archaeology projects that carried out from 2001 to 2004 in advance of construction.
-The total area includes eight open area sections that created 262 squares (4m by 4m) reaching 3,814 meter squares in total [@Chen2007].
-Most ornaments were found in situ in the context while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
-The archaeological evidence includes a rich amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures that indicates it was a settlement.
-The chronology of Kiwulan can be divided into Lower Layer Culture (700 - 1200 AD) and Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between based on a series of 32 radiocarbon dates [@Chen2007].
-This paper focuses on the Upper Layer Culture spanning from the late Iron Age that also called proto-historical period [@Liu2017], and historical period starting from the 17th century with European contact in Taiwan.
+Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was a rescue archaeology projects that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
+The total area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m2 in total [@Chen2007].
+The archaeological evidence includes a large amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
+Most ornaments were found in situ while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
+The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of excavation (\Figure 2).
+They were sampled because those units were stratigraphically intact with the least disturbance by modern construction activity, compared to excavation squares on the periphery of the site.
 
-The earliest record of direct European contact with indigenous people in Yilan can be traced back to 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
+The chronology of Kiwulan can be divided into two phases represented by a Lower Layer Culture (700 - 1200 AD) and a Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between.
+This chronology is based on the stratigraphy and a series of 32 radiocarbon dates [@Chen2007].
+This paper focuses on the Upper Layer Culture spanning from the late Iron Age, and historical period starting from the 17th century with European contact in Taiwan.
+
+Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with a hundred year intervals according to original radiocarbon dates, excavation depth, and types of diagnostic porcelain [@Hsieh2009; @Wang2011].
+The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
+However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
+For each samples square I separated excavation units into three episodes of time: pre-European contact period (L5, L6), European contact period (L4), and Chinese contact period (L2 and L1).
+
+I identify the start of European contact at Kiwulan at 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
 163].
@@
 12].
-The diagnostic artifacts of European contact found at Kiwulan including An-ping jars, stonewares that were largely introduced to Taiwan during the early 17th century.
-Those artifacts coupled with the original radiocarbon dates were used to identify the pre-European and post-European periods.
-The European colonization ended in 1662 when they were defeated by the kingdom of Tungning found by Koxinga from China.
+\Q:The start of European contact Kiwalan is indicated in the archaeological record by the presence of An-ping jars, stonewares that were largely introduced to Taiwan during the early 17th century.
+I used the presence and absence of An-ping jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.
+European colonization ended in 1662 when they were defeated by the kingdom of Tungning found by Koxinga from China.
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The evidence of Chinese migration can be identified both from Chinese official records and large amount of Chinese daily life porcelains found at Kiwulan [@Chen1963].
+I recongnise this as the Chinese colonization period, which is archaeological visible at Kiwulan by \Chinese porcelains and bricks that brought by Chinsese communities [@Hsieh2009].
+Also, the description of Chinese migration can be identified from Chinese official records [@Chen1963].
 
@@
 ```{r, KWL-sam-area, fig.cap = "Forty squares were sampled from the largest open area (in red with square ID number)", out.width= "70%", echo=FALSE}
+
+# Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
+Need to add in here that each square is 4 x 4 m, and say which way north is.
+Or add to the map and scale and north arrow.
+
 # mapping sampling area
@@
 
-The ornaments studied in this paper found from 40 adjacent square of the largest open area that was located in the middle part of excavation (Figure \@ref(fig:KWL-sam-area).
-They were sampled because those units were more stratigraphically intact with the least disturbance by modern constructions.
-The Upper Layer Culture component could be divided into six sub contexts spanning from 14 century to 19 century with a hundred year interval according to original radiocarbon dates, excavation depth, and types of diagnostic porcelain [@Hsieh2009; @Wang2011].
-The contexts were assigned numbers from 1 to 6 where L1 corresponds to the youngest context and L6, the bottom layer, refers to the oldest context.
-To obtain a more reliable chronology for comparison between episodes of culture contact,I reevaluate those contexts for each sampled square by focusing on Anping jars and stonewares that are believed manufactured and distributed in Southeast Asia through European trade route during the early 17th century [@Ketel2011].
-Those two categories of ceramic coupled with radiocarbon dates were used to identify the pre-European and post-European periods.
-Following the previous numbering principle, I separated three episodes of time including pre-European contact period(L5, L6), European contact period(L4), and Chinese contact period (L2 and L1).
+# The ornaments
 
@@
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D categories were from original excavation report.", out.width= "400px", echo=FALSE}
+The photographs of B, C, D categories were from original excavation report (Chen 2007).", out.width= "400px", echo=FALSE}
 knitr::include_graphics(here("analysis", "figures", "ornament.png"))
@@
 
-To explore whether body ornaments can reflect the colonial influences on indigenous populations, this paper examines the variety, frequency, and spatial distribution of trade ornaments across different time periods, including pre-European contact, European contact, and Chinese contact.
-Those three analytic units could represent three scales of exchange network from small to large, and three degree of contact from weak to intense.
-Because the appearance of ornaments is highly related to the original raw material, the type of raw material is the first criterion of classification that determined five major categories.
-In each category, ornaments were classified into several subtype according to their shapes when variations were present in each category.
-The variation was defined by the difference identified in at least two variables, for example dimensional variables such as length, or stylish variables such as patterns.
-In order to compare the general pattern of changes for subtypes, less frequent subtypes were left out from the analysis, but will be discussed separately in later section.
-
-Ornaments from the post-hole area accounts for the majority of the amount in total.
+Ornaments from the post-hole area accounts for the \majority of the amount in total.
 Post-hole area provides a clear contexts and basis for comparison to see if there is any spatial pattern of distribution.
-For the post-hole area, there were `r sample_size` ornaments recovered from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the same open area section.
+For the post-hole area, there were `r sample_size` ornaments recovered from the 40 sampling squares, which accounts \for `r sample_pro` % of the total number of ornaments found in the same open area section.
 Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
 The variety of subtype and their individual frequency are summarized in \@ref(tab:tidy-data-for-table) in terms of three time episodes.
-The image for each subtype in major category shows in figure \@ref(fig:ornaments-image).
+The image for each subtype in each major category is shown in Figure \@ref(fig:ornaments-image).
+
+<---include the number of ornaments from burials above--->
+
+<---Something sounds wrong here. You say ‘majority’ but then later in the paragraph you say it’s 0.5% of the total number? How can 0.5% be the majority? --->
 
 ```{r, plot-counts-by-period, fig.cap = "Frequency of each category of ornament with a separate plot showing different time periods individually"}
+### delete one plot
+
 # categories of ornaments by period
@@
 
-# Results
+# Results of hypothesis testing
 
-## Frequency between time periods
-Figure \@ref(fig:plot-counts-by-period) shows the comparison between different time periods that indicates a general pattern that most ornaments existed already before European contact and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
+In this section I report the tests of the three hypotheses about indirect impacts of European colonization on the Indigenous community at Kiwulan.
+To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European contact, and a lesser diversity after Chinese contact, I compare frequencies of ornament types between different time periods.
+To test the prediction that after European contact were new spatial patterns in the distributions of ornaments at Kiwulan, I use kernel density and point pattern analyses.
+
+## Frequency of ornaments between time periods
+Figure \@ref(fig:plot-counts-by-period) shows the comparison of ornament type frequencies between different time periods at Kiwulan.
+It indicates a general pattern that most ornament types were present before European contact.
+The frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
 This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
-However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even much less in the Chinese contact period.
+However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then their frequency decreased in the European contact period and was even much less in the Chinese contact period.
 
@@
 
-The distribution of frequency for subtypes in each major category are presented in \@ref(fig:plot-all-ornaments-count).
-The distribution shows that agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period.
+The distribution of frequency for subtypes in each major category are presented in Figure \@ref(fig:plot-altogether).
+The distribution shows that agate beads and metal rings have greater quantity and \variety of shapes compared to copper bell and glass beads during the European contact period.
 Agate beads can be divided into seven subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased significantly during the European contact and then declined in the Chinese contact.
-Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period.
-Higher variety of agate beads were found in European contact compared to other time periods.
+The small oval type shows a similar frequency before and during European contact but is not found in the Chinese contact period.
+Waxy oval beads were only found during European contact period.
+A higher variety of agate beads were found in European contact compared to other time periods.
 Similarly, metal rings can be classified into eight subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape.
-The greater varieties for those two categories might indicate the multiple origins due to global trade network stimulated by the Europeans.
-In contrast, copper bell and glass bead have less variety, but glass bead has larger number due to they were usually found in a cluster that indicates the original use as strings or necklaces.
-The common shape for copper bell is the large shape with a wide variety of human face as a motif, while the common shape for glass bead is bead with length less than 1cm that are identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].
+The greater varieties for those two categories might indicate multiple origins due to global trade networks stimulated by the Europeans.
+In contrast, copper bell and glass bead have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces.
+The common shape for copper bells is the large shape with a wide variety of human faces as a motif, while the common shape for glass beads is a bead with length less than 1 cm that are often identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].
+
+<---I think we need an addition bit of analysis in this section to show the relationship between diversity and sample size.
+We naturally expect diversity to increase as sample size increases.
+So we want to test and see if the changes in diversity that you describe here are only the natural effects of increase/decrease in sample size, or there are ‘real’ increases in diversity that cannot be explained by sample size changes.
+The peer reviewers will want to know the answers to this.
+--->
 
@@
 ```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
+
+#I don’t see any reference to this in the text.
+If you don’t reference a figure in the text, you need to delete it from the manuscript, Why not use a colour scale for frequency?
+I think the viridis scale would be perfect for this
+
 ggplot() +
@@
 ```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for each class of ornament by time periods"}
+
+# Seems like some ornament types have <5 pieces. Probably not worth to plot those. You can write in the caption ‘only those types with n > 5 are shown here’
+
 # filter golden bead
@@
 
-## Frequency for spatial pattern
-Figure \@ref(fig:plot-spatial-pattern-each) presents the spatial distribution of all categories for each time period.
-It shows that one square unearthed more ornaments than others in the pre-European contact.
-During the European contact period, ornaments was more ubiquitous with some clusters on northern part, and less presence during the Chinese contact period.
-When looking at the distribution for each major class individually, some clusters across the area were observed during the European contact including golden bead, agate bead, metal ring, and bell.
-However, it seems no clear consistent clustered pattern across those different ornaments.
+## Spatial distribution of ornament types
+
+Figure \@ref(fig:plot-spatial-pattern-each) presents the spatial distribution of the main ornament types for each time period.
+Before the European contact, the map shows a greater amount of glass beads found at the northern and middle parts of the research area, and they were more widespread during the European contact period.
+During the European contact period, ornaments were more ubiquitous with some clusters on northern part, and less abundance during the Chinese contact period.
+When looking at the distribution for each major class individually, some clusters across the area can be observed during European contact, including golden beads, agate beads, metal rings, and bells.
+However, there seems to be no clear consistent clustered pattern across those different ornaments.
 Each class shows its own pattern where the squares with higher number of ornaments distributed separately and independently.
-For example, the cluster of golden bead were found at the northern part, while the cluster of agate bead were found in the middle part.
-In contrast, there are more than one cluster of metal rings that distributed separately across the research area.
-Copper bells were usually found solely and seem randomly distributed across the area.
-Before European contact, the map shows greater amount of glass beads were found at the northern and middle part of the research area, and more widespread during the European contact period.
-In the Chinese Contact period, both the amount and density of different classes of ornaments decreased that they were only found in a few squares, for example, golden bead and glass bead were only found in a couple squares at the middle part.
+For example, the cluster of golden beads was found at the northern part, while the cluster of agate beads was found in the middle part.
+In contrast, there are multiple clusters of metal rings that are distributed separately across the research area.
+Copper bells were usually found individually and seem randomly distributed across the area.
+In the Chinese Contact period, both the amount and density of different classes of ornaments decreased that they were only found in a few squares, for example, golden beads and glass beads were only found in a few squares in the middle part.
 
@@
 
-## Spatial distribution
-To examine whether the distribution of ornament is uneven that could reflect the presence of social inequality, point pattern analysis is used to examine the distribution.
-Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some "hotspots" that imply social processes [@Bevan2016; @Ducke2015], and in this case the pattern of ornaments could be able to indicate the changes due to culture contact.
-The location where ornaments from were square-based recorded, thus the intensity approach was used that focuses on the average density of points across space to explore their distribution.
-Points were be randomly assigned to each square according to the frequency data.
-Next is to subset three groups for three time periods, and then compute the kernal density for each time period for comparison.
-Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
+## Point pattern analysis of ornament distribution 
+I used point pattern analysis  to examine whether the distribution of ornaments was uneven in non-random ways that could reflect the presence of social inequality.
+Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
+).
+In this case hotspots in the pattern of ornaments could indicate the changes in social organization due to culture contact.
+I used the intensity approach that focuses on the average density of points across space to explore their distribution.
+Ornament location points were given random coordinates within each square they they were recovered from because their exact recovery location in each square was not recorded excavation.
+The next step was to subset the orgaments into three groups for three time periods, and then compute the kernel density for each time period for comparison.
+Kernel density estimations (KDE) estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
 Here we use KDE to visualize the event of the presence of ornaments that were represented by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
-Density value is assigned to each cell, of which the unit is meter.
-The results \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period.
-Those maps present three consistent sub-region with shifting core area over time.
-The distribution might indicate different social groups who possessed more ornaments, and the multiple groups during European contact period might reflect some presence of unequal consumption.
-However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious hot spot.
-To further evaluate the pattern, we test the hypothesis that whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.
+Density values of artefacts per square meter are assigned to each cell.
+The results in Figure \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period.
+Those maps present three consistent sub-regions with a core area that shifts over time.
+The distribution might indicate different social groups who possessed more ornaments.
+The multiple groups during European contact period might reflect more unequal consumption of ornaments across the site, relative to other periods.
+However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
+To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is randomly distributed to understand the underlying social processes related to ornament distributions.
 
@@
 
+# make the blue lines thicker to be more visible
 par(mfrow = c(3, 1))
@@
 
-Here we use Monte Carlo method to test the complete spatial randomness of spatial events based on the average nearest-neighbor distance method (ANN).
-We hypothesize that the distribution of ornaments is consistent with a completely random process, and then simulate the presence of ornaments across the space for one thousand times for each time periods.
-The data was compared to the random pattern generated by simulated processes to see if our data is randomly distributed or not.
-The results \@ref(fig:hypo-test-distrubution) are represented by histogram showing one thousand simulation.
-The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments are clustered distributed.
-Similar result is also observed during the pre-European period but less extreme.
-The distribution of ornaments is more random during the Chinese contact period and about one third of the simulated values are greater than our observed ANN value.
-This testing explains that the distribution of ornaments during the pre-European and European contact period might reflect the presence of different social groups.
-Moreover, the cluster of ornaments way more significant during the European contact period that might indicate some extent of control of ornaments.
+Here we use a Monte Carlo method to test the spatial randomness of spatial events based on the average nearest-neighbor distance method (ANN).
+Our null hypothesis is that the distribution of ornaments is consistent with a completely random process, so we then simulate the locations of ornaments across the space 1000 times for each time period.
+The observed data was compared to the random pattern generated by simulated processes to see if our data is randomly distributed or not.
+Figure \@ref(fig:hypo-test-distrubution) shows the distributions of the ANN distances calculated on 1000 simulations of ornament locations.
+The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments have a non-random clustered distributions.
+A similar, but less extreme, result is also observed during the pre-European period.
+The observed distribution of ornaments is more similar to the random distributions during the Chinese contact period, with about one third of the simulated values are greater than our observed ANN value.
+This testing reveals that the clustered distributions of ornaments during the pre-European and European contact periods might reflect the presence of different social groups.
+Moreover, the clustering of ornaments during the European contact period is highly non-random, potentially indicate some social changes leasing to a concentration of power to control the distribution of ornaments at Kiwulan.
 
 # Discussion 
-
-The introduction of ornaments to Northeastern Taiwan during different periods could inform how local indigenous people perceive and distribute the trading goods in different contexts of culture contacts by examining their frequency and spatial distribution across space.
-Taiwan island had been involved in regional exchange networks of Southeast Asia since Neolithic period, a global trade network during the European contact period in the early 17th century, and later intense but smaller scale network with Han Chinese communities since 19th century [@Chen2005].
-Kiwulan demonstrates continuous occupation covers those episode of culture contacts that enables a comparative research to explore the indirect influence of culture contact especially in the global colonial context by examining prestige goods.
-These three analytic chronology at Kiwulan represent different scale of exchange network and intensity of culture contact.
-For the scale of trade network, Pre-European contact can be defined as a middle scale represented by the prehistoric regional trade network in Southeast Asia.
-International trade network during the European contact presents a global scale, while local trade network during the Chinese contact presents a small scale.
-For the intensity of culture contact, both pre-European contact and European contact period could be defined as indirect contact, but the colonial power underlying the contact with the Europeans should be also considered.
-The culture contact during the Chinese period is more direct due to the migration of Han Chinese to Kiwulan.
+<--I think you have space for a short paragraph here on  network/corporate model at the end>
 
 The results show there was a greater diversity of ornaments types and materials at Kiwulan during European contact.
-This might result from the large scale exchange network that stimulate the circulation of different ornaments classes.
-Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads, which might indicate the diverse origins.
+This might result from the large scale exchange network that stimulated the circulation of different ornament classes.
+Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads, which might indicate their diverse origins.
 However, the frequency of overall ornaments and each subtype declines significantly during the Chinese contact period.
-The reason could be related to smaller scale of network and the overall decline of Kiwulan population at that time.
-The spatial pattern of ornaments shows that there were some clusters appeared since pre-European and European contact period.
-The hypothesis testing for spatial pattern further indicate more concentration of ornaments during European contact.
-Because trade ornaments had been viewed as prestige goods in local cultural context in the prehistoric Northeastern Taiwan, this differentiation could reflect some level of social inequality appeared before European contact and then it was reinforced during European period.
+The reason could be related to smaller scale of network and the \overall \decline of Kiwulan population at that time.
+
+<----Why did it decline? What is the evidence for this? A few more sentences are needed on this-->
+
+The spatial pattern of ornaments shows that their distribution was clustered since the pre-European and European contact period.
+Hypothesis testing for spatial patterns further indicates that these clusters are non-random, and are concentrated during European contact.
+Because trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering could indicate that a degree of social inequality were present before European contact and then it was reinforced and amplified during the European contact period.
 
@@
 
-Indirect influence of colonialism has been proposed recently and some studies reveal how peripheral areas had been influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017].
-Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts.
-Foreign trade ornaments were viewed as prestige goods and played an important role in the cultural context of indigenous societies in northeastern Taiwan according to the use context as grave goods and historical documents written by foreign record keeper.
+Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017].
+Kiwulan in Northeastern Taiwan is an especially important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by tea to regional and global trade networks.
+Kiwulan can provide valuable insights into the discussion of the influence of indirect colonial contacts.
+In northeast Taiwan foreign trade ornaments were viewed as prestige goods and played an important social role in divination and funeral rituals.
 As a marker for status, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
-Greater amount and diversity ornament types during the European contact period reflects the international stimulation of exchange in a colonial context.
-Those ornaments may convey different meaning with exotic and colonizer image that increases more competitions between aggrandized individuals for prestige and wealth accumulation, which might result in a increasing social inequality.
-The spatial pattern of ornaments at Kiwulan during European contact presents multiple clusters that is significantly different from a random distribution pattern.
-This study demonstrates that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization.
+The greater amount and diversity of ornament types during the European contact period reflects international exchange in a colonial context.
+Those ornaments may carry an exotic and powerful image that signals wealth and trading connections to the inhabitants of Kiwulan.
+These symbolic values may have stimulated more competitions between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might result in a increase in social inequality.
+
+This study demonstrates that foreign ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations.
+Ornaments can give insights into the amplification of social inequality stimulated by the European colonization.
 It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system.
-During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwalan population.
-This is a case study that focuses on trade ornaments and more studies can be done to include other trade goods like porcelains.
-Also, further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.
+During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwulan population.
+Future work could extent this approach to studies of other trade goods such as ceramics.
+Future directions for understanding the dynamics of social inequality at Kiwulan could include analysis of pottery production and standisation, and mortuary practices.
+
 # Acknowledgements

### 2019-08-01 analysis/paper/paper.Rmd
update

@@
 Throughout East Asia direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
-Understanding the indirect effects of colonialism are important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
-In many parts of the world, the introduction of foreign trade goods by colonial traders into local Indigenous societies caused significant transformations of Indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
-The agency of Indigenous peoples in the colonial contexts has been broadly discussed to understand the negotiations between colonized and colonizer, and the resistance and accommodations of Indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
-Studying the consumption of foreign goods is one way to understand Indigenous experiences of colonial encounters and explore the agency of Indigenous peoples by examining how goods were distributed and used [@Dietler2005; @Mullins2011; @Scaramelli2005].
+Understanding the indirect effects of colonialism are important for detecting colonial impacts on indigenous peoples in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
+In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused significant transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
+The agency of indigenous peoples in the colonial contexts has been broadly discussed to understand the negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
+Studying the consumption of foreign goods is one way to understand indigenous experiences of colonial encounters and explore the agency of indigenous peoples by examining how goods were distributed and used [@Dietler2005; @Mullins2011; @Scaramelli2005].
 
-Taiwan provides an ideal case study of indirect effects of colonialism because it has archaeological evidence spanning its initial involvement in regional exchange networks of East Asia and Southeast Asia since the tenth century.
+Taiwan provides an ideal case study of indirect effects of colonialism because it has archaeological evidence spanning its initial involvement in regional exchange networks of East Asia and Southeast Asia since the \tenth century.
 It became part of global trade systems during the early 17th century due to western European expansions that connected Europe and Asia [@Liu2017].
-Despite colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of Indigenous responses to the European colonization.
+Despite colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonization.
 @Berrocal2018 study \the long-term record of Heping Dao\ in northern Taiwan and explore the effects of European colonization on the local populations by focusing on the environmental change and introduction of diseases, with a little mention of indigenous social complexity.
 Yilan in Northeastern Taiwan is an especially unique location for exploring indirect effects due to its physically isolated location.
-The rugged \XXX mountains to the west of the Yilan Plain that made it not easily accessible to Europeans and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-The sea provides a way through which Indigenous people in Yilan could interact with other groups from different regions, including other native Indigenous groups, foreign traders such as the Han people of China, and Europeans.
+The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
+The sea provides a way through which indigenous people in Yilan could interact with other groups from different regions, including other native Indigenous groups, foreign traders such as the Han people of China, and Europeans.
 Large-scale and frequent trade activities brought a wide variety of trade objects into Indigenous communities in Yilan through local regional exchange network.
@@
 Personal adornments are a social signal of an individual’s status [@Joyce2005; @Scaramelli2005].
-For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].
+For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
 
@@
 I focus on  the ornaments recovered from the archaeological excavations at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
-Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems, and \later encountered a large wave of Han Chinese migration in the 19th century [@Chen2005].
-Northeastern Taiwan provides a good example to explore the indirect colonial impacts on Indigenous societies by comparing archaeological evidence from different episodes of culture contact [\cf.@Lape2003].
+Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems, and encountered \several waves of Han Chinese migration in the 19th century [@Chen2005].
+Northeastern Taiwan provides a good example to explore the indirect colonial impacts on Indigenous societies by comparing archaeological evidence from different episodes of culture contact [cf.@Lape2003].
 I will describe the variety and cultural context of trade ornaments in this region, and examine their potential of the ornaments to inform on social changes relating to trade with colonial groups.
@@
 I also predict that after European contact there will be new spatial patterns in the distributions of ornaments at Kiwulan that indicates increasing social inequality.
-I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan due to \???.
+I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan due to overall population declines and acculturation\.
 
@@
 
-Prior to European colonization, the indigenous people in northern Taiwan had already established their own exchange network that has been described as "inter-insular trade" by @Chen2005 [pp.
+Prior to European colonization, the Indigenous people in northern Taiwan had already established their own exchange network that has been described as “inter-insular trade” by Chen @Chen2005 [pp.
 12] to refer to small-scale regional trade between China and Taiwan on irregular basis.
@@
 Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
-Rivers and seas provide the way for the interaction between local indigenous settlements and further \afield communities, such as the Basai in northern Taiwan and Han Chinese.
+Rivers and seas provide the way for the interaction between local indigenous settlements and further afield communities, such as the Basai in northern Taiwan and Han Chinese.
 Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
@@
 Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) in exchange for necessities with other indigenous groups.
-This form of exchange was widespread and even the Spanish soldiers learned to use agate as bargaining chips for gambling, since agate equaled money [@LiandWu2006, pp.
+This form of exchange was widespread and even the Spanish soldiers learned to use agate as bargaining chips for gambling [@LiandWu2006, pp.
 132-149].
-Father Jacinto Esquivel also recorded how Indigenous people viewed agate pieces as prestige goods in their culture.
+Esquivel also recorded how indigenous people viewed agate pieces as prestige goods in their culture.
 A indigenous man who wanted to get married had to pay agate beads to the parents of his future wife.
@@
 122-151].
-An Indigenous funeral process was recorded, documenting the use of agate beads in ritual contexts, with more agate beads, pottery and cloths placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
+An indigenous funeral process was recorded, documenting the use of agate beads in ritual contexts, with more agate beads, pottery and cloths placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
 153].
@@
 In 1642, the Dutch VOC defeated the Spanish and took over their forts in northern Taiwan.
-They introduced a feudal system in an attempt to control the Indigenous communities by asking Indigenous leaders to attend annual ceremony for demonstrate their loyalty [@Andrade2007, ch.
+They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend annual ceremony for demonstrate their loyalty [@Andrade2007, ch.
 9; @Kang2016, ch.
 4].
-The Dutch provided beads and other goods based on the demands of indigenous people to secure alliances in the annual ceremony or during traveling [@Kang2016, ch.
+The Dutch provided beads and other goods based on the demands of Indigenous people to secure alliances in the annual ceremony or during traveling [@Kang2016, ch.
 6].
@@
 
-Chinese historical records from the \Qing dynasty contain some notes on the purposes of ornaments from Yilan.
-According to those records, Indigenous people in Yilan had a custom to wear ornaments in ceremonial contexts to display their wealth and status [@Chen1963, pp.
+Chinese historical records from the year of 1829, 1837, and 1852 during the Qing dynasty [1616-1911] contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
 228, 308; @Ke1993, pp.
@@
 77].
+According to those records, indigenous people in Yilan had a custom to wear ornaments in ceremonial contexts to display their wealth and status.
 Among those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production.
@@
 Other people wore agate beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonization, @Ino2002 conducted a field survey for several indigenous groups and found that golden necklaces were not used in Yilan at that time, but elderly people still used beads.
+In 1895, at the beginning of Japanese colonization, @Ino2002 conducted a field survey for several Indigenous groups and found that golden necklaces were not used in Yilan at that time, but elderly people still used beads.
 
@@
 
-Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with a hundred year intervals according to original radiocarbon dates, excavation depth, and types of diagnostic porcelain [@Hsieh2009; @Wang2011].
+Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with a hundred year intervals according to original radiocarbon dates, excavation depth, and types of diagnostic porcelains [@Hsieh2009; @Wang2011].
 The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
 However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
-For each samples square I separated excavation units into three episodes of time: pre-European contact period (L5, L6), European contact period (L4), and Chinese contact period (L2 and L1).
+For each sampled square I separated excavation units into three episodes of time: pre-European contact period (L5, L6), European contact period (L4), and Chinese contact period (L2 and L1).
 
@@
 12].
-\Q:The start of European contact Kiwalan is indicated in the archaeological record by the presence of An-ping jars, stonewares that were largely introduced to Taiwan during the early 17th century.
+The start of European contact in Kiwalan is indicated in the archaeological record by the presence of An-ping jars, stonewares that were largely introduced to Taiwan during the early 17th century.
 I used the presence and absence of An-ping jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.

### 2019-08-09 analysis/paper/paper.Rmd
editting based on comments

@@
 )
-
 ```
@@
 
-Taiwan provides an ideal case study of indirect effects of colonialism because it has archaeological evidence spanning its initial involvement in regional exchange networks of East Asia and Southeast Asia since the \tenth century.
-It became part of global trade systems during the early 17th century due to western European expansions that connected Europe and Asia [@Liu2017].
-Despite colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonization.
-@Berrocal2018 study \the long-term record of Heping Dao\ in northern Taiwan and explore the effects of European colonization on the local populations by focusing on the environmental change and introduction of diseases, with a little mention of indigenous social complexity.
-Yilan in Northeastern Taiwan is an especially unique location for exploring indirect effects due to its physically isolated location.
+Taiwan became part of global trade systems during the early 17th century due to western European expansions that connected Europe and Asia.
+Despite the colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonization.
+@Berrocal2018 study both archaeological and historical records of Heping Dao in northern Taiwan that was the location of the colony of San Salvador founded by the Spanish in 1626 and took over by the Dutch in 1642.
+They uncovered parts of the foundation of a Spanish church and a European cemetery that indicate the evidence of European occupation.
+It shows a direct impact on the local indigenous populations by disturbing their long settlement pattern and changing local landscape.
+Different from the view of a Spanish fort, Yilan in Northeastern Taiwan is an especially unique location for exploring indirect colonial effects on local indigenous settlements due to its physically isolated location.
 The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-The sea provides a way through which indigenous people in Yilan could interact with other groups from different regions, including other native Indigenous groups, foreign traders such as the Han people of China, and Europeans.
-Large-scale and frequent trade activities brought a wide variety of trade objects into Indigenous communities in Yilan through local regional exchange network.
+The sea provides a way through which indigenous people in Yilan could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans.
+Large-scale and frequent trade activities brought a wide variety of trade objects into indigenous communities in Yilan through local regional exchange network.
 One of the most common types of trade goods in this region were personal ornaments such as glass and stone beads was one of the common foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
@@
 
-In this paper, I explore the evidence from Yilan spanning from late Iron Age to historical period to address the question of whether indirect colonial influences on Indigenous populations can be detected.
+In this paper, I explore the evidence from Yilan spanning from late Iron Age to historical period to address the question of whether indirect colonial influences on indigenous populations can be detected.
 I focus on the ornaments recovered from the archaeological excavations at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
-Prior to the contact with the Europeans in the early 17th century, northern Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems, and encountered \several waves of Han Chinese migration in the 19th century [@Chen2005].
-Northeastern Taiwan provides a good example to explore the indirect colonial impacts on Indigenous societies by comparing archaeological evidence from different episodes of culture contact [cf.@Lape2003].
+Prior to the contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since tenth century, and encountered large waves of Han Chinese immigration after 17th century since Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
+11].
+Northeastern Taiwan provides a good example to explore the indirect colonial impacts on indigenous societies by comparing archaeological evidence from different episodes of culture contact [cf.@Lape2003].
 I will describe the variety and cultural context of trade ornaments in this region, and examine their potential of the ornaments to inform on social changes relating to trade with colonial groups.
@@
 I also predict that after European contact there will be new spatial patterns in the distributions of ornaments at Kiwulan that indicates increasing social inequality.
-I hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan due to overall population declines and acculturation\.
+I hypothesize that after Chinese contact there was a decline in the use and discard of ornaments at Kiwulan due to overall population declines and acculturation in the Chinese community.
 
@@
 
-Prior to European colonization, the Indigenous people in northern Taiwan had already established their own exchange network that has been described as “inter-insular trade” by Chen @Chen2005 [pp.
-12] to refer to small-scale regional trade between China and Taiwan on irregular basis.
-This local exchange network of northeastern Taiwan was increasingly influenced by long distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since 15th century due to its location on the shipping routes.
-This was intensified in the 17th century with the arrival of the Europeans and their trade goods [@Chen2005; @Liu2017; @Wang2007].
-Due to its location on the periphery of major trade ports, \Keelung and \Tamsui, in northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
+Northeastern Taiwan was part of the local exchange network in northern Taiwan, where the indigenous people had already established their own exchange network before European arrived that has been described as “inter-insular trade” by Chen @Chen2005 [pp.
+12] to refer to small-scale regional trade between China and Taiwan on an irregular basis.
+The exchange network in northern Taiwan was increasingly influenced by long-distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since the 15th century due to its location on the shipping routes [@Chen2005].
+This was intensified in the 17th century with the arrival of the Europeans and their trade goods [@Wang2007].
+Due to its location on the periphery of major trade ports, Keelung and Tamsui, in northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
 Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
@@
 Historical documents from the Spanish and Dutch in the early 17th century provide some information about the social life and culture of indigenous people in Northern Taiwan when the Europeans arrived.
-The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purpose [@Blusse2000, pp.
+The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for a religious purpose [@Blusse2000, pp.
 343].
@@
 Esquivel also recorded how indigenous people viewed agate pieces as prestige goods in their culture.
-A indigenous man who wanted to get married had to pay agate beads to the parents of his future wife.
-Also, agate or golden beads could be used as to resolve conflicts in their daily lives [@LiandWu2006, pp.
+An indigenous man who wanted to get married had to pay agate beads to the parents of his future wife.
+Also, agate or golden beads could be used to resolve conflicts in their daily lives [@LiandWu2006, pp.
 151].
@@
 122-151].
-An indigenous funeral process was recorded, documenting the use of agate beads in ritual contexts, with more agate beads, pottery and cloths placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
+An indigenous funeral process was recorded, documenting the use of agate beads in ritual contexts, with more agate beads, pottery, and cloths placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
 153].
@@
 In 1642, the Dutch VOC defeated the Spanish and took over their forts in northern Taiwan.
-They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend annual ceremony for demonstrate their loyalty [@Andrade2007, ch.
+They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty [@Andrade2007, ch.
 9; @Kang2016, ch.
@@
 77].
-According to those records, indigenous people in Yilan had a custom to wear ornaments in ceremonial contexts to display their wealth and status.
+According to those records, indigenous people in Yilan had the custom to wear ornaments in ceremonial contexts to display their wealth and status.
 Among those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production.
@@
 Other people wore agate beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonization, @Ino2002 conducted a field survey for several Indigenous groups and found that golden necklaces were not used in Yilan at that time, but elderly people still used beads.
+In 1895, at the beginning of Japanese colonization, a field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+227-232].
 
@@
 133-134].
-Interview records mention agate beads as valuable objects that are usually passed down from mother to daughter as a heirlooms in shaman families, without knowing the exact origin.
+Interview records mention agate beads as valuable objects that are usually passed down from mother to daughter as heirlooms in shaman families, without knowing the exact origin.
 When not in use, beads were often hidden in safe places, such as on the roof beams.
@@
 In sum, we find multiple sources describing how local Indigenous people used ornaments in cultural contexts that represents some social roles or high status.
-However, compared to European contact period, there are fewer mentions of beads in Chinese contact period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
+However, compared to the European contact period, there are fewer mentions of beads in Chinese contact period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
 
@@
 
-```{r, KWL-map, fig.cap = "Map showing the relative location of Kiwulan, Keelung, and Tamsui in northeastern Taiwan", out.width= "400px", echo=FALSE}
+```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px", echo=FALSE}
+# need to add the locations of Keelung, Tamsui, and hoping dao
 library(here)
@@
 ```
-
-Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was a rescue archaeology projects that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
-The total area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m2 in total [@Chen2007].
-The archaeological evidence includes a large amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
-Most ornaments were found in situ while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
-The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of excavation (\Figure 2).
-They were sampled because those units were stratigraphically intact with the least disturbance by modern construction activity, compared to excavation squares on the periphery of the site.
-
-The chronology of Kiwulan can be divided into two phases represented by a Lower Layer Culture (700 - 1200 AD) and a Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between.
-This chronology is based on the stratigraphy and a series of 32 radiocarbon dates [@Chen2007].
-This paper focuses on the Upper Layer Culture spanning from the late Iron Age, and historical period starting from the 17th century with European contact in Taiwan.
-
-Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with a hundred year intervals according to original radiocarbon dates, excavation depth, and types of diagnostic porcelains [@Hsieh2009; @Wang2011].
-The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
-However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
-For each sampled square I separated excavation units into three episodes of time: pre-European contact period (L5, L6), European contact period (L4), and Chinese contact period (L2 and L1).
-
-I identify the start of European contact at Kiwulan at 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
-163].
-Later in 1647, the Dutch attacked the indigenous villages and forced them to accept colonial rules and economic demands by paying annual tribute [@Andrade2007].
-According to Dutch census reports in 1650, Kiwulan was the largest indigenous settlement in the Yilan Plain, with a population of 840 adults [@Nakamura1938, pp.
-12].
-The start of European contact in Kiwalan is indicated in the archaeological record by the presence of An-ping jars, stonewares that were largely introduced to Taiwan during the early 17th century.
-I used the presence and absence of An-ping jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.
-European colonization ended in 1662 when they were defeated by the kingdom of Tungning found by Koxinga from China.
-Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-I recongnise this as the Chinese colonization period, which is archaeological visible at Kiwulan by \Chinese porcelains and bricks that brought by Chinsese communities [@Hsieh2009].
-Also, the description of Chinese migration can be identified from Chinese official records [@Chen1963].
 
@@
                 na.rm = TRUE))
-
-sample_size <- nrow(ornaments_period)
-total_AD <- nrow(kwl_upper)
-sample_pro <- round(sample_size/total_AD, 1)
 ```
 
-```{r read-in-the-spatial-data, eval = TRUE}
-library(sf)
-library(cartography)
-burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
-AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
-AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
-location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
-units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
+```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number.
+Each square is 4 x 4 m", out.width= "70%", echo=FALSE}
+
+# add to the map and scale and north arrow. 
+knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.png"))
 ```
 
-```{r tidy-spatial-data, eval = TRUE}
-# 40 sampling units
-sample_units <-
-  AD_data %>% 
-  semi_join(units_40, by = c (`坑號` = "the_sq"))
+Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
+The total area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
+The archaeological evidence includes a large amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
+Most ornaments were found in situ while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
+The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of the excavation, see \@ref(fig:KWL-sam-area).
+They were sampled because those units were stratigraphically intact with the least disturbance by modern construction activity, compared to excavation squares on the periphery of the site.
 
-# join two dataset to get coordinate
-ornaments_period_join <-
-  ornaments_period %>%
-  mutate(period = as.factor(period)) %>% # seems need this line first 
-  group_by(Pit_No, period) %>% 
-  count(Categories) %>% 
-  inner_join(AD_data, by = c("Pit_No" = "坑號"))
-```
+The chronology of Kiwulan can be divided into two phases represented by a Lower Layer Culture (700 - 1200 AD) and a Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between.
+This chronology is based on the stratigraphy and a series of 32 radiocarbon dates [@Chen2007].
+This paper focuses on the Upper Layer Culture spanning from the late Iron Age to the historical period starting from the 17th century with European contact in Taiwan.
 
-```{r, KWL-sam-area, fig.cap = "Forty squares were sampled from the largest open area (in red with square ID number)", out.width= "70%", echo=FALSE}
+Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with a hundred year intervals according to original radiocarbon dates, excavation depth, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
+The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
+However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
+For each sampled square I separated excavation units into three episodes of time: pre-European contact period (L5 and L6), European contact period (L4), and Chinese contact period (L2 and L1).
 
-# Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
-Need to add in here that each square is 4 x 4 m, and say which way north is.
-Or add to the map and scale and north arrow.
-
-# mapping sampling area
-ggplot(sample_units) +
-  geom_sf(data = AD_data, fill = NA) +
-  geom_sf(color = 'red') +
-  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
-  theme_minimal() +
-  theme_void() 
-```
+I identify the start of European contact at Kiwulan at 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
+163].
+Later in 1647, the Dutch attacked the indigenous villages and forced them to accept colonial rules and economic demands by paying annual tribute [@Andrade2007].
+According to Dutch census reports in 1650, Kiwulan was the largest indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
+12].
+I take approximate indicators of the start of European contact at Kiwulan as the appearance of An-ping jars and stonewares (martavans/martaban) that were largely introduced to Taiwan during the early 17th century.
+They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages, and the jar shapes found at Kiwulan are typical of those found elsewhere in V.O.C sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
+I used the presence and absence of An-ping jars and stonewares, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.
+European colonization ended in 1662 when they were defeated by the kingdom of Tungning founded by Koxinga from China.
+Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
+I recognize this as the Chinese contact period, which is archaeologically visible at Kiwulan by a large amount and varieties of Chinese porcelains, clay stoves, and a few opium pipe-bowls that were identified as items commonly used in the daily life of Chinese and the use of bricks and tiles for architectural buildings [@Hsieh2009].
+Also, the description of the Chinese migration can be seen from the Chinese official records written in the early 19th century [@Chen1963; @Ke1993].
 
@@
 
-In general, complete ornaments with clear context were mostly found in burials that helps to understand how people wear them.
-In addition to burials, large amount of ornaments were found at household area that were identified by a series of post holes across the space.
-Here I focus on the ornaments found at post-hole area considering the overall quantity and the comparison across the space.
-Ornaments were made from a wide variety of raw materials, in which glass, agate and copper are the common materials that were thought to be introduced through trading with the Europeans and Chinese traders since no clear evidence of waste by-products in the archaeological record [@Chen2007].
-Although a possibility of small workshops that individuals were able to produce golden beads from prepared glass materials by heating [@Cheng2008], no evidence of accompanying tools or materials in proportion to the large amount of golden beads were found.
-The shape of ornaments include beads, bracelets, rings, bells, pendants, and knitted objects.
-It shows that the shape is closely related to what kind of raw material it was made from.
-For example, beads were only made of glass and agate, while copper ornaments were in the form such as bells, bracelets, and rings [@Chen2007].
-
-# Samples and Methods
-
-```{r plot-all-ornaments-count, fig.cap= "Frequency of ornaments", out.width='60%'}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan from sampling area", out.width='60%'}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 
+# making table
 knitr::kable(ornaments_shape,
              caption = "A table of ornament subtypes")
+
+# some basic stat
+sam_post_hole <- nrow(ornaments_period)
+total_post_hole <- nrow(kwl_upper)
+sample_per <- round(sam_post_hole/total_post_hole*100, 1)
+
+ornament_context <-
+  data.frame("contexts" = c("post hole area", "burials", "middens"),
+           "sampling" = c(406, 3173, 27),
+           "total_KWL" = c(873, 6769, 55))
+
+ornament_context %>% 
+  mutate(percent_context = round(sampling/sum(ornament_context$sampling)*100, 1),
+         sam_total_percent = round(sampling/total_KWL*100, 1)) %>% 
+  add_row(contexts = "sum", 
+          sampling = sum(ornament_context$sampling), 
+          total_KWL = sum(ornament_context$total_KWL))
 ```
@@
 
-Ornaments from the post-hole area accounts for the \majority of the amount in total.
-Post-hole area provides a clear contexts and basis for comparison to see if there is any spatial pattern of distribution.
-For the post-hole area, there were `r sample_size` ornaments recovered from the 40 sampling squares, which accounts \for `r sample_pro` % of the total number of ornaments found in the same open area section.
+Ornaments from 40 sampling squares were excavated in-situ in different archaeological contexts, including 406 pieces from post-hole area, 3173 pieces from burials, and 27 pieces from middens.
+Burials unearthed a large amount of ornaments because most beads were preserved well in the form of bead strings that consist of hundreds to thousand pieces.
+In this paper I focus on the post hole area where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
+Post hole area represents Kiwulan domestic space that provides clear contexts and basis for comparison to see if there is any spatial pattern of distribution.
 Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
-The variety of subtype and their individual frequency are summarized in \@ref(tab:tidy-data-for-table) in terms of three time episodes.
+The variety of subtype and their individual frequency are summarized in Table \@ref(tab:tidy-data-for-table) in terms of three time episodes.
 The image for each subtype in each major category is shown in Figure \@ref(fig:ornaments-image).
 
-<---include the number of ornaments from burials above--->
+Ornaments were found in 30 burials from the sampling area as grave goods.
+Their positions in burials in association with particular parts of human remains indicate the purpose of body adornment.
+For example, agate beads, golden beads, and small metal rings were often found near skulls that indicate the use as necklaces or earrings, while big metal rings were found wearing on radius and ulna bones as bangles.
+This pattern corresponds to the texts regarding the body decorations of indigenous culture from Chinese historical documents [@Chen1963; @Ke1993].
+Agate beads were mostly found in female burials with an average of 4 to 5 pieces, while golden beads were found in both genders with an average of 2 to 3 pieces.
 
-<---Something sounds wrong here. You say ‘majority’ but then later in the paragraph you say it’s 0.5% of the total number? How can 0.5% be the majority? --->
-
-```{r, plot-counts-by-period, fig.cap = "Frequency of each category of ornament with a separate plot showing different time periods individually"}
+```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods."}
 ### delete one plot
@@
 # histogram facetting for three periods
-orna_periods_facet <-
 ggplot(ornaments_count) +
@@
   coord_flip() +
-  labs(y = "Frequency") +
-  theme_minimal(base_size = 12) +
-  theme(axis.text.y = element_blank(),
-        axis.title.y = element_blank(),
-        axis.title.x = element_blank())
-
-# stack histogram
-orna_all_viridis <-
-ggplot(ornaments_count) +
-  geom_col(aes(Categories, n, fill = period)) +
-  labs(y = "Frequency", x = "Categories") +
-  coord_flip() +
-  theme_minimal(base_size = 12) +
-  scale_fill_viridis_d()
-
-library(cowplot)
-library(viridis)
-# add two plots together
-ggdraw() +
-  draw_plot(orna_all_viridis + 
-            theme(legend.justification = "top"), 0, 0, 1, 1) +
-  draw_plot(orna_periods_facet + scale_color_viridis(discrete = TRUE) + 
-              theme(legend.justification = "bottom"), 0.5, 0.15, 0.5, 0.5) +
-  draw_plot_label(c("A", "B"), c(0, 0.5), c(1, 0.7), size = 15)
+  labs(x = "categories", y = "Frequency") +
+  theme_minimal(base_size = 12)
 ```
@@
 
-In this section I report the tests of the three hypotheses about indirect impacts of European colonization on the Indigenous community at Kiwulan.
+In this section I report the tests of the three hypotheses about indirect impacts of European colonization on the indigenous community at Kiwulan.
 To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European contact, and a lesser diversity after Chinese contact, I compare frequencies of ornament types between different time periods.
@@
 However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then their frequency decreased in the European contact period and was even much less in the Chinese contact period.
-
-```{r scatter-plot-bead-by-perio, eval= FALSE}
-# plot for intact golden bead by period
-ornaments_intact_gold <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Golden bead')
-
-ggplot(ornaments_intact_gold) +
-  geom_point(aes(Length, Width, col = Perforation_average)) +
-  facet_wrap(~ period) +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Golden beads") +
-  coord_fixed(ratio = 1) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
-
-# plot for intact glass bead by period
-# need to deal with the original data
-
-# plot for intact agate bead by period
-ornaments_intact_agate <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete','near-complete'),
-         Categories == 'Agate bead')
-
-ggplot(ornaments_intact_agate, 
-       aes(Length, Width, col = period))+
-  geom_point() +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Agate beads") +
-  coord_fixed(ratio = 1) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
-```
 
@@
 
-```{r plot-altogether, fig.cap= "Frequency of "}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across contact periods for metal rings, agate beads, bells, and glass beads."}
+library(cowplot)
 plot_grid(ring_shape_plot,
@@
 
-```{r scatter-plot-metal-by-period, eval= FALSE}
-# plot for intact bell by period
-ornaments_intact_bell <-
-  ornaments_period %>% 
-  filter(Condition %in% c('complete', 'near-complete'),
-         Categories == 'Bell',
-         Type != 'unclassified') %>% 
-  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`))) #remove +
-
-ggplot(ornaments_intact_bell) +
-  geom_point(aes(Length, Width, col = Type)) +
-  facet_wrap(~ period) +
-  labs(x = "Length (mm)",
-       y = "Width (mm)",
-       title = "Bell") +
-  coord_fixed(ratio = 1) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust=0.5))
-```
-
 ```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
 
-#I don’t see any reference to this in the text.
-If you don’t reference a figure in the text, you need to delete it from the manuscript, Why not use a colour scale for frequency?
-I think the viridis scale would be perfect for this
+library(sf)
+# read in the spatial data
+AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
+units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
 
+# tidy spatial data for 40 sampling units
+sample_units <-
+  AD_data %>%
+  semi_join(units_40, by = c (`坑號` = "the_sq"))
+
+# join two dataset to get coordinate
+ornaments_period_join <-
+  ornaments_period %>%
+  mutate(period = as.factor(period)) %>% # seems need this line first
+  group_by(Pit_No, period) %>%
+  count(Categories) %>%
+  inner_join(AD_data, by = c("Pit_No" = "坑號"))
+
+# plot spatial pattern for all ornaments
 ggplot() +
@@
 
-```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for each class of ornament by time periods"}
-
-# Seems like some ornament types have <5 pieces. Probably not worth to plot those. You can write in the caption ‘only those types with n > 5 are shown here’
+```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
 
@@
 
-# filter metal ring
-ornaments_period_join_metal <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Metal ring')
-
-# plot by period
-dis_metal <-
-ggplot(ornaments_period_join_metal) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_metal, 
-          aes(geometry = geometry,
-              fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Metal Ring",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-# filter metal ring
-ornaments_period_join_bell <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Bell')
-
-# plot by period
-dis_bell <-
-ggplot(ornaments_period_join_bell) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_bell, 
-          aes(geometry = geometry,
-              fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Copper Bell",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
 plot_grid(dis_golden,
@@
           dis_agate, 
-          dis_metal, 
-          dis_bell,
           align = "v",
-          nrow = 5)
+          nrow = 3)
 ```
@@
 
-Figure \@ref(fig:plot-spatial-pattern-each) presents the spatial distribution of the main ornament types for each time period.
-Before the European contact, the map shows a greater amount of glass beads found at the northern and middle parts of the research area, and they were more widespread during the European contact period.
-During the European contact period, ornaments were more ubiquitous with some clusters on northern part, and less abundance during the Chinese contact period.
-When looking at the distribution for each major class individually, some clusters across the area can be observed during European contact, including golden beads, agate beads, metal rings, and bells.
+Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from research area for each time period.
+Before the European contact, the map shows a greater amount of ornaments found at the northern and middle parts of the research area, and they were more widespread during the European contact period.
+During the European contact period, ornaments were more ubiquitous with some clusters on the northern part, and less abundance during the Chinese contact period.
+Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major classes individually, some clusters across the area can be observed during the European contact, such as golden beads and agate beads.
 However, there seems to be no clear consistent clustered pattern across those different ornaments.
@@
 
-```{r plot-all-densities, fig.cap= "Kernal density map by periods"}
+```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods"}
 par(mfrow = c(1, 3))
@@
 ## Point pattern analysis of ornament distribution
+
 I used point pattern analysis to examine whether the distribution of ornaments was uneven in non-random ways that could reflect the presence of social inequality.
 Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
-).
 In this case hotspots in the pattern of ornaments could indicate the changes in social organization due to culture contact.
 I used the intensity approach that focuses on the average density of points across space to explore their distribution.
-Ornament location points were given random coordinates within each square they they were recovered from because their exact recovery location in each square was not recorded excavation.
-The next step was to subset the orgaments into three groups for three time periods, and then compute the kernel density for each time period for comparison.
+Ornament location points were given random coordinates within each square they were recovered from because their exact recovery location in each square was not recorded during excavation.
+The next step was to subset the ornaments into three groups for three time periods, and then compute the kernel density for each time period for comparison.
+
 Kernel density estimations (KDE) estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
 Here we use KDE to visualize the event of the presence of ornaments that were represented by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
-Density values of artefacts per square meter are assigned to each cell.
+Density values of artifacts per square meter are assigned to each cell.
 The results in Figure \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period.
@@
 
-```{r}
+```{r function for ANN values}
 # function for hypothesis testing
@@
 
-```{r hypo-test-distrubution, fig.cap= "Histogram of simulated ANN values with the blue line indicating the observed value"}
+```{r hypo-test-distribution, fig.cap= "Histogram of simulated ANN values with the blue line indicating the observed value"}
 
@@
 The observed data was compared to the random pattern generated by simulated processes to see if our data is randomly distributed or not.
-Figure \@ref(fig:hypo-test-distrubution) shows the distributions of the ANN distances calculated on 1000 simulations of ornament locations.
+Figure \@ref(fig:hypo-test-distribution) shows the distributions of the ANN distances calculated on 1000 simulations of ornament locations.
 The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments have a non-random clustered distributions.
@@
 This testing reveals that the clustered distributions of ornaments during the pre-European and European contact periods might reflect the presence of different social groups.
-Moreover, the clustering of ornaments during the European contact period is highly non-random, potentially indicate some social changes leasing to a concentration of power to control the distribution of ornaments at Kiwulan.
+Moreover, the clustering of ornaments during the European contact period is highly non-random, potentially indicate some social changes leading to a concentration of power to control the distribution of ornaments at Kiwulan.
 
 # Discussion
-<--I think you have space for a short paragraph here on  network/corporate model at the end>
 
@@
 This might result from the large scale exchange network that stimulated the circulation of different ornament classes.
-Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads, which might indicate their diverse origins.
+Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bells and glass beads, which might indicate their diverse origins.
 However, the frequency of overall ornaments and each subtype declines significantly during the Chinese contact period.
-The reason could be related to smaller scale of network and the \overall \decline of Kiwulan population at that time.
-
-<----Why did it decline? What is the evidence for this? A few more sentences are needed on this-->
+The reason could be related to a smaller scale of network, the overall decline of the indigenous population in Yilan, and adopting the mortuary practices of Han Chinese.
+Since the end of the 18th century, lots of indigenous people moved southwards to Hualien due to population pressure caused by the great amount of Han Chinese immigrants [@Chen2007].
+According to the Chinese census reports in 1821 and Japanese field survey in 1895, the Kiwulan population was less than 100 that dropped by nearly 90 percent compared with the last census in the 17th century [@Yao1996; @Ino1898].
+In addition, burials after 18th century unearthed fewer ornaments or even lack of them that might indicate a change in mortuary practices.
+This can be also identified by the presence of coffin-like case that was viewed as a funerary item of Han Chinese, however, with local indigenous carving patterns on the wooden surfaces [@Chen2007; @Chiu2004].
 
@@
 Hypothesis testing for spatial patterns further indicates that these clusters are non-random, and are concentrated during European contact.
-Because trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering could indicate that a degree of social inequality were present before European contact and then it was reinforced and amplified during the European contact period.
+Because trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering could indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European contact period.
+In addition, there was a burial dated to 17th century unearthed 60 golden beads above the average of 2-3 pieces that hints the possibility of increasing social inequality [@Chen2007; @Cheng2008].
+In Taiwan, most prehistoric organization were thought as egalitarian societies because of the lack of centralization of power that is often used as a major indicator of hierarchical complexity.
+However, this concept of egalitarianism has been challenged by evidence of social inequalities such as uneven distributions of prestige goods or the emergence of influential individuals in some non- hierarchical human societies [@Ames2010; @Bowles2010; @Woodburn1982].
+
+Horizontal complexity is another approach to discuss the emergence of social inequity by focusing on how people achieve and maintain power in a wide variety of societies [@Drennan2010; @Feinman2000].
+For example, the corporate/network model proposed by @Feinman2000 that expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.
+In the network mode, inequality develops when individuals accumulate wealth through individual networks and use wealth to attract factions, control resources, and monopolize trade networks.
+In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labor tasks [@Feinman2000; @Siegel1999].
+Applying the corporate/network model may give insight into the underlying pathways that led to social inequality in northeastern Taiwan.
+In the circumscribed environments of the Yilan Plain, the corporate/network continuum will provide a comparative basis to explore the impact of European contact on changes in the degree of indigenous social inequality.
+One insight emerging from this model is that the long-distance trade network introduced by the Europeans results in the appearance of a network mode due to the competitions among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
+Because of the weak direct control from the European colonizers in northeastern Taiwan, it implies that local leaders might have flexibility to manipulate European colonial image, expand personal power, and monopolize the high-value trade goods [@Kang2012].
 
@@
 Those ornaments may carry an exotic and powerful image that signals wealth and trading connections to the inhabitants of Kiwulan.
-These symbolic values may have stimulated more competitions between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might result in a increase in social inequality.
+These symbolic values may have stimulated more competitions between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might result in an increase in social inequality.
 
@@
 During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwulan population.
-Future work could extent this approach to studies of other trade goods such as ceramics.
-Future directions for understanding the dynamics of social inequality at Kiwulan could include analysis of pottery production and standisation, and mortuary practices.
+Future work could extend this approach to studies of other trade goods such as ceramics.
+Future directions for understanding the dynamics of social inequality at Kiwulan could include analysis of pottery production and standardization, and mortuary practices.
 

### 2019-08-09 analysis/paper/KWL-excavation-map .R

@@ -1,0 +1,36 @@
+library(sf)
+library(cartography)
+# read in the spatial data
+burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
+AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
+AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
+location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
+units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
+
+# tidy spatial data
+# 40 sampling units
+sample_units <-
+  AD_data %>%
+  semi_join(units_40, by = c (`坑號` = "the_sq"))
+
+# join two dataset to get coordinate
+ornaments_period_join <-
+  ornaments_period %>%
+  mutate(period = as.factor(period)) %>% # seems need this line first
+  group_by(Pit_No, period) %>%
+  count(Categories) %>%
+  inner_join(AD_data, by = c("Pit_No" = "坑號"))
+
+# mapping sampling area
+ggplot(sample_units) +
+  geom_sf(data = AD_data, fill = NA) +
+  geom_sf(color = 'red') +
+  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
+  theme_minimal() +
+  theme_void()
+
+#1. Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
+#2. Need to add in here that each square is 4 x 4 m
+#3. and say which way north is. Or add to the map and scale and north arrow.
+
+ggsave("analysis/figures/KWL-excavation-map.png")

### 2019-08-09 analysis/paper/burial.R

@@ -1,0 +1,39 @@
+# read data
+kwl_burial <-
+  readxl::read_excel(here("analysis", "data", "raw_data", "KWL_burial_ornament.xls"))
+
+# tidy up for all ornaments from burials
+kwl_burial_all_tidy <-
+  kwl_burial %>%
+  group_by(`編號/空間號`, `坑號`, `類別`) %>%
+  count() %>%
+  mutate(count = str_extract(`類別`, "[0-9]+")) %>%
+  mutate(`類別` = str_remove(`類別`, "[0-9]+")) %>%
+  mutate(count = as.numeric(count)) %>%
+  mutate(count = ifelse(is.na(count), n, count))
+
+# filter the ornaments from burials in sampling area
+kwl_burial_sampling <-
+  kwl_burial %>%
+  filter(!is.na(`坑號`)) %>%
+  group_by(`編號/空間號`, `坑號`, `類別`) %>%
+  count() %>%
+  mutate(count = str_extract(`類別`, "[0-9]+")) %>%
+  mutate(`類別` = str_remove(`類別`, "[0-9]+")) %>%
+  mutate(count = as.numeric(count)) %>%
+  mutate(count = ifelse(is.na(count), n, count))
+
+# summarize for each class, filter out the extreme value for golden bead
+KWL_all_burial <-
+  kwl_burial_all_tidy %>%
+  filter(!n == 60) %>%
+  group_by(`類別`) %>%
+  summarise(mean = mean(count))
+
+# for sampling area
+KWL_sam_burial<-
+  kwl_burial_sampling %>%
+  filter(!n == 60) %>%
+  group_by(`類別`) %>%
+  summarise(mean = mean(count))
+

### 2019-08-09 analysis/paper/map.R

@@ -88,3 +88,5 @@ ymin = 24.7,
 ymax = 25.2)
 
 ggsave("analysis/figures/kiwulan-location-map.png")
+
+# add the location of Keelung, Tamsui, and Holping dau

### 2019-08-09 analysis/paper/notes.R

@@ -34,3 +34,144 @@ anim_gold_bead_spatial <-
         axis.text.y = element_blank(),
         strip.text.x = element_text(size = 11))
 
+# categories of ornaments by period
+ornaments_count <-
+  ornaments_period %>%
+  filter(!is.na(`6-layers`),
+         Categories != "Unknown Metal") %>%
+  ungroup() %>%
+  mutate(Categories = fct_lump(Categories,
+                               n = 5)) %>%
+  mutate(Categories = fct_infreq(Categories)) %>%
+  mutate(Categories = fct_rev(Categories)) %>%
+  group_by(period, Categories) %>%
+  count()
+
+# histogram facetting for three periods
+orna_periods_facet <-
+  ggplot(ornaments_count) +
+  geom_col(aes(Categories, n)) +
+  facet_wrap(~ period) +
+  coord_flip() +
+  labs(y = "Frequency") +
+  theme_minimal(base_size = 12) +
+  theme(axis.text.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.title.x = element_blank())
+
+# stack histogram
+orna_all_viridis <-
+  ggplot(ornaments_count) +
+  geom_col(aes(Categories, n, fill = period)) +
+  labs(y = "Frequency", x = "Categories") +
+  coord_flip() +
+  theme_minimal(base_size = 12) +
+  scale_fill_viridis_d()
+
+library(cowplot)
+library(viridis)
+# add two plots together
+ggdraw() +
+  draw_plot(orna_all_viridis +
+              theme(legend.justification = "top"), 0, 0, 1, 1) +
+  draw_plot(orna_periods_facet + scale_color_viridis(discrete = TRUE) +
+              theme(legend.justification = "bottom"), 0.5, 0.15, 0.5, 0.5) +
+  draw_plot_label(c("A", "B"), c(0, 0.5), c(1, 0.7), size = 15)
+
+
+# plot for intact golden bead by period
+ornaments_intact_gold <-
+  ornaments_period %>%
+  filter(Condition %in% c('complete', 'near-complete'),
+         Categories == 'Golden bead')
+
+ggplot(ornaments_intact_gold) +
+  geom_point(aes(Length, Width, col = Perforation_average)) +
+  facet_wrap(~ period) +
+  labs(x = "Length (mm)",
+       y = "Width (mm)",
+       title = "Golden beads") +
+  coord_fixed(ratio = 1) +
+  theme_minimal() +
[…]

### 2019-08-09 analysis/paper/paper.Rmd
added chi-square and Spearman's correlation test for variety of ornaments

@@
 However, compared to the European contact period, there are fewer mentions of beads in Chinese contact period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
-
 # Excavations at Kiwulan in northeastern Taiwan
@@
 
-Kiwulan \@ref(fig:KWL-map) is located at northern Yilan and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
+Kiwulan (Figure \@ref(fig:KWL-map)) is located at northern Yilan and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
 The total area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
@@
 Most ornaments were found in situ while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
-The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of the excavation, see \@ref(fig:KWL-sam-area).
+The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of the excavation, see Figure \@ref(fig:KWL-sam-area).
 They were sampled because those units were stratigraphically intact with the least disturbance by modern construction activity, compared to excavation squares on the periphery of the site.
@@
 sample_per <- round(sam_post_hole/total_post_hole*100, 1)
-
-ornament_context <-
-  data.frame("contexts" = c("post hole area", "burials", "middens"),
-           "sampling" = c(406, 3173, 27),
-           "total_KWL" = c(873, 6769, 55))
-
-ornament_context %>% 
-  mutate(percent_context = round(sampling/sum(ornament_context$sampling)*100, 1),
-         sam_total_percent = round(sampling/total_KWL*100, 1)) %>% 
-  add_row(contexts = "sum", 
-          sampling = sum(ornament_context$sampling), 
-          total_KWL = sum(ornament_context$total_KWL))
 ```
 
-```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major category.
+```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
 A: agate beads, B: bells, C: glass beads and golden beads, D: metal rings.
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D categories were from original excavation report (Chen 2007).", out.width= "400px", echo=FALSE}
+The photographs of B, C, D classes were from original excavation report (Chen 2007).", out.width= "400px", echo=FALSE}
 knitr::include_graphics(here("analysis", "figures", "ornament.png"))
@@
 Ornaments from 40 sampling squares were excavated in-situ in different archaeological contexts, including 406 pieces from post-hole area, 3173 pieces from burials, and 27 pieces from middens.
-Burials unearthed a large amount of ornaments because most beads were preserved well in the form of bead strings that consist of hundreds to thousand pieces.
+Burials unearthed a large amount of ornaments because mostly beads were preserved well in the form of bead strings that consisted of hundreds to thousand pieces.
 In this paper I focus on the post hole area where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
@@
 The variety of subtype and their individual frequency are summarized in Table \@ref(tab:tidy-data-for-table) in terms of three time episodes.
-The image for each subtype in each major category is shown in Figure \@ref(fig:ornaments-image).
+The image for each subtype in each major class is shown in Figure \@ref(fig:ornaments-image).
 
@@
 
+```{r chi square testing}
+# prepare contingency table for chi sqaure
+ornaments_shape_chi <-
+  ornaments_period_clean %>% 
+  group_by(period, Categories) %>% 
+  tally() %>% 
+  filter(!Categories %in% c("Unknown Metal", 
+                            "Perforated plate", 
+                            "Knited item",
+                            "Hair pin",
+                            "Bone Ornament",
+                            "Metal Piece",
+                            "Glass ring")) %>% 
+  spread(Categories, n)
+
+# compute chi sqaure
+chisq_test_result <- chisq.test(ornaments_shape_chi[ , -1])
+chisq_test_result_stat <- round(chisq_test_result$statistic, 2)
+chisq_test_result_df <- round(chisq_test_result$parameter, 2)
+chisq_test_result_p_value <- (chisq_test_result$p.value)
+```
+
 # Results of hypothesis testing
@@
 ## Frequency of ornaments between time periods
-Figure \@ref(fig:plot-counts-by-period) shows the comparison of ornament type frequencies between different time periods at Kiwulan.
-It indicates a general pattern that most ornament types were present before European contact.
-The frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
+
+Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan.
+The result of chi-square test for the number of ornaments (X-squared = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = `r chisq_test_result_p_value`) shows that there is a significant difference between time period groups.
+It indicates a general pattern that most ornament types were present before European contact.The frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
 This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
@@
 
-The distribution of frequency for subtypes in each major category are presented in Figure \@ref(fig:plot-altogether).
-The distribution shows that agate beads and metal rings have greater quantity and \variety of shapes compared to copper bell and glass beads during the European contact period.
+```{r regression}
+# tidy up the data for correlation testing
+ornaments_diversity_test <-
+  ornaments_shape_count %>%
+  group_by(period) %>% 
+  mutate(sam_size = sum(n)) %>% 
+  add_count() %>% 
+  select(period, n, sam_size) %>% 
+  distinct()
+
+ornaments_diversity_test_2 <-
+  ornaments_shape_count %>%
+  unite(diversity, c(period, Categories), sep = "-", remove = FALSE) %>% 
+  group_by(diversity) %>% 
+  mutate(sam_size = sum(n)) %>% 
+  add_count() %>% 
+  select(diversity, n, sam_size) %>% 
+  separate(diversity, c("period","subtype"), sep = "-", remove = FALSE) %>% 
+  distinct()
+
+# spearman test
+spearman_result <- 
+cor.test( ~ n + sam_size, 
+         data = ornaments_diversity_test_2,
+         method = "spearman",
+         continuity = FALSE,
+         conf.level = 0.95)
+
+spearman_result_stat <- round(spearman_result$statistic, 2)
+spearman_result_rho <- unname(round(spearman_result$estimate, 2))
+spearman_result_p_value <- round(spearman_result$p.value, 2)
+```
+
+The distribution of frequency for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether).
+Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `spearman_result_rho`, p-value = `r spearman_result_p_value`).
+This indicates that the increases in diversity can be explained by the effects of culture contact instead of natural effects of sample size.
+According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period.
 Agate beads can be divided into seven subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased significantly during the European contact and then declined in the Chinese contact.
@@
 Similarly, metal rings can be classified into eight subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape.
-The greater varieties for those two categories might indicate multiple origins due to global trade networks stimulated by the Europeans.
+The greater varieties for those two classes might indicate multiple origins due to global trade networks stimulated by the Europeans.
 In contrast, copper bell and glass bead have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces.
 The common shape for copper bells is the large shape with a wide variety of human faces as a motif, while the common shape for glass beads is a bead with length less than 1 cm that are often identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].
-
-<---I think we need an addition bit of analysis in this section to show the relationship between diversity and sample size.
-We naturally expect diversity to increase as sample size increases.
-So we want to test and see if the changes in diversity that you describe here are only the natural effects of increase/decrease in sample size, or there are ‘real’ increases in diversity that cannot be explained by sample size changes.
-The peer reviewers will want to know the answers to this.
---->
 

### 2019-08-09 analysis/paper/map.R

@@ -19,6 +19,9 @@ TW_map <- fortify(TW_adm)
 
 # add site location
 Kiwulan <- data.frame(lon = c(121.7809), lat = c(24.8066))
+Keelung <- data.frame(lon = c(121.7621), lat = c(25.1573))
+Tamsui <- data.frame(lon = c(121.4349), lat = c(25.1764))
+Hualien <- data.frame(lon = c(121.6084), lat = c(24.0228))
 
 TW_SE_Asia <-
 ggplot(data = world) +

### 2019-08-09 analysis/paper/notes.R

@@ -175,3 +175,34 @@ dis_bell <-
         axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         strip.text.x = element_text(size = 11))
+
+# table for a quick look at number from excavation reports and raw data
+ornament_context <-
+  data.frame("contexts" = c("post hole area", "burials", "middens"),
+             "sampling" = c(406, 3173, 27),
+             "total_KWL" = c(873, 6769, 55))
+
+ornament_context %>%
+  mutate(percent_context = round(sampling/sum(ornament_context$sampling)*100, 1),
+         sam_total_percent = round(sampling/total_KWL*100, 1)) %>%
+  add_row(contexts = "sum",
+          sampling = sum(ornament_context$sampling),
+          total_KWL = sum(ornament_context$total_KWL))
+
+# spearman testing and plot
+ornaments_diversity_test_2 <-
+  ornaments_shape_count %>%
+  unite(diversity, c(period, Categories), sep = "-", remove = FALSE) %>%
+  group_by(diversity) %>%
+  mutate(sam_size = sum(n)) %>%
+  add_count() %>%
+  select(diversity, n, sam_size) %>%
+  separate(diversity, c("period","subtype"), sep = "-", remove = FALSE) %>%
+  distinct()
+
+# plot
+ggplot(ornaments_diversity_test_2,
+       aes(sam_size, n)) +
+  geom_point(aes(color = period)) +
+  labs(x = 'sample size', y = 'subtypes') +
+  scale_x_log10()

### 2019-08-10 analysis/paper/map.R
polished the Taiwan map

@@ -18,17 +18,17 @@ TW_adm <- st_read(here("analysis", "data", "raw_data","TWN_adm2.shp"))
 TW_map <- fortify(TW_adm)
 
 # add site location
-Kiwulan <- data.frame(lon = c(121.7809), lat = c(24.8066))
-Keelung <- data.frame(lon = c(121.7621), lat = c(25.1573))
-Tamsui <- data.frame(lon = c(121.4349), lat = c(25.1764))
-Hualien <- data.frame(lon = c(121.6084), lat = c(24.0228))
+site_location <-
+  data.frame(location = c("Kiwulan", "Keeling", "Tamsui", "Hualien"),
+             lon = c(121.7809, 121.7621, 121.4349, 121.6084),
+             lat = c(24.8066, 25.1523, 25.1764, 24.0228))
 
 TW_SE_Asia <-
 ggplot(data = world) +
   geom_sf() +
-  geom_rect(xmin = 121, xmax = 122.0, ymin = 24.3, ymax = 25.2,
+  geom_rect(xmin = 120.9, xmax = 122.2, ymin = 24.4, ymax = 25.5,
             fill = NA, colour = "red", size = 0.5) +
-  coord_sf(xlim = c(107, 135), ylim = c(13, 35), expand = FALSE, datum = NA) +
+  coord_sf(xlim = c(107, 135), ylim = c(5, 45), expand = FALSE, datum = NA) +
   theme(panel.background = element_rect(fill = "azure"))
 
 # Topographic map
@@ -37,34 +37,36 @@ library(tmaptools)
 library(shadowtext)
 # we don't want to download every time, so let's save the map locally
 # from https://stackoverflow.com/a/52710855/1036500
-tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(121.0, 24.3,
-                                                 122.0, 25.2))), zoom = 10))
+tw_map <- ggmap(get_stamenmap(rbind(as.numeric(c(120.7, 24.2,
+                                                 122.1, 25.6))), zoom = 10))
 #saveRDS(tw_map, here("analysis", "data", "raw_data", "tw_map.rds"))
 #tw_map <- readRDS(here("analysis", "data", "raw_data", "tw_map.rds"))
 pg <- ggplot_build(tw_map)
 
 TW_map_with_site <-
 tw_map +
-  geom_point(data = Kiwulan,
+  geom_point(data = site_location,
              aes(x = lon,
                  y = lat),
-             size = 3,
+             size = 2,
              color = "red") +
[…]

### 2019-08-11 analysis/paper/paper.Rmd

@@
 
-```{r chi square testing}
+```{r chi-square-testing}
 # prepare contingency table for chi sqaure

### 2019-08-13 analysis/paper/paper.Rmd
2nd edit

@@
   - Li-Ying Wang
-  - Ben Marwick
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 The direct impacts of European colonialism on indigenous communities in East Asia were much less pronounced than island Southeast Asia and Oceania.
-Throughout East Asia direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
+Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
 Understanding the indirect effects of colonialism are important for detecting colonial impacts on indigenous peoples in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
-In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused significant transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
-The agency of indigenous peoples in the colonial contexts has been broadly discussed to understand the negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
-Studying the consumption of foreign goods is one way to understand indigenous experiences of colonial encounters and explore the agency of indigenous peoples by examining how goods were distributed and used [@Dietler2005; @Mullins2011; @Scaramelli2005].
+In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
+Studying the consumption of foreign goods can improve our understanding of the negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2005; @Dietler2015; @Given2004; @Mullins2011; @Rubertone2000; @Scaramelli2005; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
+
+Prior to contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the tenth century, and encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
+11].
+The exchange network in northern Taiwan was influenced by long-distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since the 15th century due to its location on shipping routes [@Chen2005].
+This intensified in the 17th century with the arrival of the Europeans and their trade goods [@Wang2007].
 
@@
 Despite the colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonization.
-@Berrocal2018 study both archaeological and historical records of Heping Dao in northern Taiwan that was the location of the colony of San Salvador founded by the Spanish in 1626 and took over by the Dutch in 1642.
-They uncovered parts of the foundation of a Spanish church and a European cemetery that indicate the evidence of European occupation.
-It shows a direct impact on the local indigenous populations by disturbing their long settlement pattern and changing local landscape.
-Different from the view of a Spanish fort, Yilan in Northeastern Taiwan is an especially unique location for exploring indirect colonial effects on local indigenous settlements due to its physically isolated location.
+@Berrocal2018 examined both archaeological and historical records of Heping Dao in northern Taiwan.
+This was the location of the colony of San Salvador founded by the Spanish in 1626 and taken over by the Dutch in 1642.
+Berrocal et al.
+uncovered parts of the foundation of a Spanish church and a European cemetery that indicate the evidence of European occupation.
+These remains indicate a direct impact on the local indigenous populations through their modification of the local landscape.
+In contrast to Heping Dao, Yilan in Northeastern Taiwan is physically isolated and there are no traces of European settlement.
+This makes Yilan an especially unique location for exploring indirect colonial effects on local indigenous settlements because of isolation.
 The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-The sea provides a way through which indigenous people in Yilan could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans.
-Large-scale and frequent trade activities brought a wide variety of trade objects into indigenous communities in Yilan through local regional exchange network.
-One of the most common types of trade goods in this region were personal ornaments such as glass and stone beads was one of the common foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
-Personal adornments are a social signal of an individual’s status [@Joyce2005; @Scaramelli2005].
-For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
-
-In this paper, I explore the evidence from Yilan spanning from late Iron Age to historical period to address the question of whether indirect colonial influences on indigenous populations can be detected.
-I focus on the ornaments recovered from the archaeological excavations at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
-Prior to the contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since tenth century, and encountered large waves of Han Chinese immigration after 17th century since Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
-11].
-Northeastern Taiwan provides a good example to explore the indirect colonial impacts on indigenous societies by comparing archaeological evidence from different episodes of culture contact [cf.@Lape2003].
-I will describe the variety and cultural context of trade ornaments in this region, and examine their potential of the ornaments to inform on social changes relating to trade with colonial groups.
-I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to participation in a larger scale exchange network.
-I also predict that after European contact there will be new spatial patterns in the distributions of ornaments at Kiwulan that indicates increasing social inequality.
-I hypothesize that after Chinese contact there was a decline in the use and discard of ornaments at Kiwulan due to overall population declines and acculturation in the Chinese community.
-
-# Cultural context of ornaments in Yilan, northeastern Taiwan
-
-Northeastern Taiwan was part of the local exchange network in northern Taiwan, where the indigenous people had already established their own exchange network before European arrived that has been described as “inter-insular trade” by Chen @Chen2005 [pp.
-12] to refer to small-scale regional trade between China and Taiwan on an irregular basis.
-The exchange network in northern Taiwan was increasingly influenced by long-distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since the 15th century due to its location on the shipping routes [@Chen2005].
-This was intensified in the 17th century with the arrival of the Europeans and their trade goods [@Wang2007].
-Due to its location on the periphery of major trade ports, Keelung and Tamsui, in northern Taiwan, Yilan was involved in this network through the periodic exchange activities between different indigenous groups.
 Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
@@
 
+\Although mountains isolated Yilan by land, the sea provided a way through which indigenous people in Yilan could interact with groups elsewhere, such as indigenous groups, distant traders such as the Han people of China, and Europeans.
+Large-scale and frequent trading activities brought a wide variety of objects into indigenous communities in Yilan.
+\The indigenous people of Yilan established their own exchange network that has been described as “inter-insular trade” by @Chen2005 [p.
+12] to refer to small-scale regional trade between China and Taiwan on an irregular basis.
+One of the most commonly traded types of object in this region were personal ornaments such as glass and stone beads.
+These are one of the most abundant types of foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
+Personal adornments in the archaeological record are useful as social signal of an individual's status [@Joyce2005; @Scaramelli2005].
+For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
+
+In this paper, I explore archaeological evidence from Yilan spanning from the late Iron Age to the historical period to address the question of whether indirect colonial influences on the indigenous populations can be detected.
+I focus on the ornaments recovered from the archaeological excavations at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
+Northeastern Taiwan provides a good example to explore the indirect colonial impacts on indigenous societies by comparing archaeological evidence from different episodes of culture contact [cf.
+@Lape2003].
+I will describe the variety and cultural context of trade ornaments in this region, and examine the potential of ornaments to inform on social changes relating to trade with colonial groups.
+I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to participation in a larger scale exchange network.
+I also predict that after European contact there will be new patterns in the spatial distributions of ornaments at Kiwulan that indicate increasing social inequality.
+I hypothesize that after 1662 when European contact declined and Chinese contact became the dominant international tradition partner, there was a decline in the use and discard of ornaments at Kiwulan due to overall population declines and acculturation in the Chinese community.
+To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European contact, and a lesser diversity after Chinese contact, I compare frequencies of ornament types between different time periods.
+To test the prediction that after European contact were new spatial patterns in the distributions of ornaments at Kiwulan, I use kernel density and point pattern analyses.
+
+# Cultural context of ornaments in Yilan, northeastern Taiwan
+
 Historical documents from the Spanish and Dutch in the early 17th century provide some information about the social life and culture of indigenous people in Northern Taiwan when the Europeans arrived.
-The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for a religious purpose [@Blusse2000, pp.
+The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
 343].
-The report of the Dominican priest Fr.
+A report by Dominican priest Fr.
 Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) in exchange for necessities with other indigenous groups.
@@
 151].
-Other records mention that the female shamans (majuorbol) in the tribe would use agate beads as magical items in ritual practice for body healing[@Borao2009, pp.
+Other records mention that the female shamans (majuorbol) in the tribe would use agate beads as magical items in ritual practice for body healing [@Borao2009, pp.
 122-151].
-An indigenous funeral process was recorded, documenting the use of agate beads in ritual contexts, with more agate beads, pottery, and cloths placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
+Records of an indigenous funeral process document the use of agate beads in ritual contexts, with more agate beads, pottery, and cloth placed into the graves of more influential people to indicate a family’s higher prestige  [@LiandWu2006, pp.
 153].
-Those historical accounts indicate that agate beads or golden beads were already treated as prestige goods before the arrival of the Europeans.
+These historical accounts indicate that agate beads or golden beads were already treated as prestige goods before the arrival of the Europeans.
 These were likely introduced earlier from Chinese traders.
@@
 
-Chinese historical records from the year of 1829, 1837, and 1852 during the Qing dynasty [1616-1911] contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
+Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty [1616-1911] contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
 228, 308; @Ke1993, pp.
@@
 77].
-According to those records, indigenous people in Yilan had the custom to wear ornaments in ceremonial contexts to display their wealth and status.
-Among those ornaments, golden fish-shaped necklaces made of copper threads had higher value due to its delicacy and more materials invested in production.
+According to those records, indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
+Among those ornaments, golden fish-shaped necklaces made of copper threads had high value due to its delicacy and the materials invested in production.
 These were usually possessed by wealthy people.
@@
 
-Recent ethnographic research with the living members of the Kavalan tribe describes how agate beads were used in divination practices, called “subli”, by female shamans [@Liu2008, pp.
+Recent ethnographic research with the living members of the Kavalan tribe describes how agate beads were used in divination practices, called "subli", by female shamans [@Liu2008, pp.
 133-134].
@@
 Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination, consistent with observations from the European colonial period.
-It was usually the female shamans conducting divination rituals that passed the beads down from their ancestors.
-Despite most Kavalan people today not knowing the specific origin of their agate beads, the value of beads is still highlighted through oral history and by their scarcity.
+Despite most Kavalan people today not knowing the specific origin of their agate beads, the high value of beads is still indicated in oral history and by their scarcity.
 In sum, we find multiple sources describing how local Indigenous people used ornaments in cultural contexts that represents some social roles or high status.
 However, compared to the European contact period, there are fewer mentions of beads in Chinese contact period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
+
 # Excavations at Kiwulan in northeastern Taiwan
@@
 ```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px", echo=FALSE}
-# need to add the locations of Keelung, Tamsui, and hoping dao
+# need to adjust Keelung, heping dao, check it out after knitting
 library(here)
@@
 Each square is 4 x 4 m", out.width= "70%", echo=FALSE}
-
-# add to the map and scale and north arrow. 
+# add to the map and scale and north arrow. Check it out after knitting 
 knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.png"))
@@
 Kiwulan (Figure \@ref(fig:KWL-map)) is located at northern Yilan and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
-The total area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
+The total area includes eight open area sections of 262 excavation squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
 The archaeological evidence includes a large amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
@@
 
-Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with a hundred year intervals according to original radiocarbon dates, excavation depth, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
+Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
 The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
@@
 12].
-I take approximate indicators of the start of European contact at Kiwulan as the appearance of An-ping jars and stonewares (martavans/martaban) that were largely introduced to Taiwan during the early 17th century.
-They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages, and the jar shapes found at Kiwulan are typical of those found elsewhere in V.O.C sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
+In the archaeological record, the approximate indicators of the start of European contact at Kiwulan are the appearance of An-ping jars and stonewares (martavans/martaban) that were largely introduced to Taiwan during the early 17th century.
+They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
+The jar shapes found at Kiwulan are typical of those found elsewhere in V.O.C sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
 I used the presence and absence of An-ping jars and stonewares, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.
+
 European colonization ended in 1662 when they were defeated by the kingdom of Tungning founded by Koxinga from China.
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-I recognize this as the Chinese contact period, which is archaeologically visible at Kiwulan by a large amount and varieties of Chinese porcelains, clay stoves, and a few opium pipe-bowls that were identified as items commonly used in the daily life of Chinese and the use of bricks and tiles for architectural buildings [@Hsieh2009].
-Also, the description of the Chinese migration can be seen from the Chinese official records written in the early 19th century [@Chen1963; @Ke1993].
+The archaeological signature of the Chinese contact period at Kiwulan is the large amount and varieties of Chinese porcelains, clay stoves, and a few opium pipe-bowls that were identified as items commonly used in the daily life of Chinese and the use of bricks and tiles for architectural buildings [@Hsieh2009].
+Also, the description of the Chinese migration to Yilan can be seen from official Chinese records written in the early 19th century [@Chen1963; @Ke1993].
 
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan from sampling area", out.width='60%'}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan", out.width='60%'}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 knitr::kable(ornaments_shape,
-             caption = "A table of ornament subtypes")
+             caption = "Ornament subtype frequencies at Kiwulan")
 
@@
 
-Ornaments from 40 sampling squares were excavated in-situ in different archaeological contexts, including 406 pieces from post-hole area, 3173 pieces from burials, and 27 pieces from middens.
-Burials unearthed a large amount of ornaments because mostly beads were preserved well in the form of bead strings that consisted of hundreds to thousand pieces.
-In this paper I focus on the post hole area where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
-Post hole area represents Kiwulan domestic space that provides clear contexts and basis for comparison to see if there is any spatial pattern of distribution.
-Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells.
+Ornaments from 40 sampling squares were excavated in-situ in different archaeological contexts, including 406 pieces from post-hole areas, 3173 pieces from 30 burials, and 27 pieces from middens.
+Burials unearthed a large amount of ornaments in the form of bead strings that consisted of hundreds to thousands of pieces.
+In this paper I focus on the post hole areas where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
+Post hole areas represents domestic spaces suitable for testing for changes in the patterns of distribution relating to social organization.
+Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells (see also Figure 4).
 The variety of subtype and their individual frequency are summarized in Table \@ref(tab:tidy-data-for-table) in terms of three time episodes.
-The image for each subtype in each major class is shown in Figure \@ref(fig:ornaments-image).
-
-Ornaments were found in 30 burials from the sampling area as grave goods.
-Their positions in burials in association with particular parts of human remains indicate the purpose of body adornment.
-For example, agate beads, golden beads, and small metal rings were often found near skulls that indicate the use as necklaces or earrings, while big metal rings were found wearing on radius and ulna bones as bangles.
-This pattern corresponds to the texts regarding the body decorations of indigenous culture from Chinese historical documents [@Chen1963; @Ke1993].
-Agate beads were mostly found in female burials with an average of 4 to 5 pieces, while golden beads were found in both genders with an average of 2 to 3 pieces.
 
@@
 
-# Results of hypothesis testing
+# Results of hypothesis testing ?
 
-In this section I report the tests of the three hypotheses about indirect impacts of European colonization on the indigenous community at Kiwulan.
-To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European contact, and a lesser diversity after Chinese contact, I compare frequencies of ornament types between different time periods.
-To test the prediction that after European contact were new spatial patterns in the distributions of ornaments at Kiwulan, I use kernel density and point pattern analyses.
-
-## Frequency of ornaments between time periods
+## Changes in the frequencies of ornament types over time
 
 Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan.
-The result of chi-square test for the number of ornaments (X-squared = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = `r chisq_test_result_p_value`) shows that there is a significant difference between time period groups.
-It indicates a general pattern that most ornament types were present before European contact.The frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads.
-This trend can be also seen on other ornaments including agate bead, metal ring, and bell.
-However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then their frequency decreased in the European contact period and was even much less in the Chinese contact period.
+The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = `r chisq_test_result_p_value`) shows that there is a non-random difference between the three time periods.
+It indicates that most ornament types were present before European contact.
+Ornament frequencies reached a peak during European contact and then dropped during the Chinese contact period, especially for golden beads.
+This trend can be also seen on other ornaments including agate beads, metal rings, and bells.
+However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European contact period and a further decrease in the Chinese contact period.
 
@@
 The distribution of frequency for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether).
-Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `spearman_result_rho`, p-value = `r spearman_result_p_value`).
-This indicates that the increases in diversity can be explained by the effects of culture contact instead of natural effects of sample size.
+Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p-value = `r spearman_result_p_value`).
+This indicates that the increases in diversity can be explained by the effects of culture contact instead of effects of sample size.
 According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period.
-Agate beads can be divided into seven subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased significantly during the European contact and then declined in the Chinese contact.
-The small oval type shows a similar frequency before and during European contact but is not found in the Chinese contact period.
-Waxy oval beads were only found during European contact period.
-A higher variety of agate beads were found in European contact compared to other time periods.
-Similarly, metal rings can be classified into eight subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape.
-The greater varieties for those two classes might indicate multiple origins due to global trade networks stimulated by the Europeans.
-In contrast, copper bell and glass bead have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces.
+The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the Europeans.
+In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces.
 The common shape for copper bells is the large shape with a wide variety of human faces as a motif, while the common shape for glass beads is a bead with length less than 1 cm that are often identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].
@@
 ```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
-
 library(sf)
@@
 
-## Spatial distribution of ornament types
+## Changes in patterns of the spatial distribution of ornament types
 
-Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from research area for each time period.
-Before the European contact, the map shows a greater amount of ornaments found at the northern and middle parts of the research area, and they were more widespread during the European contact period.
-During the European contact period, ornaments were more ubiquitous with some clusters on the northern part, and less abundance during the Chinese contact period.
-Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major classes individually, some clusters across the area can be observed during the European contact, such as golden beads and agate beads.
-However, there seems to be no clear consistent clustered pattern across those different ornaments.
+Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from the research area for each time period.
+Before European contact, a greater amount of ornaments were found at the northern and middle parts of the research area.
+During the European contact period, ornaments were more widespread, with some clusters on the northern part.
+During the Chinese contact period the distribution is more even again.
+Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European contact, such as golden beads and agate beads.
+However, there seems to be no clear consistent pattern across those different ornaments.
 Each class shows its own pattern where the squares with higher number of ornaments distributed separately and independently.
@@
 Copper bells were usually found individually and seem randomly distributed across the area.
-In the Chinese Contact period, both the amount and density of different classes of ornaments decreased that they were only found in a few squares, for example, golden beads and glass beads were only found in a few squares in the middle part.
+In the Chinese Contact period, both the amount and density of different classes of ornaments decreased.
 
@@
 
-I used point pattern analysis to examine whether the distribution of ornaments was uneven in non-random ways that could reflect the presence of social inequality.
+I used point pattern analysis to examine whether the distribution of ornaments was uneven in non-random ways that could reflect changes in social inequality.
 Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
-In this case hotspots in the pattern of ornaments could indicate the changes in social organization due to culture contact.
-I used the intensity approach that focuses on the average density of points across space to explore their distribution.
-Ornament location points were given random coordinates within each square they were recovered from because their exact recovery location in each square was not recorded during excavation.
-The next step was to subset the ornaments into three groups for three time periods, and then compute the kernel density for each time period for comparison.
+Hotspots in the pattern of ornaments could indicate changes in social organization due to culture contact.
+To explore ornament distributions I used the intensity approach that focuses on the average density of points across space.
+During excavation, ornament locations were only recorded to the square (4 x 4 m).
+To prepare the location data for point pattern analysis, I assigned each ornament to a random coordinate pair located in square it was recovered from.
+The next step was to subset the ornaments into three groups for three time periods.
+Finally I computed the kernel densities for each time period for comparison.
 
-Kernel density estimations (KDE) estimate the probability of the density of events across space by creating a continuous, smooth density surface across space.
-Here we use KDE to visualize the event of the presence of ornaments that were represented by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
-Density values of artifacts per square meter are assigned to each cell.
-The results in Figure \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period.
+Kernel density estimations (KDE) estimate the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
+Here we use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
+Density values of artifacts per square meter were calculated for each cell.
+
+The results show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period (Figure \@ref(fig:plot-all-densities)).
 Those maps present three consistent sub-regions with a core area that shifts over time.
@@
 However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
-To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is randomly distributed to understand the underlying social processes related to ornament distributions.
+To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
 
@@
 
-Here we use a Monte Carlo method to test the spatial randomness of spatial events based on the average nearest-neighbor distance method (ANN).
-Our null hypothesis is that the distribution of ornaments is consistent with a completely random process, so we then simulate the locations of ornaments across the space 1000 times for each time period.
-The observed data was compared to the random pattern generated by simulated processes to see if our data is randomly distributed or not.
+To test this hypothesis about the randomness of spatial locations, I used a Monte Carlo method to simulate average nearest-neighbor distances (ANN).
+Our null hypothesis is that the distribution of ornaments is consistent with a completely random process.
+We simulated the locations of ornaments across the space 1000 times for each time period.
+The observed data was compared to the random patterns generated by simulated processes to determine if our data are randomly distributed or not.
 Figure \@ref(fig:hypo-test-distribution) shows the distributions of the ANN distances calculated on 1000 simulations of ornament locations.
-The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments have a non-random clustered distributions.
+The results show that 100% of the simulated values are much greater than our observed ANN value during the European contact period, which means the ornaments have non-random clustered distributions.
 A similar, but less extreme, result is also observed during the pre-European period.
@@
 This testing reveals that the clustered distributions of ornaments during the pre-European and European contact periods might reflect the presence of different social groups.
-Moreover, the clustering of ornaments during the European contact period is highly non-random, potentially indicate some social changes leading to a concentration of power to control the distribution of ornaments at Kiwulan.
+Moreover, the clustering of ornaments during the European contact period is highly non-random, potentially indicating that some of the social changes leading to a concentration of power to control the distribution of ornaments at Kiwulan.
 
@@
 This might result from the large scale exchange network that stimulated the circulation of different ornament classes.
-Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bells and glass beads, which might indicate their diverse origins.
 However, the frequency of overall ornaments and each subtype declines significantly during the Chinese contact period.
-The reason could be related to a smaller scale of network, the overall decline of the indigenous population in Yilan, and adopting the mortuary practices of Han Chinese.
-Since the end of the 18th century, lots of indigenous people moved southwards to Hualien due to population pressure caused by the great amount of Han Chinese immigrants [@Chen2007].
-According to the Chinese census reports in 1821 and Japanese field survey in 1895, the Kiwulan population was less than 100 that dropped by nearly 90 percent compared with the last census in the 17th century [@Yao1996; @Ino1898].
-In addition, burials after 18th century unearthed fewer ornaments or even lack of them that might indicate a change in mortuary practices.
-This can be also identified by the presence of coffin-like case that was viewed as a funerary item of Han Chinese, however, with local indigenous carving patterns on the wooden surfaces [@Chen2007; @Chiu2004].
+This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the mortuary practices of Han Chinese.
+Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
+According to Chinese census reports in 1821 and Japanese field surveys in 1895, the Kiwulan population was less than 100 [@Yao1996; @Ino1898].
 
-The spatial pattern of ornaments shows that their distribution was clustered since the pre-European and European contact period.
-Hypothesis testing for spatial patterns further indicates that these clusters are non-random, and are concentrated during European contact.
-Because trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering could indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European contact period.
-In addition, there was a burial dated to 17th century unearthed 60 golden beads above the average of 2-3 pieces that hints the possibility of increasing social inequality [@Chen2007; @Cheng2008].
-In Taiwan, most prehistoric organization were thought as egalitarian societies because of the lack of centralization of power that is often used as a major indicator of hierarchical complexity.
-However, this concept of egalitarianism has been challenged by evidence of social inequalities such as uneven distributions of prestige goods or the emergence of influential individuals in some non- hierarchical human societies [@Ames2010; @Bowles2010; @Woodburn1982].
+Spatial patterns of ornaments shows that their distribution was clustered during the pre-European and European contact period.
+Hypothesis testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European contact period.
+Because ethnographic evidence indications that trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering may indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European contact period.
+In addition, there was a burial dated to 17th century unearthed 60 golden beads above the average of 2-3 pieces in the pre-European period, further indicating increasing social inequality [@Chen2007; @Cheng2008].
 
-Horizontal complexity is another approach to discuss the emergence of social inequity by focusing on how people achieve and maintain power in a wide variety of societies [@Drennan2010; @Feinman2000].
-For example, the corporate/network model proposed by @Feinman2000 that expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.
-In the network mode, inequality develops when individuals accumulate wealth through individual networks and use wealth to attract factions, control resources, and monopolize trade networks.
+How might these results fit into a bigger picture of social change at periphery of colonial systems?
+We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Drennan2010; @Feinman2000; @Ames2010; @Bowles2010].
+For example, the corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.
+In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks.
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labor tasks [@Feinman2000; @Siegel1999].
-Applying the corporate/network model may give insight into the underlying pathways that led to social inequality in northeastern Taiwan.
-In the circumscribed environments of the Yilan Plain, the corporate/network continuum will provide a comparative basis to explore the impact of European contact on changes in the degree of indigenous social inequality.
-One insight emerging from this model is that the long-distance trade network introduced by the Europeans results in the appearance of a network mode due to the competitions among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
+
+We may consider that Yilan social organization went from corporate mode in the pre-European contact period, then transitioned into network mode during the European colonial contact period, then back to corporate mode in the Chinese contact period.
+The small number of ornaments, and less concentrated distribution during the pre-European period is consistent with shared power and wealth of the corporate mode.
+The long-distance trade network introduced by the Europeans resulted in the appearance of a network mode due to competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
 Because of the weak direct control from the European colonizers in northeastern Taiwan, it implies that local leaders might have flexibility to manipulate European colonial image, expand personal power, and monopolize the high-value trade goods [@Kang2012].
+During the Chinese contact period, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
+This reflects the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals accumulate wealth.
 
@@
 Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017].
-Kiwulan in Northeastern Taiwan is an especially important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by tea to regional and global trade networks.
+Kiwulan in Northeastern Taiwan is an important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
 Kiwulan can provide valuable insights into the discussion of the influence of indirect colonial contacts.
-In northeast Taiwan foreign trade ornaments were viewed as prestige goods and played an important social role in divination and funeral rituals.
-As a marker for status, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
-The greater amount and diversity of ornament types during the European contact period reflects international exchange in a colonial context.
-Those ornaments may carry an exotic and powerful image that signals wealth and trading connections to the inhabitants of Kiwulan.
-These symbolic values may have stimulated more competitions between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might result in an increase in social inequality.
+The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
+The greater amount and diversity of ornament types during the European contact period reflects international exchanges in a colonial context.
+Those ornaments may carry an exotic and powerful image that signals wealth and trading connections for the inhabitants of Kiwulan.
+These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
 
-This study demonstrates that foreign ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations.
-Ornaments can give insights into the amplification of social inequality stimulated by the European colonization.
-It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system.
-During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwulan population.
+This study demonstrates that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
+Ornaments can give insights into the amplification of social inequality stimulated by European colonization.
+It also shows the agency of indigenous people to incorporate ornaments into their social system.
 Future work could extend this approach to studies of other trade goods such as ceramics.
-Future directions for understanding the dynamics of social inequality at Kiwulan could include analysis of pottery production and standardization, and mortuary practices.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardization, and mortuary practices.
 

### 2019-08-13 analysis/paper/paper.Rmd
add map figure

--- 2019-08-11 analysis/paper/paper.Rmd
@@
 
-```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px", echo=FALSE}
+```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px"}
 # need to add the locations of Keelung, Tamsui, and hoping dao

### 2019-08-13 analysis/paper/paper.Rmd
fix conflict

--- 2019-08-13 analysis/paper/paper.Rmd
@@
 
-```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px", echo=FALSE}
-# need to adjust Keelung, heping dao, check it out after knitting
+```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px"}
+# need to add the locations of Keelung, Tamsui, and hoping dao
 library(here)

### 2019-08-13 analysis/paper/map.R
polish the map

@@ -59,13 +59,22 @@ tw_map +
                  y = lat),
              size = 2,
              color = "red") +
-  geom_shadowtext(data = site_location,
+  geom_shadowtext(data = site_location %>%
+                   filter(location != "Heping dao") ,
             aes(x = lon,
                 y = lat,
                 label = location),
-            size = 5,
+            size = 4,
             position = position_nudge(y = 0.05),
             check.overlap = TRUE) +
+  geom_shadowtext(data = site_location %>%
+                    filter(location == "Heping dao"),
+                  aes(x = lon,
+                      y = lat,
+                      label = location),
+                  size = 4,
+                  position = position_nudge(y = 0.07, x = 0.1),
+                  check.overlap = TRUE) +
   coord_sf(xlim = c(120.95, 122.05),
            ylim = c(24.4, 25.4),
            expand = FALSE) +

### 2019-08-13 analysis/paper/paper.Rmd
minor changes to introduction
try to get rid of white space around map

@@
 )
+# prepare for JAS:R
+# https://www.elsevier.com/journals/journal-of-archaeological-science-reports/2352-409x/guide-for-authors#2000
+# not exceed 6000 words excluding figures, tables, highlights 
+# iamge Single column	w = 90 mm 
+# image Double column w = 190 mm 
 ```
@@
 Large-scale and frequent trading activities brought a wide variety of objects into indigenous communities in Yilan.
-\The indigenous people of Yilan established their own exchange network that has been described as “inter-insular trade” by @Chen2005 [p.
+\The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
 12] to refer to small-scale regional trade between China and Taiwan on an irregular basis.
@@
 
-# Results of hypothesis testing ?
+# Results of hypothesis testing
 

### 2019-08-13 analysis/paper/map.R

@@ -106,11 +106,22 @@ tw_map +
         axis.title.y = element_blank())
 
 ggplot() +
-  coord_equal(xlim = c(0, 2.5), ylim = c(0, 1), expand = FALSE) +
-  annotation_custom(ggplotGrob(TW_SE_Asia), xmin = 0, xmax = 1.5, ymin = 0, ymax = 1) +
-  annotation_custom(ggplotGrob(TW_map_with_site), xmin = 1.2, xmax = 2.5, ymin = 0, ymax = 1) +
+  scale_x_continuous(limits = c(0.2,2.5), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0,1), expand = c(0, 0)) +
+  coord_equal() +
+  annotation_custom(ggplotGrob(TW_SE_Asia),
+                    xmin = 0,
+                    xmax = 1.5,
+                    ymin = 0,
+                    ymax = 1) +
+  annotation_custom(ggplotGrob(TW_map_with_site),
+                    xmin = 1.2,
+                    xmax = 2.5,
+                    ymin = 0,
+                    ymax = 1) +
   theme_void()
 
-ggsave(here("analysis", "figures", "kiwulan-location-map.png"))
+ggsave(here("analysis", "figures", "kiwulan-location-map.png"),
+       width = 190, units = "mm")
 
 # add the location of Keelung, Tamsui, and Heping dau

### 2019-08-13 analysis/paper/map.R
map size is good now (set the height also in ggsave)

@@ -106,8 +106,8 @@ tw_map +
         axis.title.y = element_blank())
 
 ggplot() +
-  scale_x_continuous(limits = c(0.2,2.5), expand = c(0, 0)) +
-  scale_y_continuous(limits = c(0,1), expand = c(0, 0)) +
+  scale_x_continuous(limits = c(0, 2.5), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
   coord_equal() +
   annotation_custom(ggplotGrob(TW_SE_Asia),
                     xmin = 0,
@@ -122,6 +122,8 @@ ggplot() +
   theme_void()
 
 ggsave(here("analysis", "figures", "kiwulan-location-map.png"),
-       width = 190, units = "mm")
+       width = 190,
+       height = 90,
+       units = "mm")
 
 # add the location of Keelung, Tamsui, and Heping dau

### 2019-08-14 analysis/paper/paper.Rmd
need to add: Figure 2: Map showing the largest section of excavation areas at Kiwulan

--- 2019-08-13 analysis/paper/paper.Rmd
@@
 )
+# prepare for JAS:R
+# https://www.elsevier.com/journals/journal-of-archaeological-science-reports/2352-409x/guide-for-authors#2000
+# not exceed 6000 words excluding figures, tables, highlights 
+image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
+image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
 ```
@@
 
-\Although mountains isolated Yilan by land, the sea provided a way through which indigenous people in Yilan could interact with groups elsewhere, such as indigenous groups, distant traders such as the Han people of China, and Europeans.
+Although mountains isolated Yilan by land, the sea provided a way through which indigenous people in Yilan could interact with groups elsewhere, such as indigenous groups, distant traders such as the Han people of China, and Europeans.
 Large-scale and frequent trading activities brought a wide variety of objects into indigenous communities in Yilan.
-\The indigenous people of Yilan established their own exchange network that has been described as “inter-insular trade” by @Chen2005 [p.
+The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
 12] to refer to small-scale regional trade between China and Taiwan on an irregular basis.
@@
 
-```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px"}
+```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", fig.width=image_width_2_col_inch}
 # need to add the locations of Keelung, Tamsui, and hoping dao
@@
 ```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number.
-Each square is 4 x 4 m", out.width= "70%", echo=FALSE}
+Each square is 4 x 4 m", fig.width=image_width_2_col_inch}
 # add to the map and scale and north arrow. Check it out after knitting 
@@
 
-Kiwulan (Figure \@ref(fig:KWL-map)) is located at northern Yilan and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
+Kiwulan is located at northern Yilan (Figure \@ref(fig:KWL-map)) and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge .
 The total area includes eight open area sections of 262 excavation squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan", out.width='60%'}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan", fig.width=image_width_1_col_inch}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 
+plot_ornaments_count <- 
 ggplot(ornaments_count,
@@
   coord_flip() +
+  scale_y_continuous(limits = c(0, 160)) +
   labs(y = "Frequency",
        x = "Categories") +
-  theme_minimal(base_size = 14)
+  theme_minimal(base_size = 10)
+
+ggsave(plot = plot_ornaments_count,
+       here("analysis", "figures",
+       "plot-ornaments-count.png"),
+       width = 90, 
+       height = 40,
+       units = "mm")
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-ornaments-count.png"))
 ```
@@
 # histogram facetting for three periods
+plot_ornaments_count_three_periods <- 
 ggplot(ornaments_count) +
@@
   coord_flip() +
-  labs(x = "categories", y = "Frequency") +
+  labs(x = "Categories",
+       y = "Frequency") +
   theme_minimal(base_size = 12)
+
+ggsave(plot = plot_ornaments_count_three_periods,
+       here("analysis", "figures",
+       "plot-ornaments-count-three-periods.png"),
+       width = 190, 
+       height = 80,
+       units = "mm")
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-ornaments-count-three-periods.png"))
 ```
@@
 Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan.
-The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = `r chisq_test_result_p_value`) shows that there is a non-random difference between the three time periods.
+The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p = $`r round(chisq_test_result_p_value, 14)`$) shows that there is a non-random difference between the three time periods.
 It indicates that most ornament types were present before European contact.
@@
 
+base_size_multi_plot <- 10
+
 agate_bead_shape_plot <- 
@@
        x = "Agate bead") +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = base_size_multi_plot) +
   theme(strip.background = element_blank(), 
@@
   facet_wrap(~ period) +
-  theme_minimal(base_size = 14) 
+  theme_minimal(base_size = base_size_multi_plot) 
 ```
@@
        x = "Bell") +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = base_size_multi_plot) +
   theme(strip.background = element_blank(), 
@@
        x = "Glass bead") +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = base_size_multi_plot) +
   theme(strip.background = element_blank(), 
@@
 library(cowplot)
+plot_multi_types_by_period <- 
 plot_grid(ring_shape_plot,
@@
           rel_heights = c(1,1,0.7,0.62))
+
+ggsave(plot = plot_multi_types_by_period,
+       here("analysis", "figures",
+       "plot-multi-types-by-period.png"),
+       width = 190, 
+       height = 150,
+       units = "mm")
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-multi-types-by-period.png"))
 ```
@@
 The distribution of frequency for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether).
-Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p-value = `r spearman_result_p_value`).
+Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
 This indicates that the increases in diversity can be explained by the effects of culture contact instead of effects of sample size.
@@
 # plot spatial pattern for all ornaments
+plot_spatial_distribution_all_ornaments_by_period <- 
 ggplot() +
@@
         strip.text.x = element_text(size = 11))
+
+ggsave(plot = plot_spatial_distribution_all_ornaments_by_period,
+       here("analysis", "figures",
+       "plot-spatial-distribution-all-ornaments-by-period.png"),
+       width = 190, 
+       height = 50,
+       units = "mm")
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-spatial-distribution-all-ornaments-by-period.png"))
 ```
 
-```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
+```{r, plot-spatial-pattern-each, fig.cap="Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
 
-# filter golden bead
-ornaments_period_join_gold <-
+ornaments_period_join_selected_types <- 
   ornaments_period_join %>% 
-  filter(Categories %in% 'Golden bead')
+  filter(Categories %in% c('Golden bead', 
+                           'Glass bead',
+                           'Agate bead'))
 
-# plot by period
-dis_golden <-
-ggplot(ornaments_period_join_gold) +
+plot_spatial_distribution_some_ornaments_by_period <- 
+ggplot() +
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, 
+  geom_sf(data = ornaments_period_join_selected_types, 
           aes(geometry = geometry,
               fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Golden Beads",
+  facet_grid(rows = vars(Categories),
+             cols = vars(period)) +
+  labs(title = "",
        fill = "frequency") +
   scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
+  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5),
@@
         axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
+        strip.text.x = element_text(size = 8))
 
-# filter glass bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Glass bead')
+ggsave(plot = plot_spatial_distribution_some_ornaments_by_period,
+       here("analysis", "figures",
+       "plot-spatial-distribution-some-ornaments-by-period.png"),
+       width = 190, 
+       height = 130,
+       units = "mm")
 
-# plot by period
-dis_glass <-
-  ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, 
-          aes(geometry = geometry,
-              fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Glass Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-spatial-distribution-some-ornaments-by-period.png"))
 
-# filter agate bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Agate bead')
-
-# plot by period
-dis_agate <-
-ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, 
-          aes(geometry = geometry, 
-              fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Agate Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-plot_grid(dis_golden,
-          dis_glass,
-          dis_agate, 
-          align = "v",
-          nrow = 3)
 ```
@@
 kernel_to_plot <- 
-  plot(kernel_tmp, ...)
+  plot(kernel_tmp, main = "")
   contour(kernel_tmp, add = TRUE)
+  title(..., line = -2, adj = 0.4)
 
@@
 ```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods"}
-par(mfrow = c(1, 3))
+
 # apply function to each data frame in the list
 titles <- names(ornaments_period_join_cen)
+
+png(
+  here("analysis", "figures",
+       "plot-kde-maps.png"), 
+    width = 190,
+    height = 63,
+    units = "mm",
+  res = 300)
+
+par(mfrow = c(1, 3),
+    oma = c(1,1,1,1) + 0.1,
+    mar = c(0,0,1,1) + 0.1)
 
@@
       ~density_analysis(.x, main = .y))
+
+invisible(dev.off())
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-kde-maps.png"))
 ```
@@
 
-```{r hypo-test-distribution, fig.cap= "Histogram of simulated ANN values with the blue line indicating the observed value"}
+```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values with the blue line indicating the observed value"}
 
 # make the blue lines thicker to be more visible
+png(
+  here("analysis", "figures",
+       "plot-kde-ann-histograms.png"), 
+    width = 90,
+    height = 150,
+    units = "mm",
+  res = 300)
 par(mfrow = c(3, 1))
@@
     ~invisible(nndist_hypo_testing(.x, 
-                         main = .y)))
-)
+                         main = .y))))
+invisible(dev.off())
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-kde-ann-histograms.png"))
 ```

### 2019-08-14 analysis/paper/paper.Rmd
make plot titles consistent

@@
   group_split() %>% 
-  set_names(c("Pre-European",
-              "European",
-              "Chinese"))
+  set_names(c("Before European Contact",
+              "European Contact",
+              "Chinese Contact"))
 ```

### 2019-08-14 analysis/paper/paper.Rmd
minor changes based on 2nd edit from Ben

--- 2019-08-13 analysis/paper/paper.Rmd
@@
 )
-# prepare for JAS:R
-# https://www.elsevier.com/journals/journal-of-archaeological-science-reports/2352-409x/guide-for-authors#2000
-# not exceed 6000 words excluding figures, tables, highlights 
-# iamge Single column	w = 90 mm 
-# image Double column w = 190 mm 
 ```
@@
 Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
-Rivers and seas provide the way for the interaction between local indigenous settlements and further afield communities, such as the Basai in northern Taiwan and Han Chinese.
+
+Although mountains isolated Yilan by land, rivers and seas provided a way for the interaction between local indigenous settlements and further afield communities, such as the Basai in northern Taiwan, and Han Chinese.
+The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
+12] to refer to small-scale regional trade between China and Taiwan on an irregular basis since 14th century.
 Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
-
-\Although mountains isolated Yilan by land, the sea provided a way through which indigenous people in Yilan could interact with groups elsewhere, such as indigenous groups, distant traders such as the Han people of China, and Europeans.
-Large-scale and frequent trading activities brought a wide variety of objects into indigenous communities in Yilan.
-\The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
-12] to refer to small-scale regional trade between China and Taiwan on an irregular basis.
+Trading activities brought a wide variety of objects into indigenous communities in Yilan.
 One of the most commonly traded types of object in this region were personal ornaments such as glass and stone beads.

### 2019-08-14 analysis/paper/paper.Rmd
2nd edit

@@
 Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
-Understanding the indirect effects of colonialism are important for detecting colonial impacts on indigenous peoples in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017].
-In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
-Studying the consumption of foreign goods can improve our understanding of the negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2005; @Dietler2015; @Given2004; @Mullins2011; @Rubertone2000; @Scaramelli2005; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
+Understanding the indirect effects of colonialism are important for detecting colonial impacts on indigenous peoples in the periphery of colonial control [@Trabert2017].
+In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
+Studying the consumption of foreign goods can improve our understanding of the negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Rubertone2000; @Scaramelli2005; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
 
@@
 In contrast to Heping Dao, Yilan in Northeastern Taiwan is physically isolated and there are no traces of European settlement.
+Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
+The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
 This makes Yilan an especially unique location for exploring indirect colonial effects on local indigenous settlements because of isolation.
-The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
 
@@
 The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
-12] to refer to small-scale regional trade between China and Taiwan on an irregular basis since 14th century.
+12] to refer to small-scale regional trade between China and Taiwan on an irregular basis before European contact.
 Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].

### 2019-08-14 analysis/paper/paper.Rmd
update figure code after it disappeared :$

@@
 ---
-title: "Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan"
+title: "Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan"
 author:
@@
 # Excavations at Kiwulan in northeastern Taiwan
-
-```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px"}
+```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", fig.width=image_width_2_col_inch}
 # need to add the locations of Keelung, Tamsui, and hoping dao
@@
 # L2 and L1 was the period of Chinese contact. 
-
 ornaments_period <- 
@@
 ```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number.
-Each square is 4 x 4 m", out.width= "70%", echo=FALSE}
+Each square is 4 x 4 m", fig.width=image_width_2_col_inch}
 # add to the map and scale and north arrow. Check it out after knitting 
+library(sf)
+library(cartography)
+# read in the spatial data
+burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
+AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
+AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
+location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
+units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
+
+# tidy spatial data
+# 40 sampling units
+sample_units <-
+  AD_data %>%
+  semi_join(units_40, by = c (`坑號` = "the_sq"))
+
+# join two dataset to get coordinate
+ornaments_period_join <-
+  ornaments_period %>%
+  mutate(period = as.factor(period)) %>% # seems need this line first
+  group_by(Pit_No, period) %>%
+  count(Categories) %>%
+  inner_join(AD_data, by = c("Pit_No" = "坑號"))
+
+# mapping sampling area
+KWL_excavation_map <- 
+ggplot(sample_units) +
+  geom_sf(data = AD_data, fill = NA) +
+  geom_sf(color = 'red') +
+  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
+  theme_minimal() +
+  theme_void()
+
+#1. Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
+#2. Need to add in here that each square is 4 x 4 m
+#3. and say which way north is. Or add to the map and scale and north arrow.
+
+ggsave(plot = KWL_excavation_map, 
+       here("analysis", "figures", "KWL-excavation-map.png"), width = 190, units = "mm")
+
 knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.png"))
@@
 # total number of ornaments for each categories 
-
 ornaments_count <-
@@
   mutate(Categories = fct_reorder(Categories, n))
-
+plot_ornaments_count <- 
 ggplot(ornaments_count,
@@
   coord_flip() +
+  scale_y_continuous(limits = c(0, 160)) +
   labs(y = "Frequency",
        x = "Categories") +
-  theme_minimal(base_size = 14)
+  theme_minimal(base_size = 10)
+ggsave(plot = plot_ornaments_count,
+       here("analysis", "figures",
+       "plot-ornaments-count.png"),
+       width = 90, 
+       height = 40,
+       units = "mm")
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-ornaments-count.png"))
 ```
@@
          Type = ifelse(Type == "large thin string", "thin large", Type))
-
 ornaments_shape <-
@@
   arrange(Categories, desc(`European Contact`))
-
 # making table
@@
              caption = "Ornament subtype frequencies at Kiwulan")
-
 # some basic stat
@@
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D classes were from original excavation report (Chen 2007).", out.width= "400px", echo=FALSE}
+The photographs of B, C, D classes were from original excavation report (Chen 2007).", echo=FALSE}
 knitr::include_graphics(here("analysis", "figures", "ornament.png"))
@@
 ### delete one plot
-
 # categories of ornaments by period
@@
   count()
-
 # histogram facetting for three periods
+plot_ornaments_count_three_periods <- 
 ggplot(ornaments_count) +
@@
   spread(Categories, n) 
-
 # compute chi sqaure
@@
 Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan.
-The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = `r chisq_test_result_p_value`) shows that there is a non-random difference between the three time periods.
+The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$) shows that there is a non-random difference between the three time periods.
 It indicates that most ornament types were present before European contact.
@@
          Type != 'other')
-
+base_size_multi_plot <- 10
 agate_bead_shape_plot <- 
@@
   mutate(Type = fct_rev(Type)) 
-
 ring_shape_plot <-
@@
          Type != 'unclassified')
-
 bell_shape_plot <-
@@
   filter(Categories == 'Glass bead')
-
 glass_bead_plot <-
@@
           rel_heights = c(1,1,0.7,0.62))
+ggsave(plot = plot_multi_types_by_period,
+       here("analysis", "figures",
+       "plot-multi-types-by-period.png"),
+       width = 190, 
+       height = 150,
+       units = "mm")
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-multi-types-by-period.png"))
 ```
@@
   distinct()
-
 ornaments_diversity_test_2 <-
@@
   distinct()
-
 # spearman test
@@
          conf.level = 0.95)
-
 spearman_result_stat <- round(spearman_result$statistic, 2)
@@
 library(sf)
+
 # read in the spatial data
@@
 units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
-
 # tidy spatial data for 40 sampling units
@@
   semi_join(units_40, by = c (`坑號` = "the_sq"))
+# join two dataset to get coordinate
 
-# join two dataset to get coordinate
 ornaments_period_join <-
@@
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
+# plot spatial pattern for all ornaments
 
-# plot spatial pattern for all ornaments
+plot_spatial_distribution_all_ornaments_by_period <- 
 ggplot() +
@@
         strip.text.x = element_text(size = 11))
+
+ggsave(plot = plot_spatial_distribution_all_ornaments_by_period,
+       here("analysis", "figures",
+       "plot-spatial-distribution-all-ornaments-by-period.png"),
+       width = 190, 
+       height = 50,
+       units = "mm")
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-spatial-distribution-all-ornaments-by-period.png"))
 ```
 
-```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
+```{r, plot-spatial-pattern-each, fig.cap="Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
+ornaments_period_join_selected_types <- 
+  ornaments_period_join %>% 
+  filter(Categories %in% c('Golden bead', 
+                           'Glass bead',
+                           'Agate bead'))
 
-# filter golden bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Golden bead')
-
-# plot by period
-dis_golden <-
-ggplot(ornaments_period_join_gold) +
+plot_spatial_distribution_some_ornaments_by_period <- 
+ggplot() +
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, 
+  geom_sf(data = ornaments_period_join_selected_types, 
           aes(geometry = geometry,
               fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Golden Beads",
+  facet_grid(rows = vars(Categories),
+             cols = vars(period)) +
+  labs(title = "",
        fill = "frequency") +
   scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
+  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5),
@@
         axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
+        strip.text.x = element_text(size = 8))
 
-# filter glass bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Glass bead')
+ggsave(plot = plot_spatial_distribution_some_ornaments_by_period,
+       here("analysis", "figures",
+       "plot-spatial-distribution-some-ornaments-by-period.png"),
+       width = 190, 
+       height = 130,
+       units = "mm")
 
-# plot by period
-dis_glass <-
-  ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, 
-          aes(geometry = geometry,
-              fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Glass Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-# filter agate bead
-ornaments_period_join_gold <-
-  ornaments_period_join %>% 
-  filter(Categories %in% 'Agate bead')
-
-# plot by period
-dis_agate <-
-ggplot(ornaments_period_join_gold) +
-  geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join_gold, 
-          aes(geometry = geometry, 
-              fill = n), alpha = 0.9) +
-  facet_wrap(~ period) +
-  labs(title = "Agate Beads",
-       fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
-
-plot_grid(dis_golden,
-          dis_glass,
-          dis_agate, 
-          align = "v",
-          nrow = 3)
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-spatial-distribution-some-ornaments-by-period.png"))
 ```
@@
          lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]]))) 
-
 ornaments_period_join_cen <- 
@@
 library(raster)
-
 #store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points, by period
-
 # use common range for each period
@@
 yrange <- range(lat_all, na.rm=T)
-
 # create a function to apply to each period
@@
 ornaments_point_sp_tmp <- as(ornaments_point_ppp_tmp, "SpatialPoints")
-
 # kernel density estimation with contour
@@
 kernel_to_plot <- 
-  plot(kernel_tmp, ...)
+  plot(kernel_tmp, main = "")
   contour(kernel_tmp, add = TRUE)
-
+  title(..., line = -2, adj = 0.4)
 return(list(ornaments_point_ppp = ornaments_point_ppp_tmp,
@@
 ```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods"}
-par(mfrow = c(1, 3))
 # apply function to each data frame in the list
 titles <- names(ornaments_period_join_cen)
-
+png(
+  here("analysis", "figures",
+       "plot-kde-maps.png"), 
+    width = 190,
+    height = 63,
+    units = "mm",
+  res = 300)
+par(mfrow = c(1, 3),
+    oma = c(1,1,1,1) + 0.1,
+    mar = c(0,0,1,1) + 0.1)
 kernels_for_plotting <- 
@@
       ~density_analysis(.x, main = .y))
+invisible(dev.off())
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-kde-maps.png"))
 ```
@@
 To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
-
 ```{r function for ANN values}
@@
 ann_tmp <- mean(nndist(x$ornaments_point_ppp, k = 1))
-
 # object for storing simulated ANN values
 ann_r_tmp <- vector(length = n) 
-
 for (i in 1:n) {
@@
 }  
-
 the_plot <- 
@@
   ...)
-
 abline(v = ann_tmp, col = "blue")
-
 return(the_plot)
@@
 
-```{r hypo-test-distribution, fig.cap= "Histogram of simulated ANN values with the blue line indicating the observed value"}
-
+```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values with the blue line indicating the observed value"}
 # make the blue lines thicker to be more visible
+png(
+  here("analysis", "figures",
+       "plot-kde-ann-histograms.png"), 
+    width = 90,
+    height = 150,
+    units = "mm",
+  res = 300)
 par(mfrow = c(3, 1))
@@
     ~invisible(nndist_hypo_testing(.x, 
-                         main = .y)))
-)
+                         main = .y))))
+invisible(dev.off())
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-kde-ann-histograms.png"))
 ```

### 2019-08-14 analysis/paper/KWL-excavation-map .R

@@ -1,36 +1,0 @@
-library(sf)
-library(cartography)
-# read in the spatial data
-burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
-AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
-AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
-location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
-units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
-
-# tidy spatial data
-# 40 sampling units
-sample_units <-
-  AD_data %>%
-  semi_join(units_40, by = c (`坑號` = "the_sq"))
-
-# join two dataset to get coordinate
-ornaments_period_join <-
-  ornaments_period %>%
-  mutate(period = as.factor(period)) %>% # seems need this line first
-  group_by(Pit_No, period) %>%
-  count(Categories) %>%
-  inner_join(AD_data, by = c("Pit_No" = "坑號"))
-
-# mapping sampling area
-ggplot(sample_units) +
-  geom_sf(data = AD_data, fill = NA) +
-  geom_sf(color = 'red') +
-  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
-  theme_minimal() +
-  theme_void()
-
-#1. Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
-#2. Need to add in here that each square is 4 x 4 m
-#3. and say which way north is. Or add to the map and scale and north arrow.
-
-ggsave("analysis/figures/KWL-excavation-map.png")

### 2019-08-15 analysis/paper/paper.Rmd
added abstract and keywords

@@
 abstract: |
-  Text of abstract
+  European expansion to Asia since 16th century made a great impact on local indigenous people, where indigenous societies faced forced contact with imperial powers that often transformed their economic, cultural, and socio-political systems.
+In East Asia, direct European colonial rule on local communities was much less pronounced than it in island Southeast Asia and Oceania.
+However, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
+Here I use northeastern Taiwan as an example to explore the indirect European colonial impacts on local societies by examining trade ornaments excavated at Kiwualn site (1400-1900 AD), a large late Iron Age settlement that encountered different episodes of culture contact.
+The cultural context, variety, and spatial distribution of trade ornaments were examined in terms of three time episodes, before European contact, European contact, and Chinese contact.
+The results show that there is a greater amount and diversity of ornament types during the European contact period, and the spatial pattern of ornaments indicates non-random clusters that hints uneven distribution.
+Those ornaments may carry an exotic and powerful image that signals wealth and trading connections with Europeans for the inhabitants of Kiwulan.
+These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation resulting in an increase in social inequality.
+This study demonstrates the use of foreign ornaments as a proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
+Ornaments can give insights into the amplification of social inequality stimulated by European colonization.
+It also shows the agency of indigenous people to incorporate ornaments into their social system.
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  European contact; trade ornaments; social inequality; indigenous people; East Asia
 highlights: |

### 2019-08-15 analysis/paper/paper.Rmd
added keywords and polished the excavation map plot

@@
 keywords: |
-  European contact; trade ornaments; social inequality; indigenous people; East Asia
+  Historical archaeology; European colonization; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
 highlights: |
@@
 library(cartography)
+#devtools::install_github('oswaldosantos/ggsn')
+library(ggsn)
 # read in the spatial data
 burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
+post <- invisible(st_read(here("analysis", "data", "raw_data", "AD_postholes.shp"), quiet = TRUE))
 AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
@@
   geom_sf(color = 'red') +
-  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
+  geom_sf_text(aes(label = `坑號`), size = 2.5, 
+               alpha = 0.9, fontface = "bold") +
+  geom_sf(data = post, color = 'blue', size = 0.2, alpha = 0.5) +
   theme_minimal() +
-  theme_void()
+  blank() +
+    north(AD_data, scale = 0.2, symbol = 3)
+  #theme_void()
 

### 2019-08-17 analysis/paper/paper.Rmd
revised abstract and added highlights, used viridis color for plots

@@
 abstract: |
-  European expansion to Asia since 16th century made a great impact on local indigenous people, where indigenous societies faced forced contact with imperial powers that often transformed their economic, cultural, and socio-political systems.
-In East Asia, direct European colonial rule on local communities was much less pronounced than it in island Southeast Asia and Oceania.
+  The direct impact of European colonization since 16th century on local indigenous people has been widely discussed.
+In East Asia, direct European colonial rule on local communities was much less pronounced.
 However, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
-Here I use northeastern Taiwan as an example to explore the indirect European colonial impacts on local societies by examining trade ornaments excavated at Kiwualn site (1400-1900 AD), a large late Iron Age settlement that encountered different episodes of culture contact.
-The cultural context, variety, and spatial distribution of trade ornaments were examined in terms of three time episodes, before European contact, European contact, and Chinese contact.
-The results show that there is a greater amount and diversity of ornament types during the European contact period, and the spatial pattern of ornaments indicates non-random clusters that hints uneven distribution.
-Those ornaments may carry an exotic and powerful image that signals wealth and trading connections with Europeans for the inhabitants of Kiwulan.
-These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation resulting in an increase in social inequality.
+Using trade ornaments recovered from Kiwualn site (1400-1900 AD) in northeastern Taiwan, this research explores the indirect European colonial impacts on local societies by examining cultural context, variety, and spatial distribution of trade ornaments in terms of three time episodes relating different scales of culture contact.
+The results indicate the diversity of ornaments was greater and their spatial distribution was non-random during the European contact period that hints the increasing social inequality in a colonial context.
 This study demonstrates the use of foreign ornaments as a proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
@@
 highlights: |
-  These are the highlights. 
+  1. Indirect impact of European colonization on indigenous peoples in the periphery of colonial control was rarely discussed.
+  2. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
+  3. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European contact.
+  4.
+Ornaments may carry an exotic and powerful image that signals wealth and trading connections with colonists that resulted in an increase in social inequality due to competition of wealth accumulation between individuals.
+  5. Ornaments can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
 ---
@@
 
-Prior to contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the tenth century, and encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
+Prior to contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the 10th century, and encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
 11].
@@
 ```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number.
+Small dots represent the location of post-holes.
 Each square is 4 x 4 m", fig.width=image_width_2_col_inch}
@@
                alpha = 0.9, fontface = "bold") +
-  geom_sf(data = post, color = 'blue', size = 0.2, alpha = 0.5) +
+  geom_sf(data = post, color = 'blue', size = 0.2, alpha = 0.4) +
   theme_minimal() +
@@
   #theme_void()
-
-#1. Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
-#2. Need to add in here that each square is 4 x 4 m
-#3. and say which way north is. Or add to the map and scale and north arrow.
 
@@
 
-Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
+Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
 The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
@@
 library(sf)
+library(viridis)
 
@@
 units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
+
 # tidy spatial data for 40 sampling units
@@
   semi_join(units_40, by = c (`坑號` = "the_sq"))
+
 # join two dataset to get coordinate
-
 ornaments_period_join <-
@@
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
+
 # plot spatial pattern for all ornaments
-
 plot_spatial_distribution_all_ornaments_by_period <- 
@@
   labs(fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
+  scale_fill_viridis(direction = -1) +
   theme_minimal() +
@@
 knitr::include_graphics(here("analysis", 
-                             "figures", 
-                             "plot-spatial-distribution-all-ornaments-by-period.png"))
+                             "figures", "plot-spatial-distribution-all-ornaments-by-period.png"))
 ```
@@
        fill = "frequency") +
-  scale_fill_gradient(high = "black", low = "grey") +
+  scale_fill_viridis(direction = -1) +
   theme_bw() +
@@
 knitr::include_graphics(here("analysis", 
-                             "figures", 
-                             "plot-spatial-distribution-some-ornaments-by-period.png"))
+                             "figures", "plot-spatial-distribution-some-ornaments-by-period.png"))
 ```
@@
 To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
+
 ```{r function for ANN values}
@@
 # Acknowledgements
+<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
 
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
 ##### pagebreak

### 2019-08-18 analysis/paper/paper.Rmd

@@
 keywords: |
-  Historical archaeology; European colonization; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
+  Historical archaeology; Iron Age; European colonization; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
 highlights: |
-  1. Indirect impact of European colonization on indigenous peoples in the periphery of colonial control was rarely discussed.
-  2. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
-  3. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European contact.
-  4.
-Ornaments may carry an exotic and powerful image that signals wealth and trading connections with colonists that resulted in an increase in social inequality due to competition of wealth accumulation between individuals.
-  5. Ornaments can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
+  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
+  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European contact.
+  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
+  4. Trade ornament can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
 ---
@@
 # Acknowledgements
+
+I would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the ornaments used in this study.
+I thank Shui-jin Chiu, Shu-hui Jian, Jiun-hao Lai, and the staff in archaeology lab for their invaluable assistance in preparing samples for analysis.
+I also thank Yu-pei Chen for the discussion regarding the site at the earliest stage of fieldwork.
+This work was supported in part by the Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
+I especially thank Ben Marwick for providing academic support during this project, helping me with polishing figures with R langauge, and providing feedback on early drafts of this manuscript.
+
 <!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->

### 2019-08-20 analysis/paper/paper.Rmd
edits to abstract

@@
 abstract: |
-  The direct impact of European colonization since 16th century on local indigenous people has been widely discussed.
-In East Asia, direct European colonial rule on local communities was much less pronounced.
-However, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
-Using trade ornaments recovered from Kiwualn site (1400-1900 AD) in northeastern Taiwan, this research explores the indirect European colonial impacts on local societies by examining cultural context, variety, and spatial distribution of trade ornaments in terms of three time episodes relating different scales of culture contact.
-The results indicate the diversity of ornaments was greater and their spatial distribution was non-random during the European contact period that hints the increasing social inequality in a colonial context.
+  In East Asia, direct European colonial rule on local communities was much less pronounced that other parts of the world.
+Ij this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
+Using trade ornaments recovered from Kiwulan (1400-1900 AD) in northeastern Taiwan, I explore the indirect impacts of European colonial activties on local societies.
+I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three time episodes relating different scales of culture contact.
+The results indicate the diversity of ornaments was greater and their spatial distribution was clustered during the European contact period.
+This hints at increasing social inequality in a colonial context.
 This study demonstrates the use of foreign ornaments as a proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.

### 2019-08-20 analysis/paper/paper.Rmd
shorten some sentences

@@
 
-Prior to contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the 10th century, and encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
+Prior to contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the 10th century.
+Taiwan encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
 11].

### 2019-08-21 analysis/paper/paper.Rmd
add reproducibility section and deleted some shapefile data

@@
   In East Asia, direct European colonial rule on local communities was much less pronounced that other parts of the world.
-Ij this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
-Using trade ornaments recovered from Kiwulan (1400-1900 AD) in northeastern Taiwan, I explore the indirect impacts of European colonial activties on local societies.
+In this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
+Using trade ornaments recovered from Kiwulan (1400-1900 AD) in northeastern Taiwan, I explore the indirect impacts of European colonial activities on local societies.
 I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three time episodes relating different scales of culture contact.
@@
 # Excavations at Kiwulan in northeastern Taiwan
-```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", fig.width=image_width_2_col_inch}
+```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text. Map data from naturalearthdata.com", fig.width=image_width_2_col_inch}
 # need to add the locations of Keelung, Tamsui, and hoping dao
@@
 The variety of subtype and their individual frequency are summarized in Table \@ref(tab:tidy-data-for-table) in terms of three time episodes.
+
+# Reproducibility and open source materials 
+To enable re-use of materials and improve reprodicibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in XXX.XXX.
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the tests reported here.
+All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
+The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.
 
@@
 I would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the ornaments used in this study.
-I thank Shui-jin Chiu, Shu-hui Jian, Jiun-hao Lai, and the staff in archaeology lab for their invaluable assistance in preparing samples for analysis.
+I thank Shui-jin Chiu, Shu-hui Jian, Jiun-yao  Lai, and the staff in archaeology lab for their invaluable assistance in preparing samples for analysis.
 I also thank Yu-pei Chen for the discussion regarding the site at the earliest stage of fieldwork.

### 2019-08-22 analysis/paper/paper.Rmd
edited based on committee's comments

@@
 abstract: |
-  In East Asia, direct European colonial rule on local communities was much less pronounced that other parts of the world.
+  In East Asia, direct European colonial rule on local communities was much less pronounced than in other parts of the world.
 In this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
@@
 I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three time episodes relating different scales of culture contact.
-The results indicate the diversity of ornaments was greater and their spatial distribution was clustered during the European contact period.
+The results indicate the diversity of ornaments was greater during the European contact period compared to previous periods and their spatial distribution was clustered.
 This hints at increasing social inequality in a colonial context.
@@
 
-The direct impacts of European colonialism on indigenous communities in East Asia were much less pronounced than island Southeast Asia and Oceania.
+The direct impacts of European colonialism on indigenous communities in East Asia were much less \pronounced than in island Southeast Asia and Oceania.
 Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
@@
 
-Prior to contact with the Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the 10th century.
-Taiwan encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007, ch.
-11].
+Prior to contact with Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the 10th century.
+Taiwan encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007].
 The exchange network in northern Taiwan was influenced by long-distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since the 15th century due to its location on shipping routes [@Chen2005].
@@
 The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-This makes Yilan an especially unique location for exploring indirect colonial effects on local indigenous settlements because of isolation.
+This makes Yilan an especially unique location for exploring \indirect colonial effects on local indigenous settlements because of isolation.
 
@@
 
-In 1642, the Dutch VOC defeated the Spanish and took over their forts in northern Taiwan.
+In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan.
 They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty [@Andrade2007, ch.
@@
 According to those records, indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
-Among those ornaments, golden fish-shaped necklaces made of copper threads had high value due to its delicacy and the materials invested in production.
+Among those ornaments, golden fish-shaped necklaces made of copper threads had high value due to their delicacy and the materials invested in production.
 These were usually possessed by wealthy people.
@@
 They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
-The jar shapes found at Kiwulan are typical of those found elsewhere in V.O.C sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
+The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
 I used the presence and absence of An-ping jars and stonewares, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan", fig.width=image_width_1_col_inch}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 knitr::kable(ornaments_shape,
-             caption = "Ornament subtype frequencies at Kiwulan")
+             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
 # some basic stat
@@
 
-```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values with the blue line indicating the observed value"}
+```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values from 1000 simulations for three time periods.
+X values represent ANN expected value under a completely random process resulted from simulated pattern.
+Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value."}
+
 # make the blue lines thicker to be more visible
@@
 
-The results show there was a greater diversity of ornaments types and materials at Kiwulan during European contact.
+The results show there was a greater diversity of ornaments types and materials at Kiwulan during European \presence.
 This might result from the large scale exchange network that stimulated the circulation of different ornament classes.

### 2019-08-22 analysis/paper/paper.Rmd
replace culture contact with presence, arrival, or interaction, change the font color of Taiwan map

@@
 ---
-title: "Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan"
+title: "Ornaments as indicators of social change before and after European arrival at Kiwulan, Northeastern Taiwan"
 author:
@@
 Using trade ornaments recovered from Kiwulan (1400-1900 AD) in northeastern Taiwan, I explore the indirect impacts of European colonial activities on local societies.
-I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three time episodes relating different scales of culture contact.
-The results indicate the diversity of ornaments was greater during the European contact period compared to previous periods and their spatial distribution was clustered.
+I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three periods relating to different cultural interactions.
+The results indicate the diversity of ornaments was greater during the European presence period compared to previous periods and their spatial distribution was clustered.
 This hints at increasing social inequality in a colonial context.
@@
   1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
-  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European contact.
+  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European presence.
   3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
@@
 
-The direct impacts of European colonialism on indigenous communities in East Asia were much less \pronounced than in island Southeast Asia and Oceania.
+The direct impacts of European colonialism on indigenous communities in East Asia were much less conspicuous than in island Southeast Asia and Oceania.
 Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
@@
 The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
-12] to refer to small-scale regional trade between China and Taiwan on an irregular basis before European contact.
+12] to refer to small-scale regional trade between China and Taiwan on an irregular basis before European were present.
 Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
@@
 I will describe the variety and cultural context of trade ornaments in this region, and examine the potential of ornaments to inform on social changes relating to trade with colonial groups.
-I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to participation in a larger scale exchange network.
-I also predict that after European contact there will be new patterns in the spatial distributions of ornaments at Kiwulan that indicate increasing social inequality.
-I hypothesize that after 1662 when European contact declined and Chinese contact became the dominant international tradition partner, there was a decline in the use and discard of ornaments at Kiwulan due to overall population declines and acculturation in the Chinese community.
-To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European contact, and a lesser diversity after Chinese contact, I compare frequencies of ornament types between different time periods.
-To test the prediction that after European contact were new spatial patterns in the distributions of ornaments at Kiwulan, I use kernel density and point pattern analyses.
+I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European arrival due to participation in a larger scale exchange network.
+I also predict that after the European presence became established there will be new patterns in the spatial distributions of ornaments at Kiwulan that indicate increasing social inequality.
+I hypothesize that after 1662 when the European presence declined and the Chinese became the dominant international tradition partner, there was a decline in the use and discard of ornaments at Kiwulan due to overall population declines and acculturation in the Chinese community.
+To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European arrival, and a lesser diversity after the Chinese presence became established, I compare frequencies of ornament types between different time periods.
+To test the prediction that after European arrival were new spatial patterns in the distributions of ornaments at Kiwulan, I use kernel density and point pattern analyses.
 
@@
 In sum, we find multiple sources describing how local Indigenous people used ornaments in cultural contexts that represents some social roles or high status.
-However, compared to the European contact period, there are fewer mentions of beads in Chinese contact period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
+However, compared to the European period, there are fewer mentions of beads in Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
 
@@
   mutate(period = case_when(
-    `6-layers` %in% 1:2 ~ "Chinese Contact",
-    `6-layers` == 4 ~ "European Contact",
+    `6-layers` %in% 1:2 ~ "Chinese Presence",
+    `6-layers` == 4 ~ "European Presence",
     `6-layers` %in% 5:6 ~ "Before European Contact",
@@
                        level = c("Before European Contact", 
-                                 "European Contact", 
-                                 "Chinese Contact"))) %>% 
+                                 "European Presence", 
+                                 "Chinese Presence"))) %>% 
   mutate(Length = as.numeric(`Length(mm)`),
@@
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates [@Chen2007].
-This paper focuses on the Upper Layer Culture spanning from the late Iron Age to the historical period starting from the 17th century with European contact in Taiwan.
+This paper focuses on the Upper Layer Culture spanning from the late Iron Age to the historical period starting from the 17th century with the European presence in Taiwan.
 
@@
 However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
-For each sampled square I separated excavation units into three episodes of time: pre-European contact period (L5 and L6), European contact period (L4), and Chinese contact period (L2 and L1).
+For each sampled square I separated excavation units into three episodes of time: pre-European period (L5 and L6), European period (L4), and Chinese period (L2 and L1).
 
-I identify the start of European contact at Kiwulan at 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
+I identify the first European arrival at Kiwulan at 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
 163].
@@
 12].
-In the archaeological record, the approximate indicators of the start of European contact at Kiwulan are the appearance of An-ping jars and stonewares (martavans/martaban) that were largely introduced to Taiwan during the early 17th century.
+In the archaeological record, the approximate indicators of the start of the European influence at Kiwulan are the appearance of An-ping jars and stonewares (martavans/martaban) that were largely introduced to Taiwan during the early 17th century.
 They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
@@
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The archaeological signature of the Chinese contact period at Kiwulan is the large amount and varieties of Chinese porcelains, clay stoves, and a few opium pipe-bowls that were identified as items commonly used in the daily life of Chinese and the use of bricks and tiles for architectural buildings [@Hsieh2009].
+The archaeological signature of the Chinese period at Kiwulan is the large amount and varieties of Chinese porcelains, clay stoves, and a few opium pipe-bowls that were identified as items commonly used in the daily life of Chinese and the use of bricks and tiles for architectural buildings [@Hsieh2009].
 Also, the description of the Chinese migration to Yilan can be seen from official Chinese records written in the early 19th century [@Chen1963; @Ke1993].
@@
                       "unclassified")) %>% 
-  arrange(Categories, desc(`European Contact`))
+  arrange(Categories, desc(`European Presence`))
 # making table
@@
 It indicates that most ornament types were present before European contact.
-Ornament frequencies reached a peak during European contact and then dropped during the Chinese contact period, especially for golden beads.
+Ornament frequencies reached a peak during European period and then dropped during the Chinese period, especially for golden beads.
 This trend can be also seen on other ornaments including agate beads, metal rings, and bells.
-However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European contact period and a further decrease in the Chinese contact period.
+However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period.
 
@@
 
-```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across contact periods for metal rings, agate beads, bells, and glass beads."}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, agate beads, bells, and glass beads."}
 library(cowplot)
@@
 Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
-This indicates that the increases in diversity can be explained by the effects of culture contact instead of effects of sample size.
-According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period.
+This indicates that the increases in diversity can be explained by the effects of culture interaction instead of effects of sample size.
+According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
 The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the Europeans.
@@
 Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from the research area for each time period.
-Before European contact, a greater amount of ornaments were found at the northern and middle parts of the research area.
-During the European contact period, ornaments were more widespread, with some clusters on the northern part.
-During the Chinese contact period the distribution is more even again.
-Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European contact, such as golden beads and agate beads.
+Before the European arrival, a greater amount of ornaments were found at the northern and middle parts of the research area.
+During the European period, ornaments were more widespread, with some clusters on the northern part.
+During the Chinese period the distribution is more even again.
+Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European period, such as golden beads and agate beads.
 However, there seems to be no clear consistent pattern across those different ornaments.
@@
 Copper bells were usually found individually and seem randomly distributed across the area.
-In the Chinese Contact period, both the amount and density of different classes of ornaments decreased.
+In the Chinese period, both the amount and density of different classes of ornaments decreased.
 
@@
   set_names(c("Before European Contact",
-              "European Contact",
-              "Chinese Contact"))
+              "European Presence",
+              "Chinese Presence"))
 ```
@@
 Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
-Hotspots in the pattern of ornaments could indicate changes in social organization due to culture contact.
+Hotspots in the pattern of ornaments could indicate changes in social organization due to culture interaction.
 To explore ornament distributions I used the intensity approach that focuses on the average density of points across space.
@@
 
-The results show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period (Figure \@ref(fig:plot-all-densities)).
+The results show that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period (Figure \@ref(fig:plot-all-densities)).
 Those maps present three consistent sub-regions with a core area that shifts over time.
 The distribution might indicate different social groups who possessed more ornaments.
-The multiple groups during European contact period might reflect more unequal consumption of ornaments across the site, relative to other periods.
+The multiple groups during European period might reflect more unequal consumption of ornaments across the site, relative to other periods.
 However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
@@
 Figure \@ref(fig:hypo-test-distribution) shows the distributions of the ANN distances calculated on 1000 simulations of ornament locations.
-The results show that 100% of the simulated values are much greater than our observed ANN value during the European contact period, which means the ornaments have non-random clustered distributions.
+The results show that 100% of the simulated values are much greater than our observed ANN value during the European period, which means the ornaments have non-random clustered distributions.
 A similar, but less extreme, result is also observed during the pre-European period.
-The observed distribution of ornaments is more similar to the random distributions during the Chinese contact period, with about one third of the simulated values are greater than our observed ANN value.
-This testing reveals that the clustered distributions of ornaments during the pre-European and European contact periods might reflect the presence of different social groups.
-Moreover, the clustering of ornaments during the European contact period is highly non-random, potentially indicating that some of the social changes leading to a concentration of power to control the distribution of ornaments at Kiwulan.
+The observed distribution of ornaments is more similar to the random distributions during the Chinese period, with about one third of the simulated values are greater than our observed ANN value.
+This testing reveals that the clustered distributions of ornaments during the pre-European and European periods might reflect the presence of different social groups.
+Moreover, the clustering of ornaments during the European period is highly non-random, potentially indicating that some of the social changes leading to a concentration of power to control the distribution of ornaments at Kiwulan.
 
@@
 
-The results show there was a greater diversity of ornaments types and materials at Kiwulan during European \presence.
+The results show there was a greater diversity of ornaments types and materials at Kiwulan during the European presence.
 This might result from the large scale exchange network that stimulated the circulation of different ornament classes.
-However, the frequency of overall ornaments and each subtype declines significantly during the Chinese contact period.
+However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period.
 This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the mortuary practices of Han Chinese.
@@
 
-Spatial patterns of ornaments shows that their distribution was clustered during the pre-European and European contact period.
-Hypothesis testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European contact period.
-Because ethnographic evidence indications that trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering may indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European contact period.
+Spatial patterns of ornaments shows that their distribution was clustered during the pre-European and European period.
+Hypothesis testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European period.
+Because ethnographic evidence indications that trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering may indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European period.
 In addition, there was a burial dated to 17th century unearthed 60 golden beads above the average of 2-3 pieces in the pre-European period, further indicating increasing social inequality [@Chen2007; @Cheng2008].
@@
 
-We may consider that Yilan social organization went from corporate mode in the pre-European contact period, then transitioned into network mode during the European colonial contact period, then back to corporate mode in the Chinese contact period.
+We may consider that Yilan social organization went from corporate mode before the European arrival, then transitioned into network mode during the European presence, then back to corporate mode in the Chinese period.
 The small number of ornaments, and less concentrated distribution during the pre-European period is consistent with shared power and wealth of the corporate mode.
@@
 Because of the weak direct control from the European colonizers in northeastern Taiwan, it implies that local leaders might have flexibility to manipulate European colonial image, expand personal power, and monopolize the high-value trade goods [@Kang2012].
-During the Chinese contact period, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
+During the Chinese period, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
 This reflects the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals accumulate wealth.
@@
 Kiwulan in Northeastern Taiwan is an important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
-Kiwulan can provide valuable insights into the discussion of the influence of indirect colonial contacts.
-The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods.
-The greater amount and diversity of ornament types during the European contact period reflects international exchanges in a colonial context.
+Kiwulan can provide valuable insights into the discussion of indirect colonial influence.
+The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture interaction periods.
+The greater amount and diversity of ornament types during the European period reflects international exchanges in a colonial context.
 Those ornaments may carry an exotic and powerful image that signals wealth and trading connections for the inhabitants of Kiwulan.
@@
 
-# Acknowledgements
+# Acknowledgments
 
@@
 This work was supported in part by the Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
+I would also like to thank Ben Fitzhugh and Peter Lape, and anonymous reviewers for their insightful comments on early drafts.
 I especially thank Ben Marwick for providing academic support during this project, helping me with polishing figures with R langauge, and providing feedback on early drafts of this manuscript.

### 2019-08-22 analysis/paper/map.R

@@ -34,8 +34,13 @@ ggplot(data = world) +
   geom_sf() +
   geom_rect(xmin = 120.7, xmax = 122.2, ymin = 24.2, ymax = 25.5,
             fill = NA, colour = "red", size = 0.5) +
-  geom_shadowtext(data= world_points, aes(x = X, y = Y, label = name),
-                  size = 4, position = position_nudge(y = - 1.7)) +
+  geom_shadowtext(data= world_points,
+                  aes(x = X, y = Y,
+                      label = name),
+                  color='black',
+                  bg.colour='white',
+                  size = 4,
+                  position = position_nudge(y = - 1.7, x = 0.5)) +
   coord_sf(xlim = c(100, 142), ylim = c(7, 43), expand = FALSE) + #add datum = NA to remove
   scale_x_continuous(breaks = seq(100, 140, by = 20)) +
   scale_y_continuous(breaks = seq(15, 40, by = 10)) +

### 2019-08-26 analysis/paper/paper.Rmd
some edits based on committee's suggestions

@@
 abstract: |
-  In East Asia, direct European colonial rule on local communities was much less pronounced than in other parts of the world.
+  In East Asia, direct European colonial rule on local communities was much less conspicuous than in other parts of the world.
 In this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
@@
 These remains indicate a direct impact on the local indigenous populations through their modification of the local landscape.
-In contrast to Heping Dao, Yilan in Northeastern Taiwan is physically isolated and there are no traces of European settlement.
+In contrast to Heping Dao, Yilan in Northeastern Taiwan is a physically isolated area adjacent to colonial centers and there are no traces of European settlement but occasional interactions with the Europeans.
 Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
 The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-This makes Yilan an especially unique location for exploring \indirect colonial effects on local indigenous settlements because of isolation.
+This makes Yilan an especially unique location for exploring indirect colonial effects on local indigenous settlements living beyond the borders of direct colonial occupation.
 
@@
 However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
-For each sampled square I separated excavation units into three episodes of time: pre-European period (L5 and L6), European period (L4), and Chinese period (L2 and L1).
+For each sampled square I separated excavation units into three episodes of time: pre-European period (L5 and L6), European period (L4), and Chinese period (L2 and L1) to indicate the dominant foreign influence and cultural interaction for different time periods.
 
@@
 Burials unearthed a large amount of ornaments in the form of bead strings that consisted of hundreds to thousands of pieces.
+The ornaments from the burial context are not analyzed here since not many burials can be assigned to a time period with confidence.
 In this paper I focus on the post hole areas where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
@@
 # Reproducibility and open source materials
+
 To enable re-use of materials and improve reprodicibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in XXX.XXX.
@@
 
-```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods"}
+```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods. Used the bandwidth based on Silverman(1986)'s rule of thumb."}
 # apply function to each data frame in the list
@@
 Kernel density estimations (KDE) estimate the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
-Here we use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
+Here I use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
 Density values of artifacts per square meter were calculated for each cell.
 
-The results show that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period (Figure \@ref(fig:plot-all-densities)).
-Those maps present three consistent sub-regions with a core area that shifts over time.
-The distribution might indicate different social groups who possessed more ornaments.
-The multiple groups during European period might reflect more unequal consumption of ornaments across the site, relative to other periods.
-However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
+The results show that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period according to the kernal density maps (Figure \@ref(fig:plot-all-densities)).
+It seems that there are three consistent sub-regions with a core area that shifts over time.
+The distribution might indicate different household-based social groups who possessed more ornaments.
+The multiple groups during European period might reflect more unequal consumption of ornaments across the site, relative to other periods, to indicate social processes caused by human behavior.
+However, it could also just a random pattern resulted from bigger sample size.
+In addition, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
 To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
@@
 This testing reveals that the clustered distributions of ornaments during the pre-European and European periods might reflect the presence of different social groups.
-Moreover, the clustering of ornaments during the European period is highly non-random, potentially indicating that some of the social changes leading to a concentration of power to control the distribution of ornaments at Kiwulan.
+Moreover, the clustering of ornaments during the European period is highly non-random, might potentially indicating that some of the social changes leading to different degrees of access foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan.
 
@@
 Kiwulan in Northeastern Taiwan is an important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
-Kiwulan can provide valuable insights into the discussion of indirect colonial influence.
-The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture interaction periods.
+Kiwulan can provide valuable insights into the discussion of indirect colonial influence on the local societies living beyond the borders of direct European colonial occupation.
+The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
 The greater amount and diversity of ornament types during the European period reflects international exchanges in a colonial context.
-Those ornaments may carry an exotic and powerful image that signals wealth and trading connections for the inhabitants of Kiwulan.
+Ornaments were viewed as prestige goods in local indigenous culture before European contact.
+After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may be intensified that further signals wealth and privileged trading connections with European colonists for the inhabitants of Kiwulan.
 These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
@@
 Ornaments can give insights into the amplification of social inequality stimulated by European colonization.
-It also shows the agency of indigenous people to incorporate ornaments into their social system.
+It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
 Future work could extend this approach to studies of other trade goods such as ceramics.
@@
 This work was supported in part by the Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
-I would also like to thank Ben Fitzhugh and Peter Lape, and anonymous reviewers for their insightful comments on early drafts.
+I would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on the early draft.
 I especially thank Ben Marwick for providing academic support during this project, helping me with polishing figures with R langauge, and providing feedback on early drafts of this manuscript.

### 2019-08-27 analysis/paper/paper.Rmd
typo

@@
 
-```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods. Used the bandwidth based on Silverman(1986)'s rule of thumb."}
+```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb."}
 # apply function to each data frame in the list

### 2019-08-27 analysis/paper/paper.Rmd
added DOI

--- 2019-08-26 analysis/paper/paper.Rmd
@@
 
-To enable re-use of materials and improve reprodicibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in XXX.XXX.
+To enable re-use of materials and improve reprodicibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in DOI 10.17605/OSF.IO/R8YGA.
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the tests reported here.
@@
 I would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on the early draft.
-I especially thank Ben Marwick for providing academic support during this project, helping me with polishing figures with R langauge, and providing feedback on early drafts of this manuscript.
+I especially thank Ben Marwick for providing academic support during this project, helping me with polishing figures with R language, and providing feedback on early drafts of this manuscript.
 

### 2019-08-27 analysis/paper/paper.Rmd

@@
 The multiple groups during European period might reflect more unequal consumption of ornaments across the site, relative to other periods, to indicate social processes caused by human behavior.
-However, it could also just a random pattern resulted from bigger sample size.
-In addition, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
-To further evaluate the pattern, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
+However, it could also just a random pattern resulting from bigger sample size.
+In addition, the generation of core areas might be biased due to small sample sizes, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
+To further evaluate these patterns, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
 

### 2019-09-19 analysis/paper/paper.Rmd
shorten the abstract, change the some words to british spelling, and change format of figures to JEPG

@@
 bibliography: references.bib
-csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
+csl: "../templates/antiquity.csl" # Insert path for the bib-style
 abstract: |
-  In East Asia, direct European colonial rule on local communities was much less conspicuous than in other parts of the world.
-In this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered.
-Using trade ornaments recovered from Kiwulan (1400-1900 AD) in northeastern Taiwan, I explore the indirect impacts of European colonial activities on local societies.
-I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three periods relating to different cultural interactions.
-The results indicate the diversity of ornaments was greater during the European presence period compared to previous periods and their spatial distribution was clustered.
-This hints at increasing social inequality in a colonial context.
-This study demonstrates the use of foreign ornaments as a proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
-Ornaments can give insights into the amplification of social inequality stimulated by European colonization.
-It also shows the agency of indigenous people to incorporate ornaments into their social system.
+  Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
+Trade ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
+The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered.
+This hints at increasing social inequality resulting from a colonial influence.
+Ornaments give insights into the increasing social inequality stimulated by European colonisation, and show the agency of indigenous people to incorporate ornaments into their social system.
 keywords: |
-  Historical archaeology; Iron Age; European colonization; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
+  Historical archaeology; Iron Age; European colonisation; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
 highlights: |
@@
 )
-# prepare for JAS:R
-# https://www.elsevier.com/journals/journal-of-archaeological-science-reports/2352-409x/guide-for-authors#2000
+# prepare for Antiquity https://antiquity.ac.uk/submit-paper
+# should be no more than 13.5cm wide and at a minimum resolution of 300dpi. 
 # not exceed 6000 words excluding figures, tables, highlights 
@@
 Taiwan became part of global trade systems during the early 17th century due to western European expansions that connected Europe and Asia.
-Despite the colonization of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonization.
+Despite the colonisation of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonisation.
 @Berrocal2018 examined both archaeological and historical records of Heping Dao in northern Taiwan.
@@
 These remains indicate a direct impact on the local indigenous populations through their modification of the local landscape.
-In contrast to Heping Dao, Yilan in Northeastern Taiwan is a physically isolated area adjacent to colonial centers and there are no traces of European settlement but occasional interactions with the Europeans.
+In contrast to Heping Dao, Yilan in Northeastern Taiwan is a physically isolated area adjacent to colonial centres and there are no traces of European settlement but occasional interactions with the Europeans.
 Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
@@
 Other people wore agate beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonization, a field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of Japanese colonisation, a field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
@@
 library(here)
-knitr::include_graphics(here("analysis", "figures", "kiwulan-location-map.png"))
+knitr::include_graphics(here("analysis", "figures", "kiwulan-location-map.jpg"))
 ```
@@
 ggsave(plot = KWL_excavation_map, 
-       here("analysis", "figures", "KWL-excavation-map.png"), width = 190, units = "mm")
+       here("analysis", "figures", "KWL-excavation-map.jpg"), 
+       width = 135,
+       dpi = 300,
+       units = "mm")
 
-knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.png"))
+knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.jpg"))
 ```
@@
 The total area includes eight open area sections of 262 excavation squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
-The archaeological evidence includes a large amount of artifacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
+The archaeological evidence includes a large amount of artefacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
 Most ornaments were found in situ while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
@@
 
-European colonization ended in 1662 when they were defeated by the kingdom of Tungning founded by Koxinga from China.
+European colonisation ended in 1662 when they were defeated by the kingdom of Tungning founded by Koxinga from China.
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artefact counts", fig.width=image_width_1_col_inch}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
        here("analysis", "figures",
-       "plot-ornaments-count.png"),
+       "plot-ornaments-count.jpg"),
        width = 90, 
        height = 40,
+       dpi = 300,
        units = "mm")
@@
                              "figures", 
-                             "plot-ornaments-count.png"))
+                             "plot-ornaments-count.jpg"))
 ```
@@
 knitr::kable(ornaments_shape,
-             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
+             caption = "Ornament subtype at Kiwulan. The numbers represent artefact counts ")
 # some basic stat
@@
 In this paper I focus on the post hole areas where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
-Post hole areas represents domestic spaces suitable for testing for changes in the patterns of distribution relating to social organization.
+Post hole areas represents domestic spaces suitable for testing for changes in the patterns of distribution relating to social organisation.
 Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells (see also Figure 4).
@@
 
-To enable re-use of materials and improve reprodicibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in DOI 10.17605/OSF.IO/R8YGA.
+To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualisations contained in this paper is included in http://doi.org/10.17605/OSF.IO/R8YGA.
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the tests reported here.
@@
        y = "Frequency") +
-  theme_minimal(base_size = 12)
+  theme_minimal(base_size = 10)
 
@@
        here("analysis", "figures",
-       "plot-ornaments-count-three-periods.png"),
-       width = 190, 
-       height = 80,
+       "plot-ornaments-count-three-periods.jpg"),
+       width = 135, 
+       dpi = 300,
+       height = 50,
        units = "mm")
@@
                              "figures", 
-                             "plot-ornaments-count-three-periods.png"))
+                             "plot-ornaments-count-three-periods.jpg"))
 ```
@@
          Type != 'other')
-base_size_multi_plot <- 10
+base_size_multi_plot <- 8
 agate_bead_shape_plot <- 
@@
        here("analysis", "figures",
-       "plot-multi-types-by-period.png"),
-       width = 190, 
-       height = 150,
+       "plot-multi-types-by-period.jpg"),
+       width = 135, 
+       dpi = 300,
+       height = 95,
        units = "mm")
@@
                              "figures", 
-                             "plot-multi-types-by-period.png"))
+                             "plot-multi-types-by-period.jpg"))
 ```
@@
   add_count() %>% 
-  select(period, n, sam_size) %>% 
+  dplyr::select(period, n, sam_size) %>% 
   distinct()
@@
   add_count() %>% 
-  select(diversity, n, sam_size) %>% 
+  dplyr::select(diversity, n, sam_size) %>% 
   separate(diversity, c("period","subtype"), sep = "-", remove = FALSE) %>% 
@@
 spearman_result_rho <- unname(round(spearman_result$estimate, 2))
-spearman_result_p_value <- round(spearman_result$p.value, 2)
+spearman_result_p_value <- round(spearman_result$p.value, 5)
 ```
@@
         axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 11))
+        strip.text.x = element_text(size = 6),
+        legend.text = element_text(size = 6),
+        legend.title = element_text(size = 6),
+        legend.key.size =  unit(3, "mm"))
 
@@
        here("analysis", "figures",
-       "plot-spatial-distribution-all-ornaments-by-period.png"),
-       width = 190, 
-       height = 50,
+       "plot-spatial-distribution-all-ornaments-by-period.jpg"),
+       width = 135, 
+       height = 31,
+       dpi = 300,
        units = "mm")
 knitr::include_graphics(here("analysis", 
-                             "figures", "plot-spatial-distribution-all-ornaments-by-period.png"))
+                             "figures", "plot-spatial-distribution-all-ornaments-by-period.jpg"))
 ```
@@
   scale_fill_viridis(direction = -1) +
-  theme_bw() +
+  theme_minimal() +
   theme(plot.title = element_text(hjust = 0.5),
@@
         axis.text.y = element_blank(),
-        strip.text.x = element_text(size = 8))
+        strip.text.x = element_text(size = 6))
 
@@
        here("analysis", "figures",
-       "plot-spatial-distribution-some-ornaments-by-period.png"),
-       width = 190, 
-       height = 130,
+       "plot-spatial-distribution-some-ornaments-by-period.jpg"),
+       width = 135, 
+       height = 82,
+       dpi = 300,
        units = "mm")
@@
 knitr::include_graphics(here("analysis", 
-                             "figures", "plot-spatial-distribution-some-ornaments-by-period.png"))
+                             "figures", "plot-spatial-distribution-some-ornaments-by-period.jpg"))
 ```
@@
   contour(kernel_tmp, add = TRUE)
-  title(..., line = -2, adj = 0.4)
+  title(..., line = -0.3, adj = 0.4, cex.main = 0.8)
 return(list(ornaments_point_ppp = ornaments_point_ppp_tmp,
@@
 
-```{r plot-all-densities, fig.cap= "Kernal density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb."}
+```{r plot-all-densities, fig.cap= "Kernel density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb."}
 # apply function to each data frame in the list
 titles <- names(ornaments_period_join_cen)
-png(
+jpeg(
   here("analysis", "figures",
-       "plot-kde-maps.png"), 
-    width = 190,
-    height = 63,
+       "plot-kde-maps.jpg"), 
+    width = 135,
+    height = 40,
     units = "mm",
@@
                              "figures", 
-                             "plot-kde-maps.png"))
+                             "plot-kde-maps.jpg"))
 ```
@@
 I used point pattern analysis to examine whether the distribution of ornaments was uneven in non-random ways that could reflect changes in social inequality.
-Point pattern analysis is a way to assess whether the distribution of sites or artifacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
-Hotspots in the pattern of ornaments could indicate changes in social organization due to culture interaction.
+Point pattern analysis is a way to assess whether the distribution of sites or artefacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
+Hotspots in the pattern of ornaments could indicate changes in social organisation due to culture interaction.
 To explore ornament distributions I used the intensity approach that focuses on the average density of points across space.
@@
 Here I use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
-Density values of artifacts per square meter were calculated for each cell.
+Density values of artefacts per square meter were calculated for each cell.
 
-The results show that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period according to the kernal density maps (Figure \@ref(fig:plot-all-densities)).
+The results show that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period according to the kernel density maps (Figure \@ref(fig:plot-all-densities)).
 It seems that there are three consistent sub-regions with a core area that shifts over time.
@@
 # make the blue lines thicker to be more visible
-png(
+jpeg(
   here("analysis", "figures",
-       "plot-kde-ann-histograms.png"), 
+       "plot-kde-ann-histograms.jpg"), 
     width = 90,
@@
                              "figures", 
-                             "plot-kde-ann-histograms.png"))
+                             "plot-kde-ann-histograms.jpg"))
 ```
@@
 
-We may consider that Yilan social organization went from corporate mode before the European arrival, then transitioned into network mode during the European presence, then back to corporate mode in the Chinese period.
+We may consider that Yilan social organisation went from corporate mode before the European arrival, then transitioned into network mode during the European presence, then back to corporate mode in the Chinese period.
 The small number of ornaments, and less concentrated distribution during the pre-European period is consistent with shared power and wealth of the corporate mode.
@@
 
-Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017].
+Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by the colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017].
 Kiwulan in Northeastern Taiwan is an important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
@@
 This study demonstrates that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
-Ornaments can give insights into the amplification of social inequality stimulated by European colonization.
+Ornaments can give insights into the amplification of social inequality stimulated by European colonisation.
 It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
 Future work could extend this approach to studies of other trade goods such as ceramics.
-We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardization, and mortuary practices.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardisation, and mortuary practices.
 

### 2019-09-21 analysis/paper/paper.Rmd
reducing word count for resubmission

@@
   comment = "#>",
+  cache = TRUE,
   fig.path = "../figures/"
@@
 Understanding the indirect effects of colonialism are important for detecting colonial impacts on indigenous peoples in the periphery of colonial control [@Trabert2017].
-In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005].
-Studying the consumption of foreign goods can improve our understanding of the negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Rubertone2000; @Scaramelli2005; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005].
+In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Silliman2005].
+Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
+Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in Taiwan, the northeast region was isolated from intensive direct contact by the Xueshan Mountains.
 
-Prior to contact with Europeans in the early 17th century, Taiwan had been involved in both regional and long-distance East and Southeast Asian trade systems since the 10th century.
-Taiwan encountered large waves of Han Chinese immigration after 17th century when Taiwan was incorporated into the territories of Qing Dynasty of China [@Liu2017; @Andrade2007].
-The exchange network in northern Taiwan was influenced by long-distance trade between Fuzhou or Quanzhou in China and Ryukyu in Japan since the 15th century due to its location on shipping routes [@Chen2005].
-This intensified in the 17th century with the arrival of the Europeans and their trade goods [@Wang2007].
+This article describes ornaments excavated from Kiwulan (700-1900 AD), the largest Iron Age settlement on the Yilan plain in northeast Taiwan.
+The first recorded European presence in Yilan was a Spanish revenge attack on indigenous villages in 1632 [@Borao2001, pp.
+163].
+In 1647 the Dutch attacked villages and forced them to accept colonial rules and pay an annual tribute [@Andrade2007].
+According to Dutch census reports in 1650, Kiwulan was the largest indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
+12].
+Following defeat of Dutch by the Chinese general Koxinga in 1661-2, the Dutch abandoned northern Taiwan.
+Direct contact with Han Chinese is indicated by Qing dynasty census reports mentioning Yilan villages in 1821 [@Yao1996].
 
-Taiwan became part of global trade systems during the early 17th century due to western European expansions that connected Europe and Asia.
-Despite the colonisation of Taiwan by the Spanish and Dutch from 1624 to 1662, there has been little discussion of the archaeology of indigenous responses to the European colonisation.
-@Berrocal2018 examined both archaeological and historical records of Heping Dao in northern Taiwan.
-This was the location of the colony of San Salvador founded by the Spanish in 1626 and taken over by the Dutch in 1642.
-Berrocal et al.
-uncovered parts of the foundation of a Spanish church and a European cemetery that indicate the evidence of European occupation.
-These remains indicate a direct impact on the local indigenous populations through their modification of the local landscape.
-In contrast to Heping Dao, Yilan in Northeastern Taiwan is a physically isolated area adjacent to colonial centres and there are no traces of European settlement but occasional interactions with the Europeans.
-Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides.
-The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012].
-This makes Yilan an especially unique location for exploring indirect colonial effects on local indigenous settlements living beyond the borders of direct colonial occupation.
-
-Although mountains isolated Yilan by land, rivers and seas provided a way for the interaction between local indigenous settlements and further afield communities, such as the Basai in northern Taiwan, and Han Chinese.
-The indigenous people of Yilan established their own exchange network that has been described as part of an "inter-insular trade" system [@Chen2005, p.
-12] to refer to small-scale regional trade between China and Taiwan on an irregular basis before European were present.
-Kavalan people offered rice, deer hides, and gold in exchange for beads, metal tools or ornaments, ironwares, porcelains, and textiles with outside traders [@Chen2005; @Hsieh2009; @LiandWu2006].
-Trading activities brought a wide variety of objects into indigenous communities in Yilan.
-One of the most commonly traded types of object in this region were personal ornaments such as glass and stone beads.
-These are one of the most abundant types of foreign goods found at the settlement sites in this region [@Chen2007; @Li2014; @NMTH2005].
-Personal adornments in the archaeological record are useful as social signal of an individual's status [@Joyce2005; @Scaramelli2005].
+One of the most commonly traded types of object in this region were personal ornaments such as glass and stone beads [@Chen2007; @Li2014; @NMTH2005].
+Personal adornments in the archaeological record are useful as signal of an individual's status [@Joyce2005; @Scaramelli2005].
 For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
-
-In this paper, I explore archaeological evidence from Yilan spanning from the late Iron Age to the historical period to address the question of whether indirect colonial influences on the indigenous populations can be detected.
-I focus on the ornaments recovered from the archaeological excavations at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan [@Chen2007].
-Northeastern Taiwan provides a good example to explore the indirect colonial impacts on indigenous societies by comparing archaeological evidence from different episodes of culture contact [cf.
-@Lape2003].
-I will describe the variety and cultural context of trade ornaments in this region, and examine the potential of ornaments to inform on social changes relating to trade with colonial groups.
-I hypothesize that there was a greater diversity of ornaments types and materials at Kiwulan after European arrival due to participation in a larger scale exchange network.
-I also predict that after the European presence became established there will be new patterns in the spatial distributions of ornaments at Kiwulan that indicate increasing social inequality.
-I hypothesize that after 1662 when the European presence declined and the Chinese became the dominant international tradition partner, there was a decline in the use and discard of ornaments at Kiwulan due to overall population declines and acculturation in the Chinese community.
-To test the hypothesis of a greater diversity of ornaments types and materials at Kiwulan after European arrival, and a lesser diversity after the Chinese presence became established, I compare frequencies of ornament types between different time periods.
-To test the prediction that after European arrival were new spatial patterns in the distributions of ornaments at Kiwulan, I use kernel density and point pattern analyses.
+In this paper, I explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the period of Chinese presence.
+I address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
 
@@
 
-Historical documents from the Spanish and Dutch in the early 17th century provide some information about the social life and culture of indigenous people in Northern Taiwan when the Europeans arrived.
 The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
 343].
-A report by Dominican priest Fr.
-Jacinto Esquivel in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, usually used cuentas (agate beads) in exchange for necessities with other indigenous groups.
+A report by a Dominican priest in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, exchanged agate beads with other indigenous groups.
 This form of exchange was widespread and even the Spanish soldiers learned to use agate as bargaining chips for gambling [@LiandWu2006, pp.
 132-149].
-Esquivel also recorded how indigenous people viewed agate pieces as prestige goods in their culture.
-An indigenous man who wanted to get married had to pay agate beads to the parents of his future wife.
-Also, agate or golden beads could be used to resolve conflicts in their daily lives [@LiandWu2006, pp.
-151].
+The use of beads as prestige goods is indicated by their role in bride price payments, and compensation to resolve disputes [@LiandWu2006, pp.
+132-149].
 Other records mention that the female shamans (majuorbol) in the tribe would use agate beads as magical items in ritual practice for body healing [@Borao2009, pp.
@@
 153].
-These historical accounts indicate that agate beads or golden beads were already treated as prestige goods before the arrival of the Europeans.
-These were likely introduced earlier from Chinese traders.
+These historical accounts indicate that agate beads were already treated as prestige goods in Yilan before the arrival of Europeans.
 
@@
 4].
-The Dutch provided beads and other goods based on the demands of Indigenous people to secure alliances in the annual ceremony or during traveling [@Kang2016, ch.
+The Dutch provided beads and other goods based on the demands of Indigenous people to secure alliances in the annual ceremony or during travelling [@Kang2016, ch.
 6].
-Thus, I assume that the Dutch feudal system resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.
+We might predict that the Dutch feudal system resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.
 
-Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty [1616-1911] contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
+Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty (1616-1911) contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
 228, 308; @Ke1993, pp.
@@
 Other people wore agate beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonisation, a field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of Japanese colonisation, an academic field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
 
-Recent ethnographic research with the living members of the Kavalan tribe describes how agate beads were used in divination practices, called "subli", by female shamans [@Liu2008, pp.
+Recent ethnographic research with the members of the Kavalan tribe, the contemporary indigenous people of the Yilan plain, describes how agate beads were used in divination practices by female shamans [@Liu2008, pp.
 133-134].
-Interview records mention agate beads as valuable objects that are usually passed down from mother to daughter as heirlooms in shaman families, without knowing the exact origin.
+Interview records mention agate beads as valuable objects that are usually passed down from mother to daughter as heirlooms in shaman families.
 When not in use, beads were often hidden in safe places, such as on the roof beams.
-Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination, consistent with observations from the European colonial period.
-Despite most Kavalan people today not knowing the specific origin of their agate beads, the high value of beads is still indicated in oral history and by their scarcity.
-In sum, we find multiple sources describing how local Indigenous people used ornaments in cultural contexts that represents some social roles or high status.
+Despite most Kavalan people today not knowing the specific origin of their agate beads, the high value of beads is still indicated in oral history.
+In sum, we find multiple sources describing how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
 However, compared to the European period, there are fewer mentions of beads in Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
 
-# Excavations at Kiwulan in northeastern Taiwan
+# Excavations at Kiwulan in northeast Taiwan
+
 ```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text. Map data from naturalearthdata.com", fig.width=image_width_2_col_inch}
@@
 
-Kiwulan (Figure \@ref(fig:KWL-map)) is located at northern Yilan and was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
+Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
 The total area includes eight open area sections of 262 excavation squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
@@
 
-The chronology of Kiwulan can be divided into two phases represented by a Lower Layer Culture (700 - 1200 AD) and a Upper Layer Culture (1400 - 1900 AD) with a sterile layer in between.
-This chronology is based on the stratigraphy and a series of 32 radiocarbon dates [@Chen2007].
-This paper focuses on the Upper Layer Culture spanning from the late Iron Age to the historical period starting from the 17th century with the European presence in Taiwan.
+The chronology of Kiwulan can be divided into two phases represented by a Lower Layer Culture (700-1200 AD) and a Upper Layer Culture (1400-1900 AD) separated by a sterile.
+This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
+This paper focuses on the Upper Layer Culture because this has a sequence spanning pre-European contact, European presence, and Chinese presence.
+Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
+Due to ambiguities in the previous chronology, I re-examined the excavation records to produce a new chronology for the sampled squares.
+For each sampled square I separated excavation units into three episodes of time: pre-European period, European period, and Chinese period to indicate the dominant foreign influence and cultural interaction for different time periods.
 
-Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
-The layers were assigned numbers from 1 to 6 where L1 corresponds to the youngest layer and L6, the bottom layer, refers to the oldest layer.
-However, because the specific methods of assigning layers to a time period was not documented in detail for the previous chronology, I reexamined the excavation records to produce a new chronology for the sampled squares.
-For each sampled square I separated excavation units into three episodes of time: pre-European period (L5 and L6), European period (L4), and Chinese period (L2 and L1) to indicate the dominant foreign influence and cultural interaction for different time periods.
-
-I identify the first European arrival at Kiwulan at 1632, when the local villages were attacked by the Spanish who took revenge on an incident happened earlier that year [@Borao2001, pp.
-163].
-Later in 1647, the Dutch attacked the indigenous villages and forced them to accept colonial rules and economic demands by paying annual tribute [@Andrade2007].
-According to Dutch census reports in 1650, Kiwulan was the largest indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
-12].
-In the archaeological record, the approximate indicators of the start of the European influence at Kiwulan are the appearance of An-ping jars and stonewares (martavans/martaban) that were largely introduced to Taiwan during the early 17th century.
+The archaeological indicators of the start of the European influence at Kiwulan are the appearance of An-ping jars and stonewares (martavans/martaban) that were introduced to Taiwan during the early 17th century.
 They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
@@
 
-European colonisation ended in 1662 when they were defeated by the kingdom of Tungning founded by Koxinga from China.
+The European colonial presence in northern Taiwan ended in 1662 when the Dutch were defeated by the kingdom of Tungning founded by Koxinga from China.
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The archaeological signature of the Chinese period at Kiwulan is the large amount and varieties of Chinese porcelains, clay stoves, and a few opium pipe-bowls that were identified as items commonly used in the daily life of Chinese and the use of bricks and tiles for architectural buildings [@Hsieh2009].
-Also, the description of the Chinese migration to Yilan can be seen from official Chinese records written in the early 19th century [@Chen1963; @Ke1993].
+The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains, clay stoves.
+Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles [@Hsieh2009].
+Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century [@Chen1963; @Ke1993].
 
@@
 
-Ornaments from 40 sampling squares were excavated in-situ in different archaeological contexts, including 406 pieces from post-hole areas, 3173 pieces from 30 burials, and 27 pieces from middens.
-Burials unearthed a large amount of ornaments in the form of bead strings that consisted of hundreds to thousands of pieces.
-The ornaments from the burial context are not analyzed here since not many burials can be assigned to a time period with confidence.
-In this paper I focus on the post hole areas where `r sam_post_hole` ornaments were recovered, which accounts for `r sample_per` % of the total number from the same context in the Upper Layer Culture at Kiwulan.
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares were excavated _in situ_ from household area of Kiwulan, indicated by numerous post-holes.
+Thirty burials contained 3173 ornaments, and 27 ornaments were recovered from middens.
+In the burials ornaments were in the form of bead strings that consisted of hundreds to thousands of pieces.
+The ornaments from the burial contexts are not analysed here since few burials can be confidently assigned to a time period.
 Post hole areas represents domestic spaces suitable for testing for changes in the patterns of distribution relating to social organisation.
@@
 Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
-According to Chinese census reports in 1821 and Japanese field surveys in 1895, the Kiwulan population was less than 100 [@Yao1996; @Ino1898].
 

### 2019-09-21 analysis/paper/scraps.rmd
reducing word count for resubmission

@Berrocal2018 examined both archaeological and historical records of Heping Dao in northern Taiwan. This was the location of the colony of San Salvador founded by the Spanish in 1626 and taken over by the Dutch in 1642. Berrocal et al. uncovered parts of the foundation of a Spanish church and a European cemetery that indicate the evidence of European occupation. These remains indicate a direct impact on the local indigenous populations through their modification of the local landscape. In contrast to Heping Dao, Yilan in Northeastern Taiwan is a physically isolated area adjacent to colonial centres and there are no traces of European settlement but occasional interactions with the Europeans. Yilan, also called "Kavalan" by local indigenous people, is an alluvial plain circumscribed by the Pacific and mountains on the sides. The rugged Hsuehshan Range to the west and Central Range to the south of the Yilan Plain that made it not easily accessible to the European colonists and the colonial control was weaker compared to other parts of this region [@Andrade2007; @Kang2012]. This makes Yilan an especially unique location for exploring indirect colonial effects on local indigenous settlements living beyond the borders of direct colonial occupation.

### 2019-11-19 analysis/paper/paper.Rmd

@@
 sample_per <- round(sam_post_hole/total_post_hole*100, 1)
+
+#
+golden_bead_n <- 
+  ornaments_count %>% 
+  filter(Categories == "Golden bead") %>% 
+  pull(n)
+
+glass_bead_n <- 
+  ornaments_count %>% 
+  filter(Categories == "Glass bead") %>% 
+  pull(n)
+
+agate_bead_n <- 
+  ornaments_count %>% 
+  filter(Categories == "Agate bead") %>% 
+  pull(n)
+
+metal_ring_n <- 
+  ornaments_count %>% 
+  filter(Categories == "Metal ring") %>% 
+  pull(n)
+
+metal_bell_n <- 
+  ornaments_count %>% 
+  filter(Categories == "Bell") %>% 
+  pull(n)
 ```
@@
 
-This study focuses on `r sam_post_hole` ornaments from 40 sampling squares were excavated _in situ_ from household area of Kiwulan, indicated by numerous post-holes.
-Thirty burials contained 3173 ornaments, and 27 ornaments were recovered from middens.
-In the burials ornaments were in the form of bead strings that consisted of hundreds to thousands of pieces.
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares that were excavated _in situ_ from the household area of Kiwulan, indicated by numerous post-holes (Figure \@ref(fig:KWL-sam-area)).
+Thirty burials contained 3,173 ornaments, and 27 ornaments were recovered from middens.
+In the burials ornaments were in the form of bead strings that consisted of hundreds to thousands of pieces per burial.
 The ornaments from the burial contexts are not analysed here since few burials can be confidently assigned to a time period.
-Post hole areas represents domestic spaces suitable for testing for changes in the patterns of distribution relating to social organisation.
-Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells (see also Figure 4).
-The variety of subtype and their individual frequency are summarized in Table \@ref(tab:tidy-data-for-table) in terms of three time episodes.
+The household area represents domestic spaces suitable for testing for changes in ornament distribution relating to social organisation.
+Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), agate beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)).
+The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
 
@@
 
-# Results of hypothesis testing
+# Results
 
@@
 
-Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan.
-The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$) shows that there is a non-random difference between the three time periods.
-It indicates that most ornament types were present before European contact.
-Ornament frequencies reached a peak during European period and then dropped during the Chinese period, especially for golden beads.
+Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequencies of the major classes of ornaments for different the time periods at Kiwulan.
+The difference in frequencies between the three time periods reflect significant differences in the use of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$).
+Most ornament types were present before European contact.
+Ornament frequencies reached a peak during the European period and then dropped during the Chinese period, especially golden beads.
 This trend can be also seen on other ornaments including agate beads, metal rings, and bells.
@@
 
-The distribution of frequency for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether).
+The distribution of frequencies for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether).
 Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
 This indicates that the increases in diversity can be explained by the effects of culture interaction instead of effects of sample size.
-According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
-The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the Europeans.
-In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces.
-The common shape for copper bells is the large shape with a wide variety of human faces as a motif, while the common shape for glass beads is a bead with length less than 1 cm that are often identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].
+Agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
+The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the European presence.
+In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in clusters that indicate their original use as strings or necklaces.
+Copper bells are typically >2 cm long with a wide variety of human faces as a motif, while glass beads are usually <0.5 cm long Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC [@Francis2002].
 
@@
 Each class shows its own pattern where the squares with higher number of ornaments distributed separately and independently.
-For example, the cluster of golden beads was found at the northern part, while the cluster of agate beads was found in the middle part.
+For example, a cluster of golden beads was found at the northern part, while a cluster of agate beads was found in the middle part.
 In contrast, there are multiple clusters of metal rings that are distributed separately across the research area.
-Copper bells were usually found individually and seem randomly distributed across the area.
+Copper bells were usually found individually and appear randomly distributed across the area.
 In the Chinese period, both the amount and density of different classes of ornaments decreased.
@@
 
-I used point pattern analysis to examine whether the distribution of ornaments was uneven in non-random ways that could reflect changes in social inequality.
-Point pattern analysis is a way to assess whether the distribution of sites or artefacts represent some “hotspots” produced by non-random processes, implying the influence of social processes [@Bevan2016; @Ducke2015].
-Hotspots in the pattern of ornaments could indicate changes in social organisation due to culture interaction.
-To explore ornament distributions I used the intensity approach that focuses on the average density of points across space.
-During excavation, ornament locations were only recorded to the square (4 x 4 m).
-To prepare the location data for point pattern analysis, I assigned each ornament to a random coordinate pair located in square it was recovered from.
-The next step was to subset the ornaments into three groups for three time periods.
+Point pattern analysis can assess whether the distribution of artefacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments resulting from social inequality stimulated by a colonial presence.
+To prepare the ornament location data for point pattern analysis, I assigned each ornament to a random coordinate pair located in 4 x 4 m square it was recovered from (ornaments lack individual point provenance).
+The next step was to divide the ornaments into three time periods.
 Finally I computed the kernel densities for each time period for comparison.
-
-Kernel density estimations (KDE) estimate the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
-Here I use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
+Kernel density estimations (KDE) compute the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
+Here I use KDE to visualize core areas of ornaments and surrounding neighbourhoods [@Bonnier2019; @Cortegoso2016].
 Density values of artefacts per square meter were calculated for each cell.
 
-The results show that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period according to the kernel density maps (Figure \@ref(fig:plot-all-densities)).
-It seems that there are three consistent sub-regions with a core area that shifts over time.
-The distribution might indicate different household-based social groups who possessed more ornaments.
-The multiple groups during European period might reflect more unequal consumption of ornaments across the site, relative to other periods, to indicate social processes caused by human behavior.
-However, it could also just a random pattern resulting from bigger sample size.
-In addition, the generation of core areas might be biased due to small sample sizes, for example, a few ornaments found at one single square during the Chinese period could create an obvious hot spot.
-To further evaluate these patterns, we test the hypothesis that the distribution of ornaments is not randomly distributed to determine if social processes might explain ornament distributions.
+Figure \@ref(fig:plot-all-densities) shows that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period.
+There are three consistent sub-regions with a core area that shifts over time.
+The distribution might indicate increase and decrease in the number of different household-based social groups who possessed more ornaments.
+The multiple groups during the European period might reflect more unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from bigger sample size.
+In addition, the generation of core areas might be biased due to small sample sizes, for example, a few ornaments found at one single square during the Chinese period could create an obvious hotspot.
 
@@
 
-To test this hypothesis about the randomness of spatial locations, I used a Monte Carlo method to simulate average nearest-neighbor distances (ANN).
-Our null hypothesis is that the distribution of ornaments is consistent with a completely random process.
-We simulated the locations of ornaments across the space 1000 times for each time period.
-The observed data was compared to the random patterns generated by simulated processes to determine if our data are randomly distributed or not.
-Figure \@ref(fig:hypo-test-distribution) shows the distributions of the ANN distances calculated on 1000 simulations of ornament locations.
-The results show that 100% of the simulated values are much greater than our observed ANN value during the European period, which means the ornaments have non-random clustered distributions.
+To test for randomness in spatial locations, I used a Monte Carlo method to simulate average nearest-neighbour distances (ANN).
+Figure \@ref(fig:hypo-test-distribution) shows the observed ANN distances with the distributions of the ANN distances calculated on 1000 simulations of random ornament locations.
+The results show that 100% of the simulated values are much greater than our observed ANN value during the European period, which means the ornaments have non-randomly clustered distributions.
 A similar, but less extreme, result is also observed during the pre-European period.
@@
 This testing reveals that the clustered distributions of ornaments during the pre-European and European periods might reflect the presence of different social groups.
-Moreover, the clustering of ornaments during the European period is highly non-random, might potentially indicating that some of the social changes leading to different degrees of access foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan.
+Moreover, the clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan.
 
@@
 
-The results show there was a greater diversity of ornaments types and materials at Kiwulan during the European presence.
-This might result from the large scale exchange network that stimulated the circulation of different ornament classes.
+An indirect colonial presence is indicated at Kiwulan by a greater diversity of ornaments types and materials during the European period.
+This likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
 However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period.
-This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the mortuary practices of Han Chinese.
+This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the practices of Han Chinese who did not use these kinds of beads in their dress.
 Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
@@
 Spatial patterns of ornaments shows that their distribution was clustered during the pre-European and European period.
-Hypothesis testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European period.
-Because ethnographic evidence indications that trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering may indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European period.
-In addition, there was a burial dated to 17th century unearthed 60 golden beads above the average of 2-3 pieces in the pre-European period, further indicating increasing social inequality [@Chen2007; @Cheng2008].
+Testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European period.
+Because ethnographic evidence indicates that trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering may indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European period.
+A further indicator of increasing social inequality is a burial dated to the 17th century that included 60 golden beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008].
 
@@
 In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks.
-In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labor tasks [@Feinman2000; @Siegel1999].
+In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
 
-We may consider that Yilan social organisation went from corporate mode before the European arrival, then transitioned into network mode during the European presence, then back to corporate mode in the Chinese period.
+Yilan social organisation went from corporate mode before the European arrival, then transitioned to network mode during the European presence, then back to corporate mode in the Chinese period.
 The small number of ornaments, and less concentrated distribution during the pre-European period is consistent with shared power and wealth of the corporate mode.
-The long-distance trade network introduced by the Europeans resulted in the appearance of a network mode due to competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
-Because of the weak direct control from the European colonizers in northeastern Taiwan, it implies that local leaders might have flexibility to manipulate European colonial image, expand personal power, and monopolize the high-value trade goods [@Kang2012].
+The long-distance trade network introduced by Europeans resulted in the appearance of a network mode due to competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
+Because of the weak direct control from the European colonizers in northeast Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
 During the Chinese period, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
-This reflects the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals accumulate wealth.
+This reflects the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth.
 
@@
 
-Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by the colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017].
-Kiwulan in Northeastern Taiwan is an important case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
-Kiwulan can provide valuable insights into the discussion of indirect colonial influence on the local societies living beyond the borders of direct European colonial occupation.
+Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017].
+Kiwulan in northeast Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
+Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation.
 The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
 The greater amount and diversity of ornament types during the European period reflects international exchanges in a colonial context.
-Ornaments were viewed as prestige goods in local indigenous culture before European contact.
-After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may be intensified that further signals wealth and privileged trading connections with European colonists for the inhabitants of Kiwulan.
+Ornaments were viewed as prestige goods in the local indigenous culture before European contact.
+After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan.
 These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
 
-This study demonstrates that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
-Ornaments can give insights into the amplification of social inequality stimulated by European colonisation.
+We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities.
+The Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
+Ornaments give insights into the amplification of social inequality stimulated by European colonisation.
 It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.

### 2019-11-21 analysis/paper/paper.Rmd
Change some terms and clarify sentenses

@@
 The archaeological evidence includes a large amount of artefacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
-Most ornaments were found in situ while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
+Most ornaments were found while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
 The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of the excavation, see Figure \@ref(fig:KWL-sam-area).
-They were sampled because those units were stratigraphically intact with the least disturbance by modern construction activity, compared to excavation squares on the periphery of the site.
+They were sampled because those units were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
 
-The chronology of Kiwulan can be divided into two phases represented by a Lower Layer Culture (700-1200 AD) and a Upper Layer Culture (1400-1900 AD) separated by a sterile.
+The chronology of Kiwulan can be divided into two phases represented by a upper component (700-1200 AD) and a lower component (1400-1900 AD) separated by a sterile.
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
-This paper focuses on the Upper Layer Culture because this has a sequence spanning pre-European contact, European presence, and Chinese presence.
+This paper focuses on the upper component because this has a sequence spanning pre-European contact, European presence, and Chinese presence.
 Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
@@
 In the burials ornaments were in the form of bead strings that consisted of hundreds to thousands of pieces per burial.
-The ornaments from the burial contexts are not analysed here since few burials can be confidently assigned to a time period.
+The ornaments from the burial contexts are not statistically analysed because most burials were assigned to 17-18 th century that offers limited quantitative comparison with other time periods.
+But the use of ornaments as body adornments in burials is discussed to understand the functions of ornaments.
 The household area represents domestic spaces suitable for testing for changes in ornament distribution relating to social organisation.

### 2019-11-22 analysis/paper/paper.Rmd
updates based on reviews

@@
 
-This study focuses on `r sam_post_hole` ornaments from 40 sampling squares that were excavated _in situ_ from the household area of Kiwulan, indicated by numerous post-holes (Figure \@ref(fig:KWL-sam-area)).
-Thirty burials contained 3,173 ornaments, and 27 ornaments were recovered from middens.
-In the burials ornaments were in the form of bead strings that consisted of hundreds to thousands of pieces per burial.
-The ornaments from the burial contexts are not statistically analysed because most burials were assigned to 17-18 th century that offers limited quantitative comparison with other time periods.
-But the use of ornaments as body adornments in burials is discussed to understand the functions of ornaments.
-The household area represents domestic spaces suitable for testing for changes in ornament distribution relating to social organisation.
-Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), agate beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)).
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aliened postholes with in-situ posts (Figure \@ref(fig:KWL-sam-area)).
+These postholes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
+At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction across the sampling area.
+There are 30 burials in the sampling area account for one thrid of the total number of burials at Kiwulan.
+Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes are 3,173.
+The use of ornaments as body adornments in burials will be discussed to understand the functions of ornaments, but not statistically analysed since the number of ornaments is highly affected by the presence of bead strings, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports that ornaments were treated as high value items at Kiwulan because of small number in the midden contexts compared to high frequency in burial contexts.
+It is important to analyse the detailed distribution of ornaments from the dwelling place as it can provide key information concerning the spatial pattens across the space as well as changes in ornament relating to social organisation across different time periods.
+
+The ornaments found in the  Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), agate beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)).
 The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).

### 2019-11-27 analysis/paper/paper.Rmd
revision based on comments

@@
   - Li-Ying Wang
+  - Ben Marwick
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 I address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
+# Historical background
 
-# Cultural context of ornaments in Yilan, northeastern Taiwan
-
+The island of Taiwan lies at the junction of mainland China, southeast Asia, and northeast Asia in the Pacific Ocean.
+The prehistory of Taiwan island could be roughly divided into three major periods, paleolithic (c.
+27,000 BP- 5000BP), Neolithic (c.
+6500- 2000BP), and Iron age (c.
+2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages (@Liu2011; @Chen2017).
+It is generally accepted that Taiwan entered historical period since the early 17th century due to the colonial presence of the Spanish and Dutch who played an important role in keeping written records about Taiwan.
 The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
@@
 This paper focuses on the upper component because this has a sequence spanning pre-European contact, European presence, and Chinese presence.
-Previous researchers divided the Upper Layer Culture component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
+Previous researchers divided the upper component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
 Due to ambiguities in the previous chronology, I re-examined the excavation records to produce a new chronology for the sampled squares.
@@
 At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction across the sampling area.
-There are 30 burials in the sampling area account for one thrid of the total number of burials at Kiwulan.
+There are 30 burials in the sampling area account for one third of the total number of burials at Kiwulan.
 Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes are 3,173.
@@
 In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports that ornaments were treated as high value items at Kiwulan because of small number in the midden contexts compared to high frequency in burial contexts.
-It is important to analyse the detailed distribution of ornaments from the dwelling place as it can provide key information concerning the spatial pattens across the space as well as changes in ornament relating to social organisation across different time periods.
+It is important to analyse the detailed distribution of ornaments from the dwelling place as it can provide key information concerning social inequality based on uneven access to the high-value items, and changes in the spatial pattern of ornament across different time periods.
 
@@
 The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
+The chemical composition analysis for glass beads shows high-lead content in the glass, coupled with the winding/folding technique of bead making, suggest a China production tradition[@Cheng2008; @Wang2018].
+The style of other ornaments, including glass beads, metal bells, agate beads, are also associate with a Chinese origin [@Chen2007].
+There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans via economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
 
@@
 There are three consistent sub-regions with a core area that shifts over time.
-The distribution might indicate increase and decrease in the number of different household-based social groups who possessed more ornaments.
+The distribution might indicate increase and decrease in the number of different social groups who possessed more ornaments.
 The multiple groups during the European period might reflect more unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from bigger sample size.
@@
 
-An indirect colonial presence is indicated at Kiwulan by a greater diversity of ornaments types and materials during the European period.
+An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornaments types and materials during the European period. The origin of ornaments
+
 This likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
@@
 Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
+
+The explanations for the increase of ornaments during the European period could be
+
+should be related to participation in global trade networks is not clarified.
+
+On the contrary, a more economical interpretation could be that they are used to vindicate the local customs and resist against European intrusion, rather than the opposite, as the author does.
+Without using foreign elements (something possible in Kiwulan, where 300+ tobacco pipes have been found, for example) it is difficult to see the influence of global networks in any particular place.
 

### 2019-11-29 analysis/paper/paper.Rmd
added sentences about origin of beads and metal ornaments

@@
 ---
-title: "Ornaments as indicators of social change before and after European arrival at Kiwulan, Northeastern Taiwan"
+title: "Ornaments as indicators of social changes and cultural practices in northeast Taiwan before and after European colonial period"
 author:
@@
   Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
-Trade ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
+Trade ornaments from Kiwulan (1400-1900 AD) in northeast Taiwan show the indirect impacts of European colonial activities on local societies.
 The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered.
@@
 Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
-Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in Taiwan, the northeast region was isolated from intensive direct contact by the Xueshan Mountains.
+Northeast Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in Taiwan, the northeast region was isolated from intensive direct contact by the Xueshan Mountains.
 
@@
 12].
-Following defeat of Dutch by the Chinese general Koxinga in 1661-2, the Dutch abandoned northern Taiwan.
+Following defeat of Dutch by the Chinese general Koxinga in 1661-1662, the Dutch abandoned northern Taiwan.
 Direct contact with Han Chinese is indicated by Qing dynasty census reports mentioning Yilan villages in 1821 [@Yao1996].
@@
 I address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
-# Historical background
+# Historical Background of complex exchange network during the late Iron Age and early historical period
 
@@
 343].
-A report by a Dominican priest in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, exchanged agate beads with other indigenous groups.
-This form of exchange was widespread and even the Spanish soldiers learned to use agate as bargaining chips for gambling [@LiandWu2006, pp.
+A report by a Dominican priest in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups.
+This form of exchange was widespread and even the Spanish soldiers learned to use carnelian as bargaining chips for gambling [@LiandWu2006, pp.
 132-149].
@@
 132-149].
-Other records mention that the female shamans (majuorbol) in the tribe would use agate beads as magical items in ritual practice for body healing [@Borao2009, pp.
+Other records mention that the female shamans (majuorbol) in the tribe would use carnelian beads as magical items in ritual practice for body healing [@Borao2009, pp.
 122-151].
-Records of an indigenous funeral process document the use of agate beads in ritual contexts, with more agate beads, pottery, and cloth placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
+Records of an indigenous funeral process document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
 153].
@@
 According to those records, indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
-Among those ornaments, golden fish-shaped necklaces made of copper threads had high value due to their delicacy and the materials invested in production.
+Among those ornaments, fish-shaped necklaces made of metal threads had high value due to their delicacy and the materials invested in production.
 These were usually possessed by wealthy people.
@@
 227-232].
+In sum, we find multiple sources describing how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
+Compared to the European period, there are fewer mentions of beads in Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
+However, we must recognize some biases in historical records that might lead to misinterpretations of the use of ornaments.
+Inconsistent descriptions from different written sources and a lack of context in texts are the main problems need to be resolved by studying the ornaments in archaeological contexts.
 
-Recent ethnographic research with the members of the Kavalan tribe, the contemporary indigenous people of the Yilan plain, describes how agate beads were used in divination practices by female shamans [@Liu2008, pp.
-133-134].
-Interview records mention agate beads as valuable objects that are usually passed down from mother to daughter as heirlooms in shaman families.
-When not in use, beads were often hidden in safe places, such as on the roof beams.
-Despite most Kavalan people today not knowing the specific origin of their agate beads, the high value of beads is still indicated in oral history.
-In sum, we find multiple sources describing how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
-However, compared to the European period, there are fewer mentions of beads in Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
+Currently the ornaments found in northeast Taiwan in early historical period around 400 hundred years ago, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metal working, or accessible local raw materials.
+The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
+Although there is wide varity of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen].
+There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
+Northeast Taiwan from the late Iron age to early historical period was involved in complex exchange network both in a regional and global scale.
 
@@
 
-The ornaments found in the Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), agate beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)).
+The ornaments found in the Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), carnelian beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)).
 The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
-The chemical composition analysis for glass beads shows high-lead content in the glass, coupled with the winding/folding technique of bead making, suggest a China production tradition[@Cheng2008; @Wang2018].
-The style of other ornaments, including glass beads, metal bells, agate beads, are also associate with a Chinese origin [@Chen2007].
-There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans via economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
 
@@
 Ornament frequencies reached a peak during the European period and then dropped during the Chinese period, especially golden beads.
-This trend can be also seen on other ornaments including agate beads, metal rings, and bells.
+This trend can be also seen on other ornaments including carnelian beads, metal rings, and bells.
 However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period.
@@
 This indicates that the increases in diversity can be explained by the effects of culture interaction instead of effects of sample size.
-Agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
-The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the European presence.
+Carnelian beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
+The greater varieties for carnelian beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the European presence.
 In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in clusters that indicate their original use as strings or necklaces.
@@
 During the Chinese period the distribution is more even again.
-Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European period, such as golden beads and agate beads.
+Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European period, such as golden beads and carnelian beads.
 However, there seems to be no clear consistent pattern across those different ornaments.
 Each class shows its own pattern where the squares with higher number of ornaments distributed separately and independently.
-For example, a cluster of golden beads was found at the northern part, while a cluster of agate beads was found in the middle part.
+For example, a cluster of golden beads was found at the northern part, while a cluster of carnelian beads was found in the middle part.
 In contrast, there are multiple clusters of metal rings that are distributed separately across the research area.

### 2019-12-03 analysis/paper/paper.Rmd
revision based on reviewer comments

@@
 I address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
+
 # Historical Background of complex exchange network during the late Iron Age and early historical period
@@
 Compared to the European period, there are fewer mentions of beads in Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
-However, we must recognize some biases in historical records that might lead to misinterpretations of the use of ornaments.
-Inconsistent descriptions from different written sources and a lack of context in texts are the main problems need to be resolved by studying the ornaments in archaeological contexts.
+However, we must recognize the inherent biases in historical records and fragmented nature of texts selected by written record keepers that might lead to misinterpretations of artifacts [@Galloway2006].
+Inconsistent and incomplete descriptions from different written sources lacking contexts are the problems need to be compared and examined by studying the ornaments in archaeological contexts.
 
@@
 The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
-Although there is wide varity of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen].
+Although there is wide varity of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011].
 There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
@@
 Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes are 3,173.
-The use of ornaments as body adornments in burials will be discussed to understand the functions of ornaments, but not statistically analysed since the number of ornaments is highly affected by the presence of bead strings, which sometimes contains thousands of beads in an individual burial [@Chen2007].
-In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports that ornaments were treated as high value items at Kiwulan because of small number in the midden contexts compared to high frequency in burial contexts.
-It is important to analyse the detailed distribution of ornaments from the dwelling place as it can provide key information concerning social inequality based on uneven access to the high-value items, and changes in the spatial pattern of ornament across different time periods.
+The use of ornaments as body adornments in burials will be discussed to understand the function of ornaments, but not statistically analysed since the amount of ornaments is highly affected by the presence of bead strands or patterned bands, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports the argument that ornaments were treated as high value items at Kiwulan based on a few finds in the midden contexts compared to high frequency in burial contexts.
+It is important to analyse the detailed distribution of ornaments from the dwelling place across different time periods as it can provide key information concerning social inequality indicated by uneven access to high-value item in relation to the foreign influences.
 
@@
 The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
+The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to their use in the burial contexts.
 
@@
 Carnelian beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
-The greater varieties for carnelian beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the European presence.
-In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in clusters that indicate their original use as strings or necklaces.
-Copper bells are typically >2 cm long with a wide variety of human faces as a motif, while glass beads are usually <0.5 cm long Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC [@Francis2002].
+The greater varieties for carnelian beads and metal rings might indicate multiple origins due to participation in large scale trade networks stimulated by the European presence.
+In contrast, copper bells and glass beads have less variety that the bells are typically >2 cm long with a wide variation of human faces as a motif, while glass beads usually <0.5 cm long Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC [@Francis2002].
 
@@
 
-An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornaments types and materials during the European period. The origin of ornaments
-
-This likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
-However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period.
+An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornaments types and materials during the European period.
+Northeast Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish.
+The origin of ornaments might be multiple places mostly in Southeast Asia and introduced into northeast Taiwan by Chinese merchants before 17th century.
+Later, the trade activities became more frequent and intense in the 17th century that might be related to the colonial involvement of the Europeans based on the similar styles of golden beads found in Luzon that was once the colony of the Spanish during this period.
+In addition to ornaments, stonewares at Kiwulan that were also found on a Spanish shipwreck and Southeast Asia in relation to the Dutch and the Spanish trading routes, and the tobacco use were also believed introduced to Kiwulan by the Europeans [@Wang2007; @Chen2007].
+The greater diversity and quantity of ornaments likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
+However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period in the early 19th century.
 This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the practices of Han Chinese who did not use these kinds of beads in their dress.
 Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
+Although historical records indicate that trade ornaments were prestige goods in prehistoric northeast Taiwan, the information might be biased that needs to be compared with archaeological evidence.
 
-The explanations for the increase of ornaments during the European period could be
-
-should be related to participation in global trade networks is not clarified.
-
-On the contrary, a more economical interpretation could be that they are used to vindicate the local customs and resist against European intrusion, rather than the opposite, as the author does.
-Without using foreign elements (something possible in Kiwulan, where 300+ tobacco pipes have been found, for example) it is difficult to see the influence of global networks in any particular place.
-
+The archaeological contexts show that ornaments were rarely found in middens and mostly preserved well as body adornments in burial contexts serving as grave goods.
+This could reflect ornaments were viewed as valuable objects such as status indicators and their distribution across the dwelling area would help to understand the social relation reflected by people who possessed them, and the reaction of the society as a whole to the colonial presence in the 17th century.
 Spatial patterns of ornaments shows that their distribution was clustered during the pre-European and European period.
 Testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European period.
-Because ethnographic evidence indicates that trade ornaments were prestige goods in prehistoric Northeastern Taiwan, this spatial clustering may indicate that a degree of social inequality was present before European contact and then it was reinforced and amplified during the European period.
+The spatial clustering during the pre-European and European period shows an increase in use of ornaments in the daily life at Kiwulan.
+It may indicate that a degree of social inequality based on uneven distribution was present before European contact and then it was reinforced and amplified during the European period.
 A further indicator of increasing social inequality is a burial dated to the 17th century that included 60 golden beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008].
@@
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
-
-Yilan social organisation went from corporate mode before the European arrival, then transitioned to network mode during the European presence, then back to corporate mode in the Chinese period.
+This provides an interpretation that Yilan social organisation went from corporate mode before the European arrival, then transitioned to network mode during the European presence.
 The small number of ornaments, and less concentrated distribution during the pre-European period is consistent with shared power and wealth of the corporate mode.
@@
 Because of the weak direct control from the European colonizers in northeast Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
+
+Another scenario could be that the increasing number and concentrated spatial pattern of ornaments present a form of resistance to European intrusion by using them to show social identity and emphasize the local custom had existed before European contact.
+Resistance could be presented in many forms, for example, @Rubertone discussed a case study in the North America that a large amount of native-made shell beads supposed to be tributes for the Europeans were found in native burial contexts.
+Similarly, the frequent presence of ornaments across the dwelling area not only exclusive in burial contexts suggests ornaments might be intentionally displayed in daily life to reinforce the cultural tradition.
 During the Chinese period, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
 This reflects the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth.
+It is also possible that some ornaments were treated as heirlooms that passed down from one generation to another generation.
+However, the spacial pattern can also reflects the trend of distribution even they were heirlooms and this might suggest that the ornaments found in Chinese period might come from earlier period and continuously used till then.
+Further analysis about the source for each type of ornaments would be helpful to understand their origin and use context.
 

### 2019-12-05 analysis/paper/paper.Rmd
revision based on the comments from reviewer 2

@@
 153].
-These historical accounts indicate that agate beads were already treated as prestige goods in Yilan before the arrival of Europeans.
+These historical accounts indicate that carnelian beads were already treated as prestige goods in Yilan before the arrival of Europeans.
 
@@
 These were usually possessed by wealthy people.
-Other people wore agate beads or glass beads on their head or neck to participate in ceremonies.
+Other people wore carnelian beads or glass beads on their head or neck to participate in ceremonies.
 In 1895, at the beginning of Japanese colonisation, an academic field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
@@
 
-Currently the ornaments found in northeast Taiwan in early historical period around 400 hundred years ago, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metal working, or accessible local raw materials.
+Currently the ornaments found in northeast Taiwan in early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metal working, or accessible local raw materials.
 The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (700-1200 AD) and a lower component (1400-1900 AD) separated by a sterile.
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD) and a lower component (700-1200 AD) separated by a sterile.
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
@@
 
-The archaeological indicators of the start of the European influence at Kiwulan are the appearance of An-ping jars and stonewares (martavans/martaban) that were introduced to Taiwan during the early 17th century.
-They are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
+The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jar, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
+Large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, that is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh\\].
+The origin of light gray glazed jars were thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007].
+Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
 The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
-I used the presence and absence of An-ping jars and stonewares, together with the radiocarbon dates, to identify excavation units associated with the pre-European and post-European periods.
+Those jars might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan.
+Either way indicates that the Europeans played an important role for introducing those jars to Taiwan.
+I used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.
 
@@
 Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains, clay stoves.
-Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles [@Hsieh2009].
-Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century [@Chen1963; @Ke1993].
+The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups.
+Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles used by Chinese [@Hsieh2009].
+Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrant in 1768 [@Chen1963; @Ke1993].
+
+```{r C14-table-data}
+C14 <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
+knitr::kable(C14,
+             caption = "Radiocarbon dates on charcoals from Kiwulan")
+
+```
 
@@
   mutate(Type = str_remove(Type, "prism/"),
-         Type = ifelse(Type == "glass bead", "medium bead", Type),
+         Type = ifelse(Type == "small bead", "small (0.5-1 cm)", Type),
+         Type = ifelse(Type == "glass bead", "medium (1-2 cm)", Type),
          Type = ifelse(Type == "small I", "plain small", Type),
@@
 ```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
-A: agate beads, B: bells, C: glass beads and golden beads, D: metal rings.
+A: carnelian beads, B: bells, C: glass beads and golden beads, D: metal rings.
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
@@
 The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
+The criteria for classification of subtype is based on their dimensional attributes that are important to understand the variation in shapes.
+This might simplify the subtypes of glass beads when they are the same size and shape but in different color or texture.
+However, their variety would be considered and discussed separately.
 The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to their use in the burial contexts.
@@
   labs(y = "",
-       x = "Agate bead") +
+       x = "Carnelian bead") +
   theme_minimal(base_size = base_size_multi_plot) +
@@
 
-```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, agate beads, bells, and glass beads."}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads."}
 library(cowplot)
@@
 The greater varieties for carnelian beads and metal rings might indicate multiple origins due to participation in large scale trade networks stimulated by the European presence.
-In contrast, copper bells and glass beads have less variety that the bells are typically >2 cm long with a wide variation of human faces as a motif, while glass beads usually <0.5 cm long Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC [@Francis2002].
+In contrast, copper bells has less variety that the bells are typically >2 cm long with a wide variation of human faces as a motif.
+Although glass beads has less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have wide variety of color or pattern mostly made by winding technique with high lead content in composition indicating possibly from China [@Cheng2008].
+Despite more researches needed to understand the specific source of ornaments, current researches suggest that glass beads and metal ornaments have similar technique and composition to those found in China [@Chen2011; @Wang2018].
+There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component (1400-1900 AD).
+However, the glass beads found from lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC and declined at the early 2nd millennium [@Wang2018; @Francis2002].
+It indicates a different bead source more related to the exchange network in the South China Sea for the lower component.
 
@@
 
-Another scenario could be that the increasing number and concentrated spatial pattern of ornaments present a form of resistance to European intrusion by using them to show social identity and emphasize the local custom had existed before European contact.
-Resistance could be presented in many forms, for example, @Rubertone discussed a case study in the North America that a large amount of native-made shell beads supposed to be tributes for the Europeans were found in native burial contexts.
-Similarly, the frequent presence of ornaments across the dwelling area not only exclusive in burial contexts suggests ornaments might be intentionally displayed in daily life to reinforce the cultural tradition.
-During the Chinese period, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
-This reflects the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth.
-It is also possible that some ornaments were treated as heirlooms that passed down from one generation to another generation.
-However, the spacial pattern can also reflects the trend of distribution even they were heirlooms and this might suggest that the ornaments found in Chinese period might come from earlier period and continuously used till then.
-Further analysis about the source for each type of ornaments would be helpful to understand their origin and use context.
+Another scenario could be that the increasing number and concentrated spatial pattern of ornaments indicate cultural resistance against the European intrusion by using them to show social identity and emphasize the local custom that had existed before European contact.
+Resistance could be presented in many forms.
+For example, @Rubertone discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands.
+Similarly, the frequent presence of ornaments across the dwelling area during the European suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition.
+During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
+This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or population decrease.
+It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next.
+This might suggest that the ornaments found in Chinese period were old item continuously used till then.
+Further analysis about the source for ornaments from different periods will be helpful to understand their origin and reginal circulation.
 

### 2019-12-06 analysis/paper/paper.Rmd
added code chunk for c14 dates

@@
 The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
-Although there is wide varity of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011].
+Although there is wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011].
 There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
@@
 ```{r C14-table-data}
-C14 <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
-knitr::kable(C14,
+kw_dates <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
+table_kw_dates <-
+  kw_dates %>% 
+  select(-se, -rce)
+
+knitr::kable(table_kw_dates,
              caption = "Radiocarbon dates on charcoals from Kiwulan")
+```
 
+```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan.
+The y-axis represents summed probability and the x-axis is the date in calendar years.
+Produced with the BChronDensity() function in program R's Bchron package (Parnell et al.
+2008).
+IntCal13 Atmospheric curve."}
+
+# subset dates with error value
+kw_dates_with_error <- 
+  kw_dates %>% 
+  filter(!is.na(se))
+
+# convert rce to numeric
+kw_dates_with_error$rce <- as.numeric(kw_dates_with_error$rce)
+
+# make some kind of ID
+kw_dates_with_error$id <- with(kw_dates_with_error,
+                               paste0(Square,"-",Layer))
+
+# Use Bchron to clibrate
+library(Bchron)
+
+kw_dates_with_error_calibrated <- 
+  with(kw_dates_with_error,
+       BchronCalibrate(
+         ages = rce,
+         ageSds = se,
+         ids = id,
+         calCurves = rep('intcal13',
+                         length(kw_dates_with_error$rce))
+       ))
+
+# have a look at the output
+summary(kw_dates_with_error_calibrated)
+par(mfrow = c(2, 3))
+plot(kw_dates_with_error_calibrated)
+
+# we can combine all the calibrated dates into a single plot, like this
+kw_ages_densities <- 
+  with(kw_dates_with_error,
+       BchronDensity(
+         ages = rce,
+         ageSds = se,
+         calCurves = rep('intcal13',
+                         length(kw_dates_with_error$rce))
+       ))
+
+#plot it
+plot(kw_ages_densities,
+     xlab= "Years BC")
+
+# to describe in your text, ?BchronDensity
+# "We used the Bchron R package to compute a non-parametric phase model 
+# on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
+# FIgure X shows the results of this analysis... indicating that the peak period of occupation is around 600 BC"
 ```
@@
 It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next.
-This might suggest that the ornaments found in Chinese period were old item continuously used till then.
+This might suggest that the ornaments found in Chinese period were old item continuously used until then.
 Further analysis about the source for ornaments from different periods will be helpful to understand their origin and reginal circulation.
@@
 
-# Acknowledgments
+# Acknowledgements
 

### 2019-12-07 analysis/paper/paper.Rmd
edited the text and revised the table

@@
   kw_dates %>% 
-  select(-se, -rce)
+  mutate(`Depth related to sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>% 
+  mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
+  mutate(Period = as.factor(Period)) %>% 
+  mutate(Period = factor(Period,
+                       level = c("pre-European", 
+                                 "European", 
+                                 "post-European"))) %>% 
+  arrange(Period) %>% 
+  select("Pit-Layer", "Depth related to sea level (cm)", Period, "Date BP", "Calibrated date", "Details")
 
@@
 The next step was to divide the ornaments into three time periods.
+All squares in the sampling area were occupied during all phases of habitation based on the continuous presence of archaeological materials across over the strata representing a time span of 600 years.
+Although the quantity of locally made pottery declines in the later period with the growth of the Chinese porcelains, the general density of the artifacts present consistent thoughout all three phases.
+
 Finally I computed the kernel densities for each time period for comparison.

### 2019-12-07 analysis/paper/paper.Rmd
fixed some typos and added sentences addressing the population issues

@@
 
-The island of Taiwan lies at the junction of mainland China, southeast Asia, and northeast Asia in the Pacific Ocean.
+The island of Taiwan lies at the junction of mainland China, Southeast Asia, and northeast Asia in the Pacific Ocean.
 The prehistory of Taiwan island could be roughly divided into three major periods, paleolithic (c.
@@
 2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages (@Liu2011; @Chen2017).
-It is generally accepted that Taiwan entered historical period since the early 17th century due to the colonial presence of the Spanish and Dutch who played an important role in keeping written records about Taiwan.
+It is generally accepted that Taiwan entered the historical period since the early 17th century due to the colonial presence of the Spanish and Dutch who played an important role in keeping written records about Taiwan.
 The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
@@
 In sum, we find multiple sources describing how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
-Compared to the European period, there are fewer mentions of beads in Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
-However, we must recognize the inherent biases in historical records and fragmented nature of texts selected by written record keepers that might lead to misinterpretations of artifacts [@Galloway2006].
+Compared to the European period, there are fewer mentions of beads in the Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
+However, we must recognize the inherent biases in historical records and the fragmented nature of texts selected by written record keepers that might lead to misinterpretations of artifacts [@Galloway2006].
 Inconsistent and incomplete descriptions from different written sources lacking contexts are the problems need to be compared and examined by studying the ornaments in archaeological contexts.
 
-Currently the ornaments found in northeast Taiwan in early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metal working, or accessible local raw materials.
+Currently the ornaments found in northeast Taiwan in the early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metalworking, or accessible local raw materials.
 The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
-Although there is wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011].
-There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
-Northeast Taiwan from the late Iron age to early historical period was involved in complex exchange network both in a regional and global scale.
+Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011].
+There is no clear evidence shows the European origin of beads, however, a large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
+Northeast Taiwan from the late Iron age to the early historical period was involved in complex exchange networks both on a regional and global scale.
 
@@
 
-The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jar, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
-Large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, that is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh\\].
-The origin of light gray glazed jars were thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007].
+The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
+A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh\\].
+The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007].
 Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
@@
 Those jars might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan.
-Either way indicates that the Europeans played an important role for introducing those jars to Taiwan.
+Either way indicates that the Europeans played an important role in introducing those jars to Taiwan.
 I used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.
@@
 There are 30 burials in the sampling area account for one third of the total number of burials at Kiwulan.
-Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes are 3,173.
-The use of ornaments as body adornments in burials will be discussed to understand the function of ornaments, but not statistically analysed since the amount of ornaments is highly affected by the presence of bead strands or patterned bands, which sometimes contains thousands of beads in an individual burial [@Chen2007].
-In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports the argument that ornaments were treated as high value items at Kiwulan based on a few finds in the midden contexts compared to high frequency in burial contexts.
-It is important to analyse the detailed distribution of ornaments from the dwelling place across different time periods as it can provide key information concerning social inequality indicated by uneven access to high-value item in relation to the foreign influences.
+Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes is 3,173.
+The use of ornaments as body adornments in burials will be discussed to understand the function of ornaments, but not statistically analysed since the number of ornaments is highly affected by the presence of bead strands or patterned bands, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports the argument that ornaments were treated as high-value items at Kiwulan based on a few finds in the midden contexts compared to high frequency in burial contexts.
+It is important to analyse the detailed distribution of ornaments from the dwelling place across different time periods as it can provide key information concerning social inequality indicated by uneven access to high-value items in relation to the foreign influences.
 
@@
 
-The distribution of frequencies for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether).
+The distribution of frequencies for subtypes in each major class is presented in Figure \@ref(fig:plot-altogether).
 Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
-This indicates that the increases in diversity can be explained by the effects of culture interaction instead of effects of sample size.
-Carnelian beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European period.
+This indicates that the increases in diversity can be explained by the effects of culture interaction instead of the effects of sample size.
+Carnelian beads and metal rings have greater quantity and variety of shapes compared to copper bells and glass beads during the European period.
 The greater varieties for carnelian beads and metal rings might indicate multiple origins due to participation in large scale trade networks stimulated by the European presence.
-In contrast, copper bells has less variety that the bells are typically >2 cm long with a wide variation of human faces as a motif.
-Although glass beads has less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have wide variety of color or pattern mostly made by winding technique with high lead content in composition indicating possibly from China [@Cheng2008].
+In contrast, copper bells have less variety typically >2 cm long with a wide variety of human faces as a motif.
+Although glass beads have less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have a wide variety of colors or patterns mostly made by winding technique with high lead content in composition indicating possibly from China [@Cheng2008].
 Despite more researches needed to understand the specific source of ornaments, current researches suggest that glass beads and metal ornaments have similar technique and composition to those found in China [@Chen2011; @Wang2018].
 There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component (1400-1900 AD).
-However, the glass beads found from lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC and declined at the early 2nd millennium [@Wang2018; @Francis2002].
+However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC and declined in the early 2nd millennium [@Wang2018; @Francis2002].
 It indicates a different bead source more related to the exchange network in the South China Sea for the lower component.
@@
 All squares in the sampling area were occupied during all phases of habitation based on the continuous presence of archaeological materials across over the strata representing a time span of 600 years.
-Although the quantity of locally made pottery declines in the later period with the growth of the Chinese porcelains, the general density of the artifacts present consistent thoughout all three phases.
-
 Finally I computed the kernel densities for each time period for comparison.
@@
 
-Figure \@ref(fig:plot-all-densities) shows that there is one major core area during the pre-European period, multiple core areas during European period, and a single core during the Chinese period.
+Figure \@ref(fig:plot-all-densities) shows that there is one major core area during the pre-European period, multiple core areas during the European period, and a single core during the Chinese period.
 There are three consistent sub-regions with a core area that shifts over time.
 The distribution might indicate increase and decrease in the number of different social groups who possessed more ornaments.
-The multiple groups during the European period might reflect more unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from bigger sample size.
+The multiple groups during the European period might reflect more unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from a bigger sample size.
 In addition, the generation of core areas might be biased due to small sample sizes, for example, a few ornaments found at one single square during the Chinese period could create an obvious hotspot.
+The proportion of ornaments is relative to the number of locally made pottery that shows the pattern of decline in the 19th century.
+Although the quantity of locally made pottery declines in the 19th century with an obvious growth of the imported Chinese porcelains, the general density of the total number of artifacts is consistent throughout all three phases.
+However, less locally made pottery and ornaments might relate to the indigenous population decline.
+Whether the observed clustering is random or non-random is crucial for the interpretation of intentional human activities.
 
@@
 
-An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornaments types and materials during the European period.
+An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornament types and materials during the European period.
 Northeast Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish.
@@
 This could reflect ornaments were viewed as valuable objects such as status indicators and their distribution across the dwelling area would help to understand the social relation reflected by people who possessed them, and the reaction of the society as a whole to the colonial presence in the 17th century.
-Spatial patterns of ornaments shows that their distribution was clustered during the pre-European and European period.
+Spatial patterns of ornaments show that their distribution was clustered during the pre-European and European period.
 Testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European period.
-The spatial clustering during the pre-European and European period shows an increase in use of ornaments in the daily life at Kiwulan.
+The spatial clustering during the pre-European and the European period shows an increase in the use of ornaments in daily life at Kiwulan.
 It may indicate that a degree of social inequality based on uneven distribution was present before European contact and then it was reinforced and amplified during the European period.
@@
 Resistance could be presented in many forms.
-For example, @Rubertone discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands.
+For example, @Rubertone2000 discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands.
 Similarly, the frequent presence of ornaments across the dwelling area during the European suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition.
@@
 It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next.
-This might suggest that the ornaments found in Chinese period were old item continuously used until then.
-Further analysis about the source for ornaments from different periods will be helpful to understand their origin and reginal circulation.
+This might suggest that the ornaments found in the Chinese period were old items continuously used until then.
+Further analysis of the source for ornaments from different periods will be helpful to understand their origin and regional circulation.
 

### 2019-12-08 analysis/paper/paper.Rmd

@@
 kw_dates <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
+cal_kw_dates <- readr::read_csv(here("analysis", "data", "raw_data", "kwl_cal_dates.csv"))
+
 table_kw_dates <-
@@
   arrange(Period) %>% 
-  select("Pit-Layer", "Depth related to sea level (cm)", Period, "Date BP", "Calibrated date", "Details")
+  left_join(cal_kw_dates, "Pit-Layer") %>% 
+  select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details")
 
 knitr::kable(table_kw_dates,
-             caption = "Radiocarbon dates on charcoals from Kiwulan")
+             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007). Recalibrated using IntCal13 Atmospheric curve.")
 ```
@@
 Produced with the BChronDensity() function in program R's Bchron package (Parnell et al.
-2008).
-IntCal13 Atmospheric curve."}
+2008)."}
 
@@
        ))
+
+# get and summarise the 95% credible interval for each sample into a dataframe
+kw_age_samples <- sampleAges(kw_dates_with_error_calibrated)
+kw_age_cali_df <- as.data.frame(t(apply(kw_age_samples, 2, 
+                                        quantile, prob=c(0.025,0.975))))
+
+# tidy the dataframe 
+kw_age_cali_df_tidy <-
+  kw_age_cali_df %>% 
+  rownames_to_column(var = "Pit-Layer") %>% 
+  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`," - ", `97.5%`))
+
+# save the dataframe
+write.csv(kw_age_cali_df_tidy, here("analysis/data/raw_data/kwl_cal_dates.csv"))
 

### 2019-12-09 analysis/paper/paper.Rmd
added sentences about the context of ornamnets and two components

@@
 For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
-In this paper, I explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the period of Chinese presence.
-I address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
+In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the period of Chinese presence.
+We address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
 
@@
 Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
-The total area includes eight open area sections of 262 excavation squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
+The excavation area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
 The archaeological evidence includes a large amount of artefacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
-Most ornaments were found while digging and some were collected through screens with 2 mm and 1.5 mm mesh.
-The ornaments studied in this paper were found from 40 adjacent squares in the largest open area, located in the middle part of the excavation, see Figure \@ref(fig:KWL-sam-area).
+Ornaments were found in different archaeological contexts including post hole area, burials, and middens while digging and some small beads were collected through screens with 2 mm and 1.5 mm mesh.
+The ornaments studied in this paper were found from 40 adjacent squares in the largest section, located in the middle part of the excavation, see Figure \@ref(fig:KWL-sam-area).
 They were sampled because those units were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
@@
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
-This paper focuses on the upper component because this has a sequence spanning pre-European contact, European presence, and Chinese presence.
-Previous researchers divided the upper component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
-Due to ambiguities in the previous chronology, I re-examined the excavation records to produce a new chronology for the sampled squares.
-For each sampled square I separated excavation units into three episodes of time: pre-European period, European period, and Chinese period to indicate the dominant foreign influence and cultural interaction for different time periods.
+The relationship between upper component and lower component is still under debate and the sterile might result from dry weather according to a pollen study [@Lin2015; @Chiu2004; @Chen2007].
+This paper focuses on the upper component covering a sequence spanning pre-European contact, European presence, and Chinese presence.
+Previous researches divided the upper component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
+To answer our research question, we re-examined and assigned previous layers to a new chronology.
+For each sampled square we separated excavation units into three episodes of time: pre-European period, European period, and Chinese period to indicate the dominant foreign influence and cultural interaction for different time periods.
 
@@
 Either way indicates that the Europeans played an important role in introducing those jars to Taiwan.
-I used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.
+We used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.
 
@@
 Point pattern analysis can assess whether the distribution of artefacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments resulting from social inequality stimulated by a colonial presence.
-To prepare the ornament location data for point pattern analysis, I assigned each ornament to a random coordinate pair located in 4 x 4 m square it was recovered from (ornaments lack individual point provenance).
+To prepare the ornament location data for point pattern analysis, we assigned each ornament to a random coordinate pair located in 4 x 4 m square it was recovered from (ornaments lack individual point provenance).
 The next step was to divide the ornaments into three time periods.
 All squares in the sampling area were occupied during all phases of habitation based on the continuous presence of archaeological materials across over the strata representing a time span of 600 years.
-Finally I computed the kernel densities for each time period for comparison.
+Finally we computed the kernel densities for each time period for comparison.
 Kernel density estimations (KDE) compute the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
-Here I use KDE to visualize core areas of ornaments and surrounding neighbourhoods [@Bonnier2019; @Cortegoso2016].
+Here we use KDE to visualize core areas of ornaments and surrounding neighbourhoods [@Bonnier2019; @Cortegoso2016].
 Density values of artefacts per square meter were calculated for each cell.
@@
 
-To test for randomness in spatial locations, I used a Monte Carlo method to simulate average nearest-neighbour distances (ANN).
+To test for randomness in spatial locations, we used a Monte Carlo method to simulate average nearest-neighbour distances (ANN).
 Figure \@ref(fig:hypo-test-distribution) shows the observed ANN distances with the distributions of the ANN distances calculated on 1000 simulations of random ornament locations.
@@
 
-I would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the ornaments used in this study.
-I thank Shui-jin Chiu, Shu-hui Jian, Jiun-yao Lai, and the staff in archaeology lab for their invaluable assistance in preparing samples for analysis.
-I also thank Yu-pei Chen for the discussion regarding the site at the earliest stage of fieldwork.
+We would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the ornaments used in this study.
+We thank Shui-jin Chiu, Shu-hui Jian, Jiun-yao Lai, and the staff in archaeology lab for their invaluable assistance in preparing samples for analysis.
+We also thank Yu-pei Chen for the discussion regarding the site at the earliest stage of fieldwork.
 This work was supported in part by the Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
-I would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on the early draft.
-I especially thank Ben Marwick for providing academic support during this project, helping me with polishing figures with R language, and providing feedback on early drafts of this manuscript.
+We would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on the early draft.
 

### 2019-12-09 analysis/paper/paper.Rmd
change the code order

@@
 
+```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan.
+The y-axis represents summed probability and the x-axis is the date in calendar years.
+Produced with the BChronDensity() function in program R's Bchron package (Parnell et al.
+2008)."}
+
+kw_dates <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
+
+# subset dates with error value
+kw_dates_with_error <- 
+  kw_dates %>% 
+  filter(!is.na(se))
+
+# convert rce to numeric
+kw_dates_with_error$rce <- as.numeric(kw_dates_with_error$rce)
+
+# make some kind of ID
+kw_dates_with_error$id <- with(kw_dates_with_error,
+                               paste0(Square,"-",Layer))
+
+# Use Bchron to clibrate
+library(Bchron)
+
+kw_dates_with_error_calibrated <- 
+  with(kw_dates_with_error,
+       BchronCalibrate(
+         ages = rce,
+         ageSds = se,
+         ids = id,
+         calCurves = rep('intcal13',
+                         length(kw_dates_with_error$rce))
+       ))
+
+# get and summarise the 95% credible interval for each sample into a dataframe
+kw_age_samples <- sampleAges(kw_dates_with_error_calibrated)
+kw_age_cali_df <- as.data.frame(t(apply(kw_age_samples, 2, 
+                                        quantile, prob=c(0.025,0.975))))
+
+# tidy the dataframe 
+kw_age_cali_df_tidy <-
+  kw_age_cali_df %>% 
+  rownames_to_column(var = "Pit-Layer") %>% 
+  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`," - ", `97.5%`))
+
+# save the dataframe
+write.csv(kw_age_cali_df_tidy, here("analysis/data/raw_data/kwl_cal_dates.csv"))
+
+# have a look at the output
+summary(kw_dates_with_error_calibrated)
+par(mfrow = c(2, 3))
+plot(kw_dates_with_error_calibrated)
+
+# we can combine all the calibrated dates into a single plot, like this
+kw_ages_densities <- 
+  with(kw_dates_with_error,
+       BchronDensity(
+         ages = rce,
+         ageSds = se,
+         calCurves = rep('intcal13',
+                         length(kw_dates_with_error$rce))
+       ))
+
+#plot it
+plot(kw_ages_densities,
+     xlab= "Years BC")
+
+# to describe in your text, ?BchronDensity
+# "We used the Bchron R package to compute a non-parametric phase model 
+# on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
+# FIgure X shows the results of this analysis... indicating that the peak period of occupation is around 600 BC"
+```
+
 ```{r C14-table-data}
-kw_dates <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
 cal_kw_dates <- readr::read_csv(here("analysis", "data", "raw_data", "kwl_cal_dates.csv"))
@@
 knitr::kable(table_kw_dates,
-             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007). Recalibrated using IntCal13 Atmospheric curve.")
-```
-
-```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan.
-The y-axis represents summed probability and the x-axis is the date in calendar years.
-Produced with the BChronDensity() function in program R's Bchron package (Parnell et al.
-2008)."}
-
-# subset dates with error value
-kw_dates_with_error <- 
-  kw_dates %>% 
-  filter(!is.na(se))
-
-# convert rce to numeric
-kw_dates_with_error$rce <- as.numeric(kw_dates_with_error$rce)
-
-# make some kind of ID
-kw_dates_with_error$id <- with(kw_dates_with_error,
-                               paste0(Square,"-",Layer))
-
-# Use Bchron to clibrate
-library(Bchron)
-
-kw_dates_with_error_calibrated <- 
-  with(kw_dates_with_error,
-       BchronCalibrate(
-         ages = rce,
-         ageSds = se,
-         ids = id,
-         calCurves = rep('intcal13',
-                         length(kw_dates_with_error$rce))
-       ))
-
-# get and summarise the 95% credible interval for each sample into a dataframe
-kw_age_samples <- sampleAges(kw_dates_with_error_calibrated)
-kw_age_cali_df <- as.data.frame(t(apply(kw_age_samples, 2, 
-                                        quantile, prob=c(0.025,0.975))))
-
-# tidy the dataframe 
-kw_age_cali_df_tidy <-
-  kw_age_cali_df %>% 
-  rownames_to_column(var = "Pit-Layer") %>% 
-  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`," - ", `97.5%`))
-
-# save the dataframe
-write.csv(kw_age_cali_df_tidy, here("analysis/data/raw_data/kwl_cal_dates.csv"))
-
-# have a look at the output
-summary(kw_dates_with_error_calibrated)
-par(mfrow = c(2, 3))
-plot(kw_dates_with_error_calibrated)
-
-# we can combine all the calibrated dates into a single plot, like this
-kw_ages_densities <- 
-  with(kw_dates_with_error,
-       BchronDensity(
-         ages = rce,
-         ageSds = se,
-         calCurves = rep('intcal13',
-                         length(kw_dates_with_error$rce))
-       ))
-
-#plot it
-plot(kw_ages_densities,
-     xlab= "Years BC")
-
-# to describe in your text, ?BchronDensity
-# "We used the Bchron R package to compute a non-parametric phase model 
-# on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
-# FIgure X shows the results of this analysis... indicating that the peak period of occupation is around 600 BC"
+             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007). Calibrated using IntCal13 Atmospheric curve.")
 ```

### 2019-12-09 analysis/paper/paper.Rmd
Historical contexts about global and regional trade network relating to local indigenous societies

@@
 ---
-title: "Ornaments as indicators of social changes and cultural practices in northeast Taiwan before and after European colonial period"
+title: "Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period"
 author:
@@
   Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
-Trade ornaments from Kiwulan (1400-1900 AD) in northeast Taiwan show the indirect impacts of European colonial activities on local societies.
+Trade ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
 The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered.
@@
 Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
-Northeast Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in Taiwan, the northeast region was isolated from intensive direct contact by the Xueshan Mountains.
-
-This article describes ornaments excavated from Kiwulan (700-1900 AD), the largest Iron Age settlement on the Yilan plain in northeast Taiwan.
+Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in Taiwan, the northeastern region was isolated from intensive direct contact by the Xueshan Mountains.
+
+This article describes ornaments excavated from Kiwulan (700-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
 The first recorded European presence in Yilan was a Spanish revenge attack on indigenous villages in 1632 [@Borao2001, pp.
@@
 
-# Historical Background of complex exchange network during the late Iron Age and early historical period
-
-The island of Taiwan lies at the junction of mainland China, Southeast Asia, and northeast Asia in the Pacific Ocean.
+# Ornaments in complex exchange network during the late Iron Age and early historical period
+
+The island of Taiwan lies at the junction of mainland China, Southeast Asia, and Northeast Asia in the Pacific Ocean.
 The prehistory of Taiwan island could be roughly divided into three major periods, paleolithic (c.
@@
 2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages (@Liu2011; @Chen2017).
-It is generally accepted that Taiwan entered the historical period since the early 17th century due to the colonial presence of the Spanish and Dutch who played an important role in keeping written records about Taiwan.
+It is generally accepted that Taiwan entered the historical period since the early 17th century due to the colonial presence of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
+The European colonial presence in Taiwan ended in 1662 when the Dutch were defeated by the kingdom of Tungning founded by Koxinga from China.
+Later in 1683, Taiwan was incorporated into Qing dynasty and a large wave of Han Chinese migrated to Yilan during the late 18th century.
+
 The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
@@
 These historical accounts indicate that carnelian beads were already treated as prestige goods in Yilan before the arrival of Europeans.
-
 In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan.
-They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty [@Andrade2007, ch.
+They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tributes [@Andrade2007, ch.
 9; @Kang2016, ch.
 4].
-The Dutch provided beads and other goods based on the demands of Indigenous people to secure alliances in the annual ceremony or during travelling [@Kang2016, ch.
+The Dutch provided beads and other goods based on the demands of indigenous people to secure alliances in the annual ceremony or during travelling [@Kang2016, ch.
 6].
@@
 
-Currently the ornaments found in northeast Taiwan in the early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metalworking, or accessible local raw materials.
+Currently the ornaments found in northeastern Taiwan in the early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metalworking, or accessible local raw materials.
 The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
-Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011].
-There is no clear evidence shows the European origin of beads, however, a large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
-Northeast Taiwan from the late Iron age to the early historical period was involved in complex exchange networks both on a regional and global scale.
-
-# Excavations at Kiwulan in northeast Taiwan
+Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicates multiple origins more likely from Southeast Asia [@Chen2011].
+There is no clear evidence shows the European origin of beads, however, a large amount of golden beads at Kiwulan might be introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
+Based on the archaeological evidence and historical records, Northeastern Taiwan was involved in the regional network with East Asia in the late Iron age that Chinese merchants traded metal materials, clothes, and beads with local indigenous people in Taiwan for local resources and later was incorporated into more international scale due to the arrival of the Europeans.
+The common finds of foreign stoneware jars from Southeast Asia in local sites were frequently found in European shipwrecks that indicates the direct or indirect interaction.
+Despite the origin of some ornaments were traced to China, it is likely the large amount of them found at the 17th century sites were traded into due to indirect economic activities stimulated by the Europeans or direct European colonial activities.
+
+# Excavations at Kiwulan in northeastern Taiwan
 
@@
 
-The European colonial presence in northern Taiwan ended in 1662 when the Dutch were defeated by the kingdom of Tungning founded by Koxinga from China.
-Later in 1683, the Qing dynasty ruled over Taiwan and a large wave of Han Chinese migrated to Yilan during the late 18th century.
 The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups.
@@
 An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornament types and materials during the European period.
-Northeast Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish.
-The origin of ornaments might be multiple places mostly in Southeast Asia and introduced into northeast Taiwan by Chinese merchants before 17th century.
+Northeastern Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish.
+The origin of ornaments might be multiple places mostly in Southeast Asia and introduced into northeastern Taiwan by Chinese merchants before 17th century.
 Later, the trade activities became more frequent and intense in the 17th century that might be related to the colonial involvement of the Europeans based on the similar styles of golden beads found in Luzon that was once the colony of the Spanish during this period.
@@
 Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
-Although historical records indicate that trade ornaments were prestige goods in prehistoric northeast Taiwan, the information might be biased that needs to be compared with archaeological evidence.
+Although historical records indicate that trade ornaments were prestige goods in prehistoric northeastern Taiwan, the information might be biased that needs to be compared with archaeological evidence.
 
@@
 The long-distance trade network introduced by Europeans resulted in the appearance of a network mode due to competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
-Because of the weak direct control from the European colonizers in northeast Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
+Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
 
@@
 Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017].
-Kiwulan in northeast Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
+Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
 Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation.

### 2019-12-10 analysis/paper/paper.Rmd
added one reference and editted the conclusion section

@@
 The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
-A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh\\].
+A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh1995].
 The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007].
@@
 The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
-The greater amount and diversity of ornament types during the European period reflects international exchanges in a colonial context.
-Ornaments were viewed as prestige goods in the local indigenous culture before European contact.
+The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context.
+Before European contact, ornaments were traded into local indigenous societies via the regional exchagne network with Chinese merchants and viewed as prestige goods in local indigenous culture according to their distribution in the archaeological contexts.
 After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan.
 These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
+This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that is part of culture tradition.
 
 We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities.
-The Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
+The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods.
+By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
 Ornaments give insights into the amplification of social inequality stimulated by European colonisation.

### 2019-12-10 analysis/paper/paper.Rmd
editted the background section

@@
 6500- 2000BP), and Iron age (c.
-2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages (@Liu2011; @Chen2017).
+2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages [@Liu2011; @Chen2017].
 It is generally accepted that Taiwan entered the historical period since the early 17th century due to the colonial presence of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
 The European colonial presence in Taiwan ended in 1662 when the Dutch were defeated by the kingdom of Tungning founded by Koxinga from China.
-Later in 1683, Taiwan was incorporated into Qing dynasty and a large wave of Han Chinese migrated to Yilan during the late 18th century.
-
-The Spanish founded Fort San Salvador at Keelung in 1626 and Fort San Domingo in 1629 at Tamsui, and sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
-343].
-A report by a Dominican priest in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups.
-This form of exchange was widespread and even the Spanish soldiers learned to use carnelian as bargaining chips for gambling [@LiandWu2006, pp.
+Later in 1683, Taiwan was incorporated into the Qing dynasty in China and a large wave of Han Chinese migrated to Taiwan during the late 18th century.
+Because of natural harbours, northeast Taiwan was involved in a regional trade network through cross culture interactions with Chinese merchants since 14th century, and later the global trade network with the Europeans in the 17th century who brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
+Although on the peripheral location of the trade center, northeastrn Taiwan- Yilan was connected to the trade networks via visits of other indigenous groups, Chineses merchants, and the Europeans via sea.
+
+The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Keelung in 1626, and Fort San Domingo in 1629 at Tamsui.
+They sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
+343] who left many records about the life of indigenous.
+Recording the trade beads, a report by a Dominican priest in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups.
+This form of exchange was widespread and even the Spanish soldiers learned to use carnelian beads as bargaining chips for gambling [@LiandWu2006, pp.
 132-149].
@@
 
-Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty (1616-1911) contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
+Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty (1616-1911) contain some notes on the purposes of ornaments from Yilan, northeast Taiwan [@Chen1963, pp.
 228, 308; @Ke1993, pp.
@@
 Compared to the European period, there are fewer mentions of beads in the Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
-However, we must recognize the inherent biases in historical records and the fragmented nature of texts selected by written record keepers that might lead to misinterpretations of artifacts [@Galloway2006].
-Inconsistent and incomplete descriptions from different written sources lacking contexts are the problems need to be compared and examined by studying the ornaments in archaeological contexts.
+However, we must recognize the inherent biases in historical records and the fragmented nature of texts generated by the writers [@Galloway2006].
+Inconsistent and incomplete descriptions from different written sources need to be compared and examined by studying the ornaments in archaeological contexts.
 
@@
 Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicates multiple origins more likely from Southeast Asia [@Chen2011].
-There is no clear evidence shows the European origin of beads, however, a large amount of golden beads at Kiwulan might be introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
-Based on the archaeological evidence and historical records, Northeastern Taiwan was involved in the regional network with East Asia in the late Iron age that Chinese merchants traded metal materials, clothes, and beads with local indigenous people in Taiwan for local resources and later was incorporated into more international scale due to the arrival of the Europeans.
-The common finds of foreign stoneware jars from Southeast Asia in local sites were frequently found in European shipwrecks that indicates the direct or indirect interaction.
-Despite the origin of some ornaments were traced to China, it is likely the large amount of them found at the 17th century sites were traded into due to indirect economic activities stimulated by the Europeans or direct European colonial activities.
+There is no clear evidence showing the European origin of beads, however, a large amount of golden beads at Kiwulan might be introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
+Both archaeological evidence and historical records indicating northeastern Taiwan was involved in the regional network with East Asia in the late Iron age that Chinese merchants traded metal materials, clothes, and beads with local indigenous people in Taiwan for local resources and later incorporated into more international scale due to the arrival of the Europeans.
+The foreign stoneware jars frequently found in European shipwrecks were also commonly found from many sites in Taiwan that suggests a direct or indirect interaction.
+Despite the origin of some ornaments were traced to China, it is likely that a large amount of ornament found at the 17th century sites were an influence of the Europeans colonial and economic activities.
 
@@
 Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
-The excavation area includes eight open area sections of 262 squares (4 m by 4 m) reaching 3,814 m^2^ in total [@Chen2007].
-The archaeological evidence includes a large amount of artefacts, burials, middens, post holes, wooden pillars, and stone structures, all of which indicates it was a settlement.
-Ornaments were found in different archaeological contexts including post hole area, burials, and middens while digging and some small beads were collected through screens with 2 mm and 1.5 mm mesh.
-The ornaments studied in this paper were found from 40 adjacent squares in the largest section, located in the middle part of the excavation, see Figure \@ref(fig:KWL-sam-area).
-They were sampled because those units were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
+The excavation includes eight open areas in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
+The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-tern settlement.
+The center of the site is the open area consists of A and D section which is also the study area where our samples selected from.
+In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
+At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction (Figure \@ref(fig:KWL-sam-area)).
 
@@
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
-The relationship between upper component and lower component is still under debate and the sterile might result from dry weather according to a pollen study [@Lin2015; @Chiu2004; @Chen2007].
+The relationship between upper component and lower component is still under debate and the sterile might result from dry weather according to pollen analyses [@Lin2015; @Chiu2004; @Chen2007].
 This paper focuses on the upper component covering a sequence spanning pre-European contact, European presence, and Chinese presence.
 Previous researches divided the upper component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
-To answer our research question, we re-examined and assigned previous layers to a new chronology.
-For each sampled square we separated excavation units into three episodes of time: pre-European period, European period, and Chinese period to indicate the dominant foreign influence and cultural interaction for different time periods.
+To answer our research question, we re-examined and assigned previous layers to a new chronology with three episodes of time: pre-European period, European period, and Chinese period to indicate the dominant foreign influence and cultural interaction for different time periods.
 
@@
 The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
-Those jars might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan.
-Either way indicates that the Europeans played an important role in introducing those jars to Taiwan.
-We used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.
-
+They might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan.
+Either way indicates that the Europeans played an important role in introducing foreign jars to Taiwan.
+Those jars were widely distributed across the site and served as indicators, together with the radiocarbon dates, to identify the excavation unites associated with the pre-European and the start of the European influence at Kiwulan.
 The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups.
@@
 
-This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aliened postholes with in-situ posts (Figure \@ref(fig:KWL-sam-area)).
-These postholes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
-At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction across the sampling area.
+Ornaments were found in different archaeological contexts including post-holes area, burials, and middens while digging and some small beads were collected through screens with 2 mm and 1.5 mm mesh.
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aliened post-holes with in-situ posts (Figure \@ref(fig:KWL-sam-area)).
+They were sampled because those units were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
 There are 30 burials in the sampling area account for one third of the total number of burials at Kiwulan.

### 2019-12-10 analysis/paper/paper.Rmd
some more edits

@@
 Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
-The excavation includes eight open areas in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
+The excavation includes eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
 The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-tern settlement.
-The center of the site is the open area consists of A and D section which is also the study area where our samples selected from.
+Based on the continuous deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from.
 In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
@@
 Those jars were widely distributed across the site and served as indicators, together with the radiocarbon dates, to identify the excavation unites associated with the pre-European and the start of the European influence at Kiwulan.
+In addition to ornaments and stoneware jars, around 300 pieces of local made pipes and a few imported pipes were viewed as a custom introduced by the Europeans.
+
 The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups.
@@
 This might simplify the subtypes of glass beads when they are the same size and shape but in different color or texture.
-However, their variety would be considered and discussed separately.
-The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to their use in the burial contexts.
+However, the variation would be considered and discussed separately.
+The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to the use in burial contexts.
 
@@
 However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period in the early 19th century.
-This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the practices of Han Chinese who did not use these kinds of beads in their dress.
+This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, or adopting the practices of Han Chinese who did not use these kinds of beads in their dress.
 Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
@@
 For example, @Rubertone2000 discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands.
-Similarly, the frequent presence of ornaments across the dwelling area during the European suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition.
-During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan.
-This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or population decrease.
+Similarly, the frequent presence of ornaments across the dwelling area during the European period suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition.
+During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the use and discard of ornaments at Kiwulan.
+This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or indeigenous population decline.
 It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next.

### 2019-12-10 analysis/paper/paper.Rmd
some edits and updates for plots after knitting

@@
 
-This article describes ornaments excavated from Kiwulan (700-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
+This article describes ornaments excavated from the upper component of Kiwulan (1400 AD-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
 The first recorded European presence in Yilan was a Spanish revenge attack on indigenous villages in 1632 [@Borao2001, pp.
@@
 Later in 1683, Taiwan was incorporated into the Qing dynasty in China and a large wave of Han Chinese migrated to Taiwan during the late 18th century.
-Because of natural harbours, northeast Taiwan was involved in a regional trade network through cross culture interactions with Chinese merchants since 14th century, and later the global trade network with the Europeans in the 17th century who brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
-Although on the peripheral location of the trade center, northeastrn Taiwan- Yilan was connected to the trade networks via visits of other indigenous groups, Chineses merchants, and the Europeans via sea.
+Because of natural harbors, northeast Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
+Although on the peripheral location of the trade center, northeastern Taiwan- Yilan was connected to the trade networks via visits of other indigenous groups, Chinese merchants, and the Europeans via sea.
 
@@
 Both archaeological evidence and historical records indicating northeastern Taiwan was involved in the regional network with East Asia in the late Iron age that Chinese merchants traded metal materials, clothes, and beads with local indigenous people in Taiwan for local resources and later incorporated into more international scale due to the arrival of the Europeans.
-The foreign stoneware jars frequently found in European shipwrecks were also commonly found from many sites in Taiwan that suggests a direct or indirect interaction.
-Despite the origin of some ornaments were traced to China, it is likely that a large amount of ornament found at the 17th century sites were an influence of the Europeans colonial and economic activities.
+The foreign stoneware jars frequently found in European shipwrecks were also commonly found from many sites in Taiwan that suggest a direct or indirect interaction.
+Despite the origin of some ornaments were traced to China, it is likely that a large amount of ornament found at the 17th century sites was an influence of the Europeans colonial and economic activities.
 
@@
   left_join(cal_kw_dates, "Pit-Layer") %>% 
-  select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details")
+  dplyr::select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details")
 

### 2019-12-10 analysis/paper/paper.Rmd
udpates c14 dating data to include more samples

@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD) and a lower component (700-1200 AD) separated by a sterile.
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BC) and a lower component (700-1200 AD, 1200-800 BC) separated by a sterile spanning 150 years according to .
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
@@
 # have a look at the output
-summary(kw_dates_with_error_calibrated)
 par(mfrow = c(2, 3))
@@
        ))
+invisible(dev.off())
 

### 2019-12-10 analysis/paper/paper.Rmd
plot dates with ggplot2

--- 2019-12-10 analysis/paper/paper.Rmd
@@
 
-# save the dataframe
-write.csv(kw_age_cali_df_tidy, here("analysis/data/raw_data/kwl_cal_dates.csv"))
-
 # have a look at the output
@@
   with(kw_dates_with_error,
+       rrtools:::quietly( # suppress the progress bar
          BchronDensity(
@@
                          length(kw_dates_with_error$rce))
-       ))
+       ,
+       store_density=TRUE))
+  )
 
 #plot it
-plot(kw_ages_densities,
-     xlab= "Years BC")
+plot(kw_ages_densities)
+
+kw_ages_densities_df <- 
+  tibble(ageGrid = kw_ages_densities$ageGrid,
+         densities = kw_ages_densities$densities)
+
+kw_ages_densities_dfs <- 
+  map_df(kw_ages_densities$calAges, 
+      ~tibble(ageGrid = .x$ageGrid, 
+              densities = .x$densities), .id = 'date') %>% 
+  mutate(dens_to_one = densities / sum(densities))
+
+lower_component <- c(1300, 800)
+upper_component <- c(600, 100)
+
+ggplot() +
+    geom_vline(xintercept = c(324,   # start of Euro contact, 1626 AD
+                            182),  # start of Chinese contact, 1768 AD
+                            colour = "red") +
+    geom_area(data = kw_ages_densities_dfs,
+            aes(ageGrid,
+                dens_to_one,
+                group = date),
+            fill = "grey80",
+            alpha = 0.4,
+            position = 'identity') +
+  geom_line(data = kw_ages_densities_df,
+            aes(ageGrid,
+                densities)) +
+  annotate("rect", xmin = 100, xmax = 600, # upper_component
+                   ymin = 0, ymax = Inf,
+           fill = "grey80",
+           alpha = 0.3) +
+    annotate("text", x = 350, y = 0.007,
+           label = "Upper component") +
+    annotate("rect", xmin = 1300, xmax = 800,
+                   ymin = 0, ymax = Inf,
+           fill = "grey80",
+           alpha = 0.3) +
+      annotate("text", x = 1050, y = 0.007,
+           label = "Lower component")  +
+  theme_minimal(base_size = 14) +
+  xlab("calibrated years BP") +
+  ylab("Density")
 

### 2019-12-10 analysis/paper/paper.Rmd
update C14 ages plot

--- 2019-12-10 analysis/paper/paper.Rmd
@@
   cache = TRUE,
+  dpi = 300,
   fig.path = "../figures/"
@@
 ```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan.
-The y-axis represents summed probability and the x-axis is the date in calendar years.
-Produced with the BChronDensity() function in program R's Bchron package (Parnell et al.
-2008)."}
+The dark line represents the summed probabilities of all radiocarbon ages, and the grey lines in the background are the probabilities of individual ages.
+Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit.
+Ages calibrated with the Bchron package (Parnell et al.
+2008).
+"}
 
@@
 
-# save the dataframe
-write.csv(kw_age_cali_df_tidy, here("analysis/data/raw_data/kwl_cal_dates.csv"))
-
-# have a look at the output
-par(mfrow = c(2, 3))
-plot(kw_dates_with_error_calibrated)
-
 # we can combine all the calibrated dates into a single plot, like this
@@
                          length(kw_dates_with_error$rce))
-       ))
-invisible(dev.off())
-
-#plot it
-plot(kw_ages_densities,
-     xlab= "Years BC")
+       ,
+       store_density=TRUE))
+  )
+
+# plot it
+# plot(kw_ages_densities)
+
+kw_ages_densities_df <- 
+  tibble(ageGrid = kw_ages_densities$ageGrid,
+         densities = kw_ages_densities$densities)
+
+kw_ages_densities_dfs <- 
+  map_df(kw_ages_densities$calAges, 
+      ~tibble(ageGrid = .x$ageGrid, 
+              densities = .x$densities), .id = 'date') %>% 
+  mutate(dens_to_one = densities / sum(densities))
+
+lower_component <- c(1300, 800)
+upper_component <- c(600, 100)
+
+lines_for_periods <- 
+  tibble(bx = c(324,   # start of Euro contact, 1626 AD
+               182),  # start of Chinese contact, 1768 AD
+          by = c(0, 0),
+          byend = c(0.00175),
+          bxend = bx)
+
+ggplot() +
+    geom_area(data = kw_ages_densities_dfs,
+            aes(ageGrid,
+                dens_to_one,
+                group = date),
+            fill = "grey80",
+            alpha = 0.4,
+            position = 'identity') +
+  geom_line(data = kw_ages_densities_df,
+            aes(ageGrid,
+                densities)) +
+  geom_segment(data = lines_for_periods,
+              aes(x = bx, 
+                  y = by, 
+                  xend = bxend, 
+                  yend = byend),
+               colour = c("red", "blue")) +
+  annotate("rect", xmin = 100, xmax = 600, 
+                   ymin = 0, ymax = lines_for_periods$byend[1],
+           fill = "grey80",
+           alpha = 0.3) +
+    annotate("text", x = 350, y = 0.0022,
+           label = "Upper component",
+           size = 2) +
+    annotate("rect", xmin = 1300, xmax = 800,
+                   ymin = 0, ymax = lines_for_periods$byend[1],
+           fill = "grey80",
+           alpha = 0.3) +
+      annotate("text", x = 1050, y = 0.00220,
+           label = "Lower component",
+           size = 2)  +
+  theme_minimal(base_size = 8) +
+  xlab("calibrated years BP") +
+  ylab("Density") +
+  coord_fixed(ratio = 400000)
 
@@
 ```{r C14-table-data}
-cal_kw_dates <- readr::read_csv(here("analysis", "data", "raw_data", "kwl_cal_dates.csv"))
-
 table_kw_dates <-
@@
   arrange(Period) %>% 
-  left_join(cal_kw_dates, "Pit-Layer") %>% 
+  left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
   dplyr::select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details")

### 2019-12-11 analysis/paper/paper.Rmd
edits to text

@@
 Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
-Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in Taiwan, the northeastern region was isolated from intensive direct contact by the Xueshan Mountains.
-
-This article describes ornaments excavated from the upper component of Kiwulan (1400 AD-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
+Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in parts of Taiwan, the northeastern region was isolated from intensive direct contact by the Xueshan Mountains.
+
+This article describes personal ornaments excavated from the upper component of Kiwulan (1400 AD-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
 The first recorded European presence in Yilan was a Spanish revenge attack on indigenous villages in 1632 [@Borao2001, pp.
@@
 
-One of the most commonly traded types of object in this region were personal ornaments such as glass and stone beads [@Chen2007; @Li2014; @NMTH2005].
+One of the most commonly traded types of object in this region were ornaments such as glass and stone beads [@Chen2007; @Li2014; @NMTH2005].
 Personal adornments in the archaeological record are useful as signal of an individual's status [@Joyce2005; @Scaramelli2005].
-For example, the consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
-In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the period of Chinese presence.
+The consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
+In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the following period of Chinese presence.
 We address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
@@
 2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages [@Liu2011; @Chen2017].
-It is generally accepted that Taiwan entered the historical period since the early 17th century due to the colonial presence of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
-The European colonial presence in Taiwan ended in 1662 when the Dutch were defeated by the kingdom of Tungning founded by Koxinga from China.
+It is generally accepted that Taiwan entered the historical period in the early 17th century due to the colonial activities of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
+The European colonial presence in Taiwan ended in 1662 when the Dutch were defeated by the kingdom of Tungning, founded by Koxinga from China.
 Later in 1683, Taiwan was incorporated into the Qing dynasty in China and a large wave of Han Chinese migrated to Taiwan during the late 18th century.
-Because of natural harbors, northeast Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
-Although on the peripheral location of the trade center, northeastern Taiwan- Yilan was connected to the trade networks via visits of other indigenous groups, Chinese merchants, and the Europeans via sea.
+Because of natural safe harbors, northeast Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
+Although located on the periphery of regional trade centers, Yilan was connected to trade networks via visits of other indigenous groups, Chinese merchants, and Europeans, via sea.
 
 The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Keelung in 1626, and Fort San Domingo in 1629 at Tamsui.
-They sent missionaries to local indigenous settlements in this region for religious purposes [@Blusse2000, pp.
-343] who left many records about the life of indigenous.
-Recording the trade beads, a report by a Dominican priest in 1632 mentioned that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups.
-This form of exchange was widespread and even the Spanish soldiers learned to use carnelian beads as bargaining chips for gambling [@LiandWu2006, pp.
+They sent missionaries to local indigenous settlements in this region [@Blusse2000, pp.
+343] and kept records about their observations of indigenous communities.
+A Dominican priest in 1632 reported that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups.
+This form of exchange was widespread and even the Spanish soldiers used carnelian beads as bargaining chips for gambling [@LiandWu2006, pp.
 132-149].
-The use of beads as prestige goods is indicated by their role in bride price payments, and compensation to resolve disputes [@LiandWu2006, pp.
+The use of beads as prestige goods is further indicated by their role in bride price payments, and compensation to resolve disputes [@LiandWu2006, pp.
 132-149].
-Other records mention that the female shamans (majuorbol) in the tribe would use carnelian beads as magical items in ritual practice for body healing [@Borao2009, pp.
+Other records mention that the female shamans in the tribe would use carnelian beads as magical items in ritual healing practices [@Borao2009, pp.
 122-151].
-Records of an indigenous funeral process document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate a family’s higher prestige [@LiandWu2006, pp.
+Records of an indigenous funeral document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp.
 153].
-These historical accounts indicate that carnelian beads were already treated as prestige goods in Yilan before the arrival of Europeans.
+While a full critical analysis of these historical accounts remains to be produced, we take them to minimally indicate that carnelian beads were already treated as prestige goods in Yilan before the arrival of Europeans.
 In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan.
@@
 4].
-The Dutch provided beads and other goods based on the demands of indigenous people to secure alliances in the annual ceremony or during travelling [@Kang2016, ch.
+The Dutch provided beads and other goods based on negotations with indigenous communities to secure alliances in the annual ceremony or during travelling [@Kang2016, ch.
 6].
-We might predict that the Dutch feudal system resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.
-
-Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty (1616-1911) contain some notes on the purposes of ornaments from Yilan, northeast Taiwan [@Chen1963, pp.
+We might predict that the activities of the Dutch feudal system to build and maintain alliances resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.
+
+Chinese historical records from 1829, 1837, and 1852 during the Qing dynasty (1616-1911) contain some notes on the purposes of ornaments from Yilan [@Chen1963, pp.
 228, 308; @Ke1993, pp.
@@
 According to those records, indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
-Among those ornaments, fish-shaped necklaces made of metal threads had high value due to their delicacy and the materials invested in production.
+Among those ornaments, fish-shaped necklaces made of metal threads had high value due to their delicacy and the exotic materials invested in production.
 These were usually possessed by wealthy people.
 Other people wore carnelian beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonisation, an academic field survey for plains indigenous groups reported that golden necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of Japanese colonisation, an academic field survey for plains indigenous groups reported that fish-shaped metal necklaces necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
-In sum, we find multiple sources describing how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
-Compared to the European period, there are fewer mentions of beads in the Chinese period and the descriptions are limited to clothing, but generally confirm the role of beads as status markers.
-However, we must recognize the inherent biases in historical records and the fragmented nature of texts generated by the writers [@Galloway2006].
-Inconsistent and incomplete descriptions from different written sources need to be compared and examined by studying the ornaments in archaeological contexts.
-
-Currently the ornaments found in northeastern Taiwan in the early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metalworking, or accessible local raw materials.
-The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
-Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicates multiple origins more likely from Southeast Asia [@Chen2011].
-There is no clear evidence showing the European origin of beads, however, a large amount of golden beads at Kiwulan might be introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007].
-Both archaeological evidence and historical records indicating northeastern Taiwan was involved in the regional network with East Asia in the late Iron age that Chinese merchants traded metal materials, clothes, and beads with local indigenous people in Taiwan for local resources and later incorporated into more international scale due to the arrival of the Europeans.
-The foreign stoneware jars frequently found in European shipwrecks were also commonly found from many sites in Taiwan that suggest a direct or indirect interaction.
-Despite the origin of some ornaments were traced to China, it is likely that a large amount of ornament found at the 17th century sites was an influence of the Europeans colonial and economic activities.
+Although these historical records are fragmentary and may contain some biases [@Galloway2006] that have not yet been studied in detail, we find consistency among multiple sources in their descriptions of how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
+Compared to the European period, there are fewer documentary mentions of beads in the Chinese period and the descriptions are limited to clothing, but these generally confirm the role of beads as status markers.
+
+Ornaments found in northeastern Taiwan in the early historical period, including glass beads, stone beads, and metal ornaments, are considered to have been imported from other regions.
+This is because of to a lack of archaeological evidence of beadmaking waste, metalworking, or accessible local raw materials.
+The chemical composition of glass beads from this region shows a high content of lead and, toghether with the winding/folding technique, these details suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
+Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common components of metal ornaments are brass and copper, with a small number made from lead and tin that indicates multiple origins that include Southeast Asia [@Chen2011].
+There is no direct evidence showing European delivery of beads, however, a large amount of the glass beads containing gold foil (hereafter, gold-foil beads) at Kiwulan might have been introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, as part of the trading route of the Spanish between 16-19th century [@Wang2007].
+Both archaeological evidence and historical records indicate northeastern Taiwan was involved inregional networks with East Asia in the late Iron age.
+These included Chinese merchants trading metal items, clothes, and beads with local indigenous people in Taiwan in exchange for local resources.
+The foreign-made large dark brown glazed stoneware jars frequently found in European shipwrecks were also commonly found from many sites in Taiwan, suggesting direct or indirect interactions.
+Despite the Chinese origin of some ornaments at Kiwulan, there is compelling evidence that a large amount of ornaments found at 17th century sites resulted from European colonial and economic activities in the region.
 
@@
 
-Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge.
-The excavation includes eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
-The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-tern settlement.
-Based on the continuous deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from.
+Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that during 2001-2004 in advance of a water diversion project and road bridge construction.
+The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
+The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement.
+Artefact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
+Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from.
 In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BC) and a lower component (700-1200 AD, 1200-800 BC) separated by a sterile spanning 150 years according to .
-This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
-The relationship between upper component and lower component is still under debate and the sterile might result from dry weather according to pollen analyses [@Lin2015; @Chiu2004; @Chen2007].
-This paper focuses on the upper component covering a sequence spanning pre-European contact, European presence, and Chinese presence.
-Previous researches divided the upper component into six layers spanning from the 14th century to the 19th century with hundred year intervals according to the radiocarbon dates, excavation depth, consistency of contexts, and types of diagnostic porcelains such as blue and white porcelains [@Hsieh2009; @Wang2011].
-To answer our research question, we re-examined and assigned previous layers to a new chronology with three episodes of time: pre-European period, European period, and Chinese period to indicate the dominant foreign influence and cultural interaction for different time periods.
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BC) and a lower component (700-1200 AD, 1200-800 BC) separated by a sterile deposit spanning c.
+150 years.
+These component divisions are based on the differences in the colour and texture of the deposit.
+The interpration of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007].
+There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
+We focus on the upper component because only this componant spans the periods of pre-European contact, European presence, and Chinese presence.
+In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases.
+Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011].
+However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artefacts to the pre-European, European, and the Chinese periods.
 
 The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
-A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh1995].
-The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007].
-Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages.
+Large dark brown glazed stoneware jars may have been made in Southeast Asia, and are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
+The earliest evidence oflight gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Hsieh2009; @Hsieh1995].
+Southeast China is assumed to be the origin of the light gray glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, such as the Zeelandia fort site in Tainan [@Wang2007].
 The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
-They might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan.
-Either way indicates that the Europeans played an important role in introducing foreign jars to Taiwan.
-Those jars were widely distributed across the site and served as indicators, together with the radiocarbon dates, to identify the excavation unites associated with the pre-European and the start of the European influence at Kiwulan.
-In addition to ornaments and stoneware jars, around 300 pieces of local made pipes and a few imported pipes were viewed as a custom introduced by the Europeans.
+We cannot be sure of the exact process that bought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by indigenous groups via regional networks in north Taiwan.
+In any case, the high volume of ceramics transported by Europeans, and their high mobility in shipping trade played an important role in introducing foreign jars to Taiwan.
+Those jars were widely distributed across the site and can serve as indicators, together with the radiocarbon dates, to identify the excavation units associated with the pre-European period and the start of European influence at Kiwulan.
+In addition to stoneware jars as indicators of European presence, around 300 pieces of locally made pipes and a few imported pipes were found at Kiwulan.
+Smoking is likely to have been introduced by  Europeans, and we found that the presence of pipes in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
 
@@
 Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles used by Chinese [@Hsieh2009].
-Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrant in 1768 [@Chen1963; @Ke1993].
+Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrants in 1768 [@Chen1963; @Ke1993].
 
@@
 
-# The ornaments
+# The personal ornaments
 
@@
   ornaments_count %>% 
-  filter(Categories == "Golden bead") %>% 
+  filter(Categories == "Gold-foil bead") %>% 
   pull(n)
@@
 ```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
-A: carnelian beads, B: bells, C: glass beads and golden beads, D: metal rings.
+A: carnelian beads, B: bells, C: glass beads and gold-foil beads, D: metal rings.
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
@@
 
-Ornaments were found in different archaeological contexts including post-holes area, burials, and middens while digging and some small beads were collected through screens with 2 mm and 1.5 mm mesh.
-This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aliened post-holes with in-situ posts (Figure \@ref(fig:KWL-sam-area)).
-They were sampled because those units were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
-There are 30 burials in the sampling area account for one third of the total number of burials at Kiwulan.
-Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes is 3,173.
-The use of ornaments as body adornments in burials will be discussed to understand the function of ornaments, but not statistically analysed since the number of ornaments is highly affected by the presence of bead strands or patterned bands, which sometimes contains thousands of beads in an individual burial [@Chen2007].
-In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports the argument that ornaments were treated as high-value items at Kiwulan based on a few finds in the midden contexts compared to high frequency in burial contexts.
-It is important to analyse the detailed distribution of ornaments from the dwelling place across different time periods as it can provide key information concerning social inequality indicated by uneven access to high-value items in relation to the foreign influences.
-
-The ornaments found in the Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), carnelian beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)).
-The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
-The criteria for classification of subtype is based on their dimensional attributes that are important to understand the variation in shapes.
-This might simplify the subtypes of glass beads when they are the same size and shape but in different color or texture.
-However, the variation would be considered and discussed separately.
-The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to the use in burial contexts.
+Ornaments were found in a variety of archaeological contexts including post-holes area, burials, and middens.
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
+Occupation floors were not identified during excavation.
+We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
+There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan.
+Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173.
+The high number of ornaments in these burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+We exclude burials from this analysis because most of them date to the pre-European period (n = 26), limiting the usefulness of comparisons between the periods.
+In addition to `r sam_post_hole`  ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts.
+We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.
 
@@
 To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualisations contained in this paper is included in http://doi.org/10.17605/OSF.IO/R8YGA.
-Also in this version-controlled compendium [@Marwick2018] are the raw data for all the tests reported here.
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualisations and tests reported here.
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 Most ornament types were present before European contact.
-Ornament frequencies reached a peak during the European period and then dropped during the Chinese period, especially golden beads.
+Ornament frequencies reached a peak during the European period and then dropped during the Chinese period, especially gold-foil beads.
 This trend can be also seen on other ornaments including carnelian beads, metal rings, and bells.
@@
                             "Glass ring",
-                            "Golden bead"))  %>% 
+                            "Gold-foil bead"))  %>% 
   ungroup() %>% 
@@
 Although glass beads have less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have a wide variety of colors or patterns mostly made by winding technique with high lead content in composition indicating possibly from China [@Cheng2008].
-Despite more researches needed to understand the specific source of ornaments, current researches suggest that glass beads and metal ornaments have similar technique and composition to those found in China [@Chen2011; @Wang2018].
-There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component (1400-1900 AD).
-However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC and declined in the early 2nd millennium [@Wang2018; @Francis2002].
-It indicates a different bead source more related to the exchange network in the South China Sea for the lower component.
+Although we are not certain of the specific origin of these beads, research suggest that these glass beads and metal ornaments have similar production techniques and composition to those found in China [@Chen2011; @Wang2018].
+There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component of Kiwulan (1400-1900 AD).
+However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC and declining during the early 2nd millennium [@Wang2018; @Francis2002].
 
@@
   ornaments_period_join %>% 
-  filter(Categories %in% c('Golden bead', 
+  filter(Categories %in% c('Gold-foil bead', 
                            'Glass bead',
@@
 During the Chinese period the distribution is more even again.
-Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European period, such as golden beads and carnelian beads.
-However, there seems to be no clear consistent pattern across those different ornaments.
-Each class shows its own pattern where the squares with higher number of ornaments distributed separately and independently.
-For example, a cluster of golden beads was found at the northern part, while a cluster of carnelian beads was found in the middle part.
+Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European period, such as gold-foil beads and carnelian beads.
+However, there seems to be no consistent pattern across those different ornaments.
+Each class shows its own pattern where the squares with higher numbers of ornaments distributed separately and independently.
+For example, a cluster of gold-foil beads was found at the northern part, while a cluster of carnelian beads was found in the middle part.
 In contrast, there are multiple clusters of metal rings that are distributed separately across the research area.
@@
 
-Point pattern analysis can assess whether the distribution of artefacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments resulting from social inequality stimulated by a colonial presence.
-To prepare the ornament location data for point pattern analysis, we assigned each ornament to a random coordinate pair located in 4 x 4 m square it was recovered from (ornaments lack individual point provenance).
+Point pattern analysis can assess whether the distribution of artefacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
+To prepare the ornament location data for point pattern analysis, because artefacts from Kiwulan lack exact piece-provance data, we assigned each ornament to a random coordinate pair in the square it was recovered from.
 The next step was to divide the ornaments into three time periods.
-All squares in the sampling area were occupied during all phases of habitation based on the continuous presence of archaeological materials across over the strata representing a time span of 600 years.
 Finally we computed the kernel densities for each time period for comparison.
@@
 There are three consistent sub-regions with a core area that shifts over time.
-The distribution might indicate increase and decrease in the number of different social groups who possessed more ornaments.
-The multiple groups during the European period might reflect more unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from a bigger sample size.
+The distribution might indicate the increase and decrease in the number of social groups who possessed ornaments.
+The multiple groups during the European period might reflect unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from a bigger sample size.
 In addition, the generation of core areas might be biased due to small sample sizes, for example, a few ornaments found at one single square during the Chinese period could create an obvious hotspot.
-The proportion of ornaments is relative to the number of locally made pottery that shows the pattern of decline in the 19th century.
-Although the quantity of locally made pottery declines in the 19th century with an obvious growth of the imported Chinese porcelains, the general density of the total number of artifacts is consistent throughout all three phases.
-However, less locally made pottery and ornaments might relate to the indigenous population decline.
 Whether the observed clustering is random or non-random is crucial for the interpretation of intentional human activities.
@@
 ```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values from 1000 simulations for three time periods.
-X values represent ANN expected value under a completely random process resulted from simulated pattern.
+X-axis values represent ANN expected value under a completely random process resulted from simulated pattern.
 Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value."}
@@
 The observed distribution of ornaments is more similar to the random distributions during the Chinese period, with about one third of the simulated values are greater than our observed ANN value.
-This testing reveals that the clustered distributions of ornaments during the pre-European and European periods might reflect the presence of different social groups.
-Moreover, the clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan.
+The Chinese period has fewer artefacts in any category, likely reflecting a smaller population at Kiwulan at this time, making spatial patterns and hotspots difficult to discern with confidence.
+This testing reveals clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
 
@@
 
-An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornament types and materials during the European period.
-Northeastern Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish.
-The origin of ornaments might be multiple places mostly in Southeast Asia and introduced into northeastern Taiwan by Chinese merchants before 17th century.
-Later, the trade activities became more frequent and intense in the 17th century that might be related to the colonial involvement of the Europeans based on the similar styles of golden beads found in Luzon that was once the colony of the Spanish during this period.
-In addition to ornaments, stonewares at Kiwulan that were also found on a Spanish shipwreck and Southeast Asia in relation to the Dutch and the Spanish trading routes, and the tobacco use were also believed introduced to Kiwulan by the Europeans [@Wang2007; @Chen2007].
+An indirect colonial influence may be indicated at Kiwulan by the greater diversity of ornament types and materials during the European period.
+Yilan was involved in complex trading networks both on a regional scale with other indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish.
+The ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century.
+Later, trade activities became more frequent and intense in the 17th century due to European activities.
 The greater diversity and quantity of ornaments likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
-However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period in the early 19th century.
-This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, or adopting the practices of Han Chinese who did not use these kinds of beads in their dress.
-Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007].
-Although historical records indicate that trade ornaments were prestige goods in prehistoric northeastern Taiwan, the information might be biased that needs to be compared with archaeological evidence.
-
-The archaeological contexts show that ornaments were rarely found in middens and mostly preserved well as body adornments in burial contexts serving as grave goods.
-This could reflect ornaments were viewed as valuable objects such as status indicators and their distribution across the dwelling area would help to understand the social relation reflected by people who possessed them, and the reaction of the society as a whole to the colonial presence in the 17th century.
-Spatial patterns of ornaments show that their distribution was clustered during the pre-European and European period.
-Testing for spatial patterns indicates that these clusters are non-random, and are most highly concentrated during the European period.
-The spatial clustering during the pre-European and the European period shows an increase in the use of ornaments in daily life at Kiwulan.
-It may indicate that a degree of social inequality based on uneven distribution was present before European contact and then it was reinforced and amplified during the European period.
-A further indicator of increasing social inequality is a burial dated to the 17th century that included 60 golden beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008].
+The frequency of overall ornaments and each subtype declines significantly after European influence fades during the Chinese period in the early 19th century.
+This may be due to a smaller scale of trading networks, the overall decline of indigenous populations in Yilan, or the adoption of Han Chinese practices.
+The decline of the population may be related to the movement of many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants at the end of the 18th century [@Chen2007].
+
+Archaeological contexts show that ornaments are especially abundant in burial contexts serving as grave goods.
+This supports the interpretation of ornaments as valuable objects functioning as status indicators.
+Spatial patterns of ornaments in dwelling contexts show that their distribution was clustered during the pre-European and European periods.
+These clusters are non-random, and are most highly concentrated during the European period.
+This may indicate that a degree of social inequality based on the uneven distribution of ornaments was already present before European contact, and then it was reinforced and amplified during the European period.
+A further indicator of increased social inequality is a burial dated to the 17th century that included 60 gold-foil beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008].
 
@@
 We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Drennan2010; @Feinman2000; @Ames2010; @Bowles2010].
-For example, the corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.
+The corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.
 In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks.
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
-This provides an interpretation that Yilan social organisation went from corporate mode before the European arrival, then transitioned to network mode during the European presence.
-The small number of ornaments, and less concentrated distribution during the pre-European period is consistent with shared power and wealth of the corporate mode.
-The long-distance trade network introduced by Europeans resulted in the appearance of a network mode due to competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
+We may be able to interpreate Yilan social organisation as moving from corporate mode before the European arrival, to a network mode during European presence.
+The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode.
+The long-distance trade network introduced by Europeans resulted in the appearance of a network mode, and the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
 Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
 
-Another scenario could be that the increasing number and concentrated spatial pattern of ornaments indicate cultural resistance against the European intrusion by using them to show social identity and emphasize the local custom that had existed before European contact.
+That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations.
+The increasing number and concentrated spatial patterns of ornaments may indicate a practice of cultural resistance against the European intrusion.
+Resistance to European economic and political demands may be inferred if ornaments were used as a display of social identity and to emphasize the local customs that had existed before European contact [cf.
+@Rubertone2000].
 Resistance could be presented in many forms.
-For example, @Rubertone2000 discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands.
-Similarly, the frequent presence of ornaments across the dwelling area during the European period suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition.
-During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the use and discard of ornaments at Kiwulan.
-This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or indeigenous population decline.
-It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next.
-This might suggest that the ornaments found in the Chinese period were old items continuously used until then.
-Further analysis of the source for ornaments from different periods will be helpful to understand their origin and regional circulation.
+Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time.
+This process would result in a natural increase in ornaments over time, unrelated to colonial influences, but has limited value in explaining the shifts in spatial patterns.
 
@@
 
-Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017].
+Examination of the archaeological record at the peripheries of colonial activities can show how remote indeignous groups were affected by major European colonial processes [@Trabert2017].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
 Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation.
-The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
+The frequency and spatial distribution of personal ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
 The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context.
@@
 After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan.
-These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
+This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
 This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that is part of culture tradition.
@@
 It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
-Future work could extend this approach to studies of other trade goods such as ceramics.
+Future work could extend this approach to sourcing the ornaments directly with geochemical methods, and studies of other trade goods such as ceramics.
 We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardisation, and mortuary practices.

### 2019-12-11 analysis/paper/paper.Rmd
added 1 data file and code for the radiocarbon table

--- 2019-12-10 analysis/paper/paper.Rmd
@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BC) and a lower component (700-1200 AD, 1200-800 BC) separated by a sterile spanning 150 years according to .
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile spanning 150 years according to .
 This chronology is based on the stratigraphy and a series of 32 radiocarbon dates previously published by @Chen2007.
@@
   mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
-  mutate(Period = as.factor(Period)) %>% 
-  mutate(Period = factor(Period,
-                       level = c("pre-European", 
-                                 "European", 
-                                 "post-European"))) %>% 
-  arrange(Period) %>% 
   left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
-  dplyr::select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details")
+  mutate(ave = (`2.5%` + `97.5%`) / 2) %>% 
+  mutate(ave = ifelse(is.na(ave), 200, ave)) %>%  
+  arrange(ave) %>% 
+  dplyr::select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details") %>% 
+  mutate(`Calibrated date (95% conf. level)` = 
+           ifelse(is.na(Calibrated date `(95% conf. level)`), "-", "Calibrated date (95% conf. level"))
 

### 2019-12-11 analysis/paper/potsherds.R

@@ -1,0 +1,112 @@
+library(tidyverse)
+
+# read data
+kwl_lp <-
+  readxl::read_excel(here("analysis", "data", "raw_data",
+                          "舊社陶坑層統計(950501)-製作坑層分佈圖用.xls"),
+                     col_names = TRUE,
+                     skip = 1)
+
+# P051 would be excluded because irregular depths
+the_pits <- c(
+  "P040",
+  "P041",
+  "P042",
+  "P051",
+  "P052",
+  "P053",
+  "P054",
+  "P058",
+  "P059",
+  "P060",
+  "P061",
+  "P062",
+  "P063",
+  "P064",
+  "P065",
+  "P066",
+  "P067",
+  "P070",
+  "P071",
+  "P072",
+  "P073",
+  "P074",
+  "P075AB",
+  "P075CD",
+  "P076",
+  "P077",
+  "P078",
+  "P079",
+  "P080",
+  "P082",
+  "P083",
+  "P084",
+  "P085",
+  "P086",
+  "P087",
+  "P088",
+  "P089",
+  "P090",
+  "P091",
+  "P092",
+  "P093"
+)
+
+## clean up
+kwl_lp_clean <-
+  kwl_lp %>%
+  mutate_at(vars("編號\n/層位", "重量小計",  "件數小計"),
+            as.numeric) %>%
+  filter(!is.na(`編號\n/層位`)) %>%
+  mutate(`編號\n/ 現象號(P)` = zoo::na.locf(`編號\n/ 現象號(P)`)) %>%
+  filter(`編號\n/ 現象號(P)` %in% the_pits) %>%
+  separate(`海拔深度\n(cm)`,
+           into = c("start", "end"),
+           sep = "~") %>%
+  mutate_at(vars(start, end),
+            as.numeric)
+
+## read ornament data
+kwl_upper <-
+  readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
+
+kwl_lp_clean_summary <-
+  kwl_lp_clean %>%
+  group_by(end) %>%
+  summarise(mean_weight= mean(重量小計, na.rm = TRUE))
+
+kwl_lp_clean_piece <-
+  kwl_lp_clean %>%
+  group_by(`編號\n/ 現象號(P)`) %>%
+  summarise(sum_p= sum(件數小計, na.rm = TRUE))
+
+# plot
+ggplot(kwl_lp_clean,
+       aes(x = end,
+           y = 重量小計)) +
+  geom_line() +
[…]

### 2019-12-11 analysis/paper/paper.Rmd

--- 2019-12-11 analysis/paper/paper.Rmd
@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BC) and a lower component (700-1200 AD, 1200-800 BC) separated by a sterile deposit spanning c.
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning c.
 150 years.

### 2019-12-11 analysis/paper/paper.Rmd
edit the data file of dates and adjust the code for that table

@@
          calCurves = rep('intcal13',
-                         length(kw_dates_with_error$rce))
-       ,
-       store_density=TRUE))
-  )
+                         length(kw_dates_with_error$rce)),
+         store_density=TRUE)))
 
@@
   mutate(`Calibrated date (95% conf. level)` = 
-           ifelse(is.na(Calibrated date `(95% conf. level)`), "-", "Calibrated date (95% conf. level"))
+           ifelse(is.na(`Calibrated date (95% conf. level)`), "-", `Calibrated date (95% conf. level)`))
 

### 2019-12-11 analysis/paper/paper.Rmd
adjust the heading in the table and edit the figure caption

@@
 Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit.
+For the upper component, the blue line indicates the start of European presence, while the red line is the Chinese presence.
 Ages calibrated with the Bchron package (Parnell et al.
@@
   kw_dates %>% 
-  mutate(`Depth related to sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>% 
+  mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>% 
   mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
@@
   arrange(ave) %>% 
-  dplyr::select("Period", "Pit-Layer", "Depth related to sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details") %>% 
+  dplyr::select("Pit-Layer", "Above mean sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details") %>% 
   mutate(`Calibrated date (95% conf. level)` = 
-           ifelse(is.na(`Calibrated date (95% conf. level)`), "-", `Calibrated date (95% conf. level)`))
+           ifelse(is.na(`Calibrated date (95% conf. level)`), "—", `Calibrated date (95% conf. level)`))
 

### 2019-12-11 analysis/paper/paper.Rmd
edit the data file, work on the code for table, get an error

@@
          calCurves = rep('intcal13',
-                         length(kw_dates_with_error$rce))
-       ))
+                         length(kw_dates_with_error$rce))))
 
@@
   arrange(ave) %>% 
-  dplyr::select("Pit-Layer", "Above mean sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details") %>% 
+  dplyr::select("Lab code", "Pit-Layer", "Above mean sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details") %>% 
   mutate(`Calibrated date (95% conf. level)` = 

### 2019-12-12 analysis/paper/paper.Rmd
change texts in the datafile and the caption in the code

@@
   left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
-  mutate(ave = (`2.5%` + `97.5%`) / 2) %>% 
+  mutate(ave = (`2.5%` + `97.5%`)/ 2) %>% 
   mutate(ave = ifelse(is.na(ave), 200, ave)) %>%  
   arrange(ave) %>% 
-  dplyr::select("Lab code", "Pit-Layer", "Above mean sea level (cm)", "Date BP", "Calibrated date (95% conf. level)", "Details") %>% 
+  dplyr::select(`Lab code`, `Pit-Layer`, `Above mean sea level (cm)`, `Date BP`, `Calibrated date (95% conf. level)`, Context) %>% 
   mutate(`Calibrated date (95% conf. level)` = 
@@
 knitr::kable(table_kw_dates,
-             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007). Calibrated using IntCal13 Atmospheric curve.")
+             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007).
+Calibrated using IntCal13 Atmospheric curve.
+In the context column, original numbering system from excavation report is used, such M009 means burial No.9 ")
 ```

### 2019-12-12 analysis/paper/paper.Rmd
make the c14 table work

@@
       fig_caption: yes
-      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+      #reference_docx: "../templates/template.docx" # Insert path for the DOCX file
 bibliography: references.bib
@@
 image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm
+
+# to make the table work, cf. https://github.com/haozhu233/kableExtra/issues/308
+options(kableExtra.auto_format = FALSE)
 ```
@@
 Ages calibrated with the Bchron package (Parnell et al.
-2008).
-"}
+2008).", cache=TRUE}
 
@@
   rownames_to_column(var = "Pit-Layer") %>% 
-  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`," - ", `97.5%`))
+  mutate_at(vars(`2.5%`, `97.5%`), round, 0) %>% 
+  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`,"-", `97.5%`))
 
@@
   kw_dates %>% 
+  mutate(`Lab code` = str_remove(`Lab code`, " ")) %>% 
   mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
@@
   arrange(ave) %>% 
-  dplyr::select(`Lab code`, `Pit-Layer`, `Above mean sea level (cm)`, `Date BP`, `Calibrated date (95% conf. level)`, Context) %>% 
+  dplyr::select(`Lab code`, `Pit-Layer`, 
+                `Above mean sea level (cm)`, 
+                `Date BP`, 
+                `Calibrated date (95% conf. level)`, 
+                Context) %>% 
   mutate(`Calibrated date (95% conf. level)` = 
-           ifelse(is.na(`Calibrated date (95% conf. level)`), "—", `Calibrated date (95% conf. level)`))
-
-knitr::kable(table_kw_dates,
-             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007).
-Calibrated using IntCal13 Atmospheric curve.
-In the context column, original numbering system from excavation report is used, such M009 means burial No.9 ")
+           ifelse(is.na(`Calibrated date (95% conf. level)`), 
+                  "", 
+                  `Calibrated date (95% conf. level)`))
+
+kableExtra::kable(table_kw_dates,
+             caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve.
+The codes in the context column refer to the excavator's numbering system, cf.
+ADD CITATION TO THE REPORT HERE")
 ```

### 2019-12-12 analysis/paper/paper.tex
make the c14 table work

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period},
            pdfauthor={Li-Ying Wang; Ben Marwick},
            pdfkeywords={Historical archaeology; Iron Age; European colonisation; Trade
ornaments; Exchange networks; Taiwan; Monte Carlo simulation},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Ornaments as indicators of social changes and cultural practices in
northeastern Taiwan before and after European colonial period}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Li-Ying Wang \\ Ben Marwick}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{11 December, 2019}

\begin{document}
\maketitle
\begin{abstract}
Long-lasting indirect impacts on indigenous peoples in the periphery of
colonial control are poorly understood, especially in East Asia. Trade
ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the
indirect impacts of European colonial activities on local societies. The
diversity of ornaments was greater during the period of European
presence compared to previous periods, and their spatial distribution
was more clustered. This hints at increasing social inequality resulting
from a colonial influence. Ornaments give insights into the increasing
social inequality stimulated by European colonisation, and show the
agency of indigenous people to incorporate ornaments into their social
system.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The direct impacts of European colonialism on indigenous communities in
East Asia were much less conspicuous than in island Southeast Asia and
Oceania. Direct European colonial rule throughout East Asia was rare and
limited, and the question of long-lasting indirect impacts on local
indigenous communities remains largely unanswered. Understanding the
indirect effects of colonialism are important for detecting colonial
impacts on indigenous peoples in the periphery of colonial control
(Trabert 2017). In many parts of the world, the introduction of foreign
trade goods by colonial traders into local indigenous societies caused
substantial transformations of indigenous economic, cultural, and
socio-political systems (Dietler 2005; Dietler 1997; Junker 1993;
Silliman 2005). Consumption patterns of foreign goods can give insights
into negotiations between colonized and colonizer, and the resistance
and accommodations of indigenous people through their daily cultural
practices (Dietler 2015; Given 2004; Mullins 2011; Scaramelli \&
Scaramelli 2005; Silliman 2001; Torrence \& Clarke 2000; Voss 2005).
Northeastern Taiwan is an ideal context to study peripheral colonial
influence because although there was a prominent Spanish and Dutch
colonial presence in parts of Taiwan, the northeastern region was
isolated from intensive direct contact by the Xueshan Mountains.

This article describes personal ornaments excavated from the upper
component of Kiwulan (1400 AD-1900 AD), the largest Iron Age settlement
on the Yilan plain in northeastern Taiwan. The first recorded European
presence in Yilan was a Spanish revenge attack on indigenous villages in
1632 (Borao 2001: 163). In 1647 the Dutch attacked villages and forced
them to accept colonial rules and pay an annual tribute (Andrade 2007).
According to Dutch census reports in 1650, Kiwulan was the largest
indigenous settlement in the plain, with a population of 840 adults
(Nakamura 1938: 12). Following defeat of Dutch by the Chinese general
Koxinga in 1661-1662, the Dutch abandoned northern Taiwan. Direct
contact with Han Chinese is indicated by Qing dynasty census reports
mentioning Yilan villages in 1821 (Yao 1996).

One of the most commonly traded types of object in this region were
ornaments such as glass and stone beads (Chen 2007; Li \& Chiu 2014;
National Musuem of Taiwan History 2005). Personal adornments in the
archaeological record are useful as signal of an individual's status
(Joyce 2005; Scaramelli \& Scaramelli 2005). The consumption of stone
beads in Southeast Asia during Iron Age is often associated with
increasing social stratification or socio-political complexity (Bellina
2014; Carter 2016; Francis 2002; Theunissen \emph{et al.} 2000; Kenoyer
2000). In this paper, we explore archaeological ornaments from Kiwulan
spanning the pre-European contact period, the period of Spanish and
Dutch presence, and the following period of Chinese presence. We address
the question of whether indirect colonial influences on the indigenous
populations can be detected through the ornament assemblages.

\hypertarget{ornaments-in-complex-exchange-network-during-the-late-iron-age-and-early-historical-period}{%
\section{Ornaments in complex exchange network during the late Iron Age
and early historical
period}\label{ornaments-in-complex-exchange-network-during-the-late-iron-age-and-early-historical-period}}

The island of Taiwan lies at the junction of mainland China, Southeast
Asia, and Northeast Asia in the Pacific Ocean. The prehistory of Taiwan
island could be roughly divided into three major periods, paleolithic
(c.~27,000 BP- 5000BP), Neolithic (c.~6500- 2000BP), and Iron age
(c.~2000- 400BP) with slightly regional differences in onset of each
period and variations in style of artifacts and assemblages (Liu 2011;
Chen 2017). It is generally accepted that Taiwan entered the historical
period in the early 17th century due to the colonial activities of the
Spanish and the Dutch who played an important role in keeping written
records about Taiwan. The European colonial presence in Taiwan ended in
1662 when the Dutch were defeated by the kingdom of Tungning, founded by
Koxinga from China. Later in 1683, Taiwan was incorporated into the Qing
dynasty in China and a large wave of Han Chinese migrated to Taiwan
during the late 18th century. Because of natural safe harbors, northeast
Taiwan was involved in a regional trade network through cross-culture
interactions with Chinese merchants since the 14th century, and later
the global trade network with the Europeans in the 17th century brought
more trade goods circulated in Southeast Asia into Taiwan (Chen 2005;
Liu \& Wang 2017). Although located on the periphery of regional trade
centers, Yilan was connected to trade networks via visits of other
indigenous groups, Chinese merchants, and Europeans, via sea.

The European presence in northern Taiwan started with the Spanish who
founded Fort San Salvador at Keelung in 1626, and Fort San Domingo in
1629 at Tamsui. They sent missionaries to local indigenous settlements
in this region (Blussé \& Everts 2000: 343) and kept records about their
observations of indigenous communities. A Dominican priest in 1632
reported that the Taparri, an indigenous tribe from northern Taiwan,
exchanged carnelian beads with other indigenous groups. This form of
exchange was widespread and even the Spanish soldiers used carnelian
beads as bargaining chips for gambling (Li \& Wu 2006: 132--49). The use
of beads as prestige goods is further indicated by their role in bride
price payments, and compensation to resolve disputes (Li \& Wu 2006:
132--49). Other records mention that the female shamans in the tribe
would use carnelian beads as magical items in ritual healing practices
(Borao 2009: 122--51). Records of an indigenous funeral document the use
of carnelian beads in ritual contexts, with more carnelian beads,
pottery, and cloth placed into the graves of more influential people to
indicate their family's higher status (Li \& Wu 2006: 153). While a full
critical analysis of these historical accounts remains to be produced,
we take them to minimally indicate that carnelian beads were already
treated as prestige goods in Yilan before the arrival of Europeans. In
1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the
Spanish and took over their forts in northern Taiwan. They introduced a
feudal system in an attempt to control the indigenous communities by
asking indigenous leaders to attend an annual ceremony for demonstrating
their loyalty and paying tributes (Andrade 2007, ch.~9; Kang 2016,
ch.~4). The Dutch provided beads and other goods based on negotations
with indigenous communities to secure alliances in the annual ceremony
or during travelling (Kang 2016, ch.~6). We might predict that the
activities of the Dutch feudal system to build and maintain alliances
resulted in an increase in the amount and diversity of ornaments in
northeastern indigenous communities during this period.

Chinese historical records from 1829, 1837, and 1852 during the Qing
dynasty (1616-1911) contain some notes on the purposes of ornaments from
Yilan (Chen 1963: 228, 308; Ke 1993: 11, 126; Yao 1996: 77). According
to those records, indigenous people in Yilan wore ornaments in
ceremonial contexts to display their wealth and status. Among those
ornaments, fish-shaped necklaces made of metal threads had high value
due to their delicacy and the exotic materials invested in production.
These were usually possessed by wealthy people. Other people wore
carnelian beads or glass beads on their head or neck to participate in
ceremonies. In 1895, at the beginning of Japanese colonisation, an
academic field survey for plains indigenous groups reported that
fish-shaped metal necklaces necklaces were not used in Yilan at that
time, but elderly people still used beads (Ino 1996: 227--32). Although
these historical records are fragmentary and may contain some biases
(Galloway 2006) that have not yet been studied in detail, we find
consistency among multiple sources in their descriptions of how
ornaments represent high status or specialised social roles in
indigenous communities in Yilan. Compared to the European period, there
are fewer documentary mentions of beads in the Chinese period and the
descriptions are limited to clothing, but these generally confirm the
role of beads as status markers.

Ornaments found in northeastern Taiwan in the early historical period,
including glass beads, stone beads, and metal ornaments, are considered
to have been imported from other regions. This is because of to a lack
of archaeological evidence of beadmaking waste, metalworking, or
accessible local raw materials. The chemical composition of glass beads
from this region shows a high content of lead and, toghether with the
winding/folding technique, these details suggest a Chinese beadmaking
tradition (Cheng 2008; Gan \emph{et al.} 2006; Wang 2018). Although
there is a wide variety of metal ornaments such as bells, bracelets,
rings, and pendants, the common components of metal ornaments are brass
and copper, with a small number made from lead and tin that indicates
multiple origins that include Southeast Asia (Chen 2011). There is no
direct evidence showing European delivery of beads, however, a large
amount of the glass beads containing gold foil (hereafter, gold-foil
beads) at Kiwulan might have been introduced by the Spanish through
economic activities because similar beads were found at Luzon, northern
Philippines, as part of the trading route of the Spanish between 16-19th
century (Wang \& Liu 2007). Both archaeological evidence and historical
records indicate northeastern Taiwan was involved inregional networks
with East Asia in the late Iron age. These included Chinese merchants
trading metal items, clothes, and beads with local indigenous people in
Taiwan in exchange for local resources. The foreign-made large dark
brown glazed stoneware jars frequently found in European shipwrecks were
also commonly found from many sites in Taiwan, suggesting direct or
indirect interactions. Despite the Chinese origin of some ornaments at
Kiwulan, there is compelling evidence that a large amount of ornaments
found at 17th century sites resulted from European colonial and economic
activities in the region.

\hypertarget{excavations-at-kiwulan-in-northeastern-taiwan}{%
\section{Excavations at Kiwulan in northeastern
Taiwan}\label{excavations-at-kiwulan-in-northeastern-taiwan}}

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/kiwulan-location-map} \caption{Map showing the location of Kiwulan, and other places in northern Taiwan named in the text. Map data from naturalearthdata.com}(\#fig:KWL-map)
\end{figure}

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/KWL-excavation-map} \caption{Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number. Small dots represent the location of post-holes. Each square is 4 x 4 m}(\#fig:KWL-sam-area)
\end{figure}

Information about Kiwulan (Figure @ref(fig:KWL-map)) comes from a rescue
archaeology project that during 2001-2004 in advance of a water
diversion project and road bridge construction. The excavations used 2
mm and 1.5 mm mesh screens and covered eight open area sections in total
of 262 squares (4 m by 4 m) reaching 3,814 m\textsuperscript{2} (Chen
2007). The nearly 2 m thick archaeological deposits reveal a large
amount of artefacts, burials, middens, post-holes, wooden pillars, and
stone structures, all of which indicates it was a long-term settlement.
Artefact locations were recorded to the 2 x 2 m sub-square they were
recovered in; they lack individual point provenance. Based on the
continuity of deposition and the frequency of artifacts, the center of
the site is the open area consisting of A and D sections, which is also
the study area where our samples selected from. In the AD area,
post-holes were found aligned in a north-south direction in some
intervals with construction marks, which were interpreted as the remains
of silt house structures. At the north margin of the dwelling place were
burials that are mostly oriented in an east-west direction (Figure
@ref(fig:KWL-sam-area)).

The chronology of Kiwulan can be divided into two phases represented by
a upper component (1400-1900 AD, 600-100 BP) and a lower component
(700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning
c.~150 years. These component divisions are based on the differences in
the colour and texture of the deposit. The interpration of the sterile
deposit is still under debate, with pollen analysis suggesting dry
weather leading to site abandonment (Lin 2015; Chiu 2004; Chen 2007).
There are 32 radiocarbon ages spanning the two components, previously
published by Chen (2007), and shown here in Figure @ref(fig:C14-data)
and Table @ref(tab:C14-table-data). We focus on the upper component
because only this componant spans the periods of pre-European contact,
European presence, and Chinese presence. In the upper component, all
excavation squares in our sampling area show signs of continuous human
occupation during each of the three phases. Previous work divided the
upper component into six analytical units, spanning from the 14th
century to the 19th century, according to the radiocarbon dates,
excavation depth, consistency of contexts, and types of chronologically
diagnostic ceramics such as blue and white porcelains (Hsieh 2009; Wang
2011). However, we found some ambiguities in the previous chronology, so
to help answer our research question, we re-examined the upper component
to devise a new chronology to assign artefacts to the pre-European,
European, and the Chinese periods.

The archaeological indicators of the start of the European influence at
Kiwulan are the appearance of light grey glazed jars, known as
``An-ping'' jars in China and Taiwan, and large dark brown glazed
stoneware jars that were introduced to Taiwan during the early 17th
century. Large dark brown glazed stoneware jars may have been made in
Southeast Asia, and are frequently found in European shipwrecks from
this period as vessels for transporting water, wine or other liquids on
long voyages. The earliest evidence oflight gray glazed jars in this
region has been found among the cargo of the Spanish shipwreck \emph{San
Diego}, which sunk in 1600 AD (Hsieh 2009; Hsieh 1995). Southeast China
is assumed to be the origin of the light gray glazed jars, however these
are commonly found at sites in Taiwan that were associated with European
activities, such as the Zeelandia fort site in Tainan (Wang \& Liu
2007). The jar shapes found at Kiwulan are typical of those found
elsewhere in VOC sites occupied during the 17th century (Berrocal
\emph{et al.} 2018: 917; Cort 2017: 282; Grave \& McNiven 2013; Ketel
2011; Klose \& Schrire 2018: 131). We cannot be sure of the exact
process that bought them to Kiwulan: they might have been directly
imported by Europeans, by Chinese merchants, or by indigenous groups via
regional networks in north Taiwan. In any case, the high volume of
ceramics transported by Europeans, and their high mobility in shipping
trade played an important role in introducing foreign jars to Taiwan.
Those jars were widely distributed across the site and can serve as
indicators, together with the radiocarbon dates, to identify the
excavation units associated with the pre-European period and the start
of European influence at Kiwulan. In addition to stoneware jars as
indicators of European presence, around 300 pieces of locally made pipes
and a few imported pipes were found at Kiwulan. Smoking is likely to
have been introduced by Europeans, and we found that the presence of
pipes in the archaeological record here is consistent with distributions
of glazed jar fragments, which are far more numerous and widespread
across the site (n = 1685).

The archaeological signature of the Chinese period at Kiwulan is the
large amount and diversity of Chinese porcelains in many forms such as
bowls, plates, and cups. Other indicators include opium pipe-bowls and
distinctive architectural bricks and tiles used by Chinese (Hsieh 2009).
Chinese migrations to Yilan were also recorded in official Chinese
records written in the early 19th century recording the first immigrants
in 1768 (Chen 1963; Ke 1993).

\begin{figure}
\centering
\includegraphics{../figures/C14-data-1.pdf}
\caption{(\#fig:C14-data)Summed probability distributions for dates from
Kiwulan. The dark line represents the summed probabilities of all
radiocarbon ages, and the grey lines in the background are the
probabilities of individual ages. Grey rectangles indicate the
approximate chronology of the major archaeological components of the
deposit. For the upper component, the blue line indicates the start of
European presence, while the red line is the Chinese presence. Ages
calibrated with the Bchron package (Parnell et al.~2008).}
\end{figure}

\begin{table}

\caption{(\#tab:C14-table-data)Radiocarbon dates on charcoals from Kiwulan (Chen 2007). Calibrated using IntCal13 Atmospheric curve. In the context column, original numbering system from excavation report is used, such M009 means burial No.9 }
\centering
\begin{tabular}[t]{l|l|l|l|l|l}
\hline
Lab code & Pit-Layer & Above mean sea level (cm) & Date BP & Calibrated date (95\% conf. level) & Context\\
\hline
NTU-3803 & P052-L7 & 0 to -10 & <200yr &  & artefact-bearing deposit\\
\hline
NTU-3925 & P051-L17 & -36 to -56 & <200yr &  & sterile deposit\\
\hline
NTU-3943 & P051-L19 & -70 to -90 & <200yr &  & sterile deposit\\
\hline
NTU-4283 & P063-L12 & -30 to -70 & <200yr &  & midden, H044\\
\hline
NTU-4293 & P089-L11 & -50 to -70 & <200yr &  & artefact-bearing deposit\\
\hline
NTU-4305 & P089-L7 & -20 to -30 & <200yr &  & artefact-bearing deposit\\
\hline
NTU-4322 & P051-L11 & 0 to -40 & <200yr &  & midden, H026\\
\hline
NTU-4323 & P070-L3 & 20 to -57 & <200yr &  & burial, M095\\
\hline
NTU-3993 & P041-L7 & -25 to -45 & 250±40 & 4-431 & artefact-bearing deposit\\
\hline
NTU-4419 & P162-L3 & -10 to -110 & 280±70 & 11-485 & midden H172\\
\hline
NTU-4311 & P052-L16 & -110 to -130 & 310±100 & 17-510 & artefact-bearing deposit\\
\hline
NTU-4320 & P168-L1 & 6 to -51 & 340±100 & 38-528 & midden, H193\\
\hline
NTU-4016 & P028-L9 & -44 to -80 & 270±40 & 145-453 & burial, M020\\
\hline
NTU-4310 & P018-L2 & -28 to -70 & 360±100 & 73-542 & burial, M039\\
\hline
NTU-3791 & P049-L11 & -20 to -30 & 340±30 & 314-482 & artefact-bearing deposit\\
\hline
NTU-4292 & P052-L6 & 4 to -56 & 510±75 & 344-648 & burial, M009\\
\hline
NTU-4304 & P066-L11 & -40 to -60 & 600±75 & 515-675 & artefact-bearing deposit\\
\hline
NTU-4423 & P144-L5 & -10 to -30 & 610±90 & 504-709 & artefact-bearing deposit\\
\hline
NTU-4315 & P248-L5 & -100 to -120 & 800±120 & 560-950 & artefact-bearing deposit\\
\hline
NTU-3926 & P041-L9 & -70 to -90 & 900±50 & 714-917 & sterile deposit\\
\hline
NTU-4421 & P162-L11 & -160 to -180 & 920±70 & 706-954 & artefact-bearing deposit\\
\hline
NTU-4319 & P154-L3 & 10 to -10 & 920±105 & 685-1053 & artefact-bearing deposit\\
\hline
NTU-4430 & P238-L10 & -130 to -150 & 1020±60 & 796-1055 & sterile deposit\\
\hline
NTU-4422 & P237-L4 & -70 to -90 & 1030±80 & 779-1156 & artefact-bearing deposit\\
\hline
NTU-3788 & P028-L15 & -130 to -150 & 1050±40 & 891-1050 & artefact-bearing deposit\\
\hline
NTU-4428 & P154-L13 & -170 to -180 & 1080±90 & 802-1222 & artefact-bearing deposit\\
\hline
NTU-4427 & P246-L8 & -160 to -180 & 1170±70 & 956-1255 & artefact-bearing deposit\\
\hline
NTU-4316 & P019-L5 & -100 to -120 & 1190±70 & 969-1263 & burial, M066\\
\hline
NTU-3792 & P041-L13 & -150 to -170 & 1240±30 & 1079-1261 & artefact-bearing deposit\\
\hline
NTU-4434 & P144-L11 & -130 to -150 & 1480±70 & 1292-1524 & artefact-bearing deposit\\
\hline
NTU-4321 & P154-L14 & -180 to -190 & 1870±110 & 1561-2085 & artefact-bearing deposit\\
\hline
\end{tabular}
\end{table}

\hypertarget{the-personal-ornaments}{%
\section{The personal ornaments}\label{the-personal-ornaments}}

\begin{figure}
\includegraphics[width=3.54in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-ornaments-count} \caption{Frequency of the major class of ornaments at Kiwulan. Frequency represents artefact counts}(\#fig:plot-all-ornaments-count)
\end{figure}

\begin{table}

\caption{(\#tab:tidy-data-for-table)Ornament subtype at Kiwulan. The numbers represent artefact counts }
\centering
\begin{tabular}[t]{l|l|r|r|r}
\hline
Categories & Type & Before European Contact & European Presence & Chinese Presence\\
\hline
Agate bead & hexagonal & 6 & 17 & 5\\
\hline
Agate bead & waxy oval & 0 & 4 & 0\\
\hline
Agate bead & small oval & 3 & 3 & 0\\
\hline
Agate bead & globular & 0 & 1 & 0\\
\hline
Agate bead & pentagonal & 0 & 1 & 0\\
\hline
Agate bead & big oval & 0 & 0 & 1\\
\hline
Agate bead & long bicone & 0 & 0 & 1\\
\hline
Agate bead & octagonal & 0 & 0 & 1\\
\hline
Bell & large & 3 & 8 & 3\\
\hline
Bell & plain small & 0 & 4 & 1\\
\hline
Bell & thin small & 0 & 1 & 1\\
\hline
Glass bead & small (0.5-1 cm) & 60 & 37 & 1\\
\hline
Glass bead & medium (1-2 cm) & 8 & 15 & 0\\
\hline
Golden bead & NA & 48 & 93 & 2\\
\hline
Metal ring & wide small & 1 & 9 & 1\\
\hline
Metal ring & thin large & 4 & 5 & 2\\
\hline
Metal ring & wide large & 0 & 5 & 0\\
\hline
Metal ring & overlapped & 0 & 2 & 0\\
\hline
Metal ring & braid & 0 & 1 & 0\\
\hline
Metal ring & entwined & 1 & 1 & 0\\
\hline
Metal ring & flat & 0 & 1 & 0\\
\hline
Metal ring & large thick string & 0 & 1 & 0\\
\hline
Metal ring & small thin string & 0 & 1 & 0\\
\hline
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=5.93in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/ornament} \caption{Subtypes of ornament in each major class. A: carnelian beads, B: bells, C: glass beads and gold-foil beads, D: metal rings. Photographs are presented in the same order as those subtypes in the table but from left to right instead. The photographs of B, C, D classes were from original excavation report (Chen 2007).}(\#fig:ornaments-image)
\end{figure}

Ornaments were found in a variety of archaeological contexts including
post-holes area, burials, and middens. This study focuses on 406
ornaments from 40 sampling squares located at the dwelling place of
Kiwulan, indicated by aligned post-holes with \emph{in-situ} posts
(Figure @ref(fig:KWL-sam-area)). Occupation floors were not identified
during excavation. We choose these units because they were
stratigraphically intact and undisturbed by modern construction
activity, compared to excavation squares on the periphery of the site.
There are 35 burials in the sampling area, one third of the total number
of burials at Kiwulan. Ornaments are commonly used as grave goods in
burials, with the total number of ornaments in burials including 3,173.
The high number of ornaments in these burials is due to the presence of
bead strands or patterned bands of beads, which sometimes contains
thousands of beads in an individual burial (Chen 2007). We exclude
burials from this analysis because most of them date to the pre-European
period (n = 26), limiting the usefulness of comparisons between the
periods. In addition to 406 ornaments from dwelling place contexts and
3,173 from burial contexts, there are 27 ornaments found in midden
contexts. We focus on ornaments from the dwelling place contexts (Figure
@ref(fig:plot-all-ornaments-count), Table @ref(tab:tidy-data-for-table))
because these give us the greatest spatial and temporal representation
across the three time periods, and so are most informative on social
inequality indicated by uneven distributions of ornaments.

\hypertarget{reproducibility-and-open-source-materials}{%
\section{Reproducibility and open source
materials}\label{reproducibility-and-open-source-materials}}

To enable re-use of materials and improve reproducibility and
transparency (Marwick 2017), the entire R code (R Core Team 2019) used
for all the analysis and visualisations contained in this paper is
included in \url{http://doi.org/10.17605/OSF.IO/R8YGA}. Also in this
version-controlled compendium (Marwick \emph{et al.} 2018) are the raw
data for all the visualisations and tests reported here. All of the
figures, tables, and statistical test results presented here can be
independently reproduced with the code and data in this repository. The
code is released under the MIT license, the data as CC-0, and figures as
CC-BY, to enable maximum re-use.

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-ornaments-count-three-periods} \caption{Frequency of the major ornament across different time periods.}(\#fig:plot-counts-by-period)
\end{figure}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{changes-in-the-frequencies-of-ornament-types-over-time}{%
\subsection{Changes in the frequencies of ornament types over
time}\label{changes-in-the-frequencies-of-ornament-types-over-time}}

Figure @ref(fig:plot-counts-by-period) shows the comparison of
frequencies of the major classes of ornaments for different the time
periods at Kiwulan. The difference in frequencies between the three time
periods reflect significant differences in the use of ornaments
(chi-square = 71.82, df = 8, p-value =
\(\ensuremath{2.136325\times 10^{-12}}\)). Most ornament types were
present before European contact. Ornament frequencies reached a peak
during the European period and then dropped during the Chinese period,
especially gold-foil beads. This trend can be also seen on other
ornaments including carnelian beads, metal rings, and bells. However,
glass beads show a different pattern that indicates a higher frequency
in the pre-European contact, and then a decrease in the European period
and a further decrease in the Chinese period.

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-multi-types-by-period} \caption{Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads.}(\#fig:plot-altogether)
\end{figure}

The distribution of frequencies for subtypes in each major class is
presented in Figure @ref(fig:plot-altogether). Spearman's correlation
test shows that there is no significant relationship between diversity
of subtypes and sample size (S = 498.09, rho = 0.11, p = \(0.69489\)).
This indicates that the increases in diversity can be explained by the
effects of culture interaction instead of the effects of sample size.
Carnelian beads and metal rings have greater quantity and variety of
shapes compared to copper bells and glass beads during the European
period. The greater varieties for carnelian beads and metal rings might
indicate multiple origins due to participation in large scale trade
networks stimulated by the European presence. In contrast, copper bells
have less variety typically \textgreater2 cm long with a wide variety of
human faces as a motif. Although glass beads have less variety in size,
small (0.5-1 cm) and medium (1-2 cm), they have a wide variety of colors
or patterns mostly made by winding technique with high lead content in
composition indicating possibly from China (Cheng 2008). Although we are
not certain of the specific origin of these beads, research suggest that
these glass beads and metal ornaments have similar production techniques
and composition to those found in China (Chen 2011; Wang 2018). There
seem to be no obvious changes in the sources of glass beads or metal
ornaments at different periods in the upper component of Kiwulan
(1400-1900 AD). However, the glass beads found from the lower component
(700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast
Asian sites since 300 BC and declining during the early 2nd millennium
(Wang 2018; Francis 2002).

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-spatial-distribution-all-ornaments-by-period} \caption{Spatial pattern of all class of ornament by time periods}(\#fig:plot-all-spatial-pattern)
\end{figure}

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-spatial-distribution-some-ornaments-by-period} \caption{Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here}(\#fig:plot-spatial-pattern-each)
\end{figure}

\hypertarget{changes-in-patterns-of-the-spatial-distribution-of-ornament-types}{%
\subsection{Changes in patterns of the spatial distribution of ornament
types}\label{changes-in-patterns-of-the-spatial-distribution-of-ornament-types}}

Figure @ref(fig:plot-all-spatial-pattern) presents the spatial
distribution of all ornaments from the research area for each time
period. Before the European arrival, a greater amount of ornaments were
found at the northern and middle parts of the research area. During the
European period, ornaments were more widespread, with some clusters on
the northern part. During the Chinese period the distribution is more
even again. Figure @ref(fig:plot-spatial-pattern-each) presents the
distribution for the major ornament classes individually, some clusters
across the area can be observed during the European period, such as
gold-foil beads and carnelian beads. However, there seems to be no
consistent pattern across those different ornaments. Each class shows
its own pattern where the squares with higher numbers of ornaments
distributed separately and independently. For example, a cluster of
gold-foil beads was found at the northern part, while a cluster of
carnelian beads was found in the middle part. In contrast, there are
multiple clusters of metal rings that are distributed separately across
the research area. Copper bells were usually found individually and
appear randomly distributed across the area. In the Chinese period, both
the amount and density of different classes of ornaments decreased.

\begin{figure}
\includegraphics[width=5.31in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-kde-maps} \caption{Kernel density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb.}(\#fig:plot-all-densities)
\end{figure}

\hypertarget{point-pattern-analysis-of-ornament-distribution}{%
\subsection{Point pattern analysis of ornament
distribution}\label{point-pattern-analysis-of-ornament-distribution}}

Point pattern analysis can assess whether the distribution of artefacts
represents hotspots produced by non-random processes (Bevan \& Lake
2016; Ducke 2015), such as concentrations of ornaments in specific
households that might result from social inequality stimulated by a
colonial presence. To prepare the ornament location data for point
pattern analysis, because artefacts from Kiwulan lack exact
piece-provance data, we assigned each ornament to a random coordinate
pair in the square it was recovered from. The next step was to divide
the ornaments into three time periods. Finally we computed the kernel
densities for each time period for comparison. Kernel density
estimations (KDE) compute the probability of the density of ornaments
across space by creating a continuous, smooth density surface across
space. Here we use KDE to visualize core areas of ornaments and
surrounding neighbourhoods (Bonnier \emph{et al.} 2019; Cortegoso
\emph{et al.} 2016). Density values of artefacts per square meter were
calculated for each cell.

Figure @ref(fig:plot-all-densities) shows that there is one major core
area during the pre-European period, multiple core areas during the
European period, and a single core during the Chinese period. There are
three consistent sub-regions with a core area that shifts over time. The
distribution might indicate the increase and decrease in the number of
social groups who possessed ornaments. The multiple groups during the
European period might reflect unequal consumption of ornaments across
the site, relative to other periods, or random patterns resulting from a
bigger sample size. In addition, the generation of core areas might be
biased due to small sample sizes, for example, a few ornaments found at
one single square during the Chinese period could create an obvious
hotspot. Whether the observed clustering is random or non-random is
crucial for the interpretation of intentional human activities.

\begin{figure}
\includegraphics[width=3.54in]{/Users/bmarwick/Desktop/kwl-ornaments/analysis/figures/plot-kde-ann-histograms} \caption{Histograms of simulated ANN values from 1000 simulations for three time periods. X-axis values represent ANN expected value under a completely random process resulted from simulated pattern. Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value.}(\#fig:hypo-test-distribution)
\end{figure}

To test for randomness in spatial locations, we used a Monte Carlo
method to simulate average nearest-neighbour distances (ANN). Figure
@ref(fig:hypo-test-distribution) shows the observed ANN distances with
the distributions of the ANN distances calculated on 1000 simulations of
random ornament locations. The results show that 100\% of the simulated
values are much greater than our observed ANN value during the European
period, which means the ornaments have non-randomly clustered
distributions. A similar, but less extreme, result is also observed
during the pre-European period. The observed distribution of ornaments
is more similar to the random distributions during the Chinese period,
with about one third of the simulated values are greater than our
observed ANN value. The Chinese period has fewer artefacts in any
category, likely reflecting a smaller population at Kiwulan at this
time, making spatial patterns and hotspots difficult to discern with
confidence. This testing reveals clustering of ornaments during the
European period is highly non-random, potentially indicating different
degrees of access to foreign ornaments or a concentration of power to
control the distribution of ornaments at Kiwulan during this period.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

An indirect colonial influence may be indicated at Kiwulan by the
greater diversity of ornament types and materials during the European
period. Yilan was involved in complex trading networks both on a
regional scale with other indigenous groups and Chinese merchants, and
at a global scale with Europeans, including the Dutch and the Spanish.
The ornaments have multiple origins, including Southeast Asia and China,
and were first introduced into northeastern Taiwan by Chinese merchants
before 17th century. Later, trade activities became more frequent and
intense in the 17th century due to European activities. The greater
diversity and quantity of ornaments likely results from participation in
large scale exchange networks that stimulated the circulation of
different ornament classes. The frequency of overall ornaments and each
subtype declines significantly after European influence fades during the
Chinese period in the early 19th century. This may be due to a smaller
scale of trading networks, the overall decline of indigenous populations
in Yilan, or the adoption of Han Chinese practices. The decline of the
population may be related to the movement of many indigenous people
moved southwards to Hualien due to population pressure caused by Han
Chinese immigrants at the end of the 18th century (Chen 2007).

Archaeological contexts show that ornaments are especially abundant in
burial contexts serving as grave goods. This supports the interpretation
of ornaments as valuable objects functioning as status indicators.
Spatial patterns of ornaments in dwelling contexts show that their
distribution was clustered during the pre-European and European periods.
These clusters are non-random, and are most highly concentrated during
the European period. This may indicate that a degree of social
inequality based on the uneven distribution of ornaments was already
present before European contact, and then it was reinforced and
amplified during the European period. A further indicator of increased
social inequality is a burial dated to the 17th century that included 60
gold-foil beads, well above the average of 2-3 pieces in the
pre-European period (Chen 2007; Cheng 2008).

How might these results fit into a bigger picture of social change at
periphery of colonial systems? We may get some insight into the general
pathways that led to social inequality in northeastern Taiwan by
considering how people have achieved and maintained power in a wide
variety of societies (Drennan \emph{et al.} 2010; Feinman 2000; Ames
2010; Bowles \emph{et al.} 2010). The corporate/network model proposed
by Feinman (2000) expands traditional hierarchical complexity to provide
a comparative basis for distinct strategies for power. In the network
mode, inequality develops when individuals accumulate wealth through
their individual networks and people use their wealth to attract
factions, control resources, and monopolize trade networks. In contrast,
the corporate mode stresses shared power across different groups and
sectors, integrative ceremonies and rituals, and large cooperative
labour tasks (Feinman 2000; Siegel 1999). We may be able to interpreate
Yilan social organisation as moving from corporate mode before the
European arrival, to a network mode during European presence. The small
number of ornaments, and less concentrated distribution during the
pre-European period appears consistent with shared power and wealth of
the corporate mode. The long-distance trade network introduced by
Europeans resulted in the appearance of a network mode, and the
emergence of competition among ambitious individuals for prestige,
wealth, or power through collecting trade goods (Brumfiel 1994; Clark \&
Blake 1994). Because of the weak direct control from the European
colonizers in northeastern Taiwan, local leaders may have had the
flexibility to manipulate European colonial images, expand personal
power, and monopolize the high-value trade goods (Kang 2012).

That said, the evidence from Kiwulan may be consistent with a variety of
scenarios of indigenous-colonial relations. The increasing number and
concentrated spatial patterns of ornaments may indicate a practice of
cultural resistance against the European intrusion. Resistance to
European economic and political demands may be inferred if ornaments
were used as a display of social identity and to emphasize the local
customs that had existed before European contact (cf.~Rubertone 2000).
Resistance could be presented in many forms. Another scenario is that
ornaments were treated as heirlooms, such as carnelian beads and
gold-foil beads, that passed from one generation to the next,
accumulating at Kiwulan over time. This process would result in a
natural increase in ornaments over time, unrelated to colonial
influences, but has limited value in explaining the shifts in spatial
patterns.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Examination of the archaeological record at the peripheries of colonial
activities can show how remote indeignous groups were affected by major
European colonial processes (Trabert 2017). Kiwulan in northeastern
Taiwan is an exceptional case study as an East Asian location that was
relatively isolated and peripheral, and yet connected by regional and
global trade networks. Kiwulan provides valuable insights into the
discussion of indirect colonial influence on local societies living
beyond the borders of direct European colonial occupation. The frequency
and spatial distribution of personal ornaments at Kiwulan present three
distinct patterns during different dominant culture interaction periods.
The greater amount and diversity of ornament types during the European
period reflects an increasing use in ornaments in a colonial context.
Before European contact, ornaments were traded into local indigenous
societies via the regional exchagne network with Chinese merchants and
viewed as prestige goods in local indigenous culture according to their
distribution in the archaeological contexts. After the arrival of the
Europeans, the exotic and powerful image carried by those ornaments may
have intensified, further signalling wealth and privileged trading
connections among the inhabitants of Kiwulan. This may have stimulated
more competition between aggrandizing individuals for prestige and
wealth accumulation at Kiwulan, which might have resulted in an increase
in social inequality. This might also indicate an act of intentional
resistance to the intrusion of the Europeans by using more ornaments
that is part of culture tradition.

We are still far from understanding the full variety of colonial impacts
on peripheral indigenous communities. The origin of ornaments and how
their sources changed over time would give more information to construct
a clear picture of complex trade networks during this periods. By
focusing on the distribution patterns in a settlement site, the Kiwulan
ornaments suggest that foreign ornaments can be a proxy to detect
indirect colonial influence on local indigenous populations. Ornaments
give insights into the amplification of social inequality stimulated by
European colonisation. It also shows the agency of indigenous people to
incorporate ornaments into their social system and use them in their
daily lives to display or intensify status differences. Future work
could extend this approach to sourcing the ornaments directly with
geochemical methods, and studies of other trade goods such as ceramics.
We have introduced here the corporate/network model for understanding
the dynamics of social inequality at Kiwulan, and future tests of this
should include analysis of pottery production and standardisation, and
mortuary practices.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank the Yilan County Cultural Affairs Bureau in
Taiwan for permitting access to the ornaments used in this study. We
thank Shui-jin Chiu, Shu-hui Jian, Jiun-yao Lai, and the staff in
archaeology lab for their invaluable assistance in preparing samples for
analysis. We also thank Yu-pei Chen for the discussion regarding the
site at the earliest stage of fieldwork. This work was supported in part
by the Fritz Graduate Fellowship and travel grants from the Department
of Anthropology at University of Washington. We would also like to thank
Ben Fitzhugh and Peter Lape for their insightful comments on the early
draft.

\hypertarget{pagebreak}{%
\subparagraph{pagebreak}\label{pagebreak}}

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Ames2010}{}%
\textsc{Ames}, K.M. 2010. On the evolution of the human capacity for
inequality and/or egalitarianism, in G. Feinman \& T.D. Price (ed.)
\emph{Pathways to power: New perspectives on the emergence of social
inequality}: 15--44. New York: Springer.

\leavevmode\hypertarget{ref-Andrade2007}{}%
\textsc{Andrade}, T. 2007. \emph{How Taiwan became chinese : Dutch,
spanish, and han colonization in the seventeenth century}. New York:
Columbia University Press.

\leavevmode\hypertarget{ref-Bellina2014}{}%
\textsc{Bellina}, B. 2014. Maritime silk roads' ornament industries:
Socio-political practices and cultural transfers in the south china sea
\emph{Cambridge Archaeological Journal} 24. Cambridge University Press:
345--77.

\leavevmode\hypertarget{ref-Berrocal2018}{}%
\textsc{Berrocal}, M.C., E.S. \textsc{Herrero}., M.G. \textsc{Moret}.,
A.U. \textsc{González}., M.T. \textsc{Pérez}., S.C. \textsc{Rodrı́guez}.,
A. \textsc{Chevalier}., F. \textsc{Valentin}. \& C.-h. \textsc{Tsang}.
2018. A comprised archaeological history of taiwan through the long-term
record of heping dao, keelung \emph{International Journal of Historical
Archaeology} 22. Springer: 905--40.

\leavevmode\hypertarget{ref-Bevan2016}{}%
\textsc{Bevan}, A. \& M. \textsc{Lake}. 2016. Intensities, interactions,
and uncertainties: Some new approaches to archaeological distributions,
in A. Bevan \& M. Lake (ed.) \emph{Computational approaches to
archaeological spaces}: 27--52. Routledge.

\leavevmode\hypertarget{ref-Blusse2000}{}%
\textsc{Blussé}, L. \& N. \textsc{Everts}. 2000. \emph{(FEII) the
formosan encounter. Notes on formosa's aboriginal society, a selection
of documents from dutch archival sources. Volume ii: 1636-1645}. Taipei:
Shung Ye Museum of Formosan Aborigines, Taipei.

\leavevmode\hypertarget{ref-Bonnier2019}{}%
\textsc{Bonnier}, A., M. \textsc{Finné}. \& E. \textsc{Weiberg}. 2019.
Examining land-use through gis-based kernel density estimation: A
re-evaluation of legacy data from the berbati-limnes survey
\emph{Journal of Field Archaeology} 44. Taylor \& Francis: 70--83.

\leavevmode\hypertarget{ref-Borao2001}{}%
\textsc{Borao}, H., J. E. 2001. \emph{(SIT, pp. 1-343). Spaniards in
taiwan, vol. I (1582--1641)}. Taipei: SMC Publishing.

\leavevmode\hypertarget{ref-Borao2009}{}%
\textsc{Borao}, J.E. 2009. \emph{The spanish experience in Taiwan,
1626-1642: The baroque ending of a renaissance endeavor}. Hong Kong:
Hong Kong University Press.

\leavevmode\hypertarget{ref-Bowles2010}{}%
\textsc{Bowles}, S., E. \textsc{Smith}. \& M. \textsc{Mulder}. 2010. The
emergence and persistence of inequality in premodern societies
\emph{Current anthropology} 51: 7--17.

\leavevmode\hypertarget{ref-Brumfiel1994}{}%
\textsc{Brumfiel}, E.M. 1994. Factional competition and political
development in the new world: An introduction, in E.M. Brumfiel \& J.
Fox (ed.) \emph{Factional competition and political development in the
new world}: 3--13. Cambridge: Cambridge University Press.

\leavevmode\hypertarget{ref-Carter2016}{}%
\textsc{Carter}, A.K. 2016. The production and exchange of glass and
stone beads in southeast asia from 500 bce to the early second
millennium ce: An assessment of the work of peter francis in light of
recent research \emph{Archaeological Research in Asia} 6. Elsevier:
16--29.

\leavevmode\hypertarget{ref-Chen2011}{}%
\textsc{Chen}, K.-t. 2011. A preliminary research of copper-based
artifacts and related remains discovered in the taiwan area \emph{Chung
Yang Yen Chiu Yuan Li Shih Yu Yen Yen Chiu So Chi K'an {[}Bulletin of
the Institute of History and Philology Academia Sinica{]}} 82: 169--259.

\leavevmode\hypertarget{ref-Chen1963}{}%
\textsc{Chen}, S. 1963. \emph{Kavalan ting zhi {[}kavalen culture
history{]}, taiwan wen xian cong kan di 106 zhong {[}taiwan literature
series: 106{]}}. Taipei: Economic Research Office, Bank ofTaiwan.

\leavevmode\hypertarget{ref-Chen2005}{}%
\textsc{Chen}, T.-j. 2005. \emph{Ji long shan yu dan shui yang : Dong ya
hai yu yu tai wan zao qi yan jiu, 1400-1700 {[}mount keelung and tai she
ocean: A study of east asian seas and the hisotry of Taiwan from 1400 to
1700{]}.} Taipei: Lian jing.

\leavevmode\hypertarget{ref-Chen2017}{}%
\textsc{Chen}, W.-c. 2017. The early occupation of taiwan, in J. Habu,
P.V. Lape, \& J.W. Olsen (ed.) \emph{Handbook of east and southeast
asian archaeology}: 277--91. Springer.

\leavevmode\hypertarget{ref-Chen2007}{}%
\textsc{Chen}, Y.-p. 2007. \emph{Qi wu lan yi zhi qiang jiu fa jue bao
gao {[}report on the archaeological excavations at ki-wu-lan site{]}}.
Yilan, Taiwan: Lanyang museum.

\leavevmode\hypertarget{ref-Cheng2008}{}%
\textsc{Cheng}, C.-F. 2008. Qi wu lan yi zhi yu she nei yi zhi chu tu bo
li zhu de xiang guan yan jiu {[}studies of glass beads excavated from
kivulan and shenei site, taiwan{]}. Master's thesis.

\leavevmode\hypertarget{ref-Chiu2004}{}%
\textsc{Chiu}, H.-L. 2004. Yilan xian jiao xi xiang qi wu lan yi zhi chu
tu mu zang yan jiu ──mai zang hang wei yu wen hua bian qian de guan cha
{[}investigations of mortuary behaviors and cultural change of the
kivulan site in i-lan county{]}. Master's thesis.

\leavevmode\hypertarget{ref-Clark1994}{}%
\textsc{Clark}, J.E. \& M. \textsc{Blake}. 1994. The power of prestige:
Competitive generosity and the emergence of rank societies in lowland
mesoamerica \emph{Factional competition and political development in the
New World}, 17--30.

\leavevmode\hypertarget{ref-Cort2017}{}%
\textsc{Cort}, L.A. 2017. Container jars from the maenam noi kilns,
thailand use and reuse along maritime trade routes in asia
\emph{Bulletin de l'École française d'Extrême-Orient} 103. École
française d'Extrême-Orient: 267--96.

\leavevmode\hypertarget{ref-Cortegoso2016}{}%
\textsc{Cortegoso}, V., R. \textsc{Barberena}., V. \textsc{Durán}. \& G.
\textsc{Lucero}. 2016. Geographic vectors of human mobility in the andes
(34--36° s): Comparative analysis of `minor'obsidian sources
\emph{Quaternary International} 422. Elsevier: 81--92.

\leavevmode\hypertarget{ref-Dietler1997}{}%
\textsc{Dietler}, M. 1997. The iron age in mediterranean france:
Colonial encounters, entanglements, and transformations \emph{Journal of
World Prehistory} 11. Springer: 269--358.

\leavevmode\hypertarget{ref-Dietler2005}{}%
--- 2005. The archaeology of colonization and the colonization of
archaeology: Theoretical challenges from an ancient mediterranean
colonial encounter, in G. Stein (ed.) \emph{The archaeology of colonial
encounters: Comparative perspectives}: 33--68. Santa Fe: NM: Sch. Am.
Res. Press.

\leavevmode\hypertarget{ref-Dietler2015}{}%
--- 2015. \emph{Archaeologies of colonialism: Consumption, entanglement,
and violence in ancient mediterranean france}. Univ of California Press.

\leavevmode\hypertarget{ref-Drennan2010}{}%
\textsc{Drennan}, R.D., C.E. \textsc{Peterson}. \& J.R. \textsc{Fox}.
2010. Degrees and kinds of inequality, in \emph{Pathways to power}:
45--76. Springer.

\leavevmode\hypertarget{ref-Ducke2015}{}%
\textsc{Ducke}, B. 2015. Spatial cluster detection in archaeology:
Current theory and practice, in J.A. Barcelo \& I. Bogdanovic (ed.)
\emph{Mathematics and archaeology}: 352--68. Barcelona, Spain: CRC Press
Boca Raton.

\leavevmode\hypertarget{ref-Feinman2000}{}%
\textsc{Feinman}, G.M. 2000. Corporate/network: New perspectives on
models of political action and the puebloan southwest \emph{Social
Theory in Archaeology, University of Utah Press, Salt Lake City},
31--51.

\leavevmode\hypertarget{ref-Francis2002}{}%
\textsc{Francis}, P. 2002. \emph{Asia's maritime bead trade: 300 bc to
the present}. University of Hawaii Press.

\leavevmode\hypertarget{ref-Galloway2006}{}%
\textsc{Galloway}, P. 2006. Material culture and text: Exploring the
spaces within and between \emph{Historical archaeology} 9. Blackwell
Publishing.

\leavevmode\hypertarget{ref-Gan2006}{}%
\textsc{Gan}, F., H. \textsc{Cheng}. \& Q. \textsc{Li}. 2006. Origin of
chinese ancient glasses---study on the earliest chinese ancient glasses
\emph{Science in China Series E: Technological Sciences} 49. Springer:
701--13.

\leavevmode\hypertarget{ref-Given2004}{}%
\textsc{Given}, M. 2004. \emph{The archaeology of the colonized}.
London; New York: Routledge.

\leavevmode\hypertarget{ref-Grave2013}{}%
\textsc{Grave}, P. \& I.J. \textsc{McNiven}. 2013. Geochemical
provenience of 16th--19th century ce asian ceramics from torres strait,
northeast australia \emph{Journal of Archaeological Science} 40.
Elsevier: 4538--51.

\leavevmode\hypertarget{ref-Hsieh2009}{}%
\textsc{Hsieh}, E. 2009. Yi lan qi wu lan yi zhi chu tu wai lai tao ci
qi zhi xiang guan yan jiu {[}the study of imported ceramics excavated at
the ki-wu-lan site, i-lan{]}. Master's thesis.

\leavevmode\hypertarget{ref-Hsieh1995}{}%
\textsc{Hsieh}, M.-L. 1995. An ping hu zou yi {[}some modest remarks on
the an-p'ing jug{]} \emph{Guo li tai wan da xue mei shu shi yan jiu ji
kan {[}Taida Journal of Art History{]}} 2. Graduate Institute Of Art
History National Taiwan University: 75--105.

\leavevmode\hypertarget{ref-Ino1996}{}%
\textsc{Ino}, K. 1996. \emph{Ping pu zu diao cha lu hang :Yi neng jia ju
(tai wan tong xin) xuan ji {[}field investigation trips at plains
indigenous peoples: Ino kanori (taiwan letters){]}}. Taipei: Yuan Liou.

\leavevmode\hypertarget{ref-Joyce2005}{}%
\textsc{Joyce}, R.A. 2005. Archaeology of the body \emph{Annual Review
of Anthropology} 34. Annual Reviews: 139--58.

\leavevmode\hypertarget{ref-Junker1993}{}%
\textsc{Junker}, L.L. 1993. Craft goods specialization and prestige
goods exchange in philippine chiefdoms of the fifteenth and sixteenth
centuries \emph{Asian Perspectives}. JSTOR, 1--35.

\leavevmode\hypertarget{ref-Kang2012}{}%
\textsc{Kang}, P. 2012. He lan dong yin du gong si zhi xia de ga ma lan
di qu te zhi {[}charateristics of kavalan under colonial rule of dutch
east India companies{]}, in M.-C. Hsu \& S.-Y. Li (ed.) \emph{Exploring
kiwulan: The ninth academic conference of yilan study}, 35: 291--317.
Yilan: Institute of Yilan County History.

\leavevmode\hypertarget{ref-Kang2016}{}%
--- 2016. \emph{Colonial imagination and local variations: The dutch
east india company and the formosan austronesians}. Lian-jing.

\leavevmode\hypertarget{ref-Ke1993}{}%
\textsc{Ke}, P. 1993. \emph{Kavalan zhi lue {[}record of kavalen{]}}.
Nantou: Historical Records Committee of Taiwan Provincial Government.

\leavevmode\hypertarget{ref-Kenoyer2000}{}%
\textsc{Kenoyer}, J.M. 2000. Wealth and socioeconomic hierarchies of the
indus valley civilization, in F.A. Janet Richards Mary Van Buren (ed.)
\emph{Order, legitimacy and wealth in early states}: 90--112. Cambridge
University Press Cambridge.

\leavevmode\hypertarget{ref-Ketel2011}{}%
\textsc{Ketel}, C. 2011. Identification of export porcelains from early
17th century voc shipwrecks and the linkage to their cultural
identification, in D. Core (ed.) \emph{The 2011 asia-pacific regional
conference on underwater cultural heritage proceedings}: 1--14. Manila,
Philippines.

\leavevmode\hypertarget{ref-Klose2018}{}%
\textsc{Klose}, J. \& C. \textsc{Schrire}. 2018. Asian ceramic
collections from voc sites at the cape, in C. Schrire (ed.)
\emph{Historical archaeology in south africa: Material culture of the
dutch east india company at the cape}: 101--41. Taylor; Francis.

\leavevmode\hypertarget{ref-Li2014}{}%
\textsc{Li}, C.-y. \& S.-j. \textsc{Chiu}. 2014. A report on excavations
in the yi-lang agricultural vocational high school site, 2000-2008
\emph{Field Archaeology of Taiwan} 17: 59--120.

\leavevmode\hypertarget{ref-LiandWu2006}{}%
\textsc{Li}, Y.-z. \& M.-z. \textsc{Wu}. 2006. \emph{Qing zai xi ban ya
ren zai tai wan, 1626-1642 {[}the spanish in Taiwan{]}}. Nantou: Taiwan
Historica.

\leavevmode\hypertarget{ref-Lin2015}{}%
\textsc{Lin}, S.-f. 2015. The relationship of prehistory cultural
development and paleoenvironmental changes: A case study of yilan plain,
ne taiwan, in Y.-c. Liu (ed.) \emph{Tai wan shi qian shi zhuan lun
{[}studies of the prehistory in taiwan{]}}: 319--45. Academia Sinica,
Lian-jing.

\leavevmode\hypertarget{ref-Liu2011}{}%
\textsc{Liu}, Y.-c. 2011. \emph{Tai wan quan zhi zhu min zhi kao gu pian
{[}the history of Taiwan: Indigenous people and archaeology{]}}. Vol. 3.
Nan-Tou: Taiwan Historica.

\leavevmode\hypertarget{ref-Liu2017}{}%
\textsc{Liu}, Y.-c. \& S.-C. \textsc{Wang}. 2017. Encountering the wider
world before the transition to history: Chinese ceramics in
proto-historic taiwan (tenth through sixteenth centuries), in M. Cruz
Berrocal \& C. Tsang (ed.) \emph{Historical archaeology of early modern
colonialism in asia-pacific: The southwest pacific and oceanian
regions}: 270--312. Florida, Gainesville: University Press of Florida.

\leavevmode\hypertarget{ref-Marwick2017}{}%
\textsc{Marwick}, B. 2017. Computational reproducibility in
archaeological research: Basic principles and a case study of their
implementation \emph{Journal of Archaeological Method and Theory} 24.
Springer: 424--50.

\leavevmode\hypertarget{ref-Marwick2018}{}%
\textsc{Marwick}, B., C. \textsc{Boettiger}. \& L. \textsc{Mullen}.
2018. Packaging data analytical work reproducibly using r (and friends)
\emph{The American Statistician} 72. Taylor \& Francis: 80--88.

\leavevmode\hypertarget{ref-Mullins2011}{}%
\textsc{Mullins}, P.R. 2011. The archaeology of consumption \emph{Annual
Review of Anthropology} 40. Annual Reviews: 133--44.

\leavevmode\hypertarget{ref-Nakamura1938}{}%
\textsc{Nakamura}. 1938. Overgekomen brieven en papieren. {[}The dutch
cencus record for indigenous peoples in taiwan{]} \emph{Southern
Anthropological Studies} 4: 12.

\leavevmode\hypertarget{ref-NMTH2005}{}%
\textsc{National Musuem of Taiwan History}, A. group in. 2005. Taiwan
under dutch and spanish: A report of historical archaeological research
in northern taiwan. Taipei: National Musuem of Taiwan History.

\leavevmode\hypertarget{ref-Rlanguage2019}{}%
\textsc{R Core Team}. 2019. \emph{R: A language and environment for
statistical computing}. Vienna, Austria: R Foundation for Statistical
Computing. \url{https://www.R-project.org}.

\leavevmode\hypertarget{ref-Rubertone2000}{}%
\textsc{Rubertone}, P.E. 2000. The historical archaeology of native
americans \emph{Annual Review of Anthropology} 29: 425--46.

\leavevmode\hypertarget{ref-Scaramelli2005}{}%
\textsc{Scaramelli}, F. \& K.T. \textsc{de Scaramelli}. 2005. The roles
of material culture in the colonization of the orinoco, venezuela
\emph{Journal of Social Archaeology} 5. Sage Publications Sage CA:
Thousand Oaks, CA: 135--68.

\leavevmode\hypertarget{ref-Siegel1999}{}%
\textsc{Siegel}, P.E. 1999. Contested places and places of contest: The
evolution of social power and ceremonial space in prehistoric Puerto
Rico \emph{Latin American Antiquity}, 209--38.

\leavevmode\hypertarget{ref-Silliman2001}{}%
\textsc{Silliman}, S. 2001. Agency, practical politics and the
archaeology of culture contact \emph{Journal of social archaeology} 1.
Sage Publications Sage CA: Thousand Oaks, CA: 190--209.

\leavevmode\hypertarget{ref-Silliman2005}{}%
\textsc{Silliman}, S.W. 2005. Culture contact or colonialism? Challenges
in the archaeology of native North America \emph{American Antiquity},
55--74.

\leavevmode\hypertarget{ref-Theunissen2000}{}%
\textsc{Theunissen}, R., P. \textsc{Grave}. \& G. \textsc{Bailey}. 2000.
Doubts on diffusion: Challenging the assumed indian origin of iron age
agate and carnelian beads in southeast Asia \emph{World archaeology} 32:
84--105.

\leavevmode\hypertarget{ref-TorrenceandClarke2000}{}%
\textsc{Torrence}, R. \& A. \textsc{Clarke}. 2000. Negotiating
difference: Practice makes theory for contemporary archaeology in
Oceania, in R. Torrence \& A. Clarke (ed.) \emph{The archaeology of
difference : Negotiating cross-cultural engagements in Oceania}: 1--31.
London; New York: Routledge.

\leavevmode\hypertarget{ref-Trabert2017}{}%
\textsc{Trabert}, S. 2017. Considering the indirect effects of
colonialism: Example from a great plains middle ground \emph{Journal of
Anthropological Archaeology} 48. Elsevier: 17--27.

\leavevmode\hypertarget{ref-Voss2005}{}%
\textsc{Voss}, B.L. 2005. From casta to californio: Social identity and
the archaeology of culture contact \emph{American Anthropologist} 107:
461--74.

\leavevmode\hypertarget{ref-Wang2018}{}%
\textsc{Wang}, K.-W. 2018. Glass beads in iron age and early modern
taiwan: An introduction \emph{BEADS: Journal of the Society of Bead
Researchers}: 16--30.

\leavevmode\hypertarget{ref-Wang2011}{}%
\textsc{Wang}, L.-Y. 2011. Yi lan qi wu lan yi zhi chu tu zhuang shi pin
zhi xiang guan yan jiu {[}a research of ornaments excavated at ki-wu-lan
site, i-lan{]}. Master's thesis.

\leavevmode\hypertarget{ref-Wang2007}{}%
\textsc{Wang}, S.-C. \& Y.-C. \textsc{Liu}. 2007. Shi qi shi ji qian hou
tai wan yan cao , yan dou yu bo li zhu shi de shu ru wang luo -yi ge xin
de jiao huan jie duan {[}the import networks of tobacco, tobacco pipes,
and glass bead ornaments into Taiwan circa the seventeenth century: A
new phase of exchange{]}. \emph{Taida Journal of Art History}: 51--83.

\leavevmode\hypertarget{ref-Yao1996}{}%
\textsc{Yao}, ying. 1996. \emph{Dong cha ji lue {[}record of taiwan{]},
taiwan wen xian cong kan di 007 zhong {[}taiwan literature series:
007{]}}. Nantou: Taiwan Historica.

\hypertarget{pagebreak-1}{%
\subparagraph{pagebreak}\label{pagebreak-1}}

\hypertarget{colophon}{%
\subsubsection{Colophon}\label{colophon}}

This report was generated on 2019-12-11 22:27:06 using the following
computational environment and dependencies:

\begin{verbatim}
#> ─ Session info ───────────────────────────────────────────────────────────────
#>  setting  value                       
#>  version  R version 3.6.0 (2019-04-26)
#>  os       macOS Mojave 10.14.6        
#>  system   x86_64, darwin15.6.0        
#>  ui       X11                         
#>  language (EN)                        
#>  collate  en_US.UTF-8                 
#>  ctype    en_US.UTF-8                 
#>  tz       America/Los_Angeles         
#>  date     2019-12-11                  
#> 
#> ─ Packages ───────────────────────────────────────────────────────────────────
#>  ! package     * version date       lib source        
#>  P assertthat    0.2.1   2019-03-21 [?] CRAN (R 3.6.0)
#>    backports     1.1.5   2019-10-02 [1] CRAN (R 3.6.0)
#>    callr         3.3.2   2019-09-22 [2] CRAN (R 3.6.0)
#>    cli           2.0.0   2019-12-09 [1] CRAN (R 3.6.0)
#>  P crayon        1.3.4   2017-09-16 [?] CRAN (R 3.6.0)
#>    desc          1.2.0   2018-05-01 [2] CRAN (R 3.6.0)
#>    devtools      2.2.1   2019-09-24 [2] CRAN (R 3.6.0)
#>    digest        0.6.23  2019-11-23 [1] CRAN (R 3.6.0)
#>    ellipsis      0.3.0   2019-09-20 [1] CRAN (R 3.6.0)
#>    evaluate      0.14    2019-05-28 [2] CRAN (R 3.6.0)
#>  P fansi         0.4.0   2018-10-05 [?] CRAN (R 3.6.0)
#>    fs            1.3.1   2019-05-06 [2] CRAN (R 3.6.0)
#>  P glue          1.3.1   2019-03-12 [?] CRAN (R 3.6.0)
#>    htmltools     0.4.0   2019-10-04 [2] CRAN (R 3.6.0)
#>    knitr         1.26    2019-11-12 [2] CRAN (R 3.6.0)
#>  P magrittr      1.5     2014-11-22 [?] CRAN (R 3.6.0)
#>    memoise       1.1.0   2017-04-21 [2] CRAN (R 3.6.0)
#>    pkgbuild      1.0.6   2019-10-09 [2] CRAN (R 3.6.0)
#>    pkgload       1.0.2   2018-10-29 [2] CRAN (R 3.6.0)
#>    prettyunits   1.0.2   2015-07-13 [2] CRAN (R 3.6.0)
#>    processx      3.4.1   2019-07-18 [2] CRAN (R 3.6.0)
#>    ps            1.3.0   2018-12-21 [2] CRAN (R 3.6.0)
#>    R6            2.4.1   2019-11-12 [1] CRAN (R 3.6.0)
#>    Rcpp          1.0.3   2019-11-08 [1] CRAN (R 3.6.0)
#>    remotes       2.1.0   2019-06-24 [2] CRAN (R 3.6.0)
#>    rlang         0.4.2   2019-11-23 [1] CRAN (R 3.6.0)
#>    rmarkdown     1.16    2019-10-01 [2] CRAN (R 3.6.0)
#>    rprojroot     1.3-2   2018-01-03 [2] CRAN (R 3.6.0)
#>    sessioninfo   1.1.1   2018-11-05 [2] CRAN (R 3.6.0)
#>  P stringi       1.4.3   2019-03-12 [?] CRAN (R 3.6.0)
#>  P stringr       1.4.0   2019-02-10 [?] CRAN (R 3.6.0)
#>    testthat      2.2.1   2019-07-25 [2] CRAN (R 3.6.0)
#>    usethis       1.5.1   2019-07-04 [2] CRAN (R 3.6.0)
#>  P withr         2.1.2   2018-03-15 [?] CRAN (R 3.6.0)
#>    xfun          0.11    2019-11-12 [2] CRAN (R 3.6.0)
#>    yaml          2.2.0   2018-07-25 [2] CRAN (R 3.6.0)
#> 
#> [1] /Users/bmarwick/Desktop/kwl-ornaments/renv/library/R-3.6/x86_64-apple-darwin15.6.0
#> [2] /Library/Frameworks/R.framework/Versions/3.6/Resources/library
#> 
#>  P ── Loaded and on-disk path mismatch.
\end{verbatim}

The current Git commit details are:

\begin{verbatim}
#> Local:    master /Users/bmarwick/Desktop/kwl-ornaments
#> Remote:   master @ origin (https://github.com/LiYingWang/kwl-ornaments)
#> Head:     [2a0bc8c] 2019-12-12: change texts in the datafile and the caption in the code
\end{verbatim}

Word count: 5263

\end{document}

### 2019-12-12 analysis/paper/paper.Rmd
updates to C14 table

@@
 ```{r C14-table-data}
+# column names should follow Radiocarbon journal style:  https://www-cambridge-org.offcampus.lib.washington.edu/core/journals/radiocarbon/information/instructions-contributors
+
 table_kw_dates <-
   kw_dates %>% 
-  mutate(`Lab code` = str_remove(`Lab code`, " ")) %>% 
+  mutate(`Lab code` = str_remove(`Lab code`, " "),
+         `Uncalibrated Age BP` = `Date BP`) %>% 
   mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
@@
   arrange(ave) %>% 
-  dplyr::select(`Lab code`, `Pit-Layer`, 
-                `Above mean sea level (cm)`, 
-                `Date BP`, 
-                `Calibrated date (95% conf. level)`, 
-                Context) %>% 
-  mutate(`Calibrated date (95% conf. level)` = 
+  mutate(`Calibrated Age BP (95% credible interval)` = 
            ifelse(is.na(`Calibrated date (95% conf. level)`), 
                   "", 
-                  `Calibrated date (95% conf. level)`))
+                  `Calibrated date (95% conf. level)`)) %>% 
+  mutate(`Uncalibrated Age BP` = str_remove_all(`Uncalibrated Age BP`, "yr")) %>% 
+  dplyr::select(`Lab code`, 
+                `Pit-Layer`, 
+                `Above mean sea level (cm)`, 
+                `Uncalibrated Age BP`, 
+                `Calibrated Age BP (95% credible interval)`, 
+                Context)
 
@@
              caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve.
-The codes in the context column refer to the excavator's numbering system, cf.
+The codes in the context column refer to the excavator's feature labels, cf.
 ADD CITATION TO THE REPORT HERE")

### 2019-12-13 analysis/paper/potsherds.R
prepare table for join

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(stringr)
 
 # read data
 kwl_lp <-
@@ -64,7 +65,11 @@ kwl_lp_clean <-
            into = c("start", "end"),
            sep = "~") %>%
   mutate_at(vars(start, end),
-            as.numeric)
+            as.numeric) %>%
+  mutate(`編號\n/層位`= ifelse(nchar(`編號\n/層位`) == 1,
+                           paste0("0", `編號\n/層位`),
+                           `編號\n/層位`)) %>%
+  mutate(join_id = paste0(`編號\n/ 現象號(P)`,"-",`編號\n/層位`))
 
 ## read ornament data
 kwl_upper <-

### 2019-12-13 analysis/paper/potsherds.R
join two tables, ornaments and potsherds

@@ -90,14 +90,14 @@ ornaments_period <-
                          level = c("Before European Contact",
                                    "European Presence",
                                    "Chinese Presence"))) %>%
-  mutate(Length = as.numeric(`Length(mm)`),
-         Thick = as.numeric(`Thick(mm)`),
-         Width = as.numeric(`Width(mm)`),
-         Perforation1 = as.numeric(`Perforation1(mm)`)) %>%
-  rowwise() %>%
-  mutate(Perforation_average =
-           mean(c(`Perforation1`,`Perforation2(mm)`),
-                na.rm = TRUE))
+  mutate(join_id = paste0(Pit_No,"-",Layer)) %>%
+  filter(Categories != "Unknown Metal") %>%
+  mutate(Categories = fct_lump(Categories,
+                               n = 5)) %>%
+  left_join(kwl_lp_clean, "join_id") %>%
+  rename(potsherds_weight = "重量小計") %>%
+  group_by(join_id, period, potsherds_weight) %>%
+  count(Layer)
 
 #
 kwl_lp_clean_summary <-

### 2019-12-13 analysis/paper/potsherds.R
test the correlation between ornaments and potsherds, add one plot

@@ -75,7 +75,7 @@ kwl_lp_clean <-
 kwl_upper <-
   readxl::read_excel(here("analysis", "data", "raw_data", "Kiwulan_Ornament_Upper.xlsx"))
 
-## clean the ornament data and assign periods
+## clean the ornament data, assign periods
 ornaments_period <-
   kwl_upper %>%
   filter(!is.na(`6-layers`)) %>%
@@ -93,50 +93,38 @@ ornaments_period <-
   mutate(join_id = paste0(Pit_No,"-",Layer)) %>%
   filter(Categories != "Unknown Metal") %>%
   mutate(Categories = fct_lump(Categories,
-                               n = 5)) %>%
+                               n = 5))
+
+## join with potsherds data
+ornaments_potsherds <-
+  ornaments_period %>%
   left_join(kwl_lp_clean, "join_id") %>%
   rename(potsherds_weight = "重量小計") %>%
+  mutate(potsherds_weight = as.numeric(potsherds_weight)) %>%
   group_by(join_id, period, potsherds_weight) %>%
-  count(Layer)
+  count(Layer) %>%
+  select(-Layer) %>%
+  rename(ornament_piece = "n") %>%
+  group_by(period) %>%
+  summarise(sum_pot= sum(potsherds_weight, na.rm = TRUE),
+            sum_orna= sum(ornament_piece, na.rm = TRUE))
 
-#
-kwl_lp_clean_summary <-
-  kwl_lp_clean %>%
-  group_by(end) %>%
-  summarise(mean_weight= mean(重量小計, na.rm = TRUE))
+## plot
+plot_ornaments_potsherds <-
+  ornaments_potsherds %>%
+  pivot_longer(-period,
+               names_to = "variable",
+               values_to = "value") %>%
+  ggplot() +
+  geom_col(aes(variable, value)) +
+  facet_wrap(~ period) +
+  scale_y_log10() +
+  labs(x = "Artifact",
+       y = "Amount") +
+  theme_minimal(base_size = 10)
 
-kwl_lp_clean_piece <-
-  kwl_lp_clean %>%
-  group_by(`編號\n/ 現象號(P)`) %>%
-  summarise(sum_p= sum(件數小計, na.rm = TRUE))
-
-# plot
-ggplot(kwl_lp_clean,
-       aes(x = end,
-           y = 重量小計)) +
-  geom_line() +
-  facet_wrap(~`編號\n/ 現象號(P)`)
-
[…]

### 2019-12-13 analysis/paper/potsherds.R
stash

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(stringr)
+library(here)
 
 # read data
 kwl_lp <-
@@ -98,6 +99,20 @@ ornaments_period <-
   rename(potsherds_weight = "重量小計") %>%
   group_by(join_id, period, potsherds_weight) %>%
   count(Layer)
+
+orn_pot_summary <-
+ornaments_period %>%
+  group_by(period) %>%
+  summarise(sum_pot = sum(potsherds_weight, na.rm = TRUE),
+            sum_orn = sum(n, na.rm = TRUE))
+
+orn_pot_summary %>%
+  uncount( sum_pot ) %>%
+  uncount(sum_orn)
+
+
+library(infer)
+
 
 #
 kwl_lp_clean_summary <-

### 2019-12-13 analysis/paper/paper.Rmd
update C14 data file and the caption in the code

@@
 The codes in the context column refer to the excavator's feature labels, cf.
-ADD CITATION TO THE REPORT HERE")
+Chen 2007")
 ```
@@
 
-Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequencies of the major classes of ornaments for different the time periods at Kiwulan.
+Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequencies of the major classes of ornaments for different time periods at Kiwulan.
 The difference in frequencies between the three time periods reflect significant differences in the use of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$).

### 2019-12-13 analysis/paper/potsherds.R
use glm to model pottery-ornament relationship

@@ -103,16 +103,16 @@ ornaments_potsherds <-
   rename(potsherds_weight = "重量小計") %>%
   mutate(potsherds_weight = as.numeric(potsherds_weight)) %>%
   group_by(join_id, period, potsherds_weight) %>%
-<<<<<<< HEAD
-  count(Layer)
-
-orn_pot_summary <-
-ornaments_period %>%
+  count(Layer) %>%
+  select(-Layer) %>%
+  rename(ornament_piece = "n") %>%
   group_by(period) %>%
-  summarise(sum_pot = sum(potsherds_weight, na.rm = TRUE),
-            sum_orn = sum(n, na.rm = TRUE))
+  summarise(sum_pot= sum(potsherds_weight, na.rm = TRUE),
+            sum_orna= sum(ornament_piece, na.rm = TRUE))
 
-orn_pot_summary %>%
+
+
+ornaments_potsherds %>%
   uncount( sum_pot ) %>%
   uncount(sum_orn)
 
@@ -158,22 +158,16 @@ ggplot() +
        x = "Depth (cm)",
        y = "Weights of Sherds") +
   theme_minimal()
-=======
-  count(Layer) %>%
-  select(-Layer) %>%
-  rename(ornament_piece = "n") %>%
-  group_by(period) %>%
-  summarise(sum_pot= sum(potsherds_weight, na.rm = TRUE),
-            sum_orna= sum(ornament_piece, na.rm = TRUE))
->>>>>>> 7cefae231a6aeb11260475c9d015e4864bdb20da
 
 ## plot
-plot_ornaments_potsherds <-
+ornaments_potsherds_long <-
   ornaments_potsherds %>%
   pivot_longer(-period,
                names_to = "variable",
-               values_to = "value") %>%
-  ggplot() +
+               values_to = "value")
+
+
+  ggplot(ornaments_potsherds_long) +
   geom_col(aes(variable, value)) +
   facet_wrap(~ period) +
   scale_y_log10() +
@@ -181,8 +175,38 @@ plot_ornaments_potsherds <-
        y = "Amount") +
   theme_minimal(base_size = 10)
 
-## test
-chisq.test(ornaments_potsherds)
-library(Hmisc)
-rcorr(ornaments_potsherds, type="pearson")
-t.test(ornaments_potsherds$sum_orna, ornaments_potsherds$sum_pot)
+  ggplot(ornaments_potsherds_long) +
+    geom_col(aes(period,
+                 fill = variable,
+                 value
+                 )) +
[…]

### 2019-12-14 analysis/paper/potsherds.R
extract p-value from glm

@@ -205,6 +205,9 @@ fit <- (glm(formula = sum_orna ~ sum_pot,
 
 summary(fit)
 
+pval <- unname(coef(summary(fit))[,'Pr(>|z|)'][2])
+# how to print with a reasonable number of significant figures:
+signif(pval, 4)
 
 
 

### 2019-12-14 analysis/paper/potsherds.R
add beta

@@ -209,6 +209,10 @@ pval <- unname(coef(summary(fit))[,'Pr(>|z|)'][2])
 # how to print with a reasonable number of significant figures:
 signif(pval, 4)
 
+beta <- unname(coef(summary(fit))[,'Estimate'][2])
+# how to print with a reasonable number of significant figures:
+signif(beta, 4)
+
 
 
 

### 2019-12-14 analysis/paper/paper.Rmd
updates on the linear model

@@
 chisq_test_result_df <- round(chisq_test_result$parameter, 2)
-chisq_test_result_p_value <- (chisq_test_result$p.value)
+chisq_test_result_p_value <- round(chisq_test_result$p.value,
+                                   abs(floor(log10(chisq_test_result$p.value)-2)))
 ```
@@
 ## Changes in the frequencies of ornament types over time
+
+```{r poisson-linear-model}
+ornaments_potsherds <- read.csv(here("analysis", "data", "raw_data", "kwl_Ornament_potsherds.csv"))
+## test: can we predict the number of ornaments using the mass of pottery?
+# yes, seems like it
+
+# poisson model
+fit <- (glm(formula = sum_orna ~ sum_pot,
+            family = "poisson",
+            data = ornaments_potsherds))
+
+linear_pvalue <- round(coef(summary(fit))[,'Pr(>|z|)'][2], 
+                       abs(floor(log10(coef(summary(fit))[,'Pr(>|z|)'][2])-2)))
+
+```
 
@@
 However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period.
+When comparing with locally made pottery that could reflect Kiwulan population, the proportion of ornaments relative to the total mass of pottery remains the same, showing the general decreasing pattern in the Chinese period (Poisson regression model, p-value = $`r linear_pvalue`$).
 
@@
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
-We may be able to interpreate Yilan social organisation as moving from corporate mode before the European arrival, to a network mode during European presence.
+We may be able to interprete Yilan social organisation as moving from corporate mode before the European arrival, to a network mode during European presence.
 The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode.

### 2019-12-14 analysis/paper/potsherds.R

@@ -110,15 +110,13 @@ ornaments_potsherds <-
   summarise(sum_pot= sum(potsherds_weight, na.rm = TRUE),
             sum_orna= sum(ornament_piece, na.rm = TRUE))
 
-
-
 ornaments_potsherds %>%
   uncount( sum_pot ) %>%
   uncount(sum_orn)
 
+write.csv(ornaments_potsherds, here("analysis", "data", "raw_data", "Kwl_Ornament_potsherds.csv"))
 
 library(infer)
-
 
 #
 kwl_lp_clean_summary <-
@@ -202,8 +200,15 @@ ornaments_potsherds_long <-
 fit <- (glm(formula = sum_orna ~ sum_pot,
             family = "poisson",
             data = ornaments_potsherds))
-
 summary(fit)
+
+fit_null <- (glm(formula = sum_orna ~ 1,
+            family = "poisson",
+            data = ornaments_potsherds))
+
+with(anova(fit_null, fit),pchisq(Deviance,Df,lower.tail=FALSE)[2])
+
+
 
 
 

### 2019-12-14 analysis/paper/paper.Rmd

@@
 We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.
-
 # Reproducibility and open source materials
@@
 
-linear_pvalue <- round(coef(summary(fit))[,'Pr(>|z|)'][2], 
-                       abs(floor(log10(coef(summary(fit))[,'Pr(>|z|)'][2])-2)))
+pval <- unname(coef(summary(fit))[,'Pr(>|z|)'][2])
+# how to print with a reasonable number of significant figures:
+pval_print <- signif(pval, 4)
+
+beta <- unname(coef(summary(fit))[,'Estimate'][2])
+# how to print with a reasonable number of significant figures:
+beta_print <- signif(beta, 4)
 
@@
 However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period.
-When comparing with locally made pottery that could reflect Kiwulan population, the proportion of ornaments relative to the total mass of pottery remains the same, showing the general decreasing pattern in the Chinese period (Poisson regression model, p-value = $`r linear_pvalue`$).
+To model the number of ornaments as a function of the mass of ceramics in each period, a Poisson GLM with a log link function was used.
+The model reveals that ceramic abundance strongly predicts the number of ornaments (β = `r beta_print`, p = `r pval_print`).
+If ceramic abundance is a suitable proxy for population at Kiuwan due to its basic role as cooking vessels, then ornament quantities per period may be influenced by the number of people living at the site.
 

### 2019-12-14 analysis/paper/paper.Rmd
change title, fix typo, and change spelling

@@
 ---
-title: "Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period"
+title: "Ornaments as indicators of social changes in northeastern Taiwan before and after European colonial period"
 author:
@@
 The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered.
-This hints at increasing social inequality resulting from a colonial influence.
-Ornaments give insights into the increasing social inequality stimulated by European colonisation, and show the agency of indigenous people to incorporate ornaments into their social system.
+This hints at increasing social inequality resulting from colonial influence.
+Ornaments give insights into the increasing social inequality stimulated by European colonization, and show the agency of indigenous people to incorporate ornaments into their social system.
 keywords: |
-  Historical archaeology; Iron Age; European colonisation; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
+  Historical archaeology; Iron Age; European colonization; Trade ornaments; Exchange networks; Taiwan; spatial pattern
 highlights: |
@@
 The island of Taiwan lies at the junction of mainland China, Southeast Asia, and Northeast Asia in the Pacific Ocean.
-The prehistory of Taiwan island could be roughly divided into three major periods, paleolithic (c.
+The prehistory of Taiwan island could be roughly divided into three major periods, palaeolithic (c.
 27,000 BP- 5000BP), Neolithic (c.
@@
 132-149].
-The use of beads as prestige goods is further indicated by their role in bride price payments, and compensation to resolve disputes [@LiandWu2006, pp.
+The use of beads as prestige goods is further indicated by their role in bride price payments, and compenzation to resolve disputes [@LiandWu2006, pp.
 132-149].
@@
 Other people wore carnelian beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonisation, an academic field survey for plains indigenous groups reported that fish-shaped metal necklaces necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of Japanese colonization, an academic field survey for plains indigenous groups reported that fish-shaped metal necklaces necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
-Although these historical records are fragmentary and may contain some biases [@Galloway2006] that have not yet been studied in detail, we find consistency among multiple sources in their descriptions of how ornaments represent high status or specialised social roles in indigenous communities in Yilan.
+Although these historical records are fragmentary and may contain some biases [@Galloway2006] that have not yet been studied in detail, we find consistency among multiple sources in their descriptions of how ornaments represent high status or specialized social roles in indigenous communities in Yilan.
 Compared to the European period, there are fewer documentary mentions of beads in the Chinese period and the descriptions are limited to clothing, but these generally confirm the role of beads as status markers.
@@
 The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
-The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement.
-Artefact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
+The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement.
+Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
 Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from.
@@
 These component divisions are based on the differences in the colour and texture of the deposit.
-The interpration of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007].
+The interpretation of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007].
 There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
-We focus on the upper component because only this componant spans the periods of pre-European contact, European presence, and Chinese presence.
+We focus on the upper component because only this component spans the periods of pre-European contact, European presence, and Chinese presence.
 In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases.
 Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011].
-However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artefacts to the pre-European, European, and the Chinese periods.
+However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artifacts to the pre-European, European, and the Chinese periods.
 
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artefact counts", fig.width=image_width_1_col_inch}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 knitr::kable(ornaments_shape,
-             caption = "Ornament subtype at Kiwulan. The numbers represent artefact counts ")
+             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
 # some basic stat
@@
 
-To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualisations contained in this paper is included in http://doi.org/10.17605/OSF.IO/R8YGA.
-Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualisations and tests reported here.
+To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in http://doi.org/10.17605/OSF.IO/R8YGA.
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 To model the number of ornaments as a function of the mass of ceramics in each period, a Poisson GLM with a log link function was used.
-The model reveals that ceramic abundance strongly predicts the number of ornaments (β = `r beta_print`, p = `r pval_print`).
+The model reveals that ceramic abundance strongly predicts the number of ornaments (β = $`r beta_print`$, p = $`r pval_print`$).
 If ceramic abundance is a suitable proxy for population at Kiuwan due to its basic role as cooking vessels, then ornament quantities per period may be influenced by the number of people living at the site.
@@
 
-Point pattern analysis can assess whether the distribution of artefacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
-To prepare the ornament location data for point pattern analysis, because artefacts from Kiwulan lack exact piece-provance data, we assigned each ornament to a random coordinate pair in the square it was recovered from.
+Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
+To prepare the ornament location data for point pattern analysis, because artifacts from Kiwulan lack exact piece-provance data, we assigned each ornament to a random coordinate pair in the square it was recovered from.
 The next step was to divide the ornaments into three time periods.
@@
 Here we use KDE to visualize core areas of ornaments and surrounding neighbourhoods [@Bonnier2019; @Cortegoso2016].
-Density values of artefacts per square meter were calculated for each cell.
+Density values of artifacts per square meter were calculated for each cell.
 
@@
 The observed distribution of ornaments is more similar to the random distributions during the Chinese period, with about one third of the simulated values are greater than our observed ANN value.
-The Chinese period has fewer artefacts in any category, likely reflecting a smaller population at Kiwulan at this time, making spatial patterns and hotspots difficult to discern with confidence.
+The Chinese period has fewer artifacts in any category, likely reflecting a smaller population at Kiwulan at this time, making spatial patterns and hotspots difficult to discern with confidence.
 This testing reveals clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
@@
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
-We may be able to interprete Yilan social organisation as moving from corporate mode before the European arrival, to a network mode during European presence.
+We may be able to interprete Yilan social organization as moving from corporate mode before the European arrival, to a network mode during European presence.
 The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode.
@@
 By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
-Ornaments give insights into the amplification of social inequality stimulated by European colonisation.
+Ornaments give insights into the amplification of social inequality stimulated by European colonization.
 It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
 Future work could extend this approach to sourcing the ornaments directly with geochemical methods, and studies of other trade goods such as ceramics.
-We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardisation, and mortuary practices.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardization, and mortuary practices.
 

### 2019-12-15 analysis/paper/paper.Rmd
adjust and save C14 age figure to figure file

@@
 132-149].
-The use of beads as prestige goods is further indicated by their role in bride price payments, and compenzation to resolve disputes [@LiandWu2006, pp.
+The use of beads as prestige goods is further indicated by their role in bride price payments, and compensation to resolve disputes [@LiandWu2006, pp.
 132-149].
@@
           by = c(0, 0),
-          byend = c(0.00175),
+          byend = c(0.0020),
           bxend = bx)
 
+kwl_ages_curve <-
 ggplot() +
@@
            alpha = 0.3) +
-    annotate("text", x = 350, y = 0.0022,
+    annotate("text", x = 350, y = 0.0026,
            label = "Upper component",
@@
            alpha = 0.3) +
-      annotate("text", x = 1050, y = 0.00220,
+    annotate("text", x = 1050, y = 0.0026,
            label = "Lower component",
@@
 
-# to describe in your text, ?BchronDensity
-# "We used the Bchron R package to compute a non-parametric phase model 
+# see ?BchronDensity. "We used the Bchron R package to compute a non-parametric phase model 
 # on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
-# FIgure X shows the results of this analysis... indicating that the peak period of occupation is around 600 BC"
+# indicating that the peak period of occupation is around 600 BC"
+
+ggsave(plot = kwl_ages_curve,
+       here("analysis", "figures",
+       "plot-c14-ages-distribution.jpg"),
+       width = 120, 
+       dpi = 300,
+       height = 60,
+       units = "mm")
+
+knitr::include_graphics(here("analysis", 
+                             "figures", 
+                             "plot-c14-ages-distribution.jpg"))
+
 ```

### 2019-12-15 analysis/paper/paper.Rmd
updates the bead name in the plots

@@
   filter(Categories != "Unknown Metal") %>%
+  mutate(Categories = ifelse(Categories == "Agate bead", "Carnelian bead", Categories),
+         Categories = ifelse(Categories == "Golden bead", "Gold-foil bead", Categories),
+         Categories = ifelse(Categories == "Bell", "Metal bell", Categories)) %>%
   ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
@@
          Type = ifelse(Type == "small II", "thin small", Type),
-         Type = ifelse(Type == "large thin string", "thin large", Type))
+         Type = ifelse(Type == "large thin string", "thin large", Type),
+         Categories = ifelse(Categories == "Agate bead", "Carnelian bead", Categories),
+         Categories = ifelse(Categories == "Golden bead", "Gold-foil bead", Categories),
+         Categories = ifelse(Categories == "Bell", "Metal bell", Categories))
 ornaments_shape <-
@@
   ornaments_count %>% 
-  filter(Categories == "Agate bead") %>% 
+  filter(Categories == "Carnelian bead") %>% 
   pull(n)
@@
   ornaments_count %>% 
-  filter(Categories == "Bell") %>% 
+  filter(Categories == "Metal bell") %>% 
   pull(n)
@@
 ornaments_count <-
-  ornaments_period %>% 
+  ornaments_period_clean %>% 
   filter(!is.na(`6-layers`),
@@
   ornaments_shape_count %>% 
-  filter(Categories == 'Agate bead',
+  filter(Categories == 'Carnelian bead',
          Type != 'other')
@@
   ornaments_shape_count %>% 
-  filter(Categories == 'Bell',
+  filter(Categories == 'Metal bell',
          Type != 'unclassified')
@@
   labs(y = "",
-       x = "Bell") +
+       x = " Metal bell") +
   theme_minimal(base_size = base_size_multi_plot) +
@@
 ornaments_period_join <-
-  ornaments_period %>%
+  ornaments_period_clean %>%
   mutate(period = as.factor(period)) %>% # seems need this line first
@@
                            'Glass bead',
-                           'Agate bead'))
+                           'Carnelian bead'))
 

### 2019-12-15 analysis/paper/paper.Rmd
fix typo in a file name

@@
 ```{r poisson-linear-model}
-ornaments_potsherds <- read.csv(here("analysis", "data", "raw_data", "kwl_Ornament_potsherds.csv"))
+ornaments_potsherds <- read.csv(here("analysis", "data", "raw_data", "Kwl_Ornament_potsherds.csv"))
 ## test: can we predict the number of ornaments using the mass of pottery?
@@
 # poisson model
-fit <- (glm(formula = sum_orna ~ sum_pot,
+fit <- glm(formula = sum_orna ~ sum_pot,
             family = "poisson",
-            data = ornaments_potsherds))
+            data = ornaments_potsherds)
 
@@
 beta_print <- signif(beta, 4)
-
 ```

### 2019-12-15 analysis/paper/paper.Rmd
remove period from fig captions, run all the code chuck

@@
 Later in 1683, Taiwan was incorporated into the Qing dynasty in China and a large wave of Han Chinese migrated to Taiwan during the late 18th century.
-Because of natural safe harbors, northeast Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
+Because of natural safe harbors, northeastern Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
 Although located on the periphery of regional trade centers, Yilan was connected to trade networks via visits of other indigenous groups, Chinese merchants, and Europeans, via sea.
@@
 There is no direct evidence showing European delivery of beads, however, a large amount of the glass beads containing gold foil (hereafter, gold-foil beads) at Kiwulan might have been introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, as part of the trading route of the Spanish between 16-19th century [@Wang2007].
-Both archaeological evidence and historical records indicate northeastern Taiwan was involved inregional networks with East Asia in the late Iron age.
+Both archaeological evidence and historical records indicate northeastern Taiwan was involved in regional networks with East Asia in the late Iron age.
 These included Chinese merchants trading metal items, clothes, and beads with local indigenous people in Taiwan in exchange for local resources.
@@
 ```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text. Map data from naturalearthdata.com", fig.width=image_width_2_col_inch}
-# need to add the locations of Keelung, Tamsui, and hoping dao
 library(here)
@@
 Ages calibrated with the Bchron package (Parnell et al.
-2008).", cache=TRUE}
+2008)", cache = TRUE}
 
@@
                              "plot-c14-ages-distribution.jpg"))
-
 ```
@@
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D classes were from original excavation report (Chen 2007).", echo=FALSE}
+The photographs of B, C, D classes were from original excavation report (Chen 2007)", echo=FALSE}
 knitr::include_graphics(here("analysis", "figures", "ornament.png"))
@@
 
-```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods."}
+```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods"}
 ### delete one plot
@@
 
-```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads."}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads"}
 library(cowplot)
@@
 
-```{r plot-all-densities, fig.cap= "Kernel density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb."}
+```{r plot-all-densities, fig.cap= "Kernel density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb"}
 # apply function to each data frame in the list
@@
 X-axis values represent ANN expected value under a completely random process resulted from simulated pattern.
-Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value."}
+Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value"}
 

### 2019-12-16 analysis/paper/paper.Rmd
solve plotting and path problems by removing cache

@@
   comment = "#>",
-  cache = TRUE,
+  cache = FALSE,
   dpi = 300,
@@
 Ages calibrated with the Bchron package (Parnell et al.
-2008)", cache = TRUE}
+2008)"}
 
@@
 
-```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads"}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads", eval = TRUE}
 library(cowplot)

### 2019-12-16 analysis/paper/paper.Rmd
getting error message while knitting

--- 2019-12-15 analysis/paper/paper.Rmd
@@
 We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.
+
 # Reproducibility and open source materials
@@
 plot_grid(ring_shape_plot,
-          agate_bead_shape_plot, 
-          bell_shape_plot, 
+          #agate_bead_shape_plot, 
+          #bell_shape_plot, 
           glass_bead_plot,

### 2019-12-16 analysis/paper/paper.Rmd
solved the knitting and ploting issue

--- 2019-12-16 analysis/paper/paper.Rmd
@@
 
-```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads", eval = TRUE}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads"}
 library(cowplot)

### 2019-12-16 analysis/paper/paper.Rmd
fix typos in text

@@
 This is because of to a lack of archaeological evidence of beadmaking waste, metalworking, or accessible local raw materials.
-The chemical composition of glass beads from this region shows a high content of lead and, toghether with the winding/folding technique, these details suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
+The chemical composition of glass beads from this region shows a high content of lead and, together with the winding/folding technique, these details suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].
 Although there is a wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common components of metal ornaments are brass and copper, with a small number made from lead and tin that indicates multiple origins that include Southeast Asia [@Chen2011].
@@
 Large dark brown glazed stoneware jars may have been made in Southeast Asia, and are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
-The earliest evidence oflight gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Hsieh2009; @Hsieh1995].
+The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Hsieh2009; @Hsieh1995].
 Southeast China is assumed to be the origin of the light gray glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, such as the Zeelandia fort site in Tainan [@Wang2007].
@@
 The model reveals that ceramic abundance strongly predicts the number of ornaments (β = $`r beta_print`$, p = $`r pval_print`$).
-If ceramic abundance is a suitable proxy for population at Kiuwan due to its basic role as cooking vessels, then ornament quantities per period may be influenced by the number of people living at the site.
+If ceramic abundance is a suitable proxy for population at Kiwuan due to its basic role as cooking vessels, then ornament quantities per period may be influenced by the number of people living at the site.
 
@@
 Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
-To prepare the ornament location data for point pattern analysis, because artifacts from Kiwulan lack exact piece-provance data, we assigned each ornament to a random coordinate pair in the square it was recovered from.
+To prepare the ornament location data for point pattern analysis, because artifacts from Kiwulan lack exact piece-provenance data, we assigned each ornament to a random coordinate pair in the square it was recovered from.
 The next step was to divide the ornaments into three time periods.
@@
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
-We may be able to interprete Yilan social organization as moving from corporate mode before the European arrival, to a network mode during European presence.
+We may be able to interpret Yilan social organization as moving from corporate mode before the European arrival, to a network mode during European presence.
 The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode.
@@
 
-Examination of the archaeological record at the peripheries of colonial activities can show how remote indeignous groups were affected by major European colonial processes [@Trabert2017].
+Examination of the archaeological record at the peripheries of colonial activities can show how remote indigenous groups were affected by major European colonial processes [@Trabert2017].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
@@
 The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context.
-Before European contact, ornaments were traded into local indigenous societies via the regional exchagne network with Chinese merchants and viewed as prestige goods in local indigenous culture according to their distribution in the archaeological contexts.
+Before European contact, ornaments were traded into local indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local indigenous culture according to their distribution in the archaeological contexts.
 After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan.

### 2019-12-17 analysis/paper/paper.Rmd
rearrange the order of figures and add cross referencing

@@
 
-The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Keelung in 1626, and Fort San Domingo in 1629 at Tamsui.
+The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Helping dao, Keelung in 1626, and Fort San Domingo in 1629 at Tamsui (Figure \@ref(fig:KWL-map)).
 They sent missionaries to local indigenous settlements in this region [@Blusse2000, pp.
@@
 
-# Excavations at Kiwulan in northeastern Taiwan
-
 ```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text. Map data from naturalearthdata.com", fig.width=image_width_2_col_inch}
@@
 ```
+
+# Excavations at Kiwulan in northeastern Taiwan
+
+Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that during 2001-2004 in advance of a water diversion project and road bridge construction.
+The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
+The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement.
+Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
+Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from (Figure \@ref(fig:KWL-sam-area)).
+In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
+At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction.
 
@@
                 na.rm = TRUE))
+
+# some basic stat for personal ornament section 
+sam_post_hole <- nrow(ornaments_period)
+total_post_hole <- nrow(kwl_upper)
+sample_per <- round(sam_post_hole/total_post_hole*100, 1)
 ```
@@
 ```
-
-Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that during 2001-2004 in advance of a water diversion project and road bridge construction.
-The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
-The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement.
-Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
-Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from.
-In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
-At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction (Figure \@ref(fig:KWL-sam-area)).
 
@@
 
+Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens.
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
+Occupation floors were not identified during excavation.
+We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
+There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan.
+Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173.
+The high number of ornaments in these burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+We exclude burials from this analysis because most of them date to the pre-European period (n = 26), limiting the usefulness of comparisons between the periods.
+In addition to `r sam_post_hole` ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts.
+We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.
+
+```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
+A: carnelian beads, B: bells, C: glass beads and gold-foil beads, D: metal rings.
+Photographs are presented in the same order as those subtypes in the table but from left to right instead.
+The photographs of B, C, D classes were from original excavation report (Chen 2007)", echo=FALSE}
+knitr::include_graphics(here("analysis", "figures", "ornament.png"))
+```
+
 ```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
@@
              caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
-# some basic stat
-sam_post_hole <- nrow(ornaments_period)
-total_post_hole <- nrow(kwl_upper)
-sample_per <- round(sam_post_hole/total_post_hole*100, 1)
 
@@
 
-```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
-A: carnelian beads, B: bells, C: glass beads and gold-foil beads, D: metal rings.
-Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D classes were from original excavation report (Chen 2007)", echo=FALSE}
-knitr::include_graphics(here("analysis", "figures", "ornament.png"))
-```
-
-Ornaments were found in a variety of archaeological contexts including post-holes area, burials, and middens.
-This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
-Occupation floors were not identified during excavation.
-We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
-There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan.
-Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173.
-The high number of ornaments in these burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
-We exclude burials from this analysis because most of them date to the pre-European period (n = 26), limiting the usefulness of comparisons between the periods.
-In addition to `r sam_post_hole`  ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts.
-We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.
-
 # Reproducibility and open source materials
@@
 The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.
+
+# Results
+
+## Changes in the frequencies of ornament types over time
 
@@
 ```
-
-# Results
-
-## Changes in the frequencies of ornament types over time
 
@@
 
+## Changes in patterns of the spatial distribution of ornament types
+
 ```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
@@
 
-## Changes in patterns of the spatial distribution of ornament types
-
 Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from the research area for each time period.
@@
 In the Chinese period, both the amount and density of different classes of ornaments decreased.
+
+## Point pattern analysis of ornament distribution
 
@@
 
-## Point pattern analysis of ornament distribution
-
 Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.

### 2019-12-17 analysis/paper/paper.Rmd
submitted to ARIA

@@
 This hints at increasing social inequality resulting from colonial influence.
-Ornaments give insights into the increasing social inequality stimulated by European colonization, and show the agency of indigenous people to incorporate ornaments into their social system.
+Ornaments give insights into the increasing social inequality stimulated by the European colonial presence, and show the agency of indigenous people to incorporate ornaments into their social system.
 keywords: |
-  Historical archaeology; Iron Age; European colonization; Trade ornaments; Exchange networks; Taiwan; spatial pattern
+  Historical archaeology; European colonial period; 17th century; Trade ornaments; Exchange networks; spatial pattern; East Asia
 highlights: |
@@
 150 years.
-These component divisions are based on the differences in the colour and texture of the deposit.
+These component divisions are based on the differences in the color and texture of the deposit.
 The interpretation of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007].

### 2020-01-28 analysis/paper/paper.Rmd
edit historical background section based on BF's comments

@@
 In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan.
-They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tributes [@Andrade2007, ch.
+They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tribute [@Andrade2007, ch.
 9; @Kang2016, ch.
 4].
-The Dutch provided beads and other goods based on negotations with indigenous communities to secure alliances in the annual ceremony or during travelling [@Kang2016, ch.
+The Dutch provided beads and other goods based on negotiations with indigenous communities to secure alliances in the annual ceremony or during their travels [@Kang2016, ch.
 6].
@@
 Other people wore carnelian beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonization, an academic field survey for plains indigenous groups reported that fish-shaped metal necklaces necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of Japanese colonization, an academic field survey for plains indigenous groups reported that fish-shaped metal necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
@@
 Ornaments found in northeastern Taiwan in the early historical period, including glass beads, stone beads, and metal ornaments, are considered to have been imported from other regions.
-This is because of to a lack of archaeological evidence of beadmaking waste, metalworking, or accessible local raw materials.
+This is because of a lack of archaeological evidence of beadmaking waste, metalworking, or accessible local raw materials.
 The chemical composition of glass beads from this region shows a high content of lead and, together with the winding/folding technique, these details suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018].

### 2020-01-28 analysis/paper/paper.Rmd
edit the title

@@
 ---
-title: "Ornaments as indicators of social changes in northeastern Taiwan before and after European colonial period"
+title: "Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period"
 author:

### 2020-04-14 analysis/paper/paper.Rmd
edits on KWL background section based on BF's comment

@@
 
-Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that during 2001-2004 in advance of a water diversion project and road bridge construction.
+Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that was conducted between 2001-2004 in advance of a water diversion project and road bridge construction.
 The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
@@
 The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning c.
-150 years.
+150 years [@Chen2007].
+The sterility may be related to dry weather leading to site abandonment according to pollen analysis [@Lin2015].
 These component divisions are based on the differences in the color and texture of the deposit.
-The interpretation of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007].
+Whether these two components represent a culture continuity from the same people is still under debate [@Chiu2004; @Chen2007].
 There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
@@
 Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit.
-For the upper component, the blue line indicates the start of European presence, while the red line is the Chinese presence.
+For the upper component, the red line indicates the start of European presence, while the blue line is the Chinese presence.
 Ages calibrated with the Bchron package (Parnell et al.
@@
                 `Pit-Layer`, 
-                `Above mean sea level (cm)`, 
+                `Depth (cm)`, 
                 `Uncalibrated Age BP`, 
@@
              caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve.
+All depth is recorded in centimeters above mean sea level.
 The codes in the context column refer to the excavator's feature labels, cf.

### 2020-04-14 analysis/paper/paper.Rmd
edits on personal ornaments section based on BF's comment

@@
 In addition to `r sam_post_hole` ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts.
-We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.
+We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
 
 ```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
-A: carnelian beads, B: bells, C: glass beads and gold-foil beads, D: metal rings.
+A: carnelian beads, B: bells, C: glass beads and gold-foil beads (far right), D: metal rings.
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
@@
 knitr::kable(ornaments_shape,
-             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
+             caption = "Frequencies of ornaments by subtype at Kiwulan. Frequency represents artifact counts ")
 

### 2020-04-14 analysis/paper/paper.Rmd
edits on Result part 1 based on BF's comment

@@
 The model reveals that ceramic abundance strongly predicts the number of ornaments (β = $`r beta_print`$, p = $`r pval_print`$).
-If ceramic abundance is a suitable proxy for population at Kiwuan due to its basic role as cooking vessels, then ornament quantities per period may be influenced by the number of people living at the site.
+If ceramic abundance is a suitable proxy for population at Kiwulan due to its basic role as cooking vessels, then ornament quantities per period may be influenced by the number of people living at the site.
 
@@
 In contrast, copper bells have less variety typically >2 cm long with a wide variety of human faces as a motif.
-Although glass beads have less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have a wide variety of colors or patterns mostly made by winding technique with high lead content in composition indicating possibly from China [@Cheng2008].
+Although glass beads have less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have a wide variety of colors or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
 Although we are not certain of the specific origin of these beads, research suggest that these glass beads and metal ornaments have similar production techniques and composition to those found in China [@Chen2011; @Wang2018].
 There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component of Kiwulan (1400-1900 AD).
-However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites since 300 BC and declining during the early 2nd millennium [@Wang2018; @Francis2002].
+However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites from 300 BC until the early 2nd millennium [@Wang2018; @Francis2002].
 

### 2020-04-16 analysis/paper/paper.Rmd
edits on result part two based on BF's comment

@@
 
-```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
+```{r plot-all-spatial-pattern, fig.cap= "Spatial densities of all class of ornament by time periods"}
 library(sf)
@@
 
-```{r, plot-spatial-pattern-each, fig.cap="Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
+```{r, plot-spatial-pattern-each, fig.cap="Spatial densities for ornament class by time periods, only those types with more than 5 pieces are shown here"}
 ornaments_period_join_selected_types <- 
@@
 Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from the research area for each time period.
-Before the European arrival, a greater amount of ornaments were found at the northern and middle parts of the research area.
-During the European period, ornaments were more widespread, with some clusters on the northern part.
-During the Chinese period the distribution is more even again.
+For deposits predating the European arrival, a greater amount of ornaments were found at the northern and middle parts of the research area.
+In European period deposits, ornaments were more widespread, with some clusters on the northern part.
+In units dating to the Chinese period the distribution is more even without clear clusters.
 Figure \@ref(fig:plot-spatial-pattern-each) presents the distribution for the major ornament classes individually, some clusters across the area can be observed during the European period, such as gold-foil beads and carnelian beads.
@@
 ```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values from 1000 simulations for three time periods.
-X-axis values represent ANN expected value under a completely random process resulted from simulated pattern.
+X-axis values represent ANN expected value under a completely random process resulted from a simulated pattern.
 Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value"}
@@
 The Chinese period has fewer artifacts in any category, likely reflecting a smaller population at Kiwulan at this time, making spatial patterns and hotspots difficult to discern with confidence.
-This testing reveals clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
+This testing reveals that clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
 

### 2020-04-16 analysis/paper/paper.Rmd
edits on discussion based on BF's comment

@@
 This may be due to a smaller scale of trading networks, the overall decline of indigenous populations in Yilan, or the adoption of Han Chinese practices.
-The decline of the population may be related to the movement of many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants at the end of the 18th century [@Chen2007].
-
-Archaeological contexts show that ornaments are especially abundant in burial contexts serving as grave goods.
+The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
+
+Archaeological contexts show that ornaments are especially abundant in burial contexts serving as grave goods [@Chen2007].
 This supports the interpretation of ornaments as valuable objects functioning as status indicators.
@@
 How might these results fit into a bigger picture of social change at periphery of colonial systems?
-We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Drennan2010; @Feinman2000; @Ames2010; @Bowles2010].
+We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Ames2010; @Bowles2010; @Drennan2010; @Feinman2000;].
 The corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.

### 2020-04-16 analysis/paper/paper.Rmd
edits on conclusion based on BF's comment

@@
 This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
-This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that is part of culture tradition.
+This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of cultural tradition.
 
@@
 We also thank Yu-pei Chen for the discussion regarding the site at the earliest stage of fieldwork.
-This work was supported in part by the Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
+This work was supported in part by the Chester Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
 We would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on the early draft.

### 2020-04-16 analysis/paper/paper.Rmd
edit excavation background based on R1 comment

@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning c.
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1350-1850 AD, 600-100 BP) and a lower component (650-1150 AD, 1200-800 BP) separated by a sterile deposit spanning c.
 150 years [@Chen2007].
@@
 Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011].
-However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artifacts to the pre-European, European, and the Chinese periods.
+Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts relative to the pre-European, European, and the Chinese periods.
+The methods we used to devise the new chronology are described in the supplementary information document.
 
@@
 Smoking is likely to have been introduced by Europeans, and we found that the presence of pipes in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
+Those diagnostic porcelains might have been continually used in later period that could affect our judgment of chronology.
+However, we focus on the high frequently in certain context that can reflect the 17th century they prevailed.
 

### 2020-04-16 analysis/paper/paper.Rmd
revise previous sentence

@@
 Smoking is likely to have been introduced by Europeans, and we found that the presence of pipes in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
-Those diagnostic porcelains might have been continually used in later period that could affect our judgment of chronology.
-However, we focus on the high frequently in certain context that can reflect the 17th century they prevailed.
+Those temporal diagnostic ceramics might have been continually used in later period that could mislead our chronology.
+Thus, we focus on identifying the certain context with high frequency of those ceramics that can reflect the early 17th century when they prevailed.
 

### 2020-04-17 analysis/paper/paper.Rmd
edit personal ornaments section based on R1 comment

@@
 We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
+Ornaments could be potential heirlooms inherited over multiple generations.
+However, there is no continuous increase over time that suggests accumulation and discard of ornaments is not constant.
+We assume that changes in the abundance of ornaments correspond to changes in the amount we see in archaeological record.
 

### 2020-04-18 analysis/paper/paper.Rmd
explain the reason for excluding burials in personal ornament section

@@
   Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
-Trade ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
+Trade ornaments from Kiwulan (1350-1850 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
 The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered.
@@
 6500- 2000BP), and Iron age (c.
-2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages [@Liu2011; @Chen2017].
+2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages [@Chen2017; @Liu2011].
 It is generally accepted that Taiwan entered the historical period in the early 17th century due to the colonial activities of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
@@
 Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173.
-The high number of ornaments in these burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
-We exclude burials from this analysis because most of them date to the pre-European period (n = 26), limiting the usefulness of comparisons between the periods.
-In addition to `r sam_post_hole` ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts.
-We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
+Besides, there are 27 ornaments found in midden contexts.
+The obvious difference in number in burials and midden contexts indicates the value of ornaments as prestige goods.
+The high number in burials is also due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+The unequal distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008.
+However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods.
+To have a comprehensive understanding, we focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
+The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
+Also, it is beyond the scope of this research.
 Ornaments could be potential heirlooms inherited over multiple generations.
@@
 There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component of Kiwulan (1400-1900 AD).
-However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites from 300 BC until the early 2nd millennium [@Wang2018; @Francis2002].
+However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites from 300 BC until the early 2nd millennium [@Francis2002; @Wang2018].
 

### 2020-04-19 analysis/paper/paper.Rmd
edits according to R3 comments

@@
 The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
-We cannot be sure of the exact process that bought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by indigenous groups via regional networks in north Taiwan.
+We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by indigenous groups via regional networks in north Taiwan.
 In any case, the high volume of ceramics transported by Europeans, and their high mobility in shipping trade played an important role in introducing foreign jars to Taiwan.
@@
          `Uncalibrated Age BP` = `Date BP`) %>% 
-  mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
+  mutate(`Depth (cm)` = paste0(Depth1, " to ", Depth2)) %>%
   mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
@@
 Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens.
+For the ornament categories, we follow the well established topology based on raw materials and shapes for the region of northeastern Taiwan [@Chen2007; @NMTH2005].
 This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
@@
 Besides, there are 27 ornaments found in midden contexts.
-The obvious difference in number in burials and midden contexts indicates the value of ornaments as prestige goods.
-The high number in burials is also due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
+The obvious difference in number in burials and midden contexts indicates the property of ornaments as prestige goods.
+The high number in burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
 The unequal distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008.
@@
 The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
-Also, it is beyond the scope of this research.
 Ornaments could be potential heirlooms inherited over multiple generations.
 However, there is no continuous increase over time that suggests accumulation and discard of ornaments is not constant.
-We assume that changes in the abundance of ornaments correspond to changes in the amount we see in archaeological record.
+Thus, we assume that changes in the abundance of ornaments reflect changes in the amount in archaeological record.
+Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
 
 ```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
-A: carnelian beads, B: bells, C: glass beads and gold-foil beads (far right), D: metal rings.
+A: carnelian beads, B: bells, C: glass beads (the far left and the middle) and gold-foil beads (the far right), D: metal rings.
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
@@
   facet_wrap(~ period) +
+  scale_y_continuous(breaks=c(0,2,4,6,8)) +
   theme_minimal(base_size = base_size_multi_plot) 
@@
 Later, trade activities became more frequent and intense in the 17th century due to European activities.
-The greater diversity and quantity of ornaments likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
+The greater diversity and quantity of ornaments likely resulted from participation in large scale exchange networks that stimulated the circulation of different ornament classes.
 The frequency of overall ornaments and each subtype declines significantly after European influence fades during the Chinese period in the early 19th century.
 This may be due to a smaller scale of trading networks, the overall decline of indigenous populations in Yilan, or the adoption of Han Chinese practices.
-The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
+The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
 
@@
 We may be able to interpret Yilan social organization as moving from corporate mode before the European arrival, to a network mode during European presence.
-The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode.
-The long-distance trade network introduced by Europeans resulted in the appearance of a network mode, and the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994].
+The changes from a less concentrated to a high concentrated distribution before and after the presence of Europeans appears consistent with the shift from shared power and wealth to accumulated wealth and monopolization.
+One possible explanation for the shift could be the long-distance trade network introduced by Europeans.
+The rarity and the image of colonial power of foreign trade goods resulted in the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting them [@Boone1992; @Brumfiel1994; @Clark1994].
 Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
@@
 That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations.
-The increasing number and concentrated spatial patterns of ornaments may indicate a practice of cultural resistance against the European intrusion.
+The increasing number and concentrated spatial patterns of ornaments may also indicate a practice of cultural resistance against the European intrusion.
 Resistance to European economic and political demands may be inferred if ornaments were used as a display of social identity and to emphasize the local customs that had existed before European contact [cf.

### 2020-04-19 analysis/paper/paper.Rmd
reorganize the conclusion and shorten the future work

@@
 Thus, we assume that changes in the abundance of ornaments reflect changes in the amount in archaeological record.
-Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
 
@@
 The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
+Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
 
@@
 The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future analysis of pottery production and standardization, and mortuary practices would provide more information.
 By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
@@
 It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
-Future work could extend this approach to sourcing the ornaments directly with geochemical methods, and studies of other trade goods such as ceramics.
-We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardization, and mortuary practices.
 

### 2020-04-21 analysis/paper/paper.Rmd
update the ornament image

@@
 ```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class.
-A: carnelian beads, B: bells, C: glass beads (the far left and the middle) and gold-foil beads (the far right), D: metal rings.
+A: carnelian beads, B: glass beads, C: gold-foil beads, D: bells, E: metal rings.
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D classes were from original excavation report (Chen 2007)", echo=FALSE}
-knitr::include_graphics(here("analysis", "figures", "ornament.png"))
+The photographs of B, C, D, E classes were from original excavation report (Chen 2007)", echo=FALSE}
+knitr::include_graphics(here("analysis", "figures", "ornament_edit.png"))
 ```

### 2020-04-22 analysis/paper/paper.Rmd
add pipe reference and fix some typos

@@
 In addition to stoneware jars as indicators of European presence, around 300 pieces of locally made pipes and a few imported pipes were found at Kiwulan.
-Smoking is likely to have been introduced by Europeans, and we found that the presence of pipes in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
+Smoking is likely to have been introduced by Europeans that was widely adopted in European countries in the 16th century and spread to other regions through interactions [@Uccar2019].
+We found that the presence of pipes in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
 Those temporal diagnostic ceramics might have been continually used in later period that could mislead our chronology.
@@
 Kernel density estimations (KDE) compute the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
-Here we use KDE to visualize core areas of ornaments and surrounding neighbourhoods [@Bonnier2019; @Cortegoso2016].
+Here we use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
 Density values of artifacts per square meter were calculated for each cell.
@@
 Before European contact, ornaments were traded into local indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local indigenous culture according to their distribution in the archaeological contexts.
-After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan.
+After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signaling wealth and privileged trading connections among the inhabitants of Kiwulan.
 This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.

### 2020-04-22 analysis/paper/paper.Rmd
replace indigenous with capital I

@@
 abstract: |
-  Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
+  Long-lasting indirect impacts on Indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
 Trade ornaments from Kiwulan (1350-1850 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
@@
 This hints at increasing social inequality resulting from colonial influence.
-Ornaments give insights into the increasing social inequality stimulated by the European colonial presence, and show the agency of indigenous people to incorporate ornaments into their social system.
+Ornaments give insights into the increasing social inequality stimulated by the European colonial presence, and show the agency of Indigenous people to incorporate ornaments into their social system.
 keywords: |
@@
   3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
-  4. Trade ornament can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
+  4. Trade ornament can be a good proxy to detect indirect colonial influence on local Indigenous populations and inform on social changes relating to trade with colonial groups.
 ---
@@
 
-The direct impacts of European colonialism on indigenous communities in East Asia were much less conspicuous than in island Southeast Asia and Oceania.
-Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local indigenous communities remains largely unanswered.
-Understanding the indirect effects of colonialism are important for detecting colonial impacts on indigenous peoples in the periphery of colonial control [@Trabert2017].
-In many parts of the world, the introduction of foreign trade goods by colonial traders into local indigenous societies caused substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Silliman2005].
-Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
+The direct impacts of European colonialism on Indigenous communities in East Asia were much less conspicuous than in island Southeast Asia and Oceania.
+Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local Indigenous communities remains largely unanswered.
+Understanding the indirect effects of colonialism are important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@Trabert2017].
+In many parts of the world, the introduction of foreign trade goods by colonial traders into local Indigenous societies caused substantial transformations of Indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Silliman2005].
+Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of Indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
 Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in parts of Taiwan, the northeastern region was isolated from intensive direct contact by the Xueshan Mountains.
@@
 This article describes personal ornaments excavated from the upper component of Kiwulan (1400 AD-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
-The first recorded European presence in Yilan was a Spanish revenge attack on indigenous villages in 1632 [@Borao2001, pp.
+The first recorded European presence in Yilan was a Spanish revenge attack on Indigenous villages in 1632 [@Borao2001, pp.
 163].
 In 1647 the Dutch attacked villages and forced them to accept colonial rules and pay an annual tribute [@Andrade2007].
-According to Dutch census reports in 1650, Kiwulan was the largest indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
+According to Dutch census reports in 1650, Kiwulan was the largest Indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
 12].
@@
 In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the following period of Chinese presence.
-We address the question of whether indirect colonial influences on the indigenous populations can be detected through the ornament assemblages.
+We address the question of whether indirect colonial influences on the Indigenous populations can be detected through the ornament assemblages.
 
@@
 Because of natural safe harbors, northeastern Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
-Although located on the periphery of regional trade centers, Yilan was connected to trade networks via visits of other indigenous groups, Chinese merchants, and Europeans, via sea.
+Although located on the periphery of regional trade centers, Yilan was connected to trade networks via visits of other Indigenous groups, Chinese merchants, and Europeans, via sea.
 
 The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Helping dao, Keelung in 1626, and Fort San Domingo in 1629 at Tamsui (Figure \@ref(fig:KWL-map)).
-They sent missionaries to local indigenous settlements in this region [@Blusse2000, pp.
-343] and kept records about their observations of indigenous communities.
-A Dominican priest in 1632 reported that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups.
+They sent missionaries to local Indigenous settlements in this region [@Blusse2000, pp.
+343] and kept records about their observations of Indigenous communities.
+A Dominican priest in 1632 reported that the Taparri, an Indigenous tribe from northern Taiwan, exchanged carnelian beads with other Indigenous groups.
 This form of exchange was widespread and even the Spanish soldiers used carnelian beads as bargaining chips for gambling [@LiandWu2006, pp.
@@
 122-151].
-Records of an indigenous funeral document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp.
+Records of an Indigenous funeral document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp.
 153].
@@
 In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan.
-They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tribute [@Andrade2007, ch.
+They introduced a feudal system in an attempt to control the Indigenous communities by asking Indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tribute [@Andrade2007, ch.
 9; @Kang2016, ch.
 4].
-The Dutch provided beads and other goods based on negotiations with indigenous communities to secure alliances in the annual ceremony or during their travels [@Kang2016, ch.
+The Dutch provided beads and other goods based on negotiations with Indigenous communities to secure alliances in the annual ceremony or during their travels [@Kang2016, ch.
 6].
-We might predict that the activities of the Dutch feudal system to build and maintain alliances resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.
+We might predict that the activities of the Dutch feudal system to build and maintain alliances resulted in an increase in the amount and diversity of ornaments in northeastern Indigenous communities during this period.
 
@@
 77].
-According to those records, indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
+According to those records, Indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
 Among those ornaments, fish-shaped necklaces made of metal threads had high value due to their delicacy and the exotic materials invested in production.
@@
 Other people wore carnelian beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonization, an academic field survey for plains indigenous groups reported that fish-shaped metal necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of Japanese colonization, an academic field survey for plains Indigenous groups reported that fish-shaped metal necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
-Although these historical records are fragmentary and may contain some biases [@Galloway2006] that have not yet been studied in detail, we find consistency among multiple sources in their descriptions of how ornaments represent high status or specialized social roles in indigenous communities in Yilan.
+Although these historical records are fragmentary and may contain some biases [@Galloway2006] that have not yet been studied in detail, we find consistency among multiple sources in their descriptions of how ornaments represent high status or specialized social roles in Indigenous communities in Yilan.
 Compared to the European period, there are fewer documentary mentions of beads in the Chinese period and the descriptions are limited to clothing, but these generally confirm the role of beads as status markers.
@@
 Both archaeological evidence and historical records indicate northeastern Taiwan was involved in regional networks with East Asia in the late Iron age.
-These included Chinese merchants trading metal items, clothes, and beads with local indigenous people in Taiwan in exchange for local resources.
+These included Chinese merchants trading metal items, clothes, and beads with local Indigenous people in Taiwan in exchange for local resources.
 The foreign-made large dark brown glazed stoneware jars frequently found in European shipwrecks were also commonly found from many sites in Taiwan, suggesting direct or indirect interactions.
@@
 The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
-We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by indigenous groups via regional networks in north Taiwan.
+We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by Indigenous groups via regional networks in north Taiwan.
 In any case, the high volume of ceramics transported by Europeans, and their high mobility in shipping trade played an important role in introducing foreign jars to Taiwan.
@@
 An indirect colonial influence may be indicated at Kiwulan by the greater diversity of ornament types and materials during the European period.
-Yilan was involved in complex trading networks both on a regional scale with other indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish.
+Yilan was involved in complex trading networks both on a regional scale with other Indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish.
 The ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century.
@@
 The frequency of overall ornaments and each subtype declines significantly after European influence fades during the Chinese period in the early 19th century.
-This may be due to a smaller scale of trading networks, the overall decline of indigenous populations in Yilan, or the adoption of Han Chinese practices.
-The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
+This may be due to a smaller scale of trading networks, the overall decline of Indigenous populations in Yilan, or the adoption of Han Chinese practices.
+The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
 Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
@@
 
-Examination of the archaeological record at the peripheries of colonial activities can show how remote indigenous groups were affected by major European colonial processes [@Trabert2017].
+Examination of the archaeological record at the peripheries of colonial activities can show how remote Indigenous groups were affected by major European colonial processes [@Trabert2017].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
@@
 The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context.
-Before European contact, ornaments were traded into local indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local indigenous culture according to their distribution in the archaeological contexts.
+Before European contact, ornaments were traded into local Indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local Indigenous culture according to their distribution in the archaeological contexts.
 After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signaling wealth and privileged trading connections among the inhabitants of Kiwulan.
@@
 
-We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities.
+We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.
 The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods.
 We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future analysis of pottery production and standardization, and mortuary practices would provide more information.
-By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations.
+By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local Indigenous populations.
 Ornaments give insights into the amplification of social inequality stimulated by European colonization.
-It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
+It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
 

### 2020-04-22 analysis/paper/paper.Rmd
shorten the future work

@@
 We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.
-The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods.
-We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future analysis of pottery production and standardization, and mortuary practices would provide more information.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments, and mortuary practices would provide more information to construct a clear picture of complex trade networks during this periods.
 By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local Indigenous populations.

### 2020-04-23 analysis/paper/paper.Rmd
correct the AD age

@@
 
-This article describes personal ornaments excavated from the upper component of Kiwulan (1400 AD-1900 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
+This article describes personal ornaments excavated from the upper component of Kiwulan (1350 AD-1850 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
 The first recorded European presence in Yilan was a Spanish revenge attack on Indigenous villages in 1632 [@Borao2001, pp.
@@
 Although we are not certain of the specific origin of these beads, research suggest that these glass beads and metal ornaments have similar production techniques and composition to those found in China [@Chen2011; @Wang2018].
-There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component of Kiwulan (1400-1900 AD).
-However, the glass beads found from the lower component (700-1200 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites from 300 BC until the early 2nd millennium [@Francis2002; @Wang2018].
+There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component of Kiwulan (1350-1850 AD).
+However, the glass beads found from the lower component (650-1150 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites from 300 BC until the early 2nd millennium [@Francis2002; @Wang2018].
 

### 2020-04-24 analysis/paper/paper.Rmd
add suggested references from R2

@@
 
-Examination of the archaeological record at the peripheries of colonial activities can show how remote Indigenous groups were affected by major European colonial processes [@Trabert2017].
+Examination of the archaeological record at the peripheries of colonial activities can show how remote Indigenous groups were affected by major European colonial processes [@Trabert2017], and similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Jane2015].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.

### 2020-04-26 analysis/paper/paper.Rmd
small edits on personal ornament, discussion and conclusion section

@@
 The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
-Ornaments could be potential heirlooms inherited over multiple generations.
+Ornaments could be potential heirlooms inherited over multiple generations and preserved well for a long time.
 However, there is no continuous increase over time that suggests accumulation and discard of ornaments is not constant.
@@
 Yilan was involved in complex trading networks both on a regional scale with other Indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish.
-The ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century.
+The ornaments are found at many Iron Age sites in Yilan but only Kiwulan shows clear stratigraphic contexts from pre-European period to modern time.
+Those trade ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century.
 Later, trade activities became more frequent and intense in the 17th century due to European activities.
@@
 
-We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.
-We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments, and mortuary practices would provide more information to construct a clear picture of complex trade networks during this periods.
 By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local Indigenous populations.
@@
 It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
+We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments such as X-ray fluorescence analysis would provide more information to construct a clear picture of complex trade networks during this periods.
 

### 2020-04-28 analysis/paper/paper.Rmd
add sentences about spatial analysis in point pattern section

@@
 
+The distribution and density of prestige goods across the residential area can provide information of social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes by statistical methods [@Kintigh1982].
 Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
@@
 The next step was to divide the ornaments into three time periods.
-Finally we computed the kernel densities for each time period for comparison.
-Kernel density estimations (KDE) compute the probability of the density of ornaments across space by creating a continuous, smooth density surface across space.
-Here we use KDE to visualize core areas of ornaments and surrounding neighborhoods [@Bonnier2019; @Cortegoso2016].
+Finally we computed the density maps for each time period for comparison.
 Density values of artifacts per square meter were calculated for each cell.
+Here we use Kernel density estimations (KDE) for visualization and identification of spatial clusters [@Baxter1997], the core areas of ornaments and surrounding neighborhoods in this case.
+KDE is a approach of spatial analysis that computes the probability of the density of ornaments across space by creating a continuous, smooth density surface across space [@Bonnier2019; @Cortegoso2016].
 

### 2020-04-30 analysis/paper/paper.Rmd
add references and a sentence about social status
update the chronology para in excavation section and add two more references

@@
 Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local Indigenous communities remains largely unanswered.
-Understanding the indirect effects of colonialism are important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@Trabert2017].
+Understanding the indirect effects of colonialism are important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@Acabado2017; @Trabert2017].
 In many parts of the world, the introduction of foreign trade goods by colonial traders into local Indigenous societies caused substantial transformations of Indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Silliman2005].
@@
 The sterility may be related to dry weather leading to site abandonment according to pollen analysis [@Lin2015].
-These component divisions are based on the differences in the color and texture of the deposit.
+These component divisions are based on the differences in color and texture of the deposit and the content of artifacts.
 Whether these two components represent a culture continuity from the same people is still under debate [@Chiu2004; @Chen2007].
@@
 Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts relative to the pre-European, European, and the Chinese periods.
-The methods we used to devise the new chronology are described in the supplementary information document.
+The methods we used to devise the new chronology are described in the Supplementary Information document.
 
@@
 Large dark brown glazed stoneware jars may have been made in Southeast Asia, and are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
-The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Hsieh2009; @Hsieh1995].
+The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Dizon2016; @Hsieh1995].
 Southeast China is assumed to be the origin of the light gray glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, such as the Zeelandia fort site in Tainan [@Wang2007].
@@
 In any case, the high volume of ceramics transported by Europeans, and their high mobility in shipping trade played an important role in introducing foreign jars to Taiwan.
+
 Those jars were widely distributed across the site and can serve as indicators, together with the radiocarbon dates, to identify the excavation units associated with the pre-European period and the start of European influence at Kiwulan.
-In addition to stoneware jars as indicators of European presence, around 300 pieces of locally made pipes and a few imported pipes were found at Kiwulan.
-Smoking is likely to have been introduced by Europeans that was widely adopted in European countries in the 16th century and spread to other regions through interactions [@Uccar2019].
-We found that the presence of pipes in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
-Those temporal diagnostic ceramics might have been continually used in later period that could mislead our chronology.
-Thus, we focus on identifying the certain context with high frequency of those ceramics that can reflect the early 17th century when they prevailed.
-
-The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups.
+In addition to stoneware jars as indicators of European presence, around 300 pieces of locally made clay pipes and a few imported pipes were found at Kiwulan.
+Smoking is likely to have been introduced by Europeans.
+This custom was widely adopted in many European countries in the 16th century and spread to other regions through interactions [@Uccar2019].
+We found that the presence of pipe bowls in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
+It should be noted that the temporal indicators might have been introduced before the European arrival by Chinese traders that could mislead the interpretation of the start of European phase.
+However, the archaeological evidence shows that the layers with abundant trade ceramics match the 17th century according to excavation report and previous studies [@Hsieh2009; @Wang2011].
+Thus, we focus on identifying the context with high frequency of those ceramics indicating the early 17th century when they prevailed.
+
+The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many styles and forms such as bowls, plates, and cups.
 Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles used by Chinese [@Hsieh2009].
 Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrants in 1768 [@Chen1963; @Ke1993].
+The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Information data.
 
@@
 The distribution and density of prestige goods across the residential area can provide information of social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes by statistical methods [@Kintigh1982].
+The distribution pattern of artifacts can reveal social differences among the societies in many archaeological contexts [@Orser1988; @Pearson1993; @Wason2004].
 Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.

### 2020-04-30 analysis/paper/paper.Rmd
update analysis of ornament distribution section with two more references

@@
 The distribution and density of prestige goods across the residential area can provide information of social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes by statistical methods [@Kintigh1982].
-The distribution pattern of artifacts can reveal social differences among the societies in many archaeological contexts [@Orser1988; @Pearson1993; @Wason2004].
+The differential accumulation of artifacts especially high value goods can reflect social differentiation or hierarchy in a society in many archaeological contexts [@Halstead1993; @Orser1988; @Pearson1993; @Trubitt2003; @Wason2004].
 Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.

### 2020-05-14 analysis/paper/paper.Rmd
editing for re-submission after peer reviews

--- 2020-04-30 analysis/paper/paper.Rmd
@@
 This article describes personal ornaments excavated from the upper component of Kiwulan (1350 AD-1850 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
+Ornaments are found at many Iron Age sites in Yilan, but only Kiwulan shows clear stratigraphic contexts from pre-European period to modern time.
 The first recorded European presence in Yilan was a Spanish revenge attack on Indigenous villages in 1632 [@Borao2001, pp.
 163].
-In 1647 the Dutch attacked villages and forced them to accept colonial rules and pay an annual tribute [@Andrade2007].
+In 1647 the Dutch attacked villages and forced them to accept colonial rule and pay an annual tribute [@Andrade2007].
 According to Dutch census reports in 1650, Kiwulan was the largest Indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
@@
 The island of Taiwan lies at the junction of mainland China, Southeast Asia, and Northeast Asia in the Pacific Ocean.
-The prehistory of Taiwan island could be roughly divided into three major periods, palaeolithic (c.
+The prehistory of Taiwan island could be roughly divided into three major periods, Palaeolithic (c.
 27,000 BP- 5000BP), Neolithic (c.
 6500- 2000BP), and Iron age (c.
-2000- 400BP) with slightly regional differences in onset of each period and variations in style of artifacts and assemblages [@Chen2017; @Liu2011].
+2000- 400BP) with slight regional differences in onset of each period and variations in style of artifacts and assemblages [@Chen2017; @Liu2011].
 It is generally accepted that Taiwan entered the historical period in the early 17th century due to the colonial activities of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
@@
 
-The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Helping dao, Keelung in 1626, and Fort San Domingo in 1629 at Tamsui (Figure \@ref(fig:KWL-map)).
+The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Heping Dao, Keelung in 1626, and Fort San Domingo in 1629 at Tamsui (Figure \@ref(fig:KWL-map)).
 They sent missionaries to local Indigenous settlements in this region [@Blusse2000, pp.
@@
 132-149].
-Other records mention that the female shamans in the tribe would use carnelian beads as magical items in ritual healing practices [@Borao2009, pp.
+Other records mention that the women shamans in the tribe would use carnelian beads as magical items in ritual healing practices [@Borao2009, pp.
 122-151].
-Records of an Indigenous funeral document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp.
+Records of an Indigenous funeral describe the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp.
 153].
@@
 
-Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that was conducted between 2001-2004 in advance of a water diversion project and road bridge construction.
+Archaeological ornaments from Kiwulan (Figure \@ref(fig:KWL-map)) come from a rescue archaeology project that was conducted between 2001-2004 in advance of a water diversion project and road bridge construction.
 The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
@@
 Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
-Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from (Figure \@ref(fig:KWL-sam-area)).
-In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures.
+Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of the A and D sections, which is also the study area where our samples come from (Figure \@ref(fig:KWL-sam-area)).
+In the AD area, post-holes were found aligned in a north-south direction in intervals with construction marks, which have been interpreted as the remains of stilt house structures.
 At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction.
@@
 
-```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number.
-Small dots represent the location of post-holes.
+```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID numbers.
+Small dots represent the locations of post-holes.
 Each square is 4 x 4 m", fig.width=image_width_2_col_inch}
@@
 150 years [@Chen2007].
-The sterility may be related to dry weather leading to site abandonment according to pollen analysis [@Lin2015].
+The sterility may be related to dry weather, according to pollen analysis, leading to site abandonment [@Lin2015].
 These component divisions are based on the differences in color and texture of the deposit and the content of artifacts.
@@
 Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011].
-Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts relative to the pre-European, European, and the Chinese periods.
-The methods we used to devise the new chronology are described in the Supplementary Information document.
+Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods.
+The methods we used to devise the new chronology are described in the Supplementary Online Materials.
 
 The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
-Large dark brown glazed stoneware jars may have been made in Southeast Asia, and are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
+Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
 The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Dizon2016; @Hsieh1995].
@@
 We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by Indigenous groups via regional networks in north Taiwan.
-In any case, the high volume of ceramics transported by Europeans, and their high mobility in shipping trade played an important role in introducing foreign jars to Taiwan.
+In any case, the high volume of ceramics transported by Europeans, and their high mobility in the shipping trade played an important role in introducing foreign jars to Taiwan.
 
@@
 We found that the presence of pipe bowls in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
-It should be noted that the temporal indicators might have been introduced before the European arrival by Chinese traders that could mislead the interpretation of the start of European phase.
-However, the archaeological evidence shows that the layers with abundant trade ceramics match the 17th century according to excavation report and previous studies [@Hsieh2009; @Wang2011].
-Thus, we focus on identifying the context with high frequency of those ceramics indicating the early 17th century when they prevailed.
+It should be noted that these temporal indicators might have been introduced before direct European contact by Chinese traders, and this could result in some uncertainty in identifying the start of European phase.
+However, the archaeological evidence shows that the layers with abundant trade ceramics match the 17th century according to the excavation report and previous studies [@Hsieh2009; @Wang2011].
+Thus, we focus on identifying the contexts with high frequencies of those ceramics as indicators the early 17th century.
 
@@
 Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrants in 1768 [@Chen1963; @Ke1993].
-The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Information data.
+The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Online Materials.
 
@@
 Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens.
-For the ornament categories, we follow the well established topology based on raw materials and shapes for the region of northeastern Taiwan [@Chen2007; @NMTH2005].
-This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
+For the ornament categories, we follow the well-established topology based on raw materials and shapes for the region of northeastern Taiwan [@Chen2007; @NMTH2005].
+This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the main habitation areas of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
 Occupation floors were not identified during excavation.
@@
 Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173.
-Besides, there are 27 ornaments found in midden contexts.
+There are also 27 ornaments found in midden contexts.
 The obvious difference in number in burials and midden contexts indicates the property of ornaments as prestige goods.
-The high number in burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007].
-The unequal distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008.
-However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods.
-To have a comprehensive understanding, we focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
+The high number in burials is due to the presence of bead strands or patterned bands of beads, which sometimes contain thousands of beads in an individual burial [@Chen2007].
+The highly uneven distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008.
+However, @Hsieh2012 suggests a relatively egalitarian society from the comparative analysis of frequency and proportion of burial goods.
+
+We focus on ornaments from the habitation contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
 The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
-Ornaments could be potential heirlooms inherited over multiple generations and preserved well for a long time.
-However, there is no continuous increase over time that suggests accumulation and discard of ornaments is not constant.
-Thus, we assume that changes in the abundance of ornaments reflect changes in the amount in archaeological record.
+A possible limitation to our chronological resolution is that ornaments could be heirlooms inherited over multiple generations and well-preserved for a long time.
+This is difficult to rule out completely, but we consider that because there is no continuous increase in ornament frequency over time, we conclude that accumulation and discard of ornaments is not constant.
+Thus, we assume that changes in the abundance of ornaments reflect relatively continuous discard behaviours rather than accumulations due to collecting of heirlooms.
 
@@
 Photographs are presented in the same order as those subtypes in the table but from left to right instead.
-The photographs of B, C, D, E classes were from original excavation report (Chen 2007)", echo=FALSE}
+The photographs of B, C, D, E classes are from the original excavation report (Chen 2007)", echo=FALSE}
 knitr::include_graphics(here("analysis", "figures", "ornament_edit.png"))
@@
 
-To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in http://doi.org/10.17605/OSF.IO/R8YGA.
+To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at http://doi.org/10.17605/OSF.IO/R8YGA.
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
@@
 
-```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods"}
+```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament types across different time periods."}
 ### delete one plot
@@
 However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period.
-To model the number of ornaments as a function of the mass of ceramics in each period, a Poisson GLM with a log link function was used.
+To model the number of ornaments as a function of the mass of ceramics in each period, we used a Poisson GLM with a log link function.
 The model reveals that ceramic abundance strongly predicts the number of ornaments (β = $`r beta_print`$, p = $`r pval_print`$).
@@
 
-```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads"}
+```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across time periods for metal rings, carnelian beads, bells, and glass beads."}
 library(cowplot)
@@
 
-The distribution of frequencies for subtypes in each major class is presented in Figure \@ref(fig:plot-altogether).
+Figure \@ref(fig:plot-altogether) shows the distribution of frequencies for subtypes in each major class.
 Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
@@
 The greater varieties for carnelian beads and metal rings might indicate multiple origins due to participation in large scale trade networks stimulated by the European presence.
-In contrast, copper bells have less variety typically >2 cm long with a wide variety of human faces as a motif.
-Although glass beads have less variety in size, small (0.5-1 cm) and medium (1-2 cm), they have a wide variety of colors or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
+In contrast, copper bells have less variety, typically >2 cm long with a wide variety of human faces as a motif.
+Although glass beads have less variety in size, presenting as small (0.5-1 cm) or medium (1-2 cm), they have a wide variety of colors or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
 Although we are not certain of the specific origin of these beads, research suggest that these glass beads and metal ornaments have similar production techniques and composition to those found in China [@Chen2011; @Wang2018].
@@
 
-```{r plot-all-spatial-pattern, fig.cap= "Spatial densities of all class of ornament by time periods"}
+```{r plot-all-spatial-pattern, fig.cap= "Spatial densities of all class of ornament by time periods."}
 library(sf)
@@
 
-```{r, plot-spatial-pattern-each, fig.cap="Spatial densities for ornament class by time periods, only those types with more than 5 pieces are shown here"}
+```{r, plot-spatial-pattern-each, fig.cap="Spatial densities for ornament class by time periods, only those types with more than 5 pieces are shown here."}
 ornaments_period_join_selected_types <- 
@@
 
-```{r plot-all-densities, fig.cap= "Kernel density map for ornaments by periods. Used the bandwidth based on Silverman (1986)'s rule of thumb"}
+```{r plot-all-densities, fig.cap= "Kernel density map for ornaments by periods, using a bandwidth based on Silverman (1986)'s rule of thumb"}
 # apply function to each data frame in the list
@@
 
-The distribution and density of prestige goods across the residential area can provide information of social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes by statistical methods [@Kintigh1982].
-The distribution pattern of artifacts can reveal social differences among the societies in many archaeological contexts [@Orser1988; @Pearson1993; @Wason2004].
-Point pattern analysis can assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
-To prepare the ornament location data for point pattern analysis, because artifacts from Kiwulan lack exact piece-provenance data, we assigned each ornament to a random coordinate pair in the square it was recovered from.
+The distribution and density of prestige goods across the residential area can provide information on social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes [@Kintigh1982].
+The differential accumulation of artifacts, especially high value goods, in many archaeological contexts can reflect social differentiation or hierarchy in a society  [@Halstead1993; @Orser1988; @Pearson1993; @Trubitt2003; @Wason2004].
+We used point pattern analysis to assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
+To prepare the ornament location data for point pattern analysis, we assigned each ornament to a random coordinate pair in the square it was recovered from because artifacts from Kiwulan lack exact piece-provenance data.
 The next step was to divide the ornaments into three time periods.
@@
 Density values of artifacts per square meter were calculated for each cell.
-Here we use Kernel density estimations (KDE) for visualization and identification of spatial clusters [@Baxter1997], the core areas of ornaments and surrounding neighborhoods in this case.
-KDE is a approach of spatial analysis that computes the probability of the density of ornaments across space by creating a continuous, smooth density surface across space [@Bonnier2019; @Cortegoso2016].
+Here we use kernel density estimation (KDE) for visualization and identification of spatial clusters [@Baxter1997], in this case the core areas of ornaments and surrounding neighborhoods.
+KDE is a method of spatial analysis that computes the probability of the density of ornaments across space by creating a continuous, smooth density surface across space [@Bonnier2019; @Cortegoso2016].
 
@@
 There are three consistent sub-regions with a core area that shifts over time.
-The distribution might indicate the increase and decrease in the number of social groups who possessed ornaments.
+The distribution might indicate an increase and decrease in the number of social groups who possessed ornaments.
 The multiple groups during the European period might reflect unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from a bigger sample size.
 In addition, the generation of core areas might be biased due to small sample sizes, for example, a few ornaments found at one single square during the Chinese period could create an obvious hotspot.
-Whether the observed clustering is random or non-random is crucial for the interpretation of intentional human activities.
+Whether the observed clustering is random or non-random is crucial for making reliable interpretations of intentional human activities at Kiwulan.
 
@@
 ```{r hypo-test-distribution, fig.cap= "Histograms of simulated ANN values from 1000 simulations for three time periods.
-X-axis values represent ANN expected value under a completely random process resulted from a simulated pattern.
+X-axis values represent ANN expected values under a completely random process resulting from a simulated pattern.
 Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value"}
@@
 The Chinese period has fewer artifacts in any category, likely reflecting a smaller population at Kiwulan at this time, making spatial patterns and hotspots difficult to discern with confidence.
-This testing reveals that clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
+Our Monte Carlo testing reveals that clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
 
@@
 Yilan was involved in complex trading networks both on a regional scale with other Indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish.
-The ornaments are found at many Iron Age sites in Yilan but only Kiwulan shows clear stratigraphic contexts from pre-European period to modern time.
 Those trade ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century.
@@
 This may be due to a smaller scale of trading networks, the overall decline of Indigenous populations in Yilan, or the adoption of Han Chinese practices.
-The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
+The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to the increasing numbers of Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
 Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
@@
 How might these results fit into a bigger picture of social change at periphery of colonial systems?
-We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Ames2010; @Bowles2010; @Drennan2010; @Feinman2000;].
-The corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power.
+We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Ames2010; @Bowles2010; @Drennan2010; @Feinman2000].
+The corporate/network model proposed expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power [@Feinman2000].
 In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks.
 In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
-We may be able to interpret Yilan social organization as moving from corporate mode before the European arrival, to a network mode during European presence.
-The changes from a less concentrated to a high concentrated distribution before and after the presence of Europeans appears consistent with the shift from shared power and wealth to accumulated wealth and monopolization.
+
+The Kiwulan ornament data may be interpreted as indicating that Yilan social organization moved from a corporate mode, before the European arrival, to a network mode during European presence.
+The changes from a less concentrated to a more concentrated distribution of ornaments before and after the presence of Europeans appears consistent with the shift from shared power and wealth to accumulated wealth and monopolization.
 One possible explanation for the shift could be the long-distance trade network introduced by Europeans.
@@
 @Rubertone2000].
-Resistance could be presented in many forms.
+Resistance could be presented in many forms, but we have no specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
 Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time.
-This process would result in a natural increase in ornaments over time, unrelated to colonial influences, but has limited value in explaining the shifts in spatial patterns.
+This process would result in a natural increase in ornaments over time, unrelated to colonial influences.
+As noted above, this pattern is not a good fit for the ornament distribution at Kiwulan, and adds little value in explaining the shifts in spatial patterns.
 
@@
 
-Examination of the archaeological record at the peripheries of colonial activities can show how remote Indigenous groups were affected by major European colonial processes [@Trabert2017], and similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Jane2015].
+Examination of the archaeological record at the peripheries of colonial activity shows how remote Indigenous groups were affected by major European colonial processes [@Trabert2017], and similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Jane2015].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
-Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation.
+Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the reach of direct European colonial occupation.
 The frequency and spatial distribution of personal ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
 The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context.
-Before European contact, ornaments were traded into local Indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local Indigenous culture according to their distribution in the archaeological contexts.
+Before European contact, ornaments were traded into local Indigenous societies via the regional exchange network with Chinese merchants, and viewed as prestige goods in the local Indigenous culture.
 After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signaling wealth and privileged trading connections among the inhabitants of Kiwulan.
 This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
-This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of cultural tradition.
+This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of cultural tradition, but additional evidence is required to confirm this.
 
@@
 This work was supported in part by the Chester Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
-We would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on the early draft.
-
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-
-##### pagebreak
+We would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.
 
@@
 
-##### pagebreak
-
 ### Colophon

### 2020-05-15 analysis/paper/paper.Rmd
fixed one missing in-text citation
revision submitted to ARA

@@
        dpi = 300,
-       height = 60,
+       height = 62,
        units = "mm")
@@
 This work was supported in part by the Chester Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
-We would also like to thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.
+We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.
+We also acknowledge comments from anonymous reviewers that have greatly improved this manuscript.
 

### 2020-07-01 analysis/paper/paper.Rmd
corrected references and added one para for chronology

@@
 The consumption of stone beads in Southeast Asia during Iron Age is often associated with increasing social stratification or socio-political complexity [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000; @Kenoyer2000].
-In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, the period of Spanish and Dutch presence, and the following period of Chinese presence.
+In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, period after the Spanish and Dutch presence, and the following period after the Chinese presence.
 We address the question of whether indirect colonial influences on the Indigenous populations can be detected through the ornament assemblages.
@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1350-1850 AD, 600-100 BP) and a lower component (650-1150 AD, 1200-800 BP) separated by a sterile deposit spanning c.
-150 years [@Chen2007].
-The sterility may be related to dry weather, according to pollen analysis, leading to site abandonment [@Lin2015].
-These component divisions are based on the differences in color and texture of the deposit and the content of artifacts.
-Whether these two components represent a culture continuity from the same people is still under debate [@Chiu2004; @Chen2007].
-There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
-We focus on the upper component because only this component spans the periods of pre-European contact, European presence, and Chinese presence.
-In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases.
-Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011].
+## Chronology
+
+The chronology of Kiwulan can be divided into two phases represented by a upper component (1350-1850 AD, 600-100 BP) and a lower component (650-1150 AD, 1200-800 BP) based on 32 radiocarbon ages, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
+There is a sterile deposit between two components spanning c.
+150 years that may be related to dry weather, according to pollen analysis, leading to site abandonment [@Chen2007; @Lin2015].
+These component divisions present the differences in color and texture of the deposit, and the content of artifacts such as pottery types.
+Whether these two components suggest a culture continuity from the same people is still under debate [@Chiu2004; @Chen2007].
+We focus on the upper component because this component spans the periods when local residents interacted with the Europeans and the Chinese.
+In our study area, all excavation squares demonstrate signs of continuous human occupation throughout the upper component.
+Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the types of chronologically diagnostic ceramics, excavation depth, consistency of contexts, and radiocarbon dates [@Hsieh2009; @Wang2011].
+To explore the changes in ornament assemblages over time associated with foreign impacts, we assigned the original excavation levels into three phases: before the European contact, after the European presence, and the presence of large Chinese immigrants.
+
+Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade network could go beyond the temporal boundaries and last longer.
+This is because the use of foreign goods in many Indigenous societies often go through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012].
+Also, we acknowledged that a fine-grained identification of decades of the European presences from the archaeological record is challenging in our research based on some reasons.
+The current radiocarbon ages that are unable to provide a high resolution of time scale, the circulation of foreign goods might start earlier before the European presences, and the use of the foreign goods in Indigenous societies could last longer after the official end of European colonization [@Mitchell2010].
+It is more suitable to use pericolonial approaches that consider a long-term indirect consequences of foreign impacts rather than a focus on specific time frame [@Trabert2018; @Scheiber2012].
+
 Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods.
@@
 Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
-
 That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations.
@@
 
-Examination of the archaeological record at the peripheries of colonial activity shows how remote Indigenous groups were affected by major European colonial processes [@Trabert2017], and similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Jane2015].
+Micro-scale analysis of the archaeological record at the peripheries of colonial activity offers a opportunity to understand pervasive but not necessarily dominant European colonial processes at remote Indigenous groups [@Trabert2018].
+Similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Zheng2015].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.

### 2020-07-02 analysis/paper/paper.Rmd
edits based on R2 comments

@@
 The current radiocarbon ages that are unable to provide a high resolution of time scale, the circulation of foreign goods might start earlier before the European presences, and the use of the foreign goods in Indigenous societies could last longer after the official end of European colonization [@Mitchell2010].
-It is more suitable to use pericolonial approaches that consider a long-term indirect consequences of foreign impacts rather than a focus on specific time frame [@Trabert2018; @Scheiber2012].
+It is more suitable to use pericolonial approaches that consider indirect and wide-ranging effects of the colonial process such as trading activities [@Trabert2018; @Scheiber2012].
 
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts totalizaing three periods", fig.width=image_width_1_col_inch}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 A further indicator of increased social inequality is a burial dated to the 17th century that included 60 gold-foil beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008].
+@Cheng2008 proposed a more hierarchical structure of Kiwulan based on the found of burial with 60 gold-foil beads that were treated as prestige goods in local community.
+However, @Hsieh's research on data from burial contexts suggested a more egalitarian society since burials with rich goods are elders that indicates accumulated wealth over time instead of inherited.
+Our results focusing on the residential area demonstrated an uneven distribution of prestige ornaments during the European presence that inclined an appearance of social inequality.
+Nevertheless, we still need a comprehensive exploration of other archaeological records to discuss the social structure of Kiwulan.
 
@@
 Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
+
 That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations.

### 2020-07-03 analysis/paper/paper.Rmd
edits based on R2 comments

@@
 Although glass beads have less variety in size, presenting as small (0.5-1 cm) or medium (1-2 cm), they have a wide variety of colors or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
-Although we are not certain of the specific origin of these beads, research suggest that these glass beads and metal ornaments have similar production techniques and composition to those found in China [@Chen2011; @Wang2018].
-There seem to be no obvious changes in the sources of glass beads or metal ornaments at different periods in the upper component of Kiwulan (1350-1850 AD).
-However, the glass beads found from the lower component (650-1150 AD) are mostly Indo-Pacific beads, widespread in Southeast Asian sites from 300 BC until the early 2nd millennium [@Francis2002; @Wang2018].
+Although we are not certain of the specific origin of all kinds of beads, research suggest that glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
+There seem to be no obvious changes in the sources of glass beads or metal ornaments at different phases in the upper component of Kiwulan (1350-1850 AD).
+However, the glass beads from the lower component (650-1150 AD) demonstrate the composition of m-Na-Al glass and soda plant ash glass that are subtypes of the Indo-Pacific beads, frequently seen in Southeast Asia[@Carter2016; @Francis2002; @Wang2014].
+A transition in the origin of glass beads from mainly Southeast Asia to multiple regions including China might indicate different trading networks.
+A regional network in South China Sea in the earlier period and later involved in larger scale trading network stimulated by the Europeans and the Chinese [@Hung2016; @Wang2007].
 
@@
 plot_spatial_distribution_all_ornaments_by_period <- 
+  ornaments_period_join %>% 
+  group_by(period, Pit_No, geometry) %>% 
+  summarize(n = sum(n)) %>% 
   ggplot() +
   geom_sf(data = sample_units, fill = NA) +
-  geom_sf(data = ornaments_period_join, 
+  geom_sf(data = ornaments_period_join_all_type_across_periods, 
           aes(geometry = geometry, 

### 2020-07-06 analysis/paper/paper.Rmd
edit chronology section based on R1 comments

@@
 This is because the use of foreign goods in many Indigenous societies often go through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012].
-Also, we acknowledged that a fine-grained identification of decades of the European presences from the archaeological record is challenging in our research based on some reasons.
-The current radiocarbon ages that are unable to provide a high resolution of time scale, the circulation of foreign goods might start earlier before the European presences, and the use of the foreign goods in Indigenous societies could last longer after the official end of European colonization [@Mitchell2010].
-It is more suitable to use pericolonial approaches that consider indirect and wide-ranging effects of the colonial process such as trading activities [@Trabert2018; @Scheiber2012].
-
-Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods.
-The methods we used to devise the new chronology are described in the Supplementary Online Materials.
-
-The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
+Also, we acknowledged that a fine-grained identification of decades of the European presence from the archaeological record is challenging for this study based on three reasons.
+The current radiocarbon ages that are unable to provide a high resolution of time breaks, the circulation of foreign goods might start earlier before the official European presence in Taiwan, and the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010].
+However, those limitations can be minimized with pericolonial approaches that emphasize the wide-ranging effects of the colonial process, allowing discussion of indirect impacts on Indigenous societies with a more practical and dynamic framework not necessarily defined by a fixed colonial period with a start and an end time [@Trabert2018; @Scheiber2012].
+
+We acknowledged the approach to study a continuum of changes of archaeological records over time to avoid the pitfall of temporal divides [@Scheiber2012]; however, we also admitted the convenience of using separated time periods for quantitative comparison between temporal groups of imported ornaments.
+To mitigate the conflict, we separated the ornaments based on the archaeological context that can reflect the acceptance of diagnostic imported goods by their relatively high frequencies as an indicator of indirect colonial impacts.
+That is, we separated the time periods to before and after a certain foreign presence using the complete archaeological context that contains higher frequencies of time indicators instead of their first appearance in a given layer.
+This method is reasonable because it allows not only the flexibility of imported goods entering Indigenous societies depending on the agency of local people, but also the application of quantitative analysis to temporal and spatial changes of imported ornaments.
+
+Based on the copies of original excavation records, we have made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations.
+First, we determined the context with indirect colonial impacts including the Europeans and the Chinese according to the higher frequency of time indicators.
+Second, we examined the description of color and texture changes in strategrphic data to identify the end of the context.
+Third, we checked whether there are radiocarbon ages from the context and if there are any conflicts or ambiguities to decide the usefulness of the ages.
+Fourth, we evaluated if there are any post-depositional issues to exclude problematic data.
+Fifth, we cross evaluated all fourty squares to make sure a consistent stratigraphic context across them.
+Sixth, we cross evaluated our chronology by checking previous studies of assigning chronology to make a final decision [@Hsieh2009; @Wang2011].
+We also provide Supplementary Online Materials that include the details of stratigraphic data, the high frequency of time indicators, and the decision making during the assigning process.
+
+The archaeological indicators of the indirect impact of the Europeans at Kiwulan are the higher frequency of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
 Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.

### 2020-07-07 analysis/paper/paper.Rmd
edit the chronology section

@@
 
-Based on the copies of original excavation records, we have made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations.
-First, we determined the context with indirect colonial impacts including the Europeans and the Chinese according to the higher frequency of time indicators.
-Second, we examined the description of color and texture changes in strategrphic data to identify the end of the context.
-Third, we checked whether there are radiocarbon ages from the context and if there are any conflicts or ambiguities to decide the usefulness of the ages.
-Fourth, we evaluated if there are any post-depositional issues to exclude problematic data.
-Fifth, we cross evaluated all fourty squares to make sure a consistent stratigraphic context across them.
-Sixth, we cross evaluated our chronology by checking previous studies of assigning chronology to make a final decision [@Hsieh2009; @Wang2011].
-We also provide Supplementary Online Materials that include the details of stratigraphic data, the high frequency of time indicators, and the decision making during the assigning process.
+## Method for assigning ornaments
+
+Based on the copies of original excavation records and fieldwork notes, we have made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations.
+First, we determined the artificial layer (mostly 10 cm) with indirect colonial impacts according to the finding of high frequency of time indicators representing the European and the Chinese period.
+Second, we identified the complete context (normally consists of 2-3 layers) where the layer belongs to based on the description of soil color, soil texture, and unearthed artifacts using original strategrphic data.
+Third, we checked if there are radiocarbon ages associated with the context as a cross validation.
+Since we noted that the radiocarbon ages normally represent long time ranges, the ages are not considered if they cannot provide useful information.
+Fourth, we examined the original fieldwork notes to identify any post-depositional issues that might affect the deposition of time indicators to ensure an appropriate assignment.
+Fifth, we cross evaluated all adjacent squares to confirm a consistent, and continuous context between them.
+Sixth, we compared our chronology with six analytic units classified by previous studies to make a final decision [@Hsieh2009; @Wang2011].
+More detailed data information is in Supplementary Online Materials.
+After assigning the contexts into the pre-European contact, the European presence, and the Chinese presence, we decided the chronology of ornaments by examining in which context they were found.
+Through those assigning steps with validation rules, we believe our chronology is reasonable and suitable for discussing the indirect colonial impacts led by trade networks.
 

### 2020-07-08 analysis/paper/paper.Rmd
edit discussion section based on R3 comment

@@
 There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan.
-Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173.
-There are also 27 ornaments found in midden contexts.
-The obvious difference in number in burials and midden contexts indicates the property of ornaments as prestige goods.
-The high number in burials is due to the presence of bead strands or patterned bands of beads, which sometimes contain thousands of beads in an individual burial [@Chen2007].
-The highly uneven distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008.
-However, @Hsieh2012 suggests a relatively egalitarian society from the comparative analysis of frequency and proportion of burial goods.
+Intact ornaments are commonly found in burials used as personal adornment, indicating the property of ornaments as prestige goods.
+In general, there are some differences between the number of gold-foil beads, carnelian beads, and glass beads across burials.
+For example, over a thousand beads are found in a few burials, but none is found in some burials.
+This is also because of the presence of bead strands or patterned bands of beads, which sometimes contain thousands of beads in an individual burial [@Chen2007].
+The uneven distribution of beads indicate some social differentiation, but a full discussion of the burial data is beyond the scope of this paper.
 
@@
 Although glass beads have less variety in size, presenting as small (0.5-1 cm) or medium (1-2 cm), they have a wide variety of colors or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
+
 Although we are not certain of the specific origin of all kinds of beads, research suggest that glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
@@
 A transition in the origin of glass beads from mainly Southeast Asia to multiple regions including China might indicate different trading networks.
-A regional network in South China Sea in the earlier period and later involved in larger scale trading network stimulated by the Europeans and the Chinese [@Hung2016; @Wang2007].
+A regional network in South China Sea in the earlier period and later involved in a larger scale trading network stimulated by the Europeans and the Chinese [@Hung2016; @Wang2007].
 
@@
 
+# create df for all types of ornaments by units and period
+ornaments_period_join_all_type_across_periods <-
+  ornaments_period_join %>% 
+  group_by(period, Pit_No, geometry) %>% 
+  summarize(n = sum(n))
+
 # plot spatial pattern for all ornaments
 plot_spatial_distribution_all_ornaments_by_period <- 
-  ornaments_period_join %>% 
-  group_by(period, Pit_No, geometry) %>% 
-  summarize(n = sum(n)) %>% 
   ggplot() +
@@
 A further indicator of increased social inequality is a burial dated to the 17th century that included 60 gold-foil beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008].
-@Cheng2008 proposed a more hierarchical structure of Kiwulan based on the found of burial with 60 gold-foil beads that were treated as prestige goods in local community.
-However, @Hsieh's research on data from burial contexts suggested a more egalitarian society since burials with rich goods are elders that indicates accumulated wealth over time instead of inherited.
-Our results focusing on the residential area demonstrated an uneven distribution of prestige ornaments during the European presence that inclined an appearance of social inequality.
-Nevertheless, we still need a comprehensive exploration of other archaeological records to discuss the social structure of Kiwulan.
+Based on the finding of that burial, @Cheng2008 proposed a more hierarchical structure of Kiwulan.
+However, @Hsieh's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicates accumulated wealth over time instead of inherited status.
+Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are opposite.
+Our results can provide an additional insight by focusing on the residential area that demonstrated an uneven distribution of ornaments during the European presence, indicating an increasing social inequality.
+Nevertheless, we still need to explore and compare with other archaeological records comprehensively discuss the social structure of Kiwulan.
 
@@
 @Rubertone2000].
-Resistance could be presented in many forms, but we have no specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
+We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century [].
+Also, an ethnographic photograph clearly presents the weaving practice by a local woman, who wore a string of beads as a necklace [@Mackey1895, pp.
+206-307].
+Displaying ornaments in social events or daily live could be a cultural symbol, and may be reinforced during the foreign presence.
+This could be a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
 Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time.
@@
 
-Micro-scale analysis of the archaeological record at the peripheries of colonial activity offers a opportunity to understand pervasive but not necessarily dominant European colonial processes at remote Indigenous groups [@Trabert2018].
+Analyses of the archaeological record at the peripheries of colonial activity offers a opportunity to understand pervasive but not necessarily dominant European colonial processes at remote Indigenous groups [@Trabert2018].
 Similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Zheng2015].
@@
 Ornaments give insights into the amplification of social inequality stimulated by European colonization.
-It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
+It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their lives or social occasions to display or intensify status differences.
 We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.

### 2020-07-08 analysis/paper/paper.Rmd
minor edit and knit

@@
 Based on the finding of that burial, @Cheng2008 proposed a more hierarchical structure of Kiwulan.
-However, @Hsieh's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicates accumulated wealth over time instead of inherited status.
+However, @Hsieh2012's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicates accumulated wealth over time instead of inherited status.
 Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are opposite.
 Our results can provide an additional insight by focusing on the residential area that demonstrated an uneven distribution of ornaments during the European presence, indicating an increasing social inequality.
-Nevertheless, we still need to explore and compare with other archaeological records comprehensively discuss the social structure of Kiwulan.
+Nevertheless, we still need to explore and compare with other archaeological records to make a better inference to social structure of Kiwulan.
 
@@
 @Rubertone2000].
-We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century [].
-Also, an ethnographic photograph clearly presents the weaving practice by a local woman, who wore a string of beads as a necklace [@Mackey1895, pp.
-206-307].
-Displaying ornaments in social events or daily live could be a cultural symbol, and may be reinforced during the foreign presence.
-This could be a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
+We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century [@Chen1963; @Ke1993].
+Also, an ethnographic photo photographed by @Mackey1895, pp.
+306-307 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece.
+This indicates displaying ornaments is part of their culture, and may be also used in their daily live according to the photo and the archaeological context where ornaments commonly found in the residential area.
+This custom might be reinforced during the foreign presence as a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
 Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time.
@@
 Ornaments give insights into the amplification of social inequality stimulated by European colonization.
-It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their lives or social occasions to display or intensify status differences.
+It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their life or social occasions to display or intensify status differences.
 We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.

### 2020-07-09 analysis/paper/paper.Rmd
Capitalize place name

@@
 
-Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade network could go beyond the temporal boundaries and last longer.
-This is because the use of foreign goods in many Indigenous societies often go through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012].
+Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade networks could go beyond the temporal boundaries and last longer.
+This is because the use of foreign goods in many Indigenous societies often goes through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012].
 Also, we acknowledged that a fine-grained identification of decades of the European presence from the archaeological record is challenging for this study based on three reasons.
@@
 To mitigate the conflict, we separated the ornaments based on the archaeological context that can reflect the acceptance of diagnostic imported goods by their relatively high frequencies as an indicator of indirect colonial impacts.
-That is, we separated the time periods to before and after a certain foreign presence using the complete archaeological context that contains higher frequencies of time indicators instead of their first appearance in a given layer.
-This method is reasonable because it allows not only the flexibility of imported goods entering Indigenous societies depending on the agency of local people, but also the application of quantitative analysis to temporal and spatial changes of imported ornaments.
+That is, we separated the time periods into a certain period with foreign presence using the complete archaeological context that contains higher frequencies of time indicators instead of the first appearance of those indicators in a given layer.
+This method is reasonable because it allows a flexibility of imported goods entering Indigenous societies depending on the agency of local people, and a time framework for quantitative analysis for temporal and spatial changes of imported ornaments.
 
@@
 Based on the copies of original excavation records and fieldwork notes, we have made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations.
-First, we determined the artificial layer (mostly 10 cm) with indirect colonial impacts according to the finding of high frequency of time indicators representing the European and the Chinese period.
-Second, we identified the complete context (normally consists of 2-3 layers) where the layer belongs to based on the description of soil color, soil texture, and unearthed artifacts using original strategrphic data.
-Third, we checked if there are radiocarbon ages associated with the context as a cross validation.
-Since we noted that the radiocarbon ages normally represent long time ranges, the ages are not considered if they cannot provide useful information.
-Fourth, we examined the original fieldwork notes to identify any post-depositional issues that might affect the deposition of time indicators to ensure an appropriate assignment.
+First, for each unit, we determined the artificial layer (mostly 10 cm) with indirect colonial impacts in terms of the higher frequency of time indicators representing the European and the Chinese period.
+Second, we identified the archaeological context (normally consists of 2-3 layers) where the layer belongs to based on the description of soil color, soil texture, and density of potsherds and charcoals.
+We took into account potsherds and charcoals based on the assumption that a larger population can be reflected by the dense distribution of potsherds and charcoals, and Kiwulan was the biggest settlement with a larger population in 1950 [@Nakamura1938, pp.
+12].
+Third, we checked if there are radiocarbon ages associated with the context as a cross-validation.
+Since some radiocarbon ages represent long time ranges, we did not consider the ages that cannot provide useful information.
+Fourth, we explored any post-depositional issues that might affect the deposition of time indicators to ensure an appropriate assignment based on the fieldwork notes and excavation report.
 Fifth, we cross evaluated all adjacent squares to confirm a consistent, and continuous context between them.
 Sixth, we compared our chronology with six analytic units classified by previous studies to make a final decision [@Hsieh2009; @Wang2011].
-More detailed data information is in Supplementary Online Materials.
+More detailed data information is provided in Supplementary Online Materials.
 After assigning the contexts into the pre-European contact, the European presence, and the Chinese presence, we decided the chronology of ornaments by examining in which context they were found.
@@
 This is difficult to rule out completely, but we consider that because there is no continuous increase in ornament frequency over time, we conclude that accumulation and discard of ornaments is not constant.
-Thus, we assume that changes in the abundance of ornaments reflect relatively continuous discard behaviours rather than accumulations due to collecting of heirlooms.
+Thus, we assume that changes in the abundance of ornaments reflect relatively continuous discard behaviors rather than accumulations due to collecting of heirlooms.
 
@@
 
-Although we are not certain of the specific origin of all kinds of beads, research suggest that glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
+Although we are not certain of the specific origin of all kinds of beads, research suggests that glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
 There seem to be no obvious changes in the sources of glass beads or metal ornaments at different phases in the upper component of Kiwulan (1350-1850 AD).
@@
 A transition in the origin of glass beads from mainly Southeast Asia to multiple regions including China might indicate different trading networks.
-A regional network in South China Sea in the earlier period and later involved in a larger scale trading network stimulated by the Europeans and the Chinese [@Hung2016; @Wang2007].
+A regional network in the South China Sea in the earlier period and later involved in a larger scale trading network stimulated by the Europeans and the Chinese [@Hung2016; @Wang2007].
 
@@
 Yilan was involved in complex trading networks both on a regional scale with other Indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish.
-Those trade ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century.
+Those trade ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before the 17th century.
 Later, trade activities became more frequent and intense in the 17th century due to European activities.
@@
 The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to the increasing numbers of Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
-Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
+Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper
 
@@
 Based on the finding of that burial, @Cheng2008 proposed a more hierarchical structure of Kiwulan.
-However, @Hsieh2012's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicates accumulated wealth over time instead of inherited status.
-Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are opposite.
+However, @Hsieh2012's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicate accumulated wealth over time instead of inherited status.
+Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are the opposite.
 Our results can provide an additional insight by focusing on the residential area that demonstrated an uneven distribution of ornaments during the European presence, indicating an increasing social inequality.
-Nevertheless, we still need to explore and compare with other archaeological records to make a better inference to social structure of Kiwulan.
+Nevertheless, we still need to explore and compare with other archaeological records to make a better inference to the social structure of Kiwulan.
 
@@
 In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks.
-In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999].
+In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labor tasks [@Feinman2000; @Siegel1999].
 
@@
 306-307 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece.
-This indicates displaying ornaments is part of their culture, and may be also used in their daily live according to the photo and the archaeological context where ornaments commonly found in the residential area.
+This indicates displaying ornaments is part of their culture, and likely to be also used in their daily life according to the photo and the archaeological context where ornaments commonly found in the residential area.
 This custom might be reinforced during the foreign presence as a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
@@
 
-Analyses of the archaeological record at the peripheries of colonial activity offers a opportunity to understand pervasive but not necessarily dominant European colonial processes at remote Indigenous groups [@Trabert2018].
+Analyses of the archaeological record at the peripheries of colonial activity offers an opportunity to understand pervasive but not necessarily dominant European colonial processes at remote Indigenous groups [@Trabert2018].
 Similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Zheng2015].

### 2020-07-09 analysis/paper/paper.Rmd
knits on BM's computer!

@@
   count(Categories) %>%
+  ungroup %>% 
   inner_join(AD_data, by = c("Pit_No" = "坑號"))
@@
   ornaments_period_join %>% 
-  group_by(period, Pit_No, geometry) %>% 
-  summarize(n = sum(n))
+  group_by(period, 
+           Pit_No
+           #,geometry
+          ) %>% 
+  summarize(n = sum(n)) %>% 
+  ungroup %>% 
+  inner_join(AD_data %>% 
+               select(坑號, geometry), by = c("Pit_No" = "坑號"))
 

### 2020-07-11 analysis/paper/paper.Rmd
finished editing

@@
 
-The chronology of Kiwulan can be divided into two phases represented by a upper component (1350-1850 AD, 600-100 BP) and a lower component (650-1150 AD, 1200-800 BP) based on 32 radiocarbon ages, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
-There is a sterile deposit between two components spanning c.
+Studying change over time at a site like Kiwulan has some distinct challenges.
+First, the archaeological deposits are relatively thin.
+This makes it difficult for excavators to clearly identify distinct chronological phase boundaries.
+In thin deposits, site formation processes can quickly blur the boundaries or completely mix deposits from different time periods.
+Second, the time periods we aim to compare are relatively short, just a few human generations.
+This limits the usefulness of common chronometric methods such as radiocarbon dating, where the age errors are relatively large compared to the periods we are studying.
+Most of the previously obtained radiocarbon ages from Kiwulan do not span the time period we are investigating here (Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data)).
+Third, pericolonial contexts such as Kiwulan can involve non-Western notions and experiences of time.
+For example, a Western concept of time seeks clear temporal divides, and may create material signature of this, but Indigenous experiences of time may prioritize continuity and continuums of gradual change, and their material culture traces may be reflect that intentional experience of time [@Trabert2018; @Scheiber2012].
+Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade networks could go beyond these temporal boundaries and last longer.
+This is because the use of foreign goods in many Indigenous societies often goes through processes of negotiations resistance, and transformation.
+A static time frame may not capture the processes well [@Scheiber2012].
+Fourth, the circulation of foreign goods that we use as chronological indicators might start earlier than the historically documented European presence in Taiwan.
+Similarly, the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010].
+This limits the usefulness of specific artifact types as chronological markers, and implies they have an error range similar to radiocarbon ages.
+Unfortunately, we do not have a way to accurately estimate these errors at Kiwulan.
+Ideally, we would divide the deposits at Kiwulan into many time slices to investigate the possibility of a gradual versus sudden change, and assess errors in the accuracy of artefacts as temporal markers.
+However, this would result in many analytical units with no finds, so we have grouped excavation units together to create minimum sample sizes suitable for addressing our research questions.
+The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Online Materials.
+
+A first step to outlining chronology of Kiwulan is indicated by the stratigraphy.
+There is an upper component (1350-1850 AD, 600-100 BP) and a lower component (650-1150 AD, 1200-800 BP) based on 32 radiocarbon ages, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
+There is a sterile deposit between these two components spanning c.
 150 years that may be related to dry weather, according to pollen analysis, leading to site abandonment [@Chen2007; @Lin2015].
 These component divisions present the differences in color and texture of the deposit, and the content of artifacts such as pottery types.
-Whether these two components suggest a culture continuity from the same people is still under debate [@Chiu2004; @Chen2007].
+Whether these two components suggest a countinuity of culture from the same people is still under debate [@Chiu2004; @Chen2007].
 We focus on the upper component because this component spans the periods when local residents interacted with the Europeans and the Chinese.
@@
 Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the types of chronologically diagnostic ceramics, excavation depth, consistency of contexts, and radiocarbon dates [@Hsieh2009; @Wang2011].
-To explore the changes in ornament assemblages over time associated with foreign impacts, we assigned the original excavation levels into three phases: before the European contact, after the European presence, and the presence of large Chinese immigrants.
-
-Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade networks could go beyond the temporal boundaries and last longer.
-This is because the use of foreign goods in many Indigenous societies often goes through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012].
-Also, we acknowledged that a fine-grained identification of decades of the European presence from the archaeological record is challenging for this study based on three reasons.
-The current radiocarbon ages that are unable to provide a high resolution of time breaks, the circulation of foreign goods might start earlier before the official European presence in Taiwan, and the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010].
-However, those limitations can be minimized with pericolonial approaches that emphasize the wide-ranging effects of the colonial process, allowing discussion of indirect impacts on Indigenous societies with a more practical and dynamic framework not necessarily defined by a fixed colonial period with a start and an end time [@Trabert2018; @Scheiber2012].
-
-We acknowledged the approach to study a continuum of changes of archaeological records over time to avoid the pitfall of temporal divides [@Scheiber2012]; however, we also admitted the convenience of using separated time periods for quantitative comparison between temporal groups of imported ornaments.
-To mitigate the conflict, we separated the ornaments based on the archaeological context that can reflect the acceptance of diagnostic imported goods by their relatively high frequencies as an indicator of indirect colonial impacts.
-That is, we separated the time periods into a certain period with foreign presence using the complete archaeological context that contains higher frequencies of time indicators instead of the first appearance of those indicators in a given layer.
-This method is reasonable because it allows a flexibility of imported goods entering Indigenous societies depending on the agency of local people, and a time framework for quantitative analysis for temporal and spatial changes of imported ornaments.
-
-## Method for assigning ornaments
-
-Based on the copies of original excavation records and fieldwork notes, we have made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations.
+To achieve sample sizes suitable for exploring changes in ornament assemblages over time associated with foreign impacts, we assigned the original excavation levels into three phases: before the European contact, after the European presence, and the presence of large Chinese immigrants.
+
+## Chronological markers in the artefact assemblages
+
+We made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations using copies of original excavation records and fieldwork notes.
 First, for each unit, we determined the artificial layer (mostly 10 cm) with indirect colonial impacts in terms of the higher frequency of time indicators representing the European and the Chinese period.
-Second, we identified the archaeological context (normally consists of 2-3 layers) where the layer belongs to based on the description of soil color, soil texture, and density of potsherds and charcoals.
-We took into account potsherds and charcoals based on the assumption that a larger population can be reflected by the dense distribution of potsherds and charcoals, and Kiwulan was the biggest settlement with a larger population in 1950 [@Nakamura1938, pp.
-12].
+Second, we identified the archaeological context (normally 2-3 layers) where the layer belongs to based on the description of soil color, soil texture, and density of potsherds and charcoal.
+We assumed that a larger population would be reflected by denser distributions of potsherds and charcoal.
 Third, we checked if there are radiocarbon ages associated with the context as a cross-validation.
-Since some radiocarbon ages represent long time ranges, we did not consider the ages that cannot provide useful information.
+Since some radiocarbon ages represent long time ranges, we did not consider the ages could determine a phase by themselves, but only as a cross-validation.
 Fourth, we explored any post-depositional issues that might affect the deposition of time indicators to ensure an appropriate assignment based on the fieldwork notes and excavation report.
-Fifth, we cross evaluated all adjacent squares to confirm a consistent, and continuous context between them.
+Fifth, we cross evaluated all adjacent squares to confirm a consistent and continuous context between them.
 Sixth, we compared our chronology with six analytic units classified by previous studies to make a final decision [@Hsieh2009; @Wang2011].
 More detailed data information is provided in Supplementary Online Materials.
-After assigning the contexts into the pre-European contact, the European presence, and the Chinese presence, we decided the chronology of ornaments by examining in which context they were found.
-Through those assigning steps with validation rules, we believe our chronology is reasonable and suitable for discussing the indirect colonial impacts led by trade networks.
-
-The archaeological indicators of the indirect impact of the Europeans at Kiwulan are the higher frequency of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
+After assigning the contexts into three analytical units, pre-European contact, the European presence, and the Chinese presence, we assigned ornaments to one of these units according to their find context.
+We did not use the ornaments themselves as temporal markers because many of the ornaments have been found throughout long time periods in Taiwan.
+Through those assigning steps with these validation rules, we believe our chronology is reasonable and suitable for discussing the indirect colonial impacts led by trade networks.
+
+Archaeological indicators of indirect impacts of the Europeans at Kiwulan are the higher frequency of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
 Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
@@
 Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrants in 1768 [@Chen1963; @Ke1993].
-The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Online Materials.
 
@@
 Intact ornaments are commonly found in burials used as personal adornment, indicating the property of ornaments as prestige goods.
+The majority burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
 In general, there are some differences between the number of gold-foil beads, carnelian beads, and glass beads across burials.
@@
 This is also because of the presence of bead strands or patterned bands of beads, which sometimes contain thousands of beads in an individual burial [@Chen2007].
-The uneven distribution of beads indicate some social differentiation, but a full discussion of the burial data is beyond the scope of this paper.
+The uneven distribution of beads likely indicates some social differentiation.
+However, the complexity of the taphonomy and chronology of the burial features at Kiwulan mean that a full discussion of these is beyond the scope of this paper.
 
 We focus on ornaments from the habitation contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.
-The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
 A possible limitation to our chronological resolution is that ornaments could be heirlooms inherited over multiple generations and well-preserved for a long time.
-This is difficult to rule out completely, but we consider that because there is no continuous increase in ornament frequency over time, we conclude that accumulation and discard of ornaments is not constant.
-Thus, we assume that changes in the abundance of ornaments reflect relatively continuous discard behaviors rather than accumulations due to collecting of heirlooms.
+This is difficult to rule out completely, but we consider that because there is no continuous increase in ornament frequency over time, we conclude that accumulation and discard of ornaments is not constant, but was affected by contemporary conditions.
+Thus, we assume that changes in the abundance of ornaments reflect changes to otherwise relatively continuous discard behaviors rather than accumulations due to collecting of heirlooms.
 
@@
 
-Although we are not certain of the specific origin of all kinds of beads, research suggests that glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
+Although we are not certain of the specific origin of the kinds of beads found at Kiwulan, the glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
 There seem to be no obvious changes in the sources of glass beads or metal ornaments at different phases in the upper component of Kiwulan (1350-1850 AD).
-However, the glass beads from the lower component (650-1150 AD) demonstrate the composition of m-Na-Al glass and soda plant ash glass that are subtypes of the Indo-Pacific beads, frequently seen in Southeast Asia[@Carter2016; @Francis2002; @Wang2014].
+However, the glass beads from the lower component (650-1150 AD) demonstrate the composition of m-Na-Al glass and soda plant ash glass that are subtypes of the Indo-Pacific beads, frequently seen in Southeast Asia [@Carter2016; @Francis2002; @Wang2014].
 A transition in the origin of glass beads from mainly Southeast Asia to multiple regions including China might indicate different trading networks.
@@
 
-Archaeological contexts show that ornaments are especially abundant in burial contexts serving as grave goods [@Chen2007].
+Archaeological contexts at Kiwulan show that ornaments are especially abundant in burial contexts serving as grave goods [@Chen2007].
 This supports the interpretation of ornaments as valuable objects functioning as status indicators.
@@
 However, @Hsieh2012's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicate accumulated wealth over time instead of inherited status.
-Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are the opposite.
-Our results can provide an additional insight by focusing on the residential area that demonstrated an uneven distribution of ornaments during the European presence, indicating an increasing social inequality.
-Nevertheless, we still need to explore and compare with other archaeological records to make a better inference to the social structure of Kiwulan.
+Although both Cheng and Hsieh use the same burial data, their inferences about Kiwulan social structure are the opposite.
+Our results provide an additional insight by focusing on the residential area which demonstrates an uneven distribution of ornaments during the European presence, indicating increasing social inequality, supporting Cheng’s conclusions.
+Nevertheless, we still need to explore and compare with other archaeological records to make more robust inferences about the social structure of Kiwulan.
 
@@
 
-That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations.
-The increasing number and concentrated spatial patterns of ornaments may also indicate a practice of cultural resistance against the European intrusion.
+That said, the evidence from Kiwulan may be consistent with a variety of scenarios of Indigenous-colonial relations.
+The increasing number and concentrated spatial patterns of ornaments may also suggest a practice of cultural resistance against the European intrusion.
 Resistance to European economic and political demands may be inferred if ornaments were used as a display of social identity and to emphasize the local customs that had existed before European contact [cf.
@@
 We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century [@Chen1963; @Ke1993].
-Also, an ethnographic photo photographed by @Mackey1895, pp.
-306-307 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece.
-This indicates displaying ornaments is part of their culture, and likely to be also used in their daily life according to the photo and the archaeological context where ornaments commonly found in the residential area.
-This custom might be reinforced during the foreign presence as a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
-Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time.
-This process would result in a natural increase in ornaments over time, unrelated to colonial influences.
-As noted above, this pattern is not a good fit for the ornament distribution at Kiwulan, and adds little value in explaining the shifts in spatial patterns.
+Also, an ethnographic photo photographed by @Mackey1895 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece.
+This indicates that the ornaments commonly found in the residential area were part of their daily life and customary displays.
+This custom might be reinforced during the foreign presence as a form of resistance, but we need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
 

### 2020-07-13 analysis/paper/paper.Rmd
changed title

@@
 ---
-title: "Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period"
+title: "Ornaments as indicators of social changes resulting from indirect effects of colonialism in northeastern Taiwan"
 author:
@@
 Third, pericolonial contexts such as Kiwulan can involve non-Western notions and experiences of time.
-For example, a Western concept of time seeks clear temporal divides, and may create material signature of this, but Indigenous experiences of time may prioritize continuity and continuums of gradual change, and their material culture traces may be reflect that intentional experience of time [@Trabert2018; @Scheiber2012].
+For example, a Western concept of time seeks clear temporal divides, and may create material signature of this, but Indigenous experiences of time may prioritize continuity and continuums of gradual change, and their material culture traces may reflect that intentional experience of time [@Trabert2018; @Scheiber2012].
 Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade networks could go beyond these temporal boundaries and last longer.
-This is because the use of foreign goods in many Indigenous societies often goes through processes of negotiations resistance, and transformation.
+This is because the use of foreign goods in many Indigenous societies often goes through processes of negotiations, resistance, and transformation.
 A static time frame may not capture the processes well [@Scheiber2012].

### 2020-09-21 analysis/paper/paper.Rmd
udpate C14 calibration method

@@
 
-                               # Use Bchron to clibrate
+                               # Use Bchron to calibrate
 library(Bchron)
@@
          ids = id,
-         calCurves = rep('intcal13',
+         calCurves = rep('intcal20',
                          length(kw_dates_with_error$rce))))
@@
          ageSds = se,
-         calCurves = rep('intcal13',
+         calCurves = rep('intcal20',
                          length(kw_dates_with_error$rce)),
@@
           by = c(0, 0),
-          byend = c(0.0020),
+          byend = c(0.002),
           bxend = bx)
@@
            label = "Upper component",
-           size = 2) +
+           size = 2.5) +
     annotate("rect", xmin = 1300, xmax = 800,
@@
            label = "Lower component",
-           size = 2)  +
+           size = 2.5)  +
   theme_minimal(base_size = 8) +

### 2020-09-21 analysis/paper/paper.Rmd
updated caption, acknowledgements, and added references

@@
 kableExtra::kable(table_kw_dates,
-             caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve.
+             caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal20 Atmospheric curve.
 All depth is recorded in centimeters above mean sea level.
@@
 We also thank Yu-pei Chen for the discussion regarding the site at the earliest stage of fieldwork.
-This work was supported in part by the Chester Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington.
+This work was supported in part by the Chester Fritz Graduate Fellowship and travel grants from the Department of Anthropology at University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
 We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.

### 2020-09-22 analysis/paper/paper.Rmd
replace figure 3 to avoid text overlapping with curve

@@
 # Use Bchron to calibrate
+# need github version to get intcal20
+# if(!require(remotes)) install.packages('remotes')
+# remotes::install_github('andrewcparnell/Bchron')
 library(Bchron)
@@
        dpi = 300,
-       height = 62,
+       height = 100,
        units = "mm")

### 2020-09-24 analysis/paper/paper.Rmd
udpate abstract and intro based on the proof version

@@
   Long-lasting indirect impacts on Indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia.
-Trade ornaments from Kiwulan (1350-1850 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies.
-The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered.
-This hints at increasing social inequality resulting from colonial influence.
-Ornaments give insights into the increasing social inequality stimulated by the European colonial presence, and show the agency of Indigenous people to incorporate ornaments into their social system.
+Trade ornaments from Kiwulan (1350–1850 CE) in northeastern Taiwan show indirect impacts of European colonial activities on local societies.
+The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered in the residential area.
+The change reflects an increased use of ornaments and an uneven distribution of ornaments in a pericolonial context.
+This may hint at increased social inequality resulting from European colonial influence.
+Trade ornaments give insights into the increased social inequality stimulated by a European colonial presence, and show the agency of Indigenous people to incorporate ornaments into their social system.
 keywords: |
-  Historical archaeology; European colonial period; 17th century; Trade ornaments; Exchange networks; spatial pattern; East Asia
+  Historical archaeology; European colonial period; Trade ornaments; Exchange networks; spatial pattern; East Asia; Taiwan
 highlights: |
@@
 Direct European colonial rule throughout East Asia was rare and limited, and the question of long-lasting indirect impacts on local Indigenous communities remains largely unanswered.
-Understanding the indirect effects of colonialism are important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@Acabado2017; @Trabert2017].
+Understanding these indirect effects is important for detecting colonial impacts on Indigenous peoples in the periphery of colonial control [@Acabado2017; @Trabert2017].
 In many parts of the world, the introduction of foreign trade goods by colonial traders into local Indigenous societies caused substantial transformations of Indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Silliman2005].
 Consumption patterns of foreign goods can give insights into negotiations between colonized and colonizer, and the resistance and accommodations of Indigenous people through their daily cultural practices [@Dietler2015; @Given2004; @Mullins2011; @Scaramelli2005; @Silliman2001; @TorrenceandClarke2000; @Voss2005].
-Northeastern Taiwan is an ideal context to study peripheral colonial influence because although there was a prominent Spanish and Dutch colonial presence in parts of Taiwan, the northeastern region was isolated from intensive direct contact by the Xueshan Mountains.
-
-This article describes personal ornaments excavated from the upper component of Kiwulan (1350 AD-1850 AD), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
+Northeastern Taiwan is an ideal context to study pericolonial influence because although there was a prominent Spanish and Dutch colonial presence in parts of Taiwan, the northeastern region was isolated from intensive direct contact by the Xueshan Mountains.
+
+This article describes personal ornaments excavated from the upper component of Kiwulan (1350 CE-1850 CE), the largest Iron Age settlement on the Yilan plain in northeastern Taiwan.
 Ornaments are found at many Iron Age sites in Yilan, but only Kiwulan shows clear stratigraphic contexts from pre-European period to modern time.
@@
 163].
-In 1647 the Dutch attacked villages and forced them to accept colonial rule and pay an annual tribute [@Andrade2007].
+In 1647 the Dutch attacked villages and forced them to accept colonial rule and pay an annual tribute without strict enforcement of colonial power that allowed flexibility for Indigenous people in Yilan [@Andrade2007; @Kang2016].
 According to Dutch census reports in 1650, Kiwulan was the largest Indigenous settlement in the plain, with a population of 840 adults [@Nakamura1938, pp.
@@
 Figure \@ref(fig:plot-altogether) shows the distribution of frequencies for subtypes in each major class.
-Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
+Spearman's correlation test shows that there is a significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
 This indicates that the increases in diversity can be explained by the effects of culture interaction instead of the effects of sample size.
@@
 The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to the increasing numbers of Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
-Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper
+Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
 

### 2020-09-24 analysis/paper/paper.Rmd

@@
 ```{r tidy-data, eval = TRUE}
-# summarise the ornaments by layers
+# summarize the ornaments by layers
 # L5 and L6 represent pre-European contact period, 

### 2020-09-25 analysis/paper/paper.Rmd
edit based on proof

@@
 ---
-title: "Ornaments as indicators of social changes resulting from indirect effects of colonialism in northeastern Taiwan"
+title: "Trade ornaments as indicators of social changes resulting from indirect effects of colonialism in northeastern Taiwan"
 author:
@@
 In this paper, we explore archaeological ornaments from Kiwulan spanning the pre-European contact period, period after the Spanish and Dutch presence, and the following period after the Chinese presence.
-We address the question of whether indirect colonial influences on the Indigenous populations can be detected through the ornament assemblages.
-
-# Ornaments in complex exchange network during the late Iron Age and early historical period
+We address the question of whether indirect colonial influences on the local Indigenous populations can be detected through the ornament assemblages.
+
+# Ornaments in exchange networks during the late Iron Age and early historical period
 
 The island of Taiwan lies at the junction of mainland China, Southeast Asia, and Northeast Asia in the Pacific Ocean.
-The prehistory of Taiwan island could be roughly divided into three major periods, Palaeolithic (c.
-27,000 BP- 5000BP), Neolithic (c.
-6500- 2000BP), and Iron age (c.
-2000- 400BP) with slight regional differences in onset of each period and variations in style of artifacts and assemblages [@Chen2017; @Liu2011].
+The prehistory of Taiwan can be roughly divided into three major periods, Palaeolithic (c.
+27,000- 5000 BP), Neolithic (c.
+6500- 2000 BP), and Iron age (c.
+2000- 400 BP) with slight regional differences in the onset of each period and variations in style of artifacts and assemblages [@Chen2017; @Liu2011).
 It is generally accepted that Taiwan entered the historical period in the early 17th century due to the colonial activities of the Spanish and the Dutch who played an important role in keeping written records about Taiwan.
@@
 Later in 1683, Taiwan was incorporated into the Qing dynasty in China and a large wave of Han Chinese migrated to Taiwan during the late 18th century.
-Because of natural safe harbors, northeastern Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later the global trade network with the Europeans in the 17th century brought more trade goods circulated in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
-Although located on the periphery of regional trade centers, Yilan was connected to trade networks via visits of other Indigenous groups, Chinese merchants, and Europeans, via sea.
-
-The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Heping Dao, Keelung in 1626, and Fort San Domingo in 1629 at Tamsui (Figure \@ref(fig:KWL-map)).
+Because of natural safe harbors, northeastern Taiwan was involved in a regional trade network through cross-culture interactions with Chinese merchants since the 14th century, and later global trade networks with Europeans in the 17th century brought more trade goods circulating in Southeast Asia into Taiwan [@Chen2005; @Liu2017].
+Although located on the periphery of regional trade centers, Yilan was connected by sea to trade networks via visits of other Indigenous groups, Chinese merchants, and Europeans.
+
+The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador in 1626 at Heping Dao, Keelung [@Berrocal2020], and Fort San Domingo in 1629 at Tamsui (Figure \@ref(fig:KWL-map)).
 They sent missionaries to local Indigenous settlements in this region [@Blusse2000, pp.
@@
 According to those records, Indigenous people in Yilan wore ornaments in ceremonial contexts to display their wealth and status.
-Among those ornaments, fish-shaped necklaces made of metal threads had high value due to their delicacy and the exotic materials invested in production.
+Among those ornaments, fish-shaped necklaces made of metal threads had high value due to their delicacy and the exotic materials used in production.
 These were usually possessed by wealthy people.
 Other people wore carnelian beads or glass beads on their head or neck to participate in ceremonies.
-In 1895, at the beginning of Japanese colonization, an academic field survey for plains Indigenous groups reported that fish-shaped metal necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
+In 1895, at the beginning of the Japanese colonization of Taiwan, an academic field survey for Indigenous groups reported that fish-shaped metal necklaces were not used in Yilan at that time, but elderly people still used beads [@Ino1996, pp.
 227-232].
@@
 There is no direct evidence showing European delivery of beads, however, a large amount of the glass beads containing gold foil (hereafter, gold-foil beads) at Kiwulan might have been introduced by the Spanish through economic activities because similar beads were found at Luzon, northern Philippines, as part of the trading route of the Spanish between 16-19th century [@Wang2007].
-Both archaeological evidence and historical records indicate northeastern Taiwan was involved in regional networks with East Asia in the late Iron age.
+Both archaeological evidence and historical records indicate northeastern Taiwan was involved in regional networks with East and Southeast Asia in the late Iron age.
 These included Chinese merchants trading metal items, clothes, and beads with local Indigenous people in Taiwan in exchange for local resources.
@@
 
-```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text. Map data from naturalearthdata.com", fig.width=image_width_2_col_inch}
+```{r, KWL-map, fig.cap = "Map (from Wang and Marwick, 2020) showing the location of Kiwulan, and other places in northern Taiwan named in the text.
+Map data from naturalearthdata.com", fig.width=image_width_2_col_inch}
 library(here)
@@
 
-Archaeological ornaments from Kiwulan (Figure \@ref(fig:KWL-map)) come from a rescue archaeology project that was conducted between 2001-2004 in advance of a water diversion project and road bridge construction.
+Archaeological ornaments from Kiwulan (Figure \@ref(fig:KWL-map)) come from a rescue archaeology project that was conducted between 2001 and 2004 in advance of a water diversion project and road bridge construction.
 The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007].
-The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement.
-Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance.
+The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicate it was a long-term settlement.
+Artefact locations were mostly recorded to the 2 x 2 m sub-square they were recovered in.
+Except for a handful of finds, artefacts lack individual point provenance information.
 Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of the A and D sections, which is also the study area where our samples come from (Figure \@ref(fig:KWL-sam-area)).
@@
 
-```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID numbers.
+```{r, KWL-sam-area, fig.cap = "Map (from Wang and Marwick, 2020) showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID numbers.
 Small dots represent the locations of post-holes.
@@
 Similarly, the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010].
-This limits the usefulness of specific artifact types as chronological markers, and implies they have an error range similar to radiocarbon ages.
+This limits the usefulness of specific artefact types as chronological markers, and implies they have an error range similar to radiocarbon ages.
 Unfortunately, we do not have a way to accurately estimate these errors at Kiwulan.
@@
 A first step to outlining chronology of Kiwulan is indicated by the stratigraphy.
-There is an upper component (1350-1850 AD, 600-100 BP) and a lower component (650-1150 AD, 1200-800 BP) based on 32 radiocarbon ages, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
+There is an upper component (1350-1850 CE, 600-100 BP) and a lower component (650-1150 CE, 1200-800 BP) based on 32 radiocarbon ages, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data).
 There is a sterile deposit between these two components spanning c.
 150 years that may be related to dry weather, according to pollen analysis, leading to site abandonment [@Chen2007; @Lin2015].
-These component divisions present the differences in color and texture of the deposit, and the content of artifacts such as pottery types.
-Whether these two components suggest a countinuity of culture from the same people is still under debate [@Chiu2004; @Chen2007].
+These component divisions present the differences in the colours and textures of the deposit, and the content of artifacts such as pottery types.
+Whether these two stratigraphic components suggest a countinuity of culture from the same people is still under debate [@Chiu2004; @Chen2007].
 We focus on the upper component because this component spans the periods when local residents interacted with the Europeans and the Chinese.
@@
 Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the types of chronologically diagnostic ceramics, excavation depth, consistency of contexts, and radiocarbon dates [@Hsieh2009; @Wang2011].
-To achieve sample sizes suitable for exploring changes in ornament assemblages over time associated with foreign impacts, we assigned the original excavation levels into three phases: before the European contact, after the European presence, and the presence of large Chinese immigrants.
+However, to achieve sample sizes suitable for exploring changes in ornament assemblages over time associated with foreign impacts, we assigned the original excavation levels into three phases: before the European contact, after the European presence, and the presence of large Chinese immigrants.
 
@@
 
-We made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations using copies of original excavation records and fieldwork notes.
-First, for each unit, we determined the artificial layer (mostly 10 cm) with indirect colonial impacts in terms of the higher frequency of time indicators representing the European and the Chinese period.
-Second, we identified the archaeological context (normally 2-3 layers) where the layer belongs to based on the description of soil color, soil texture, and density of potsherds and charcoal.
+We made refinements of the original chronology to assign contexts into the pre-European, European, and Chinese periods by following six steps of assigning and evaluating contexts using copies of original excavation records and fieldwork notes.
+First, for each excavation unit, we assigned the arbitrary excavation layer (mostly 10 cm) to a period based on frequencies of indicators of the European and the Chinese periods.
+Second, we identified the archaeological context (normally 2-3 layers) that the layer belongs to based on the description of soil colour, texture, and density of potsherds and charcoal.
 We assumed that a larger population would be reflected by denser distributions of potsherds and charcoal.
 Third, we checked if there are radiocarbon ages associated with the context as a cross-validation.
-Since some radiocarbon ages represent long time ranges, we did not consider the ages could determine a phase by themselves, but only as a cross-validation.
+Since most radiocarbon ages represent long time ranges, we did not consider the ages could determine a period by themselves, but only as a cross-validation.
 Fourth, we explored any post-depositional issues that might affect the deposition of time indicators to ensure an appropriate assignment based on the fieldwork notes and excavation report.
-Fifth, we cross evaluated all adjacent squares to confirm a consistent and continuous context between them.
-Sixth, we compared our chronology with six analytic units classified by previous studies to make a final decision [@Hsieh2009; @Wang2011].
-More detailed data information is provided in Supplementary Online Materials.
-After assigning the contexts into three analytical units, pre-European contact, the European presence, and the Chinese presence, we assigned ornaments to one of these units according to their find context.
+Fifth, we cross-evaluated all adjacent squares to confirm a consistent and continuous context between them.
+Sixth, we compared our chronology with the six analytic units classified by previous studies to make a final decision [@Hsieh2009; @Wang2011].
+More detailed data information is provided in our Supplementary Online Materials.
+After assigning the contexts into three analytical units, pre-European contact, European presence, and Chinese presence, we assigned ornaments to one of these units according to their find context.
 We did not use the ornaments themselves as temporal markers because many of the ornaments have been found throughout long time periods in Taiwan.
-Through those assigning steps with these validation rules, we believe our chronology is reasonable and suitable for discussing the indirect colonial impacts led by trade networks.
-
-Archaeological indicators of indirect impacts of the Europeans at Kiwulan are the higher frequency of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
-Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages.
-The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Dizon2016; @Hsieh1995].
-Southeast China is assumed to be the origin of the light gray glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, such as the Zeelandia fort site in Tainan [@Wang2007].
+Through those six assigning steps and validation rules, we believe our chronology is reasonable and suitable for discussing the indirect colonial impacts led by trade networks.
+
+The primary archaeological indicators of indirect impacts of Europeans at Kiwulan are the high frequencies of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century.
+Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages [@Cort2017; @Ketel2011].
+The earliest evidence of light grey glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 CE [@Dizon2016; @Hsieh1995].
+Southeast China is assumed to be the origin of the light grey glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, especially at An-ping in Tainan, where the Zeelandia fort site is located [@Wang2007].
 The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131].
-We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by Indigenous groups via regional networks in north Taiwan.
+We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by Indigenous groups via regional networks in northern Taiwan.
 In any case, the high volume of ceramics transported by Europeans, and their high mobility in the shipping trade played an important role in introducing foreign jars to Taiwan.
@@
 Smoking is likely to have been introduced by Europeans.
-This custom was widely adopted in many European countries in the 16th century and spread to other regions through interactions [@Uccar2019].
+This custom was widely adopted in many European countries in the 16th century and spread to other regions [@Uccar2019].
 We found that the presence of pipe bowls in the archaeological record here is consistent with distributions of glazed jar fragments, which are far more numerous and widespread across the site (n = 1685).
@@
 However, the archaeological evidence shows that the layers with abundant trade ceramics match the 17th century according to the excavation report and previous studies [@Hsieh2009; @Wang2011].
-Thus, we focus on identifying the contexts with high frequencies of those ceramics as indicators the early 17th century.
+Thus, we focus on identifying the contexts with high frequencies of those ceramics as indicators of the early 17th century.
 
@@
 Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles used by Chinese [@Hsieh2009].
-Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrants in 1768 [@Chen1963; @Ke1993].
+Chinese migrations to Yilan were also recorded in official Chinese documents written in the early 19th century recording the first immigrants in 1768 [@Chen1963; @Ke1993].
 
@@
 This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the main habitation areas of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)).
-Occupation floors were not identified during excavation.
+Occupation floors were roughly identified according to the position and depth of posts during excavation.
 We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site.
 There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan.
-Intact ornaments are commonly found in burials used as personal adornment, indicating the property of ornaments as prestige goods.
+Intact ornaments are commonly found in burials used as personal adornment, indicating the property of ornaments as prestige goods [@Wang2011].
 The majority burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods.
-In general, there are some differences between the number of gold-foil beads, carnelian beads, and glass beads across burials.
-For example, over a thousand beads are found in a few burials, but none is found in some burials.
+In general, there are large differences in the number of gold-foil beads, carnelian beads, and glass beads across burials.
+For example, over a thousand beads are found in a few burials, but none are found in some burials.
 This is also because of the presence of bead strands or patterned bands of beads, which sometimes contain thousands of beads in an individual burial [@Chen2007].
@@
 A: carnelian beads, B: glass beads, C: gold-foil beads, D: bells, E: metal rings.
-Photographs are presented in the same order as those subtypes in the table but from left to right instead.
+Photographs are presented in the same order as in Table 2.
 The photographs of B, C, D, E classes are from the original excavation report (Chen 2007)", echo=FALSE}
@@
 
-```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts totalizaing three periods", fig.width=image_width_1_col_inch}
+```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artefact counts for all three periods", fig.width=image_width_1_col_inch}
 # need to reorder, try https://forcats.tidyverse.org/reference/index.html
@@
 knitr::kable(ornaments_shape,
-             caption = "Frequencies of ornaments by subtype at Kiwulan. Frequency represents artifact counts ")
+             caption = "Frequencies of ornaments by subtype at Kiwulan. Frequency represents artefact counts ")
 
@@
 
-To enable re-use of materials and improve reproducibility and transparency [@Marwick2017], the entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at http://doi.org/10.17605/OSF.IO/R8YGA.
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at http://doi.org/10.17605/OSF.IO/R8YGA to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
@@
 Spearman's correlation test shows that there is a significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$).
-This indicates that the increases in diversity can be explained by the effects of culture interaction instead of the effects of sample size.
+This indicates that the increases in diversity can be explained by a combination of the effects of culture interaction and the effects of sample size.
 Carnelian beads and metal rings have greater quantity and variety of shapes compared to copper bells and glass beads during the European period.
@@
 In contrast, copper bells have less variety, typically >2 cm long with a wide variety of human faces as a motif.
-Although glass beads have less variety in size, presenting as small (0.5-1 cm) or medium (1-2 cm), they have a wide variety of colors or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
+Although glass beads have less variety in size, presenting as small (0.5-1 cm) or medium (1-2 cm), they have a wide variety of colours or patterns mostly made by a winding technique and with high lead content indicating possibly from China [@Cheng2008].
 
 Although we are not certain of the specific origin of the kinds of beads found at Kiwulan, the glass beads and metal ornaments have similar production techniques and composition to those found in China, while gold-foil beads may relate to the economic activities of the Spanish in the South China Sea region [@Chen2011; @Wang2018].
-There seem to be no obvious changes in the sources of glass beads or metal ornaments at different phases in the upper component of Kiwulan (1350-1850 AD).
-However, the glass beads from the lower component (650-1150 AD) demonstrate the composition of m-Na-Al glass and soda plant ash glass that are subtypes of the Indo-Pacific beads, frequently seen in Southeast Asia [@Carter2016; @Francis2002; @Wang2014].
+There seem to be no obvious changes in the sources of glass beads or metal ornaments at different phases in the upper component of Kiwulan (1350-1850 CE).
+However, the glass beads from the lower component (650-1150 CE) demonstrate the composition of m-Na-Al glass and soda plant ash glass that are subtypes of the Indo-Pacific beads, frequently seen in Southeast Asia [@Carter2016; @Francis2002; @Wang2014].
 A transition in the origin of glass beads from mainly Southeast Asia to multiple regions including China might indicate different trading networks.
@@
 Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of all ornaments from the research area for each time period.
-For deposits predating the European arrival, a greater amount of ornaments were found at the northern and middle parts of the research area.
+For deposits predating European arrival, a greater amount of ornaments was found at the northern and middle parts of the research area.
 In European period deposits, ornaments were more widespread, with some clusters on the northern part.
@@
 However, there seems to be no consistent pattern across those different ornaments.
-Each class shows its own pattern where the squares with higher numbers of ornaments distributed separately and independently.
+Each class shows its own pattern where the squares with higher numbers of ornaments are distributed separately and independently.
 For example, a cluster of gold-foil beads was found at the northern part, while a cluster of carnelian beads was found in the middle part.
@@
 
-```{r plot-all-densities, fig.cap= "Kernel density map for ornaments by periods, using a bandwidth based on Silverman (1986)'s rule of thumb"}
+```{r plot-all-densities, fig.cap= "Kernel density map showing hotspots for ornaments by periods, using a bandwidth based on Silverman (1986)'s rule of thumb"}
 # apply function to each data frame in the list
@@
 
-The distribution and density of prestige goods across the residential area can provide information on social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes [@Kintigh1982].
+The distribution and density of prestige goods across the residential area can provide information on consumption, implying social structure based on the assumption that distribution patterns observed from spatial data can reveal cultural processes [@Kintigh1982].
 The differential accumulation of artifacts, especially high value goods, in many archaeological contexts can reflect social differentiation or hierarchy in a society  [@Halstead1993; @Orser1988; @Pearson1993; @Trubitt2003; @Wason2004].
-We used point pattern analysis to assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific households that might result from social inequality stimulated by a colonial presence.
-To prepare the ornament location data for point pattern analysis, we assigned each ornament to a random coordinate pair in the square it was recovered from because artifacts from Kiwulan lack exact piece-provenance data.
+We used point pattern analysis to assess whether the distribution of artifacts represents hotspots produced by non-random processes [@Bevan2016; @Ducke2015], such as concentrations of ornaments in specific locations in the residential area that might result from social inequality stimulated by a colonial presence.
+To prepare the ornament location data for point pattern analysis, we assigned each ornament to a random coordinate pair in the square it was recovered from, because most artifacts from Kiwulan lack exact piece-provenance data.
 The next step was to divide the ornaments into three time periods.
@@
 Density values of artifacts per square meter were calculated for each cell.
-Here we use kernel density estimation (KDE) for visualization and identification of spatial clusters [@Baxter1997], in this case the core areas of ornaments and surrounding neighborhoods.
+Here we used kernel density estimation (KDE) for visualization and identification of spatial clusters [@Baxter1997], in this case the core areas of ornaments and surrounding neighborhoods.
 KDE is a method of spatial analysis that computes the probability of the density of ornaments across space by creating a continuous, smooth density surface across space [@Bonnier2019; @Cortegoso2016].
@@
 There are three consistent sub-regions with a core area that shifts over time.
-The distribution might indicate an increase and decrease in the number of social groups who possessed ornaments.
+The distribution might indicate an increase and decrease in the number of social groups who possessed more ornaments.
 The multiple groups during the European period might reflect unequal consumption of ornaments across the site, relative to other periods, or random patterns resulting from a bigger sample size.
@@
 X-axis values represent ANN expected values under a completely random process resulting from a simulated pattern.
-Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value"}
+Each distribution presents the null hypothesis with the blue line indicating the observed ANN value for the sample from each period"}
 
@@
 A similar, but less extreme, result is also observed during the pre-European period.
-The observed distribution of ornaments is more similar to the random distributions during the Chinese period, with about one third of the simulated values are greater than our observed ANN value.
+The observed distribution of ornaments is more similar to the random distributions during the Chinese period, with about one third of the simulated values greater than our observed ANN value.
 The Chinese period has fewer artifacts in any category, likely reflecting a smaller population at Kiwulan at this time, making spatial patterns and hotspots difficult to discern with confidence.
-Our Monte Carlo testing reveals that clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or a concentration of power to control the distribution of ornaments at Kiwulan during this period.
+Our Monte Carlo testing reveals that clustering of ornaments during the European period is highly non-random, potentially indicating different degrees of access to foreign ornaments or an intention to accumulate ornaments at Kiwulan during this period.
 
@@
 This may be due to a smaller scale of trading networks, the overall decline of Indigenous populations in Yilan, or the adoption of Han Chinese practices.
-The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to the increasing numbers of Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
+The decline of the population at Kiwulan may be related to the movement of many Indigenous people southwards to Hualien due to increasing numbers of Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].
 Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.
@@
 Based on the finding of that burial, @Cheng2008 proposed a more hierarchical structure of Kiwulan.
-However, @Hsieh2012's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicate accumulated wealth over time instead of inherited status.
+However, @Hsieh2012's research on burial data suggested a more egalitarian society since the few burials with rich goods are elders that indicate accumulated wealth over time instead of an inherited status.
 Although both Cheng and Hsieh use the same burial data, their inferences about Kiwulan social structure are the opposite.
 Our results provide an additional insight by focusing on the residential area which demonstrates an uneven distribution of ornaments during the European presence, indicating increasing social inequality, supporting Cheng’s conclusions.
-Nevertheless, we still need to explore and compare with other archaeological records to make more robust inferences about the social structure of Kiwulan.
-
-How might these results fit into a bigger picture of social change at periphery of colonial systems?
+Nevertheless, we still need to explore and compare with other archaeological evidence to make more robust inferences about the social structure of Kiwulan.
+
+How might these results fit into a bigger picture of social change in a pericolonial context?
 We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Ames2010; @Bowles2010; @Drennan2010; @Feinman2000].
-The corporate/network model proposed expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power [@Feinman2000].
+Feinman's corporate/network model expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power [@Feinman2000].
 In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks.
@@
 
-The Kiwulan ornament data may be interpreted as indicating that Yilan social organization moved from a corporate mode, before the European arrival, to a network mode during European presence.
+The Kiwulan ornament data may be interpreted as indicating that Yilan social organization moved from a corporate mode, before European arrival, to a network mode during European presence.
 The changes from a less concentrated to a more concentrated distribution of ornaments before and after the presence of Europeans appears consistent with the shift from shared power and wealth to accumulated wealth and monopolization.
 One possible explanation for the shift could be the long-distance trade network introduced by Europeans.
-The rarity and the image of colonial power of foreign trade goods resulted in the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting them [@Boone1992; @Brumfiel1994; @Clark1994].
-Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
+The rarity and the image of colonial power of foreign trade goods may have resulted in the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting them [@Boone1992; @Brumfiel1994; @Clark1994].
+Because of weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
 
@@
 @Rubertone2000].
-We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century [@Chen1963; @Ke1993].
-Also, an ethnographic photo photographed by @Mackey1895 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece.
+We know of the use of ornaments as a social signal from the custom of wearing ornaments for attending ceremonial events, as documented by Chinese historical records in the 19th century [@Chen1963; @Ke1993].
+Also, an ethnographic photo taken by @Mackey1895 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece.
 This indicates that the ornaments commonly found in the residential area were part of their daily life and customary displays.
-This custom might be reinforced during the foreign presence as a form of resistance, but we need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence.
+This custom might have been reinforced during the foreign presence as a form of resistance.
+Analysis of pottery from Kiwulan shows subtle changes in shape, but not in standardization from the pre-European to European periods, this consistency may have been another way that people of Kiwulan demonstrated resistance to foreign influence [@WangandMarwick2020].
+However, more specific evidence from Kiwulan is necessary to prefer resistance, instead of colonial influence, as the primary mechanism behind the distribution of ornaments.
 
@@
 
-Analyses of the archaeological record at the peripheries of colonial activity offers an opportunity to understand pervasive but not necessarily dominant European colonial processes at remote Indigenous groups [@Trabert2018].
-Similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Zheng2015].
+Analyses of the archaeological record at the peripheries of colonial activity offer an opportunity to understand pervasive, but not necessarily dominant, European colonial processes at remote Indigenous groups [@Trabert2018].
 Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks.
-Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the reach of direct European colonial occupation.
+Kiwulan provides valuable insights into the discussion of pericolonial situations of local societies living beyond the reach of direct European colonial occupation.
 The frequency and spatial distribution of personal ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods.
-The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context.
-Before European contact, ornaments were traded into local Indigenous societies via the regional exchange network with Chinese merchants, and viewed as prestige goods in the local Indigenous culture.
+The greater amount and diversity of ornament types during the European period reflects an increased use of ornaments in a colonial context.
+Before European contact, ornaments were traded into local Indigenous societies via regional exchange networks with Chinese merchants, and viewed as prestige goods in the local Indigenous culture.
 After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signaling wealth and privileged trading connections among the inhabitants of Kiwulan.
 This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.
-This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of cultural tradition, but additional evidence is required to confirm this.
+This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of the local cultural tradition, but additional evidence is required to confirm this.
 
@@
 Ornaments give insights into the amplification of social inequality stimulated by European colonization.
-It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their life or social occasions to display or intensify status differences.
+They also show the agency of Indigenous people to incorporate ornaments into their social system and use them to display or intensify status differences.
 We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities.
-We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments such as X-ray fluorescence analysis would provide more information to construct a clear picture of complex trade networks during this periods.
+We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments such as X-ray fluorescence analysis may provide more information to construct a clearer picture of trade networks during these periods.
 
